### 2020-04-02 material/analysis/ioi.Rmd
check non-expressive performances
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'IOIs Analysis for Non-Expressive performances'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for a tempo (IOIs).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
df_ioi <- fread("./filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

# assign RowNr
df_ioi$RowNr <- rep(1:72, nrow(df_ioi)/72)
```

```{r normIOIs, include = FALSE}
# calculate normIOIs
df_ioi$IOI <- diff(c(0, df_ioi$TimeStamp))

# convert bpm to ms
df_ioi[Tempo == 120]$Tempo <- 250
df_ioi[Tempo == 110]$Tempo <- 273
df_ioi[Tempo == 100]$Tempo <- 300

# normalise IOIs
df_ioi$normIOI <- df_ioi$IOI/df_ioi$Tempo

# remove the first note
df_ioi <- df_ioi[RowNr != 1]

# assign Interval
df_ioi$Interval <- rep(1:71, nrow(df_ioi)/71)

# remove irrelevant intervals
ls_irrelevant <- list(c(25:28), c(32:35), c(51:52), c(56:57), c(65:71))
df_ioi$EighthNote <- "Yes"

for (phrase in 1:length(ls_irrelevant)){
  for (note in 1:length(ls_irrelevant[[phrase]])){
    df_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
  }
}
```

# IOI
## 1. All - individual
- Average normIOIs for each participant.

```{r ioi, echo = FALSE}
# For each individual
ioi <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
print(ioi)
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the ideal tempo

```{r ioi_bar,  echo = FALSE}
p_ioi <- ggplot(data = ioi, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Non-Expression")
p_ioi
```

## 2. All - group
- Average normIOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi_all, echo = FALSE}
# Group mean
ioi_all <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI))]
print(ioi_all)
```

## Sequence plot
```{r seq, fig.width = 10, echo = FALSE}
# For each individual
ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), .(SubNr, Interval)]

p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, colour = SubNr)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
p_ioi_seq
```

## Exclude outlier participants
```{r outlier, echo = FALSE}
excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*sd(ioi$SD) & Mean > mean(ioi$Mean)-3*sd(ioi$SD)]
print(excluded_ioi)
```

### 2020-04-02 material/analysis/ioi.Rmd
exclude participant3
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*sd(ioi$SD) & Mean > mean(ioi$Mean)-3*sd(ioi$SD)]
+excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*mean(ioi$SD) & Mean > mean(ioi$Mean)-3*mean(ioi$SD)]
 print(excluded_ioi)

### 2020-04-02 material/preprocessor/function.R
replace dplyr descriptions with data.table
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -11,7 +11,7 @@ pitch_remover <- function(data, ideal){
   for (subnr in unique(data$SubNr)){ #set # of participants
     print(sprintf("---SubNr %i---", subnr))
     for (trial in unique(unique(data[SubNr == subnr]$TrialNr))){
-      current <- data %>% dplyr::filter(SubNr == subnr &  TrialNr == trial)
+      current <- data[SubNr == subnr & TrialNr == trial]
       if (nrow(current) != 0){ #if current data is not empty
         if (length(ideal$Pitch) != length(current$Pitch)){ #if # of onsets/offsets is not equal to ideal performance
           ls_removed <- c(ls_removed, list(c(subnr, trial, "NoteNr error")))

### 2020-04-02 material/analysis/ioi_onset.Rmd
this is only for onsets
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-02 material/analysis/ioi.Rmd
@@
 ---
-title: 'IOIs Analysis for Non-Expressive performances'
+title: 'IOIs Analysis for Non-Expressive performances (ONSET)'
 output:
@@
 ```{r extract, include = FALSE}
-df_ioi <- fread("./filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+df_ioi <- fread("./../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 ```
+
+```{r export, include = FALSE}
+# export df_ioi
+write.csv(df_ioi, file = "./data_ioi_onset.csv", row.names = F)
+```

### 2020-04-02 material/analysis/ioi_offset.Rmd
checking IOI offsets
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-02 material/analysis/ioi.Rmd
@@
 ---
-title: 'IOIs Analysis for Non-Expressive performances'
+title: 'IOIs Analysis for Non-Expressive performances (OFFSET)'
 output:
@@
 ```{r extract, include = FALSE}
-df_ioi <- fread("./filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+df_ioi <- fread("./../preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 ```
+
+```{r export, include = FALSE}
+# export df_ioi
+write.csv(df_ioi, file = "./data_ioi_offset.csv", row.names = F)
+```

### 2020-04-07 material/analysis/stim_n/ioi_duration.Rmd
edit .gitignore
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis for Non-Expressive performances (DURATION)'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a rough overview of non-expressive performance from teaching-v2.0 (duration).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
# read a filtered csv
df_onset <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
df_offset <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")

# SubNr as a factor
df_onset$SubNr <- as.factor(df_onset$SubNr)
df_offset$SubNr <- as.factor(df_offset$SubNr)

# assign RowNr
df_onset$RowNr <- rep(1:72, nrow(df_onset)/72)
df_offset$RowNr <- rep(1:72, nrow(df_offset)/72)
```

```{r normDus, include = FALSE}
# extract trials which have both onset and offset data
df_onset$Eval <- "NA"
for (subnr in unique(df_offset$SubNr)){
  for (trial in unique(df_offset[SubNr == subnr]$Trial)){
    df_onset[SubNr == subnr & TrialNr == trial]$Eval <- "Yes"
  }
}
df_ioi <- df_onset[Eval == "Yes"]

# check whether each row corresponds correctly
df_ioi$Eval2 <- "NA"
for (i in 1:nrow(df_ioi)){
  if (df_ioi[i]$SubNr == df_offset[i]$SubNr & df_ioi[i]$TrialNr == df_offset[i]$TrialNr & df_ioi[i]$Pitch == df_offset[i]$Pitch){
    df_ioi[i]$Eval2 <- "SAME"
  } else {
    df_ioi[i]$Eval2 <- "DIFF"
  }
}
print(unique(df_ioi$Eval2))

# calculate duration
df_ioi$Duration <- df_offset$TimeStamp - df_ioi$TimeStamp

# convert bpm to ms
df_ioi[Tempo == 120]$Tempo <- 250
df_ioi[Tempo == 110]$Tempo <- 273
df_ioi[Tempo == 100]$Tempo <- 300

# normalised Duration
df_ioi$normDu <- df_ioi$Duration/df_ioi$Tempo

# remove irrelevant intervals
ls_irrelevant <- list(c(27:28), c(33:35), 52, 57, c(66:72))
df_ioi$EighthNote <- "Yes"

for (phrase in 1:length(ls_irrelevant)){
  for (note in 1:length(ls_irrelevant[[phrase]])){
    df_ioi[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
  }
}
```

## Sequence plot
```{r seq, fig.width = 10, echo = FALSE}
# For each individual
ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), .(SubNr, RowNr)]

p_ioi_seq <- ggplot(data = ioi_seq, aes(x = RowNr, y = Mean, colour = SubNr)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "RowNr", y = "Normalised Duration") + scale_x_continuous(breaks=seq(1,66,1))
p_ioi_seq
```

# IOI
## 1. All - individual
- Average normDu for each participant.

```{r ioi, echo = FALSE}
# For each individual
ioi <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr)]
print(ioi)
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the ideal tempo

```{r ioi_bar,  echo = FALSE}
p_ioi <- ggplot(data = ioi, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Normalised Duration", subtitle = "Duration: Non-Expression")
p_ioi
```

## 2. All - group
- Average normDus for each condition.

```{r ioi_all, echo = FALSE}
# Group mean
ioi_all <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu))]
print(ioi_all)
```

## Exclude outlier participants
```{r outlier, echo = FALSE}
excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*mean(ioi$SD) & Mean > mean(ioi$Mean)-3*mean(ioi$SD)]
print(excluded_ioi)
```

```{r export, include = FALSE}
# export df_ioi
write.csv(df_ioi, file = "./data_duration.csv", row.names = F)
```

### 2020-04-07 material/analysis/stim_n/ioi_offset.Rmd
edit .gitignore
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-02 material/analysis/ioi_offset.Rmd
@@
 ---
-title: 'IOIs Analysis for Non-Expressive performances (OFFSET)'
+title: 'Analysis for Non-Expressive performances (OFFSET)'
 output:
@@
 
-Description: This is a summary of all analyses for a tempo (IOIs).
+Description: This is a rough overview of non-expressive performance from teaching-v2.0 (offsets).
 
@@
 ```{r extract, include = FALSE}
-df_ioi <- fread("./../preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+df_ioi <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 ## 2. All - group
-- Average normIOIs for each condition.
-- Hypothesis: Participants will play slower when teaching.
+- Average normIOIs
 

### 2020-04-07 material/analysis/stim_n/ioi_onset.Rmd
edit .gitignore
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-02 material/analysis/ioi_onset.Rmd
@@
 ---
-title: 'IOIs Analysis for Non-Expressive performances (ONSET)'
+title: 'Analysis for Non-Expressive performances (ONSET)'
 output:
@@
 
-Description: This is a summary of all analyses for a tempo (IOIs).
+Description: This is a rough overview of non-expressive performance from teaching-v2.0 (onsets).
 
@@
 ```{r extract, include = FALSE}
-df_ioi <- fread("./../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+df_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 

### 2020-04-08 material/analysis/stim_n/ioi_duration.Rmd
move seq plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## Sequence plot
-```{r seq, fig.width = 10, echo = FALSE}
-# For each individual
-ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), .(SubNr, RowNr)]
-
-p_ioi_seq <- ggplot(data = ioi_seq, aes(x = RowNr, y = Mean, colour = SubNr)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "RowNr", y = "Normalised Duration") + scale_x_continuous(breaks=seq(1,66,1))
-p_ioi_seq
-```
-
-# IOI
+# Duration
 ## 1. All - individual
@@
 
+## Sequence plot
+```{r seq, fig.width = 10, echo = FALSE}
+# For each individual
+ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), .(SubNr, RowNr)]
+
+p_ioi_seq <- ggplot(data = ioi_seq, aes(x = RowNr, y = Mean, colour = SubNr)) +
+  geom_line() +
+  geom_point() +
+  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
+                position = position_dodge(.05)) + 
+  labs(x = "RowNr", y = "Normalised Duration") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_seq
+```
+
 ## 2. All - group

### 2020-04-08 material/analysis/stim_n/ioi_duration.Rmd
change data.table name
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 }
-df_ioi <- df_onset[Eval == "Yes"]
+df_duration <- df_onset[Eval == "Yes"]
 
 # check whether each row corresponds correctly
-df_ioi$Eval2 <- "NA"
-for (i in 1:nrow(df_ioi)){
-  if (df_ioi[i]$SubNr == df_offset[i]$SubNr & df_ioi[i]$TrialNr == df_offset[i]$TrialNr & df_ioi[i]$Pitch == df_offset[i]$Pitch){
-    df_ioi[i]$Eval2 <- "SAME"
+df_duration$Eval2 <- "NA"
+for (i in 1:nrow(df_duration)){
+  if (df_duration[i]$SubNr == df_offset[i]$SubNr & df_duration[i]$TrialNr == df_offset[i]$TrialNr & df_duration[i]$Pitch == df_offset[i]$Pitch){
+    df_duration[i]$Eval2 <- "SAME"
   } else {
-    df_ioi[i]$Eval2 <- "DIFF"
+    df_duration[i]$Eval2 <- "DIFF"
   }
 }
-print(unique(df_ioi$Eval2))
+print(unique(df_duration$Eval2))
 
 # calculate duration
-df_ioi$Duration <- df_offset$TimeStamp - df_ioi$TimeStamp
+df_duration$Duration <- df_offset$TimeStamp - df_duration$TimeStamp
 
 # convert bpm to ms
-df_ioi[Tempo == 120]$Tempo <- 250
-df_ioi[Tempo == 110]$Tempo <- 273
-df_ioi[Tempo == 100]$Tempo <- 300
+df_duration[Tempo == 120]$Tempo <- 250
+df_duration[Tempo == 110]$Tempo <- 273
+df_duration[Tempo == 100]$Tempo <- 300
 
 # normalised Duration
-df_ioi$normDu <- df_ioi$Duration/df_ioi$Tempo
+df_duration$normDu <- df_duration$Duration/df_duration$Tempo
 
@@
 ls_irrelevant <- list(c(27:28), c(33:35), 52, 57, c(66:72))
-df_ioi$EighthNote <- "Yes"
+df_duration$EighthNote <- "Yes"
 
@@
   for (note in 1:length(ls_irrelevant[[phrase]])){
-    df_ioi[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
+    df_duration[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
   }
@@
 
-```{r ioi, echo = FALSE}
+```{r duration, echo = FALSE}
 # For each individual
-ioi <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr)]
-print(ioi)
+duration <- df_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr)]
+print(duration)
 ```
@@
 
-```{r ioi_bar,  echo = FALSE}
-p_ioi <- ggplot(data = ioi, aes(x = SubNr, y = Mean)) +
+```{r duration_bar,  echo = FALSE}
+p_duration <- ggplot(data = duration, aes(x = SubNr, y = Mean)) +
   geom_bar(stat = "identity", position = position_dodge()) +
@@
   labs(y = "Normalised Duration", subtitle = "Duration: Non-Expression")
-p_ioi
+p_duration
 ```
@@
 # For each individual
-ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), .(SubNr, RowNr)]
+duration_seq <- df_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), .(SubNr, RowNr)]
 
-p_ioi_seq <- ggplot(data = ioi_seq, aes(x = RowNr, y = Mean, colour = SubNr)) +
+p_duration_seq <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, colour = SubNr)) +
   geom_line() +
@@
   labs(x = "RowNr", y = "Normalised Duration") + scale_x_continuous(breaks=seq(1,66,1))
-p_ioi_seq
+p_duration_seq
 ```
@@
 
-```{r ioi_all, echo = FALSE}
+```{r duration_all, echo = FALSE}
 # Group mean
-ioi_all <- df_ioi[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu))]
-print(ioi_all)
+duration_all <- df_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu))]
+print(duration_all)
 ```
@@
 ```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*mean(ioi$SD) & Mean > mean(ioi$Mean)-3*mean(ioi$SD)]
-print(excluded_ioi)
+excluded_duration <- duration[Mean < mean(duration$Mean)+3*mean(duration$SD) & Mean > mean(duration$Mean)-3*mean(duration$SD)]
+print(excluded_duration)
 ```
@@
 ```{r export, include = FALSE}
-# export df_ioi
-write.csv(df_ioi, file = "./data_duration.csv", row.names = F)
+# export df_duration
+write.csv(df_duration, file = "./data_duration.csv", row.names = F)
 ```

### 2020-04-08 material/analysis/stim_n/ioi_onset.Rmd
add distribution plot and variability (CV)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
 # SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
 
 # assign RowNr
-df_ioi$RowNr <- rep(1:72, nrow(df_ioi)/72)
+dt_ioi$RowNr <- rep(1:72, nrow(dt_ioi)/72)
 ```
@@
 # calculate normIOIs
-df_ioi$IOI <- diff(c(0, df_ioi$TimeStamp))
+dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))
 
 # convert bpm to ms
-df_ioi[Tempo == 120]$Tempo <- 250
-df_ioi[Tempo == 110]$Tempo <- 273
-df_ioi[Tempo == 100]$Tempo <- 300
+dt_ioi[Tempo == 120]$Tempo <- 250
+dt_ioi[Tempo == 110]$Tempo <- 273
+dt_ioi[Tempo == 100]$Tempo <- 300
 
 # normalise IOIs
-df_ioi$normIOI <- df_ioi$IOI/df_ioi$Tempo
+dt_ioi$normIOI <- dt_ioi$IOI/dt_ioi$Tempo
 
 # remove the first note
-df_ioi <- df_ioi[RowNr != 1]
+dt_ioi <- dt_ioi[RowNr != 1]
 
 # assign Interval
-df_ioi$Interval <- rep(1:71, nrow(df_ioi)/71)
+dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)
 
@@
 ls_irrelevant <- list(c(25:28), c(32:35), c(51:52), c(56:57), c(65:71))
-df_ioi$EighthNote <- "Yes"
+dt_ioi$EighthNote <- "Yes"
 
@@
   for (note in 1:length(ls_irrelevant[[phrase]])){
-    df_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
+    dt_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
   }
@@
 # For each individual
-ioi <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
+ioi <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
 print(ioi)
@@
 
+## Sequence plot
+Right wrap: Subject Number
+
+```{r ioi_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+# For each individual
+ioi_seq <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]
+
+p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, group = TrialNr)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  facet_grid(SubNr ~ .) +
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_seq
+```
+
+## Distribution plot
+Right wrap: Trial Number
+Upper wrap: Subject Number
+
+```{r ioi_dis, fig.width = 14, fig.height = 8, echo = FALSE}
+p_ioi_dis <- ggplot(dt_ioi[EighthNote == "Yes"], aes(x = normIOI)) +
+  geom_histogram(binwidth = .02, alpha = .5, position = "identity") +
+  facet_grid(TrialNr ~ SubNr) +
+  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Non-Expression")
+p_ioi_dis
+```
+
 ## 2. All - group
@@
 # Group mean
-ioi_all <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI))]
+ioi_all <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI))]
 print(ioi_all)
-```
-
-## Sequence plot
-```{r seq, fig.width = 10, echo = FALSE}
-# For each individual
-ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), .(SubNr, Interval)]
-
-p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, colour = SubNr)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
-p_ioi_seq
 ```
@@
 
+# IOI Variability
+```{r cv, echo = FALSE}
+# for each participant
+cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI), CV = sd(normIOI)/mean(normIOI)), by = .(SubNr, TrialNr)]
+cv
+```
+
+```{r cv_box, echo = FALSE}
+p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
+p_cv_box
+```
+
 ```{r export, include = FALSE}
-# export df_ioi
-write.csv(df_ioi, file = "./data_ioi_onset.csv", row.names = F)
+# export dt_ioi
+write.csv(dt_ioi, file = "./data_ioi_onset.csv", row.names = F)
 ```

### 2020-04-08 material/analysis/stim_n/ioi_duration.Rmd
add distribution plot and variability (CV)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # read a filtered csv
-df_onset <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
-df_offset <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+dt_onset <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
 
 # SubNr as a factor
-df_onset$SubNr <- as.factor(df_onset$SubNr)
-df_offset$SubNr <- as.factor(df_offset$SubNr)
+dt_onset$SubNr <- as.factor(dt_onset$SubNr)
+dt_offset$SubNr <- as.factor(dt_offset$SubNr)
 
 # assign RowNr
-df_onset$RowNr <- rep(1:72, nrow(df_onset)/72)
-df_offset$RowNr <- rep(1:72, nrow(df_offset)/72)
+dt_onset$RowNr <- rep(1:72, nrow(dt_onset)/72)
+dt_offset$RowNr <- rep(1:72, nrow(dt_offset)/72)
 ```
@@
 # extract trials which have both onset and offset data
-df_onset$Eval <- "NA"
-for (subnr in unique(df_offset$SubNr)){
-  for (trial in unique(df_offset[SubNr == subnr]$Trial)){
-    df_onset[SubNr == subnr & TrialNr == trial]$Eval <- "Yes"
+dt_onset$Eval <- "NA"
+for (subnr in unique(dt_offset$SubNr)){
+  for (trial in unique(dt_offset[SubNr == subnr]$Trial)){
+    dt_onset[SubNr == subnr & TrialNr == trial]$Eval <- "Yes"
   }
 }
-df_duration <- df_onset[Eval == "Yes"]
+dt_duration <- dt_onset[Eval == "Yes"]
 
 # check whether each row corresponds correctly
-df_duration$Eval2 <- "NA"
-for (i in 1:nrow(df_duration)){
-  if (df_duration[i]$SubNr == df_offset[i]$SubNr & df_duration[i]$TrialNr == df_offset[i]$TrialNr & df_duration[i]$Pitch == df_offset[i]$Pitch){
-    df_duration[i]$Eval2 <- "SAME"
+dt_duration$Eval2 <- "NA"
+for (i in 1:nrow(dt_duration)){
+  if (dt_duration[i]$SubNr == dt_offset[i]$SubNr & dt_duration[i]$TrialNr == dt_offset[i]$TrialNr & dt_duration[i]$Pitch == dt_offset[i]$Pitch){
+    dt_duration[i]$Eval2 <- "SAME"
   } else {
-    df_duration[i]$Eval2 <- "DIFF"
+    dt_duration[i]$Eval2 <- "DIFF"
   }
 }
-print(unique(df_duration$Eval2))
+print(unique(dt_duration$Eval2))
 
 # calculate duration
-df_duration$Duration <- df_offset$TimeStamp - df_duration$TimeStamp
+dt_duration$Duration <- dt_offset$TimeStamp - dt_duration$TimeStamp
 
 # convert bpm to ms
-df_duration[Tempo == 120]$Tempo <- 250
-df_duration[Tempo == 110]$Tempo <- 273
-df_duration[Tempo == 100]$Tempo <- 300
+dt_duration[Tempo == 120]$Tempo <- 250
+dt_duration[Tempo == 110]$Tempo <- 273
+dt_duration[Tempo == 100]$Tempo <- 300
 
 # normalised Duration
-df_duration$normDu <- df_duration$Duration/df_duration$Tempo
+dt_duration$normDu <- dt_duration$Duration/dt_duration$Tempo
 
@@
 ls_irrelevant <- list(c(27:28), c(33:35), 52, 57, c(66:72))
-df_duration$EighthNote <- "Yes"
+dt_duration$EighthNote <- "Yes"
 
@@
   for (note in 1:length(ls_irrelevant[[phrase]])){
-    df_duration[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
+    dt_duration[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
   }
@@
 # For each individual
-duration <- df_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr)]
+duration <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr)]
 print(duration)
@@
 - Error bars show SD of the mean
-- A dashed line represents the ideal tempo
 
@@
 ## Sequence plot
-```{r seq, fig.width = 10, echo = FALSE}
-# For each individual
-duration_seq <- df_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), .(SubNr, RowNr)]
+```{r duration_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+# for each individual
+dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
+duration_seq <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
 
-p_duration_seq <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, colour = SubNr)) +
+p_duration_seq <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr)) +
   geom_line() +
   geom_point() +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "RowNr", y = "Normalised Duration") + scale_x_continuous(breaks=seq(1,66,1))
+  facet_grid(SubNr ~ .) +
+  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Non-Expression") +
+  scale_x_continuous(breaks=seq(1,66,1))
 p_duration_seq
+```
+
+## Distribution plot
+```{r duration_dis, fig.width = 12, echo = FALSE}
+p_duration_dis <- ggplot(dt_duration[EighthNote == "Yes"], aes(x = normDu)) +
+  geom_histogram(binwidth=.1, alpha=.5, position="identity") +
+  facet_grid(TrialNr ~ SubNr) +
+  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Non-Expression")
+p_duration_dis
 ```
@@
 ```{r duration_all, echo = FALSE}
-# Group mean
-duration_all <- df_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu))]
+# group mean
+duration_all <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu))]
 print(duration_all)
@@
 ```{r export, include = FALSE}
-# export df_duration
-write.csv(df_duration, file = "./data_duration.csv", row.names = F)
+# export dt_duration
+write.csv(dt_duration, file = "./data_duration.csv", row.names = F)
 ```
+
+# Duration Variability
+```{r cv, echo = FALSE}
+# for each participant
+cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
+cv
+```
+
+```{r cv_box, echo = FALSE}
+p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
+p_cv_box
+```

### 2020-04-09 material/analysis/stim_n/ioi_onset.Rmd
add variability to onset analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## Sequence plot
-Right wrap: Subject Number
+#### Sequence plot
+- Right wrap: Subject Number
+- A dashed line represents the ideal tempo
 
@@
 
-## Distribution plot
-Right wrap: Trial Number
-Upper wrap: Subject Number
+#### Distribution plot
+- Right wrap: Trial Number
+- Upper wrap: Subject Number
 
@@
 # Group mean
-ioi_all <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI))]
+ioi_all <- ioi[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean))]
 print(ioi_all)
@@
 ```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*mean(ioi$SD) & Mean > mean(ioi$Mean)-3*mean(ioi$SD)]
+excluded_ioi <- ioi[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
 print(excluded_ioi)
@@
 # IOI Variability
+## Individual
 ```{r cv, echo = FALSE}
 # for each participant
-cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI), CV = sd(normIOI)/mean(normIOI)), by = .(SubNr, TrialNr)]
+cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
 cv
@@
 
-```{r export, include = FALSE}
-# export dt_ioi
-write.csv(dt_ioi, file = "./data_ioi_onset.csv", row.names = F)
+## Exclude outlier participants
+```{r cv_outlier, echo = FALSE}
+cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
+
+excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
+print(excluded_cv)
 ```

### 2020-04-09 material/analysis/stim_n/ioi_duration.Rmd
add variability to duration analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r normDus, include = FALSE}
+```{r normDu, include = FALSE}
 # extract trials which have both onset and offset data
@@
 ```{r duration, echo = FALSE}
-# For each individual
+# for each individual
 duration <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr)]
@@
 
-## Sequence plot
+#### Sequence plot
 ```{r duration_seq, fig.width = 14, fig.height = 12, echo = FALSE}
@@
 
-## Distribution plot
+#### Distribution plot
 ```{r duration_dis, fig.width = 12, echo = FALSE}
@@
 ## 2. All - group
-- Average normDus for each condition.
 
@@
 # group mean
-duration_all <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu))]
+duration_all <- duration[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean))]
 print(duration_all)
@@
 
-## Exclude outlier participants
+### Exclude outlier participants
 ```{r outlier, echo = FALSE}
-excluded_duration <- duration[Mean < mean(duration$Mean)+3*mean(duration$SD) & Mean > mean(duration$Mean)-3*mean(duration$SD)]
+excluded_duration <- duration[Mean < duration_all$Mean+3*duration_all$SD & Mean > duration_all$Mean-3*duration_all$SD]
 print(excluded_duration)
-```
-
-```{r export, include = FALSE}
-# export dt_duration
-write.csv(dt_duration, file = "./data_duration.csv", row.names = F)
 ```
@@
 cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
-cv
+print(cv)
 ```
@@
 ```
+
+### Exclude outlier participants
+```{r cv_outlier, echo = FALSE}
+# for all
+cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
+print(cv_all)
+
+excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
+print(excluded_cv)
+```

### 2020-04-09 material/analysis/stim_n/ioi_onset.Rmd
modify layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-## Exclude outlier participants
+### Exclude outlier participants
 ```{r outlier, echo = FALSE}
@@
 # IOI Variability
-## Individual
 ```{r cv, echo = FALSE}
@@
 
-## Exclude outlier participants
+### Exclude outlier participants
 ```{r cv_outlier, echo = FALSE}
+# for all
 cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
+print(cv_all)
 

### 2020-04-10 material/analysis/stim_n/ioi_duration.Rmd
correct some comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-### Exclude outlier participants
+### Exclude outlier performacnes
 ```{r cv_outlier, echo = FALSE}

### 2020-04-10 material/analysis/stim_n/ioi_onset.Rmd
correct some comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-### Exclude outlier participants
+### Exclude outlier performances
 ```{r cv_outlier, echo = FALSE}

### 2020-04-14 material/analysis/stim_n/ioi_offset.Rmd
update offset analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-df_ioi <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_ioi <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
 # SubNr as a factor
-df_ioi$SubNr <- as.factor(df_ioi$SubNr)
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
 
 # assign RowNr
-df_ioi$RowNr <- rep(1:72, nrow(df_ioi)/72)
+dt_ioi$RowNr <- rep(1:72, nrow(dt_ioi)/72)
 ```
@@
 
+#### Sequence plot
+- Right wrap: Subject Number
+- A dashed line represents the ideal tempo
+
+```{r ioi_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+# For each individual
+ioi_seq <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]
+
+p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, group = TrialNr)) +
+  geom_line() +
+  geom_point() +
+  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
+  facet_grid(SubNr ~ .) +
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+p_ioi_seq
+```
+
+#### Distribution plot
+- Right wrap: Trial Number
+- Upper wrap: Subject Number
+
+```{r ioi_dis, fig.width = 14, fig.height = 8, echo = FALSE}
+p_ioi_dis <- ggplot(dt_ioi[EighthNote == "Yes"], aes(x = normIOI)) +
+  geom_histogram(binwidth = .02, alpha = .5, position = "identity") +
+  facet_grid(TrialNr ~ SubNr) +
+  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Non-Expression")
+p_ioi_dis
+```
+
 ## 2. All - group
@@
 # Group mean
-ioi_all <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI))]
+ioi_all <- ioi[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean))]
 print(ioi_all)
@@
 
-## Sequence plot
-```{r seq, fig.width = 10, echo = FALSE}
-# For each individual
-ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), .(SubNr, Interval)]
-
-p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, colour = SubNr)) +
-  geom_line() +
-  geom_point() +
-  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
-                position = position_dodge(.05)) + 
-  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
-p_ioi_seq
-```
-
-## Exclude outlier participants
+### Exclude outlier participants
 ```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*mean(ioi$SD) & Mean > mean(ioi$Mean)-3*mean(ioi$SD)]
+excluded_ioi <- ioi[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
 print(excluded_ioi)
@@
 
-```{r export, include = FALSE}
-# export df_ioi
-write.csv(df_ioi, file = "./data_ioi_offset.csv", row.names = F)
+# IOI Variability
+```{r cv, echo = FALSE}
+# for each participant
+cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+cv
 ```
+
+```{r cv_box, echo = FALSE}
+p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
+p_cv_box
+```
+
+### Exclude outlier performances
+```{r cv_outlier, echo = FALSE}
+# for all
+cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
+print(cv_all)
+
+excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
+print(excluded_cv)
+```

### 2020-04-14 material/analysis/stim_n/kv.Rmd
upload key velocity analysis for stim_n
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis for Non-Expressive performances (Key Velocity)'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---
_
Description: This is a rough overview of non-expressive performance from teaching-v2.0 (KV).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
dt_kv <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv

# SubNr as a factor
dt_kv$SubNr <- as.factor(dt_kv$SubNr)

# assign RowNr
dt_kv$RowNr <- rep(1:72, nrow(dt_kv)/72)
```

```{r Velocitys, include = FALSE}
# # convert bpm to ms
# dt_kv[Tempo == 120]$Tempo <- 250
# dt_kv[Tempo == 110]$Tempo <- 273
# dt_kv[Tempo == 100]$Tempo <- 300

# remove irrelevant intervals
ls_irrelevant <- list(c(27:28), c(33:35), 52, 57, c(66:72))
dt_kv$EighthNote <- "Yes"

for (phrase in 1:length(ls_irrelevant)){
  for (note in 1:length(ls_irrelevant[[phrase]])){
    dt_kv[RowNr == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
  }
}
```

# KV
## 1. All - individual
- Average KV for each participant.

```{r kv, echo = FALSE}
# For each individual
kv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr)]
print(kv)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kv_bar,  echo = FALSE}
p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Key Velocity", subtitle = "KV: Non-Expression")
p_kv
```

#### Sequence plot
- Right wrap: Subject Number

```{r kv_seq, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
kv_seq <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq <- ggplot(data = kv_seq, aes(x = RowNr, y = Mean, group = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
p_kv_seq
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

```{r kv_dis, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis <- ggplot(dt_kv[EighthNote == "Yes"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Non-Expression")
p_kv_dis
```

## 2. All - group
- Average Velocity

```{r kv_all, echo = FALSE}
# Group mean
kv_all <- kv[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean))]
print(kv_all)
```

### Exclude outlier participants
```{r outlier, echo = FALSE}
excluded_kv <- kv[Mean < kv_all$Mean+3*kv_all$SD & Mean > kv_all$Mean-3*kv_all$SD]
print(excluded_kv)
```

# KV Variability
```{r cv, echo = FALSE}
# for each participant
cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)]
cv
```

```{r cv_box, echo = FALSE}
p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
  geom_boxplot() +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
p_cv_box
```

### Exclude outlier performances
```{r cv_outlier, echo = FALSE}
# for all
cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
print(cv_all)

excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
print(excluded_cv)
```

### 2020-04-15 material/analysis/expression/stim_a/ioi_onset.Rmd
add ioi-analysis for performing-articulation
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-10 material/analysis/stim_n/ioi_onset.Rmd
@@
 ```{r extract, include = FALSE}
-dt_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_ioi_original <- fread("../trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+# Include only articulation
+dt_ioi <- dt_ioi_original[Condition == "performing" & Skill == "articulation"]
 
@@
 dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
-
-# assign RowNr
-dt_ioi$RowNr <- rep(1:72, nrow(dt_ioi)/72)
-```
-
-```{r normIOIs, include = FALSE}
-# calculate normIOIs
-dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))
-
-# convert bpm to ms
-dt_ioi[Tempo == 120]$Tempo <- 250
-dt_ioi[Tempo == 110]$Tempo <- 273
-dt_ioi[Tempo == 100]$Tempo <- 300
-
-# normalise IOIs
-dt_ioi$normIOI <- dt_ioi$IOI/dt_ioi$Tempo
-
-# remove the first note
-dt_ioi <- dt_ioi[RowNr != 1]
-
-# assign Interval
-dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)
-
-# remove irrelevant intervals
-ls_irrelevant <- list(c(25:28), c(32:35), c(51:52), c(56:57), c(65:71))
-dt_ioi$EighthNote <- "Yes"
-
-for (phrase in 1:length(ls_irrelevant)){
-  for (note in 1:length(ls_irrelevant[[phrase]])){
-    dt_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
-  }
-}
+dt_ioi$TrialNr <- as.factor(dt_ioi$TrialNr)
 ```
@@
 # For each individual
-ioi <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
+ioi <- dt_ioi[, .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
 print(ioi)
@@
   geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
-  labs(y = "Normalised IOI", subtitle = "IOI: Non-Expression")
+  labs(y = "Normalised IOI", subtitle = "IOI: Performing-Articulation")
 p_ioi
@@
 
-```{r ioi_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+```{r ioi_seq, fig.width = 14, fig.height = 20, echo = FALSE}
 # For each individual
-ioi_seq <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]
+ioi_seq <- dt_ioi[, .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]
 
-p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, group = TrialNr)) +
+p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, group = TrialNr, colour = TrialNr)) +
   geom_line() +
@@
 ```{r ioi_dis, fig.width = 14, fig.height = 8, echo = FALSE}
-p_ioi_dis <- ggplot(dt_ioi[EighthNote == "Yes"], aes(x = normIOI)) +
+p_ioi_dis <- ggplot(dt_ioi, aes(x = normIOI)) +
   geom_histogram(binwidth = .02, alpha = .5, position = "identity") +
   facet_grid(TrialNr ~ SubNr) +
-  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Non-Expression")
+  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Performing-Articulation")
 p_ioi_dis
@@
 # for each participant
-cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+cv <- dt_ioi[, .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
 cv
@@
   geom_boxplot() +
-  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.3)
 p_cv_box

### 2020-04-16 material/analysis/expression/stim_a/ioi_onset.Rmd
remove performances containing NA
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-dt_ioi_original <- fread("../trimmed/data_ioi_1.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-# Include only articulation
-dt_ioi <- dt_ioi_original[Condition == "performing" & Skill == "articulation"]
+dt_onset <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_error <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
+# assign RowNr
+dt_onset$RowNr <- rep(1:72, nrow(dt_onset)/72)
+# sort by SubNr, BlockNr, TrialNr and NoteNr
+dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
+# include only articulation
+dt_ioi <- dt_onset[Condition == "performing" & Skill == "articulation"]
 
@@
 dt_ioi$TrialNr <- as.factor(dt_ioi$TrialNr)
+
+# remove performances which contain NA
+dt_ioi$Error <- 0
+for (error in 1:nrow(dt_error)){
+  dt_ioi[SubNr == dt_error[error]$SubNr & BlockNr == dt_error[error]$BlockNr & TrialNr == dt_error[error]$TrialNr]$Error <- 1
+}
+dt_ioi_ori <- dt_ioi # save original data
+dt_ioi <- dt_ioi_ori[Error == 0]
+```
+
+```{r normIOIs, include = FALSE}
+# calculate normIOIs
+dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))
+
+# convert bpm to ms
+dt_ioi[Tempo == 120]$Tempo <- 250
+dt_ioi[Tempo == 110]$Tempo <- 273
+dt_ioi[Tempo == 100]$Tempo <- 300
+
+# normalise IOIs
+dt_ioi$normIOI <- dt_ioi$IOI/dt_ioi$Tempo
+
+# remove the first note
+dt_ioi <- dt_ioi[RowNr != 1]
+
+# assign Interval
+dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)
+
+# remove irrelevant intervals
+ls_irrelevant <- list(c(25:28), c(32:35), c(51:52), c(56:57), c(65:71))
+dt_ioi$EighthNote <- "Yes"
+
+for (phrase in 1:length(ls_irrelevant)){
+  for (note in 1:length(ls_irrelevant[[phrase]])){
+    dt_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
+  }
+}
 ```
@@
 # For each individual
-ioi <- dt_ioi[, .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
+ioi <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
 print(ioi)
@@
 # For each individual
-ioi_seq <- dt_ioi[, .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]
+ioi_seq <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI)), by = .(SubNr, TrialNr, Interval)]
 
@@
 ```{r ioi_dis, fig.width = 14, fig.height = 8, echo = FALSE}
-p_ioi_dis <- ggplot(dt_ioi, aes(x = normIOI)) +
-  geom_histogram(binwidth = .02, alpha = .5, position = "identity") +
+p_ioi_dis <- ggplot(dt_ioi[EighthNote == "Yes"], aes(x = normIOI)) +
+  geom_histogram(binwidth = .05, alpha = .5, position = "identity") +
   facet_grid(TrialNr ~ SubNr) +
@@
   geom_boxplot() +
-  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.3)
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5)
 p_cv_box

### 2020-04-16 material/analysis/expression/stim_d/kv.Rmd
upload analysis for kv performing condition
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis for Performing-Dynamics performances (Key Velocity)'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a rough overview of non-expressive performance from teaching-v2.0 (KV).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
# onset
dt_kv <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv

# SubNr as a factor
dt_kv$SubNr <- as.factor(dt_kv$SubNr)

# assign RowNr
dt_kv$RowNr <- rep(1:72, nrow(dt_kv)/72)

# include only articulation
dt_kv <- dt_kv[Condition == "performing" & Skill == "dynamics"]

# remove performances which contain NA
dt_kv$Error <- 0
for (error in 1:nrow(dt_error_on)){
  dt_kv[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
}
dt_kv_ori <- dt_kv # save original data
dt_kv <- dt_kv_ori[Error == 0]
```

```{r Velocitys, include = FALSE}
# # convert bpm to ms
# dt_kv[Tempo == 120]$Tempo <- 250
# dt_kv[Tempo == 110]$Tempo <- 273
# dt_kv[Tempo == 100]$Tempo <- 300

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (duration/velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)

# assign subcomponents
dt_kv$Subcomponent <- "NA"

# for forte
for (phrase in 1:length(ls_forte2)){
  for (note in 1:length(ls_forte2[[phrase]])){
    dt_kv[RowNr == ls_forte2[[phrase]][note]]$Subcomponent <- "Forte"
  }
}

# for piano
for (phrase in 1:length(ls_piano2)){
  for (note in 1:length(ls_piano2[[phrase]])){
    dt_kv[RowNr == ls_piano2[[phrase]][note]]$Subcomponent <- "Piano"
  }
}
```

# KV
## 1. All - individual
- Average KV for each participant.

```{r kv, echo = FALSE}
# For each individual
kv <- dt_kv[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Subcomponent)]
print(kv)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kv_bar,  echo = FALSE}
p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Key Velocity", subtitle = "KV: Performing-Dynamics")
p_kv
```

#### Sequence plot
- Right wrap: Subject Number

#### Forte
```{r kv_seq_for, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq_for <- dt_kv[Subcomponent == "Forte", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq_for <- ggplot(data = kv_seq_for, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
p_kv_seq_for
```

#### Piano
```{r kv_seq_pia, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq_pia <- dt_kv[Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq_pia <- ggplot(data = kv_seq_pia, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
p_kv_seq_pia
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

```{r kv_dis_for, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis_for <- ggplot(dt_kv[Subcomponent == "Forte"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Forte)")
p_kv_dis_for
```

```{r kv_dis_pia, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis_pia <- ggplot(dt_kv[Subcomponent == "Piano"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Piano)")
p_kv_dis_pia
```

## 2. All - group
- Average Velocity

```{r kv_all, echo = FALSE}
# Group mean
kv_all <- kv[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean)), by = .(Subcomponent)]
print(kv_all)
```

### Exclude outlier participants
#### Forte
```{r outlier_for, echo = FALSE}
# for forte
excluded_duration_for <- kv[Subcomponent == "Forte" & Mean < kv_all[Subcomponent == "Forte"]$Mean+3*kv_all[Subcomponent == "Forte"]$SD & Mean > kv_all[Subcomponent == "Forte"]$Mean-3*kv_all$SD]
print(excluded_duration_for)
```

#### Piano
```{r outlier_pia, echo = FALSE}
# for piano
excluded_duration_pia <- kv[Subcomponent == "Piano" & Mean < kv_all[Subcomponent == "Piano"]$Mean+3*kv_all[Subcomponent == "Piano"]$SD & Mean > kv_all[Subcomponent == "Piano"]$Mean-3*kv_all$SD]
print(excluded_duration_pia)
```

<!-- # KV Variability -->
<!-- ```{r cv, echo = FALSE} -->
<!-- # for each participant -->
<!-- cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)] -->
<!-- cv -->
<!-- ``` -->

<!-- ```{r cv_box, echo = FALSE} -->
<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
<!-- p_cv_box -->
<!-- ``` -->

<!-- ### Exclude outlier performances -->
<!-- ```{r cv_outlier, echo = FALSE} -->
<!-- # for all -->
<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
<!-- print(cv_all) -->

<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
<!-- print(excluded_cv) -->
<!-- ``` -->

### 2020-04-16 material/analysis/expression/stim_a/ioi_duration.Rmd
upload analysis for duration performing condition
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis for Performing-Articulation performances (DURATION)'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a rough overview of performing performance from teaching-v2.0 (duration).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
# onset
dt_onset <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
# assign RowNr
dt_onset$RowNr <- rep(1:72, nrow(dt_onset)/72)
# sort by SubNr, BlockNr, TrialNr and NoteNr
dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
# include only articulation
dt_onset <- dt_onset[Condition == "performing" & Skill == "articulation"]

# SubNr as a factor
dt_onset$SubNr <- as.factor(dt_onset$SubNr)
dt_onset$TrialNr <- as.factor(dt_onset$TrialNr)

# remove performances which contain NA
dt_onset$Error <- 0
for (error in 1:nrow(dt_error_on)){
  dt_onset[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
}
dt_onset_ori <- dt_onset # save original data
dt_onset <- dt_onset_ori[Error == 0]

# offset
dt_offset <- fread("../filtered/data_correct_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
dt_error_off <- fread("../filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".") # read an error csv
# assign RowNr
dt_offset$RowNr <- rep(1:72, nrow(dt_offset)/72)
# sort by SubNr, BlockNr, TrialNr and NoteNr
dt_offset <- dt_offset[order(SubNr, BlockNr, TrialNr, RowNr),]
# include only articulation
dt_offset <- dt_offset[Condition == "performing" & Skill == "articulation"]

# SubNr as a factor
dt_offset$SubNr <- as.factor(dt_offset$SubNr)
dt_offset$TrialNr <- as.factor(dt_offset$TrialNr)

# remove performances which contain NA
dt_offset$Error <- 0
for (error in 1:nrow(dt_error_off)){
  dt_offset[SubNr == dt_error_off[error]$SubNr & BlockNr == dt_error_off[error]$BlockNr & TrialNr == dt_error_off[error]$TrialNr]$Error <- 1
}
dt_offset_ori <- dt_offset # save original data
dt_offset <- dt_offset_ori[Error == 0]
```

```{r normDu, include = FALSE}
# extract trials which have both onset and offset data
dt_onset$Eval <- "NA"
for (subnr in unique(dt_offset$SubNr)){
  for (trial in unique(dt_offset[SubNr == subnr]$Trial)){
    dt_onset[SubNr == subnr & TrialNr == trial]$Eval <- "Yes"
  }
}
dt_duration <- dt_onset[Eval == "Yes"]

# check whether each row corresponds correctly
dt_duration$Eval2 <- "NA"
for (i in 1:nrow(dt_duration)){
  if (dt_duration[i]$SubNr == dt_offset[i]$SubNr & dt_duration[i]$TrialNr == dt_offset[i]$TrialNr & dt_duration[i]$Pitch == dt_offset[i]$Pitch){
    dt_duration[i]$Eval2 <- "SAME"
  } else {
    dt_duration[i]$Eval2 <- "DIFF"
  }
}
print(unique(dt_duration$Eval2))

# TimeStamp as an integer (only offsets - I don't know why 16.04.20)
dt_offset$TimeStamp <- as.integer(dt_offset$TimeStamp)

# calculate duration
dt_duration$Duration <- dt_offset$TimeStamp - dt_duration$TimeStamp

# convert bpm to ms
dt_duration[Tempo == 120]$Tempo <- 250
dt_duration[Tempo == 110]$Tempo <- 273
dt_duration[Tempo == 100]$Tempo <- 300

# normalised Duration
dt_duration$normDu <- dt_duration$Duration/dt_duration$Tempo

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (duration/velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)

# assign subcomponents
dt_duration$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato2)){
  for (note in 1:length(ls_legato2[[phrase]])){
    dt_duration[RowNr == ls_legato2[[phrase]][note]]$Subcomponent <- "Legato"
  }
}
# Staccato
for (phrase in 1:length(ls_staccato2)){
  for (note in 1:length(ls_staccato2[[phrase]])){
    dt_duration[RowNr == ls_staccato2[[phrase]][note]]$Subcomponent <- "Staccato"
  }
}
```

# Duration
## 1. All - individual
- Average normDu for each participant.

```{r duration, echo = FALSE}
# for each individual
duration <- dt_duration[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, Subcomponent)]
print(duration)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r duration_bar,  echo = FALSE}
p_duration <- ggplot(data = duration, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Normalised Duration", subtitle = "Duration: Performing-Articulation")
p_duration
```

#### Sequence plot
#### Legato
```{r duration_seq_leg, fig.width = 14, fig.height = 12, echo = FALSE}
# for each individual
dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
duration_seq <- dt_duration[Subcomponent == "Legato", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]

p_duration_seq_leg <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Articulation(Legato)") +
  scale_x_continuous(breaks=seq(1,66,1))
p_duration_seq_leg
```

#### Staccato
```{r duration_seq_sta, fig.width = 14, fig.height = 12, echo = FALSE}
# for each individual
dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
duration_seq <- dt_duration[Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]

p_duration_seq_sta <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Articulation(Staccato)") +
  scale_x_continuous(breaks=seq(1,66,1))
p_duration_seq_sta
```

#### Distribution plot
#### Legato
```{r duration_dis_leg, fig.width = 12, echo = FALSE}
p_duration_dis_leg <- ggplot(dt_duration[Subcomponent == "Legato"], aes(x = normDu)) +
  geom_histogram(binwidth = .1, alpha=.5, position="identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Articulation(Legato)")
p_duration_dis_leg
```

#### Staccato
```{r duration_dis_sta, fig.width = 12, echo = FALSE}
p_duration_dis_sta <- ggplot(dt_duration[Subcomponent == "Staccato"], aes(x = normDu)) +
  geom_histogram(binwidt = .1, alpha = .5, position="identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Articulation(Staccato)")
p_duration_dis_sta
```

## 2. All - group

```{r duration_all, echo = FALSE}
# group mean
duration_all <- duration[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean)), by = .(Subcomponent)]
print(duration_all)
```

### Exclude outlier participants
#### Legato
```{r outlier_leg, echo = FALSE}
# for legato
excluded_duration_leg <- duration[Subcomponent == "Legato" & Mean < duration_all[Subcomponent == "Legato"]$Mean+3*duration_all[Subcomponent == "Legato"]$SD & Mean > duration_all[Subcomponent == "Legato"]$Mean-3*duration_all$SD]
print(excluded_duration_leg)
```

#### Staccato
```{r outlier_sta, echo = FALSE}
# for staccato
excluded_duration_sta <- duration[Subcomponent == "Staccato" & Mean < duration_all[Subcomponent == "Staccato"]$Mean+3*duration_all[Subcomponent == "Staccato"]$SD & Mean > duration_all$Mean-3*duration_all[Subcomponent == "Staccato"]$SD]
print(excluded_duration_sta)
```

<!-- # Duration Variability -->
<!-- ```{r cv, echo = FALSE} -->
<!-- # for each participant -->
<!-- cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)] -->
<!-- print(cv) -->
<!-- ``` -->

<!-- ```{r cv_box, echo = FALSE} -->
<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
<!-- p_cv_box -->
<!-- ``` -->

<!-- ### Exclude outlier performacnes -->
<!-- ```{r cv_outlier, echo = FALSE} -->
<!-- # for all -->
<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
<!-- print(cv_all) -->

<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
<!-- print(excluded_cv) -->
<!-- ``` -->

### 2020-04-23 material/analysis/stim_n/ioi_duration.Rmd
export duration_seq csv
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 duration_seq <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
+# save data.table for comparison with stim_a
+fwrite(duration_seq, file = "./seq.csv")
 
@@
   labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Non-Expression") +
-  scale_x_continuous(breaks=seq(1,66,1))
+  scale_x_continuous(breaks=seq(1,66,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
 p_duration_seq

### 2020-04-23 material/analysis/expression/stim_a/ioi_duration.Rmd
rescale y axis for seq plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-#### Sequence plot
+### Sequence plot
+#### All
+```{r duration_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+# for each individual
+dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
+duration_seq <- dt_duration[Subcomponent != "NA", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
+
+duration_seq_stim_n <- fread("../../stim_n/seq.csv", header = T, sep = ",", dec = ".")
+duration_seq_stim_n$SubNr <- factor(duration_seq_stim_n$SubNr)
+duration_seq_stim_n$TrialNr <- factor(duration_seq_stim_n$TrialNr)
+duration_seq_stim_n$TrialNr <- "Practice"
+
+duration_seq <- rbind(duration_seq, duration_seq_stim_n)
+
+p_duration_seq <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
+  geom_line() +
+  geom_point() +
+  facet_grid(SubNr ~ .) +
+  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Articulation(All)") +
+  scale_x_continuous(breaks=seq(1,66,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
+p_duration_seq
+```
+
 #### Legato
@@
 dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
-duration_seq <- dt_duration[Subcomponent == "Legato", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
+duration_seq_leg <- dt_duration[Subcomponent == "Legato", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
 
-p_duration_seq_leg <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
+p_duration_seq_leg <- ggplot(data = duration_seq_leg, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
   geom_line() +
@@
 p_duration_dis_sta <- ggplot(dt_duration[Subcomponent == "Staccato"], aes(x = normDu)) +
-  geom_histogram(binwidt = .1, alpha = .5, position="identity") +
+  geom_histogram(binwidth = .1, alpha = .5, position="identity") +
   facet_grid(TrialNr ~ SubNr) +

### 2020-05-06 material/analysis/stim_n/duration.Rmd
export info about valid performance
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-10 material/analysis/stim_n/ioi_duration.Rmd
@@
 ---
-title: 'Analysis for Non-Expressive performances (DURATION)'
+title: 'Duration - Analysis for Non-Expressive performances'
 output:
@@
 
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (duration).
+Description: This is a rough overview of non-expressive performance (duration) from teaching-v2.0.
 
@@
 ```
+
+```{r export, include = FALSE}
+fwrite(excluded_cv, "./duration_valid.csv")
+```

### 2020-05-06 material/analysis/stim_n/ioi_offset.Rmd
export info about valid performance
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Analysis for Non-Expressive performances (OFFSET)'
+title: 'IOI-Offset Analysis for Non-Expressive performances (OFFSET)'
 output:
@@
 
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (offsets).
+Description: This is a rough overview of non-expressive performance (IOI-offset) from teaching-v2.0.
 
@@
 ```
+
+```{r export, include = FALSE}
+fwrite(excluded_cv, "./offset_valid.csv")
+```

### 2020-05-06 material/analysis/stim_n/analysis/ioi_onset.Rmd
move files to analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-10 material/analysis/stim_n/ioi_onset.Rmd
@@
 ---
-title: 'Analysis for Non-Expressive performances (ONSET)'
+title: 'IOI-Onset Analysis for Non-Expressive performances'
 output:
@@
 
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (onsets).
+Description: This is a rough overview of non-expressive performance (IOI-Onset) from teaching-v2.0.
 
@@
 ```
+
+```{r export, include = FALSE}
+fwrite(excluded_cv, "./onset_valid.csv")
+```

### 2020-05-06 material/analysis/stim_n/analysis/kv.Rmd
move files to analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-14 material/analysis/stim_n/kv.Rmd
@@
 ---
-title: 'Analysis for Non-Expressive performances (Key Velocity)'
+title: 'KV Analysis for Non-Expressive performances'
 output:
@@
 _
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (KV).
+Description: This is a rough overview of non-expressive performance (KV) from teaching-v2.0.
 
@@
 ```
+
+```{r export, include = FALSE}
+fwrite(excluded_cv, "./kv_valid.csv")
+```

### 2020-05-06 material/analysis/stim_n/analysis/duration.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-06 material/analysis/stim_n/duration.Rmd
@@
 # read a filtered csv
-dt_onset <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
-dt_offset <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+dt_onset <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
 

### 2020-05-06 material/analysis/stim_n/analysis/duration.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # for each participant
-cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
+cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
 print(cv)

### 2020-05-12 material/analysis/stim_n/analysis/duration.Rmd
add # of excluded participants / modify layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of non-expressive performance (duration) from teaching-v2.0.
+Description: This is a rough overview of non-expressive performance (Duration) from teaching-v2.0.
 
@@
 ## 1. All - individual
-- Average normDu for each participant.
+- Average normDu for each participant
 
@@
 ## 2. All - group
+- Average normDu
 
@@
 
-### Exclude outlier participants
+### Valid performances (exclude outliers)
 ```{r outlier, echo = FALSE}
@@
 ```
+- The number of excluded performance: `r duration_all$N-nrow(excluded_duration)`
 
 # Duration Variability
+## 1. All - individual
+- Average CV for each participant
+
 ```{r cv, echo = FALSE}
@@
 
-### Exclude outlier performacnes
-```{r cv_outlier, echo = FALSE}
+## 2. All - group
+- Average CV
+
+```{r cv_all, echo = FALSE}
 # for all
@@
 print(cv_all)
+```
 
+### Valid performances (exclude outliers)
+```{r cv_outlier, echo = FALSE}
 excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
@@
 ```
+- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`
 

### 2020-05-12 material/analysis/stim_n/analysis/ioi_offset.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add # of excluded participants / modify layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-06 material/analysis/stim_n/ioi_offset.Rmd
@@
 ---
-title: 'IOI-Offset Analysis for Non-Expressive performances (OFFSET)'
+title: 'IOI-Offset Analysis for Non-Expressive performances'
 output:
@@
 ```{r extract, include = FALSE}
-dt_ioi <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_ioi <- fread("../preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 ## 1. All - individual
-- Average normIOIs for each participant.
+- Average normIOIs for each participant
 
@@
 ```
+- The number of excluded performance: `r ioi_all$N-nrow(excluded_ioi)`
 
 # IOI Variability
+## 1. All - individual
+- Average CV for each participant
+
 ```{r cv, echo = FALSE}
@@
 
-### Exclude outlier performances
-```{r cv_outlier, echo = FALSE}
+## 2. All - group
+- Average CV
+
+```{r cv_all, echo = FALSE}
 # for all

### 2020-05-12 material/analysis/stim_n/analysis/ioi_onset.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add # of excluded participants / modify layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r extract, include = FALSE}
-dt_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_ioi <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 ```
+## 2. All - group
+- Average CV
 
-### Exclude outlier performances
-```{r cv_outlier, echo = FALSE}
+```{r cv_all, echo = FALSE}
 # for all

### 2020-05-12 material/analysis/stim_n/analysis/kv.Rmd
change file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add # of excluded participants / modify layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-_
+
 Description: This is a rough overview of non-expressive performance (KV) from teaching-v2.0.
@@
 ```{r extract, include = FALSE}
-dt_kv <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_kv <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 ```
+- The number of excluded performance: `r kv_all$N-nrow(excluded_kv)`
 
 # KV Variability
+## 1. All - individual
+- Average CV for each participant
+
 ```{r cv, echo = FALSE}

### 2020-05-12 material/analysis/stim_n/duration.Rmd
change file location / file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```
+- The number of excluded performance: `r duration_all$N-nrow(excluded_duration)`
 
 # Duration Variability
+## 1. All - individual
+- Average CV for each participant
+
 ```{r cv, echo = FALSE}
 # for each participant
-cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
+cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
 print(cv)

### 2020-05-12 material/analysis/stim_n/ioi_onset.Rmd
change file location / file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-09 material/analysis/stim_n/ioi_onset.Rmd
@@
 
-## Exclude outlier participants
+### Valid performances (exclude outliers)
 ```{r outlier, echo = FALSE}
@@
 ```
+- The number of excluded performance: `r ioi_all$N-nrow(excluded_ioi)`
 
 # IOI Variability
-## Individual
+## 1. All - individual
+- Average CV for each participant
+
 ```{r cv, echo = FALSE}
@@
 ```
+## 2. All - group
+- Average CV
 
-## Exclude outlier participants
+```{r cv_all, echo = FALSE}
+# for all
+cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
+print(cv_all)
+```
+
+### Valid performances (exclude outliers)
 ```{r cv_outlier, echo = FALSE}
-cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))]
-
 excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
@@
 ```
+- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`
+
+```{r export, include = FALSE}
+fwrite(excluded_cv, "./onset_valid.csv")
+```

### 2020-05-12 material/analysis/stim_n/kv.Rmd
change file location / file path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-_
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (KV).
+
+Description: This is a rough overview of non-expressive performance (KV) from teaching-v2.0.
 
@@
 ```
+- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`
+
+```{r export, include = FALSE}
+fwrite(excluded_cv, "./kv_valid.csv")
+```

### 2020-05-12 material/stimuli/average.R
generate averaged performance / add function.R
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,98 @@
+# /usr/local/bin/r
+# Average onset, duration, key velocity for normative performance
+# Created: 06/05/2020
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+
+# setup
+## ggplots
+theme_set(theme_classic())
+theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
+## function
+source("./function.R")
+
+# read csv
+dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
+#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
+duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
+kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
+
+# duration and kv
+valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
+valid_du_kv$Duplicated <- duplicated(valid_du_kv)
+
+# valid performances
+valid <- valid_du_kv[Duplicated == TRUE]
+
+### create 6 instances! ###
+valid$Sample <- sample(c(1:31), replace = FALSE)
+print(valid)
+fwrite(valid, "./valid.txt")
+
+# 1. average IOIs
+dt_ioi_instance  <- data.table()
+counter = 0
+for (i in c(1:6)){
+  stim <- combining_onset(valid, i)
+  
+  # calculate normIOI
+  stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
+  stim$normIOI <- stim$IOI/stim$Tempo
+  
+  # average normIOI
+  stim_average <- stim[, .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(RowNr)]
+  # label instance no
[…]

### 2020-05-12 material/stimuli/function.R

@@ -1,0 +1,59 @@
+# averaging function (create 6 instances)
+# packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# valid: information(data.table) about valid performances
+# number: loop number (if outside loop, enter 0)
+
+combining_onset <- function(valid, number){
+  SubNr1 <- valid[Sample == number+counter]$SubNr
+  SubNr2 <- valid[Sample == number+1+counter]$SubNr
+  SubNr3 <- valid[Sample == number+2+counter]$SubNr
+  SubNr4 <- valid[Sample == number+3+counter]$SubNr
+  SubNr5 <- valid[Sample == number+4+counter]$SubNr
+  TrialNr1 <- valid[Sample == number+counter]$TrialNr
+  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
+  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
+  TrialNr4 <- valid[Sample == number+3+counter]$TrialNr
+  TrialNr5 <- valid[Sample == number+4+counter]$TrialNr
+  stim_1 <- dt_onset[SubNr == SubNr1 & TrialNr == TrialNr1]
+  stim_2 <- dt_onset[SubNr == SubNr2 & TrialNr == TrialNr2]
+  stim_3 <- dt_onset[SubNr == SubNr3 & TrialNr == TrialNr3]
+  stim_4 <- dt_onset[SubNr == SubNr4 & TrialNr == TrialNr4]
+  stim_5 <- dt_onset[SubNr == SubNr5 & TrialNr == TrialNr5]
+  outcome <- rbind(stim_1, stim_2, stim_3, stim_4, stim_5)
+  outcome$RowNr <- rep(c(1:72), 5)
+  
+  # tempo labels
+  outcome[Tempo == 120]$Tempo <- 250
+  outcome[Tempo == 110]$Tempo <- 273
+  outcome[Tempo == 100]$Tempo <- 300
+  return(outcome)
+}
+
+combining_offset <- function(valid, number){
+  SubNr1 <- valid[Sample == number+counter]$SubNr
+  SubNr2 <- valid[Sample == number+1+counter]$SubNr
+  SubNr3 <- valid[Sample == number+2+counter]$SubNr
+  SubNr4 <- valid[Sample == number+3+counter]$SubNr
[…]

### 2020-05-12 material/stimuli/average.R
create txt files for playback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -15,13 +15,14 @@ theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 C
 ## function
 source("./function.R")
 
-# read csv
+# read csv/txt
 dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
 dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
 #onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
 #offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
 duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
 kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
+dt_ideal <- fread("./ideal.txt", header = F)
 
 # duration and kv
 valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
@@ -38,12 +39,14 @@ fwrite(valid, "./valid.txt")
 # 1. average IOIs
 dt_ioi_instance  <- data.table()
 counter = 0
-for (i in c(1:6)){
+for (i in 1:6){
   stim <- combining_onset(valid, i)
   
   # calculate normIOI
   stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
   stim$normIOI <- stim$IOI/stim$Tempo
+  # IOI = 0 if RowNr == 1
+  stim[RowNr == 1]$normIOI <- 0
   
   # average normIOI
   stim_average <- stim[, .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(RowNr)]
@@ -96,3 +99,42 @@ for (i in c(1:6)){
 }
 
 ### create playback data! ###
+# 1. determine onsets/offsets
+# onsets
+dt_playback_onset <- dt_ioi_instance[, c("Instance", "RowNr")]
+dt_playback_onset$Key_OnOff <- 1
+for (i in 1:nrow(dt_playback_onset)){
+  if (dt_playback_onset$RowNr[i] == 1){
[…]

### 2020-05-13 material/composer_art.py
remove data manipulation part and only playback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,96 +1,0 @@
-#!/usr/local/bin/python
-####################
-# Created: 29/01/2019
-# This script generates experimental stimuli for adaptation-v1.0 (articulation)
-
-# %% import packages
-import csv, time, mido
-import pandas as pd
-import numpy as np
-
-#%% open port (play sounds with Max MSP)
-port = mido.open_output('to Max 1')
-
-# %% read csv
-df = pd.read_csv('./17_art_original.txt', delimiter = " ", header = None)
-
-# change column names
-df.columns = ["NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "Tempo", "SubNr", "BlockNr", "TrialNr", "Skill", "Condition", "Image"]
-
-# %% remove irrelevant characters
-df["NoteNr"] = df["NoteNr"].map(lambda x: x.rstrip(","))
-df["Image"] = df["Image"].map(lambda x: x.rstrip(";"))
-
-# %% convert str to int
-df["NoteNr"] = pd.to_numeric(df["NoteNr"])
-
-# %%  use only relevant notes (remove metronomes)
-df_original = df
-df = df[df['NoteNr'] >= 17]
-
-# divide the dataframe to Note On/Off
-df_on = df[df["Key_OnOff"] == 1]
-df_off = df[df["Key_OnOff"] == 0]
-
-# %% generate normal distribution
-mu = np.mean([30.6598, 31.7097, -196.3454, -204.8299])
-sigma = np.mean([17.9120, 18.3470, 22.8536, 28.5505	])
-norm = np.random.normal(mu, sigma, 1000)
-
-# extract values within 1SD
-condition1 = norm < mu+sigma
-condition2 = norm > mu-sigma
-
-norm_new = list(set(np.concatenate([np.extract(condition1, norm), np.extract(condition2, norm)])))
-
-# %% manipulate Velocity values
-for i in range(len(df_on["TimeStamp"])):
-    if i < len(df_on["TimeStamp"])-1:
-        df_off["TimeStamp"].iloc[i] = int(np.random.choice(norm_new, 1)[0]) + df_on["TimeStamp"].iloc[i+1]
-
-
-# %% remove .0
-df_on["TimeStamp"] = df_on["TimeStamp"].astype(int)
-df_off["TimeStamp"] = df_off["TimeStamp"].astype(int)
-
-# %% combine two dataframe again
-df = pd.concat([df_on, df_off])
[…]

### 2020-05-13 material/composer_dyn.py

@@ -1,115 +1,0 @@
-#!/usr/local/bin/python
-####################
-# Created: 20/01/2019
-# This script generates experimental stimuli for adaptation-v1.0
-
-# %% import packages
-import csv, time, mido
-import pandas as pd
-import numpy as np
-
-#%% open port (play sounds with Max MSP)
-port = mido.open_output('to Max 1')
-
-# %% read csv
-df = pd.read_csv('./17_dyn_original.txt', delimiter = " ", header = None)
-
-# change column names
-df.columns = ["NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "Tempo", "SubNr", "BlockNr", "TrialNr", "Skill", "Condition", "Image"]
-
-# %% remove irrelevant characters
-df["NoteNr"] = df["NoteNr"].map(lambda x: x.rstrip(","))
-df["Image"] = df["Image"].map(lambda x: x.rstrip(";"))
-
-# %% convert str to int
-df["NoteNr"] = pd.to_numeric(df["NoteNr"])
-
-# %%  use only relevant notes (remove metronomes)
-df_original = df
-df =  df[df['NoteNr'] >= 17]
-
-# %%　for each note (velocity)
-""" ls_forte = [range(1,6,1), range(9,18,1), range(21,27,1), range(40,48,1)]
-ls_piano = [range(6,9,1), range(29,33,1), range(36,40,1), range(48,52,1), range(58,66,1)]
- """
-# %% create Subcomponent column with NA
-# df["Subcomponent"] = "NA"
-
-# divide the dataframe to Note On/Off
-df_on = df[df["Key_OnOff"] == 1]
-df_off = df[df["Key_OnOff"] == 0]
-# assign SubNoteNr for each dataframe
-df_on["SubNoteNr"] = range(1, len(df_on)+1)
-df_off["SubNoteNr"] = range(1, len(df_off)+1)
-
-# %% assign subcomponents
-""" for phrase in ls_forte:
-    for note in phrase:
-        df_on["Subcomponent"][df_on["SubNoteNr"] == note] = "Forte"
-        df_off["Subcomponent"][df_off["SubNoteNr"] == note] = "Forte"
-
-for phrase in ls_piano:
-    for note in phrase:
-        df_on["Subcomponent"][df_on["SubNoteNr"] == note] = "Piano"
[…]

### 2020-05-13 material/playback.py

@@ -1,0 +1,43 @@
+#!/usr/local/bin/python
+####################
+# Created: 20/04/2020
+# This script converts a txt file to midi data and playbacks on Max MSP.
+
+# %% import packages
+import csv, time, mido
+import pandas as pd
+import numpy as np
+
+#%% open port (play sounds with Max MSP)
+port = mido.open_output('to Max 1')
+
+#　%% playback
+with open("./stim_n/6_instance.txt") as csvfile:
+    current = csv.reader(csvfile, delimiter = ",")
+    next(current) # skip first row
+    counter = 0
+    for row in current:
+        if counter == 0:
+            previousTime = 0
+            counter += 1
+        else:
+            previousTime = currentTime
+        currentTime = int(row[3])*0.001
+        currentPitch = int(row[5])
+        currentVelocity = int(row[4])
+        currentOnOff = int(row[2])
+        # assign midi values
+        if currentOnOff == 1:
+            print(currentOnOff)
+            msg = mido.Message('note_on', note = currentPitch, velocity = currentVelocity, time = currentTime)
+            print(msg)
+            time.sleep(currentTime-previousTime)
+            print(msg.time)
+            port.send(msg)
+        elif currentOnOff == 0:
+            msg = mido.Message('note_off', note = currentPitch, velocity = currentVelocity, time = currentTime)
+            print(msg)
+            time.sleep(currentTime-previousTime)
+            print(msg.time)
+            port.send(msg)
+\

### 2020-05-13 material/playback.py
remove typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -40,4 +40,3 @@ with open("./stim_n/6_instance.txt") as csvfile:
             time.sleep(currentTime-previousTime)
             print(msg.time)
             port.send(msg)
-\

### 2020-05-13 material/stim_n_norm/average.R
duration from normal distribuion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -14,27 +14,35 @@ theme_set(theme_classic())
 theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
 ## function
 source("./function.R")
+# set seed
+set.seed(124)
+
+# # read csv/txt
+# dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+# dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+# #onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
+# #offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
+# duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
+# kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
+# dt_ideal <- fread("./ideal.txt", header = F)
+# 
+# # duration and kv
+# valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
+# valid_du_kv$Duplicated <- duplicated(valid_du_kv)
+# 
+# # valid performances
+# valid <- valid_du_kv[Duplicated == TRUE]
+# 
+# ### create 6 instances! ###
+# valid$Sample <- sample(c(1:31), replace = FALSE)
+# print(valid)
+# fwrite(valid, "./valid.txt")
 
 # read csv/txt
 dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
 dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
-#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
-#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
[…]

### 2020-05-13 material/stim_n_original/average.R
change folder
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,140 @@
+# /usr/local/bin/r
+# Average onset, duration, key velocity for normative performance
+# Created: 06/05/2020
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+
+# setup
+## ggplots
+theme_set(theme_classic())
+theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
+## function
+source("./function.R")
+
+# read csv/txt
+dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
+#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
+duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
+kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
+dt_ideal <- fread("./ideal.txt", header = F)
+
+# duration and kv
+valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
+valid_du_kv$Duplicated <- duplicated(valid_du_kv)
+
+# valid performances
+valid <- valid_du_kv[Duplicated == TRUE]
+
+### create 6 instances! ###
+valid$Sample <- sample(c(1:31), replace = FALSE)
+print(valid)
+fwrite(valid, "./valid.txt")
+
+# 1. average IOIs
+dt_ioi_instance  <- data.table()
+counter = 0
+for (i in 1:6){
+  stim <- combining_onset(valid, i)
+  
+  # calculate normIOI
+  stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
+  stim$normIOI <- stim$IOI/stim$Tempo
+  # IOI = 0 if RowNr == 1
[…]

### 2020-05-13 material/analysis/expression/stim_a/ioi_duration.Rmd
edit outlines
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ### Sequence plot
+- Right wrap: Subject Number
+
 #### All

### 2020-05-13 material/analysis/expression/stim_d/kv.Rmd
add seq plot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-#### Sequence plot
+### Sequence plot
 - Right wrap: Subject Number
+
+#### All
+```{r kv_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+# for each individual
+dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
+kv_seq <- dt_kv[Subcomponent != "NA", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]
+
+p_kv_seq <- ggplot(data = kv_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
+  geom_line() +
+  geom_point() +
+  facet_grid(SubNr ~ .) +
+  labs(x = "RowNr", y = "Velocity", subtitle = "KV: Performing-Dynamics(All)") +
+  scale_x_continuous(breaks=seq(1,66,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
+p_kv_seq
+```
 

### 2020-05-19 material/analysis/expression/stim_a/ioi_onset.Rmd
update layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Analysis for Non-Expressive performances (ONSET)'
+title: 'Analysis for Performing-Articulation performances (ONSET)'
 output:
@@
 
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (onsets).
+Description: This is a rough overview of performing performance (IOI-Onset) from teaching-v2.0.
 
@@
 ```
+- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`

### 2020-05-19 material/analysis/expression/stim_a/ioi_duration.Rmd
update layout
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Analysis for Performing-Articulation performances (DURATION)'
+title: 'Duration Analysis for Performing-Articulation performances'
 output:
@@
 
-duration_seq_stim_n <- fread("../../stim_n/seq.csv", header = T, sep = ",", dec = ".")
-duration_seq_stim_n$SubNr <- factor(duration_seq_stim_n$SubNr)
-duration_seq_stim_n$TrialNr <- factor(duration_seq_stim_n$TrialNr)
-duration_seq_stim_n$TrialNr <- "Practice"
-
-duration_seq <- rbind(duration_seq, duration_seq_stim_n)
-
 p_duration_seq <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
@@
 
-#### Distribution plot
+### Distribution plot
 #### Legato
@@
 
-### Exclude outlier participants
+### Valid performances (exclude outliers)
 #### Legato
@@
 ```
+- The number of excluded performance: `r duration_all[Subcomponent == "Legato"]$N-nrow(excluded_duration_leg)`
 
@@
 print(excluded_duration_sta)
+```
+- The number of excluded performance: `r duration_all[Subcomponent == "Staccato"]$N-nrow(excluded_duration_sta)`
+
+```{r export, include = FALSE}
+duration_valid <- dt_duration[, .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr)]
+fwrite(duration_valid, "./duration_valid.csv")
 ```

### 2020-05-19 material/analysis/expression/stim_a/ioi_duration.Rmd
add comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance from teaching-v2.0 (duration).
+Description: This is a rough overview of performing performance (duration) from teaching-v2.0.
 
@@
 
-### Sequence plot
+#### Sequence plot
 - Right wrap: Subject Number
@@
 
-### Distribution plot
+#### Distribution plot
+- Right wrap: Trial Number
+- Upper wrap: Subject Number
+
 #### Legato

### 2020-05-19 material/analysis/stim_n/duration.Rmd
add comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Duration - Analysis for Non-Expressive performances'
+title: 'Duration Analysis for Non-Expressive performances'
 output:
@@
 #### Sequence plot
+- Right wrap: Subject Number
+
 ```{r duration_seq, fig.width = 14, fig.height = 12, echo = FALSE}
@@
 #### Distribution plot
+- Right wrap: Trial Number
+- Upper wrap: Subject Number
+
 ```{r duration_dis, fig.width = 12, echo = FALSE}

### 2020-05-19 material/analysis/stim_n/duration.Rmd
correc x axis ticks
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Non-Expression") +
-  scale_x_continuous(breaks=seq(1,66,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
+  scale_x_continuous(breaks=seq(1,72,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
 p_duration_seq

### 2020-05-19 material/analysis/stim_n/ioi_onset.Rmd
correc x axis ticks
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   facet_grid(SubNr ~ .) +
-  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,71,1))
 p_ioi_seq

### 2020-05-19 material/analysis/stim_n/kv.Rmd
correc x axis ticks
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,72,1))
 p_kv_seq

### 2020-05-20 material/analysis/expression/stim_d/kv.Rmd
export valid trials
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-04-16 material/analysis/expression/stim_d/kv.Rmd
@@
 ---
-title: 'Analysis for Performing-Dynamics performances (Key Velocity)'
+title: 'KV Analysis for Performing-Dynamics performances'
 output:
@@
 
-Description: This is a rough overview of non-expressive performance from teaching-v2.0 (KV).
+Description: This is a rough overview of performing performance (KV) from teaching-v2.0.
 
@@
 
+#### All
+```{r kv_seq, fig.width = 14, fig.height = 12, echo = FALSE}
+# for each individual
+dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
+kv_seq <- dt_kv[Subcomponent != "NA", .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr, RowNr)]
+
+p_kv_seq <- ggplot(data = kv_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
+  geom_line() +
+  geom_point() +
+  facet_grid(SubNr ~ .) +
+  labs(x = "RowNr", y = "Velocity", subtitle = "KV: Performing-Dynamics(All)") +
+  scale_x_continuous(breaks=seq(1,72,1))
+p_kv_seq
+```
+
 #### Forte
@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,72,1))
 p_kv_seq_for
@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
+  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,72,1))
 p_kv_seq_pia
@@
 
+#### Forte
 ```{r kv_dis_for, fig.width = 14, fig.height = 8, echo = FALSE}
@@
 
+#### Piano
 ```{r kv_dis_pia, fig.width = 14, fig.height = 8, echo = FALSE}
@@
 
-### Exclude outlier participants
+### Valid performances (exclude outliers)
 #### Forte
@@
 # for forte
-excluded_duration_for <- kv[Subcomponent == "Forte" & Mean < kv_all[Subcomponent == "Forte"]$Mean+3*kv_all[Subcomponent == "Forte"]$SD & Mean > kv_all[Subcomponent == "Forte"]$Mean-3*kv_all$SD]
-print(excluded_duration_for)
+excluded_kv_for <- kv[Subcomponent == "Forte" & Mean < kv_all[Subcomponent == "Forte"]$Mean+3*kv_all[Subcomponent == "Forte"]$SD & Mean > kv_all[Subcomponent == "Forte"]$Mean-3*kv_all$SD]
+print(excluded_kv_for)
 ```
+- The number of excluded performance: `r kv_all[Subcomponent == "Forte"]$N-nrow(excluded_kv_for)`
 
@@
 # for piano
-excluded_duration_pia <- kv[Subcomponent == "Piano" & Mean < kv_all[Subcomponent == "Piano"]$Mean+3*kv_all[Subcomponent == "Piano"]$SD & Mean > kv_all[Subcomponent == "Piano"]$Mean-3*kv_all$SD]
-print(excluded_duration_pia)
+excluded_kv_pia <- kv[Subcomponent == "Piano" & Mean < kv_all[Subcomponent == "Piano"]$Mean+3*kv_all[Subcomponent == "Piano"]$SD & Mean > kv_all[Subcomponent == "Piano"]$Mean-3*kv_all$SD]
+print(excluded_kv_pia)
+```
+- The number of excluded performance: `r kv_all[Subcomponent == "Piano"]$N-nrow(excluded_kv_pia)`
+
+```{r export, include = FALSE}
+kv_valid <- dt_kv[, .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr)]
+fwrite(kv_valid, "./kv_valid.csv")
 ```

### 2020-05-20 material/analysis/expression/stim_d/ioi_onset.Rmd
add ioi-onset, duration analyses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/expression/stim_a/ioi_onset.Rmd
@@
 ---
-title: 'Analysis for Performing-Articulation performances (ONSET)'
+title: 'IOI-Onset Analysis for Performing-Dynamics performances'
 output:
@@
 dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
-# include only articulation
-dt_ioi <- dt_onset[Condition == "performing" & Skill == "articulation"]
+# include only dynamics
+dt_ioi <- dt_onset[Condition == "performing" & Skill == "dynamics"]
 
@@
   geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
-  labs(y = "Normalised IOI", subtitle = "IOI: Performing-Articulation")
+  labs(y = "Normalised IOI", subtitle = "IOI: Performing-Dynamics")
 p_ioi
@@
   facet_grid(TrialNr ~ SubNr) +
-  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Performing-Articulation")
+  labs(x = "Normalised IOI", y = "Count", subtitle = "IOI: Performing-Dynamics")
 p_ioi_dis

### 2020-05-20 material/stim_n_original/average.R
adjust scripts to 8 instances
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -5,15 +5,13 @@
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-# plot
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
 
-# setup
-## ggplots
-theme_set(theme_classic())
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
-## function
+# function
 source("./function.R")
+
+# create a folder to store stimuli
+foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
+dir.create(foldername)
 
 # read csv/txt
 dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
@@ -31,15 +29,15 @@ valid_du_kv$Duplicated <- duplicated(valid_du_kv)
 # valid performances
 valid <- valid_du_kv[Duplicated == TRUE]
 
-### create 6 instances! ###
+### create 8 instances! ###
 valid$Sample <- sample(c(1:31), replace = FALSE)
 print(valid)
-fwrite(valid, "./valid.txt")
+fwrite(valid, paste(foldername, "valid.txt", sep = ""))
 
 # 1. average IOIs
 dt_ioi_instance  <- data.table()
 counter = 0
-for (i in 1:6){
+for (i in 1:8){
   stim <- combining_onset(valid, i)
   
   # calculate normIOI
@@ -56,13 +54,16 @@ for (i in 1:6){
   dt_ioi_instance <- rbind(dt_ioi_instance, stim_average)
   
   # next instance
-  counter = counter+4
+  counter = counter+2
   }
+
+# export dt_ioi_instance
+fwrite(dt_ioi_instance, paste(foldername, "dt_ioi_instance.txt", sep = ""))
 
 # 2. average duration
 dt_du_instance <- data.table()
 counter = 0
-for (i in c(1:6)){
+for (i in c(1:8)){
   stim <- combining_onset(valid, i) #onset
   stim_offset <- combining_offset(valid, i)
   
@@ -78,13 +79,16 @@ for (i in c(1:6)){
   dt_du_instance <- rbind(dt_du_instance, stim_average)
   
   # next instance
-  counter = counter+4
+  counter = counter+2
 }
+
+# export dt_du_instance
[…]

### 2020-05-20 material/stim_n/20_21/average.R
change location
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -5,49 +5,40 @@
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-# plot
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
 
-# setup
-## ggplots
-theme_set(theme_classic())
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
-## function
+# function
 source("./function.R")
-# set seed
-set.seed(124)
 
-# # read csv/txt
-# dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
-# dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
-# #onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
-# #offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
-# duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
-# kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
-# dt_ideal <- fread("./ideal.txt", header = F)
-# 
-# # duration and kv
-# valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
-# valid_du_kv$Duplicated <- duplicated(valid_du_kv)
-# 
-# # valid performances
-# valid <- valid_du_kv[Duplicated == TRUE]
-# 
-# ### create 6 instances! ###
-# valid$Sample <- sample(c(1:31), replace = FALSE)
-# print(valid)
-# fwrite(valid, "./valid.txt")
+# create a folder to store stimuli
+foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
+dir.create(foldername)
 
 # read csv/txt
 dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
[…]

### 2020-05-20 material/stim_n/original/average.R

@@ -30,6 +30,7 @@ valid_du_kv$Duplicated <- duplicated(valid_du_kv)
 valid <- valid_du_kv[Duplicated == TRUE]
 
 ### create 8 instances! ###
+# random sampling
 valid$Sample <- sample(c(1:31), replace = FALSE)
 print(valid)
 fwrite(valid, paste(foldername, "valid.txt", sep = ""))

### 2020-05-20 material/stim_a/function.R
update function.R for averaging three performances
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,4 +1,4 @@
-# averaging function (create 6 instances)
+# averaging function (create 8 instances - averaging 3 performances)
 # packages
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 
@@ -9,20 +9,14 @@ combining_onset <- function(valid, number){
   SubNr1 <- valid[Sample == number+counter]$SubNr
   SubNr2 <- valid[Sample == number+1+counter]$SubNr
   SubNr3 <- valid[Sample == number+2+counter]$SubNr
-  SubNr4 <- valid[Sample == number+3+counter]$SubNr
-  SubNr5 <- valid[Sample == number+4+counter]$SubNr
   TrialNr1 <- valid[Sample == number+counter]$TrialNr
   TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
   TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
-  TrialNr4 <- valid[Sample == number+3+counter]$TrialNr
-  TrialNr5 <- valid[Sample == number+4+counter]$TrialNr
   stim_1 <- dt_onset[SubNr == SubNr1 & TrialNr == TrialNr1]
   stim_2 <- dt_onset[SubNr == SubNr2 & TrialNr == TrialNr2]
   stim_3 <- dt_onset[SubNr == SubNr3 & TrialNr == TrialNr3]
-  stim_4 <- dt_onset[SubNr == SubNr4 & TrialNr == TrialNr4]
-  stim_5 <- dt_onset[SubNr == SubNr5 & TrialNr == TrialNr5]
-  outcome <- rbind(stim_1, stim_2, stim_3, stim_4, stim_5)
-  outcome$RowNr <- rep(c(1:72), 5)
+  outcome <- rbind(stim_1, stim_2, stim_3)
+  outcome$RowNr <- rep(c(1:72), 3)
   
   # tempo labels
   outcome[Tempo == 120]$Tempo <- 250
@@ -35,20 +29,14 @@ combining_offset <- function(valid, number){
   SubNr1 <- valid[Sample == number+counter]$SubNr
   SubNr2 <- valid[Sample == number+1+counter]$SubNr
   SubNr3 <- valid[Sample == number+2+counter]$SubNr
-  SubNr4 <- valid[Sample == number+3+counter]$SubNr
[…]

### 2020-05-20 material/stim_a/average.R
averaging script for duration (use stim_n data for ioi, kv)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,101 @@
+# /usr/local/bin/r
+# Average onset, duration, key velocity for normative performance
+# Created: 06/05/2020
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# function
+source("./function.R")
+
+# create a folder to store stimuli
+foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
+dir.create(foldername)
+
+# read csv/txt
+dt_onset <- fread("../analysis/expression/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../analysis/expression/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+# only for performing/articulation
+dt_onset <- dt_onset[Condition == "performing" & Skill == "articulation"]
+dt_offset <- dt_offset[Condition == "performing" & Skill == "articulation"]
+# use normative performance for ioi, kv
+dt_ioi_instance <- fread("../stim_n/20_21/1589957466-200520/dt_ioi_instance.txt", header = T, sep = ",", dec = ".")
+dt_kv_instance <- fread("../stim_n/20_21/1589957466-200520/dt_kv_instance.txt", header = T, sep = ",", dec = ".")
+# use performing performance for duration
+duration_valid <- fread("../analysis/expression/stim_a/duration_valid.csv", header = T, sep = ",", dec = ".")
+dt_ideal <- fread("./ideal.txt", header = F)
+
+# valid duration performances
+valid <- duration_valid[, c("SubNr", "TrialNr")]
+
+### create 8 instances! ###
+valid$Sample <- sample(c(1:nrow(valid)), replace = FALSE)
+print(valid)
+fwrite(valid, paste(foldername, "valid.txt", sep = ""))
+
+# 1. average duration
+dt_du_instance <- data.table()
+counter = 0
+for (i in c(1:8)){
+  stim <- combining_onset(valid, i) #onset
+  stim_offset <- combining_offset(valid, i)
+  
+  # calculate Duration
+  stim$Duration <- stim_offset$TimeStamp - stim$TimeStamp
+  stim$normDu <- stim$Duration/stim$Tempo
+  
+  # average normDu
[…]

### 2020-05-20 material/stim_d/average.R
upload averaging scripts for kv
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,96 @@
+# /usr/local/bin/r
+# Average onset, duration, key velocity for normative performance
+# Created: 06/05/2020
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# function
+source("./function.R")
+
+# create a folder to store stimuli
+foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
+dir.create(foldername)
+
+# read csv/txt
+dt_onset <- fread("../analysis/expression/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../analysis/expression/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+# only for performing/articulation
+dt_onset <- dt_onset[Condition == "performing" & Skill == "dynamics"]
+dt_offset <- dt_offset[Condition == "performing" & Skill == "dynamics"]
+# use normative performance for ioi, kv
+dt_ioi_instance <- fread("../stim_n/20_21/1589957466-200520/dt_ioi_instance.txt", header = T, sep = ",", dec = ".")
+dt_du_instance <- fread("../stim_n/20_21/1589957466-200520/dt_du_instance.txt", header = T, sep = ",", dec = ".")
+# use performing performance for duration
+kv_valid <- fread("../analysis/expression/stim_d/kv_valid.csv", header = T, sep = ",", dec = ".")
+dt_ideal <- fread("./ideal.txt", header = F)
+
+# valid kv performances
+valid <- kv_valid[, c("SubNr", "TrialNr")]
+
+### create 8 instances! ###
+valid$Sample <- sample(c(1:nrow(valid)), replace = FALSE)
+print(valid)
+fwrite(valid, paste(foldername, "valid.txt", sep = ""))
+
+# 1. average kv
+dt_kv_instance <- data.table()
+counter = 0
+for (i in c(1:8)){
+  stim <- combining_onset(valid, i)
+  
+  # average KV
+  stim_average <- stim[, .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(RowNr)]
+  # label instance no
+  stim_average$Instance <- as.character(i)
+  # add to dt_du_instance
[…]

### 2020-05-20 material/stim_d/function.R

@@ -1,0 +1,47 @@
+# averaging function (create 8 instances - averaging 3 performances)
+# packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# valid: information(data.table) about valid performances
+# number: loop number (if outside loop, enter 0)
+
+combining_onset <- function(valid, number){
+  SubNr1 <- valid[Sample == number+counter]$SubNr
+  SubNr2 <- valid[Sample == number+1+counter]$SubNr
+  SubNr3 <- valid[Sample == number+2+counter]$SubNr
+  TrialNr1 <- valid[Sample == number+counter]$TrialNr
+  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
+  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
+  stim_1 <- dt_onset[SubNr == SubNr1 & TrialNr == TrialNr1]
+  stim_2 <- dt_onset[SubNr == SubNr2 & TrialNr == TrialNr2]
+  stim_3 <- dt_onset[SubNr == SubNr3 & TrialNr == TrialNr3]
+  outcome <- rbind(stim_1, stim_2, stim_3)
+  outcome$RowNr <- rep(c(1:72), 3)
+  
+  # tempo labels
+  outcome[Tempo == 120]$Tempo <- 250
+  outcome[Tempo == 110]$Tempo <- 273
+  outcome[Tempo == 100]$Tempo <- 300
+  return(outcome)
+}
+
+combining_offset <- function(valid, number){
+  SubNr1 <- valid[Sample == number+counter]$SubNr
+  SubNr2 <- valid[Sample == number+1+counter]$SubNr
+  SubNr3 <- valid[Sample == number+2+counter]$SubNr
+  TrialNr1 <- valid[Sample == number+counter]$TrialNr
+  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
+  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
+  stim_1 <- dt_offset[SubNr == SubNr1 & TrialNr == TrialNr1]
+  stim_2 <- dt_offset[SubNr == SubNr2 & TrialNr == TrialNr2]
+  stim_3 <- dt_offset[SubNr == SubNr3 & TrialNr == TrialNr3]
+  outcome <- rbind(stim_1, stim_2, stim_3)
+  outcome$RowNr <- rep(c(1:72), 3)
[…]

### 2020-05-20 material/playback.py
save playback.py
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -12,7 +12,7 @@ import numpy as np
 port = mido.open_output('to Max 1')
 
 #　%% playback
-with open("./stim_n/6_instance.txt") as csvfile:
+with open("./stim_d/1589959883-200520/3_instance.txt") as csvfile:
     current = csv.reader(csvfile, delimiter = ",")
     next(current) # skip first row
     counter = 0
@@ -40,3 +40,6 @@ with open("./stim_n/6_instance.txt") as csvfile:
             time.sleep(currentTime-previousTime)
             print(msg.time)
             port.send(msg)
+
+
+# %%

### 2020-05-22 material/playback.py
implemeting automatic saving function for midi TBC
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -11,35 +11,49 @@ import numpy as np
 #%% open port (play sounds with Max MSP)
 port = mido.open_output('to Max 1')
 
-#　%% playback
-with open("./stim_d/1589959883-200520/3_instance.txt") as csvfile:
-    current = csv.reader(csvfile, delimiter = ",")
-    next(current) # skip first row
-    counter = 0
-    for row in current:
-        if counter == 0:
-            previousTime = 0
-            counter += 1
-        else:
-            previousTime = currentTime
-        currentTime = int(row[3])*0.001
-        currentPitch = int(row[5])
-        currentVelocity = int(row[4])
-        currentOnOff = int(row[2])
-        # assign midi values
-        if currentOnOff == 1:
-            print(currentOnOff)
-            msg = mido.Message('note_on', note = currentPitch, velocity = currentVelocity, time = currentTime)
-            print(msg)
-            time.sleep(currentTime-previousTime)
-            print(msg.time)
-            port.send(msg)
-        elif currentOnOff == 0:
-            msg = mido.Message('note_off', note = currentPitch, velocity = currentVelocity, time = currentTime)
-            print(msg)
-            time.sleep(currentTime-previousTime)
-            print(msg.time)
-            port.send(msg)
+# %% read files in a directory
+files = [f for f in os.listdir("./stim_n/20_21/1589957466-200520") if not f.startswith("dt_") and not f.startswith("valid")]
+
+#　%% playback and save mid.file
+
+for file in files:
+    with open("./stim_d/1589959883-200520/"+file) as csvfile:
+        current = csv.reader(csvfile, delimiter = ",")
+        next(current) # skip first row
+        mid = mido.MidiFile()
+        track = mido.MidiTrack()
+        mid.tracks.append(track)
+        track.append(mido.Message('program_change', program=12, time=0))
+
+        counter = 0
+        for row in current:
+            if counter == 0:
+                previousTime = 0
+                counter += 1
+            else:
+                previousTime = currentTime
+            currentTime = int(row[3])*0.001
+            currentPitch = int(row[5])
+            currentVelocity = int(row[4])
+            currentOnOff = int(row[2])
+            # assign midi values
+            if currentOnOff == 1:
+                print(currentOnOff)
+                msg = mido.Message('note_on', note = currentPitch, velocity = currentVelocity, time = currentTime)
+                track.append(msg)
+                print(msg)
+                time.sleep(currentTime-previousTime)
+                print(msg.time)
[…]

### 2020-08-07 material/stimuli/playback.py
implement mid export
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -23,8 +23,22 @@ folders = [folder_low, folder_art, folder_dyn, folder_high]
 #%% playback
 for folder in folders:
     for instance in range(1, 17):
-        filename = folder + instance + "_instance.txt"
+        filename = folder + str(instance) + "_instance.txt"
         with open(filename) as csvfile:
+            mid = mido.MidiFile()
+            track = mido.MidiTrack()
+            mid.tracks.append(track)
+            if folder == "./low/1596207520-310720/":
+                midname = "./mid/low_" + str(instance) + ".mid"
+            elif folder == "./art/1596379937-020820/":
+                midname = "./mid/art_" + str(instance) + ".mid"
+            elif folder == "./dyn/1596379899-020820/":
+                midname = "./mid/dyn_" + str(instance) + ".mid"
+            elif folder == "./high/1596380946-020820/":
+                midname = "./mid/high_" + str(instance) + ".mid"
+
+            track.append(mido.Message('program_change', program=12, time=0))
+
             current = csv.reader(csvfile, delimiter = ",")
             next(current) # skip first row
             counter = 0
@@ -34,61 +48,19 @@ for folder in folders:
                     counter += 1
                 else:
                     previousTime = currentTime
-                currentTime = int(row[3])*0.001+3 # 3 sec delay at the beginning
+                currentTime = int(mido.second2tick(int(row[3])*0.001+3, 480, 500000))
+                print(currentTime-previousTime)
                 currentPitch = int(row[5])
                 currentVelocity = int(row[4])
                 currentOnOff = int(row[2])
                 # assign midi values
                 if currentOnOff == 1:
-                    print(currentOnOff)
-                    msg = mido.Message('note_on', note = currentPitch, velocity = currentVelocity, time = currentTime)
-                    print(msg)
-                    time.sleep(currentTime-previousTime)
-                    print(msg.time)
-                    port.send(msg)
+                    #print(currentOnOff)
+                    track.append(mido.Message('note_on', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
                 elif currentOnOff == 0:
-                    msg = mido.Message('note_off', note = currentPitch, velocity = currentVelocity, time = currentTime)
-                    print(msg)
-                    time.sleep(currentTime-previousTime)
-                    print(msg.time)
-                    port.send(msg)
[…]

### 2020-08-07 material/stimuli/mid_export.py
rename
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -61,6 +61,4 @@ for folder in folders:
                     #print(currentOnOff)
                     track.append(mido.Message('note_off', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
             mid.save(midname)
-
-
 # %%

### 2020-08-07 material/stimuli/mid_export.py
change program to 0 (Acoustic Grand Piano)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -8,8 +8,7 @@ import os, csv, time, mido
 import pandas as pd
 import numpy as np
 
-#%% open port (play sounds with Max MSP)
-port = mido.open_output('to Max 1')
+#%% create mid folder if not existed
 if not os.path.exists("mid"):
     os.mkdir("mid")
 
@@ -20,14 +19,18 @@ folder_dyn = "./dyn/1596379899-020820/"
 folder_high = "./high/1596380946-020820/"
 folders = [folder_low, folder_art, folder_dyn, folder_high]
 
-#%% playback
+#%% mid export
 for folder in folders:
     for instance in range(1, 17):
         filename = folder + str(instance) + "_instance.txt"
         with open(filename) as csvfile:
+            # create mid file
             mid = mido.MidiFile()
             track = mido.MidiTrack()
             mid.tracks.append(track)
+            track.append(mido.Message('program_change', program=0, time=0)) #program 0 = Acoustic Grand Piano
+
+            # name for mid file
             if folder == "./low/1596207520-310720/":
                 midname = "./mid/low_" + str(instance) + ".mid"
             elif folder == "./art/1596379937-020820/":
@@ -36,9 +39,7 @@ for folder in folders:
                 midname = "./mid/dyn_" + str(instance) + ".mid"
             elif folder == "./high/1596380946-020820/":
                 midname = "./mid/high_" + str(instance) + ".mid"
-
-            track.append(mido.Message('program_change', program=12, time=0))
-
+                
             current = csv.reader(csvfile, delimiter = ",")
             next(current) # skip first row
             counter = 0
@@ -49,16 +50,15 @@ for folder in folders:
                 else:
                     previousTime = currentTime
                 currentTime = int(mido.second2tick(int(row[3])*0.001+3, 480, 500000))
-                print(currentTime-previousTime)
                 currentPitch = int(row[5])
                 currentVelocity = int(row[4])
                 currentOnOff = int(row[2])
                 # assign midi values
                 if currentOnOff == 1:
-                    #print(currentOnOff)
                     track.append(mido.Message('note_on', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
                 elif currentOnOff == 0:
-                    #print(currentOnOff)
                     track.append(mido.Message('note_off', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
             mid.save(midname)
+
+print("Done :D")
 # %%

### 2020-08-07 material/stimuli/averaging/low/old/20_21/average.R
move files to averaging
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,163 @@
+# /usr/local/bin/r
+# Average onset, duration, key velocity for normative performance
+# Created: 06/05/2020
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# function
+source("./function.R")
+
+# create a folder to store stimuli
+foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
+dir.create(foldername)
+
+# read csv/txt
+dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
+#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
+duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
+kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
+dt_ideal <- fread("./ideal.txt", header = F)
+
+# duration and kv
+valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
+valid_du_kv$Duplicated <- duplicated(valid_du_kv)
+
+# valid performances
+valid <- valid_du_kv[Duplicated == TRUE]
+
+### create 8 instances! ###
+# random sampling
+valid$Sample <- sample(c(1:31), replace = FALSE)
+print(valid)
+fwrite(valid, paste(foldername, "valid.txt", sep = ""))
+
+# 1. average IOIs
+dt_ioi_instance  <- data.table()
+counter = 0
+for (i in 1:8){
+  stim <- combining_onset(dt_onset, valid, i)
+  
+  # calculate normIOI
+  stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
+  stim$normIOI <- stim$IOI/stim$Tempo
+  # IOI = 0 if RowNr == 1
+  stim[RowNr == 1]$normIOI <- 0
+  
[…]

### 2020-08-07 material/stimuli/averaging/low/old/original/average.R

@@ -1,0 +1,148 @@
+# /usr/local/bin/r
+# Average onset, duration, key velocity for normative performance
+# Created: 06/05/2020
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# function
+source("./function.R")
+
+# create a folder to store stimuli
+foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
+dir.create(foldername)
+
+# read csv/txt
+dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
+dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
+#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
+#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
+duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
+kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
+dt_ideal <- fread("./ideal.txt", header = F)
+
+# duration and kv
+valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
+valid_du_kv$Duplicated <- duplicated(valid_du_kv)
+
+# valid performances
+valid <- valid_du_kv[Duplicated == TRUE]
+
+### create 8 instances! ###
+# random sampling
+valid$Sample <- sample(c(1:31), replace = FALSE)
+print(valid)
+fwrite(valid, paste(foldername, "valid.txt", sep = ""))
+
+# 1. average IOIs
+dt_ioi_instance  <- data.table()
+counter = 0
+for (i in 1:8){
+  stim <- combining_onset(dt_onset, valid, i)
+  
+  # calculate normIOI
+  stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
+  stim$normIOI <- stim$IOI/stim$Tempo
+  # IOI = 0 if RowNr == 1
+  stim[RowNr == 1]$normIOI <- 0
+  
[…]

### 2020-08-07 material/stimuli/averaging/low/old/original/function.R

@@ -1,0 +1,47 @@
+# averaging function (create 8 instances - averaging 3 performances)
+# packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# valid: information(data.table) about valid performances
+# number: loop number (if outside loop, enter 0)
+
+combining_onset <- function(dt_onset, valid, number){
+  SubNr1 <- valid[Sample == number+counter]$SubNr
+  SubNr2 <- valid[Sample == number+1+counter]$SubNr
+  SubNr3 <- valid[Sample == number+2+counter]$SubNr
+  TrialNr1 <- valid[Sample == number+counter]$TrialNr
+  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
+  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
+  stim_1 <- dt_onset[SubNr == SubNr1 & TrialNr == TrialNr1]
+  stim_2 <- dt_onset[SubNr == SubNr2 & TrialNr == TrialNr2]
+  stim_3 <- dt_onset[SubNr == SubNr3 & TrialNr == TrialNr3]
+  outcome <- rbind(stim_1, stim_2, stim_3)
+  outcome$RowNr <- rep(c(1:72), 3)
+  
+  # tempo labels
+  outcome[Tempo == 120]$Tempo <- 250
+  outcome[Tempo == 110]$Tempo <- 273
+  outcome[Tempo == 100]$Tempo <- 300
+  return(outcome)
+}
+
+combining_offset <- function(dt_offset, valid, number){
+  SubNr1 <- valid[Sample == number+counter]$SubNr
+  SubNr2 <- valid[Sample == number+1+counter]$SubNr
+  SubNr3 <- valid[Sample == number+2+counter]$SubNr
+  TrialNr1 <- valid[Sample == number+counter]$TrialNr
+  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
+  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
+  stim_1 <- dt_offset[SubNr == SubNr1 & TrialNr == TrialNr1]
+  stim_2 <- dt_offset[SubNr == SubNr2 & TrialNr == TrialNr2]
+  stim_3 <- dt_offset[SubNr == SubNr3 & TrialNr == TrialNr3]
+  outcome <- rbind(stim_1, stim_2, stim_3)
[…]

### 2020-08-07 material/stimuli/mid_export.py

@@ -13,10 +13,10 @@ if not os.path.exists("mid"):
     os.mkdir("mid")
 
 #%% folder names
-folder_low = "./low/1596207520-310720/"
-folder_art = "./art/1596379937-020820/"
-folder_dyn = "./dyn/1596379899-020820/"
-folder_high = "./high/1596380946-020820/"
+folder_low = "./averaging/low/1596207520-310720/"
+folder_art = "./averaging/art/1596379937-020820/"
+folder_dyn = "./averaging/dyn/1596379899-020820/"
+folder_high = "./averaging/high/1596380946-020820/"
 folders = [folder_low, folder_art, folder_dyn, folder_high]
 
 #%% mid export
@@ -31,13 +31,13 @@ for folder in folders:
             track.append(mido.Message('program_change', program=0, time=0)) #program 0 = Acoustic Grand Piano
 
             # name for mid file
-            if folder == "./low/1596207520-310720/":
+            if folder == "./averaging/low/1596207520-310720/":
                 midname = "./mid/low_" + str(instance) + ".mid"
-            elif folder == "./art/1596379937-020820/":
+            elif folder == "./averaging/art/1596379937-020820/":
                 midname = "./mid/art_" + str(instance) + ".mid"
-            elif folder == "./dyn/1596379899-020820/":
+            elif folder == "/averaging/dyn/1596379899-020820/":
                 midname = "./mid/dyn_" + str(instance) + ".mid"
-            elif folder == "./high/1596380946-020820/":
+            elif folder == "./averaging/high/1596380946-020820/":
                 midname = "./mid/high_" + str(instance) + ".mid"
                 
             current = csv.reader(csvfile, delimiter = ",")

### 2020-08-08 material/stimuli/analysis/expression/stim_a/kv.Rmd
move analysis into stimuli folder
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-20 material/analysis/expression/stim_d/kv.Rmd
@@
 ---
-title: 'KV Analysis for Performing-Dynamics performances'
+title: 'KV Analysis for Performing-Articulation performances'
 output:
@@
 # include only articulation
-dt_kv <- dt_kv[Condition == "performing" & Skill == "dynamics"]
+dt_kv <- dt_kv[Condition == "performing" & Skill == "articulation"]
 
@@
 
-# for forte
-for (phrase in 1:length(ls_forte2)){
-  for (note in 1:length(ls_forte2[[phrase]])){
-    dt_kv[RowNr == ls_forte2[[phrase]][note]]$Subcomponent <- "Forte"
+# for legato
+for (phrase in 1:length(ls_legato2)){
+  for (note in 1:length(ls_legato2[[phrase]])){
+    dt_kv[RowNr == ls_legato2[[phrase]][note]]$Subcomponent <- "Legato"
   }
@@
 
-# for piano
-for (phrase in 1:length(ls_piano2)){
-  for (note in 1:length(ls_piano2[[phrase]])){
-    dt_kv[RowNr == ls_piano2[[phrase]][note]]$Subcomponent <- "Piano"
+# for staccato
+for (phrase in 1:length(ls_staccato2)){
+  for (note in 1:length(ls_staccato2[[phrase]])){
+    dt_kv[RowNr == ls_staccato2[[phrase]][note]]$Subcomponent <- "Staccato"
   }
@@
 # For each individual
-kv <- dt_kv[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Subcomponent)]
+kv <- dt_kv[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Subcomponent)]
 print(kv)
@@
   facet_grid(Subcomponent ~ .) +
-  labs(y = "Key Velocity", subtitle = "KV: Performing-Dynamics")
+  labs(y = "Key Velocity", subtitle = "KV: Performing-Articulation")
 p_kv
@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Velocity", subtitle = "KV: Performing-Dynamics(All)") +
+  labs(x = "RowNr", y = "Velocity", subtitle = "KV: Performing-Articulation(All)") +
   scale_x_continuous(breaks=seq(1,72,1))
@@
 
-#### Forte
+#### Legato
 ```{r kv_seq_for, fig.width = 14, fig.height = 12, echo = FALSE}
@@
 dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
-kv_seq_for <- dt_kv[Subcomponent == "Forte", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]
+kv_seq_for <- dt_kv[Subcomponent == "Legato", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]
 
@@
 
-#### Piano
+#### Staccato
 ```{r kv_seq_pia, fig.width = 14, fig.height = 12, echo = FALSE}
@@
 dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
-kv_seq_pia <- dt_kv[Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]
+kv_seq_pia <- dt_kv[Subcomponent == "Staccato", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]
 
@@
 
-#### Forte
+#### Legato
 ```{r kv_dis_for, fig.width = 14, fig.height = 8, echo = FALSE}
-p_kv_dis_for <- ggplot(dt_kv[Subcomponent == "Forte"], aes(x = Velocity)) +
+p_kv_dis_for <- ggplot(dt_kv[Subcomponent == "Legato"], aes(x = Velocity)) +
   geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
   facet_grid(TrialNr ~ SubNr) +
-  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Forte)")
+  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Articulation(Legato)")
 p_kv_dis_for
@@
 
-#### Piano
+#### Staccato
 ```{r kv_dis_pia, fig.width = 14, fig.height = 8, echo = FALSE}
-p_kv_dis_pia <- ggplot(dt_kv[Subcomponent == "Piano"], aes(x = Velocity)) +
+p_kv_dis_pia <- ggplot(dt_kv[Subcomponent == "Staccato"], aes(x = Velocity)) +
   geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
   facet_grid(TrialNr ~ SubNr) +
-  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Piano)")
+  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Articulation(Staccato)")
 p_kv_dis_pia
@@
 ### Valid performances (exclude outliers)
-#### Forte
-```{r outlier_for, echo = FALSE}
-# for forte
-excluded_kv_for <- kv[Subcomponent == "Forte" & Mean < kv_all[Subcomponent == "Forte"]$Mean+3*kv_all[Subcomponent == "Forte"]$SD & Mean > kv_all[Subcomponent == "Forte"]$Mean-3*kv_all$SD]
-print(excluded_kv_for)
+#### Legato
+```{r outlier_leg, echo = FALSE}
+# for legato
+excluded_kv_leg <- kv[Subcomponent == "Legato" & Mean < kv_all[Subcomponent == "Legato"]$Mean+3*kv_all[Subcomponent == "Legato"]$SD & Mean > kv_all[Subcomponent == "Legato"]$Mean-3*kv_all$SD]
+print(excluded_kv_leg)
 ```
-- The number of excluded performance: `r kv_all[Subcomponent == "Forte"]$N-nrow(excluded_kv_for)`
+- The number of excluded performance: `r kv_all[Subcomponent == "Legato"]$N-nrow(excluded_kv_leg)`
 
-#### Piano
-```{r outlier_pia, echo = FALSE}
-# for piano
-excluded_kv_pia <- kv[Subcomponent == "Piano" & Mean < kv_all[Subcomponent == "Piano"]$Mean+3*kv_all[Subcomponent == "Piano"]$SD & Mean > kv_all[Subcomponent == "Piano"]$Mean-3*kv_all$SD]
-print(excluded_kv_pia)
+#### Staccato
+```{r outlier_sta, echo = FALSE}
+# for staccato
+excluded_kv_sta <- kv[Subcomponent == "Staccato" & Mean < kv_all[Subcomponent == "Staccato"]$Mean+3*kv_all[Subcomponent == "Staccato"]$SD & Mean > kv_all[Subcomponent == "Staccato"]$Mean-3*kv_all$SD]
+print(excluded_kv_sta)
 ```
-- The number of excluded performance: `r kv_all[Subcomponent == "Piano"]$N-nrow(excluded_kv_pia)`
+- The number of excluded performance: `r kv_all[Subcomponent == "Staccato"]$N-nrow(excluded_kv_sta)`
 
-```{r export, include = FALSE}
-kv_valid <- dt_kv[, .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr)]
-fwrite(kv_valid, "./kv_valid.csv")
-```
+<!-- ```{r export, include = FALSE} -->
+<!-- kv_valid <- dt_kv[, .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr)] -->
+<!-- fwrite(kv_valid, "./kv_valid.csv") -->
+<!-- ``` -->
 

### 2020-08-08 material/stimuli/mid_export.py
correct comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,7 +1,7 @@
 #!/usr/local/bin/python
 ####################
 # Created: 20/04/2020
-# This script converts a txt file to midi data and playbacks on Max MSP.
+# This script converts txt files to mid files.
 
 #%% import packages
 import os, csv, time, mido
@@ -28,8 +28,7 @@ for folder in folders:
             mid = mido.MidiFile()
             track = mido.MidiTrack()
             mid.tracks.append(track)
-            track.append(mido.Message('program_change', program=0, time=0)) #program 0 = Acoustic Grand Piano
-
+            track.append(mido.Message('program_change', program=0, time=0)) # program 0 = Acoustic Grand Piano
             # name for mid file
             if folder == "./averaging/low/1596207520-310720/":
                 midname = "./mid/low_" + str(instance) + ".mid"
@@ -39,7 +38,7 @@ for folder in folders:
                 midname = "./mid/dyn_" + str(instance) + ".mid"
             elif folder == "./averaging/high/1596380946-020820/":
                 midname = "./mid/high_" + str(instance) + ".mid"
-                
+            # read current track data (txt file)   
             current = csv.reader(csvfile, delimiter = ",")
             next(current) # skip first row
             counter = 0
@@ -58,7 +57,7 @@ for folder in folders:
                     track.append(mido.Message('note_on', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
                 elif currentOnOff == 0:
                     track.append(mido.Message('note_off', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
-            mid.save(midname)
+            mid.save(midname) # save track
 
 print("Done :D")
 # %%

### 2020-08-08 script/psychopy/rating.py
test psychopy
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,19 @@
+from psychopy import visual, core
+
+# create a window
+
+mywin = visual.Window([800,600], monitor="testMonitor", units="deg")
+
+# create some stimuli
+grating = visual.GratingStim(win=mywin, mask="circle", size=3, pos=[-4,0], sf=3)
+fixation = visual.GratingStim(win=mywin, size=0.5, pos=[0,0], sf=0, rgb=-1)
+
+# draw the stimuli and update the window
+for frameN in range(200):
+    grating.setPhase(0.05, '+')
+    grating.draw()
+    fixation.draw()
+    mywin.update()
+
+# pause for 5 secs
+core.wait(5.0)

### 2020-08-08 script/psychopy/rating.py
implement first part
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,19 +1,49 @@
-from psychopy import visual, core
+"""
+This script is based on tutorials on Psychopy3 website.
+"""
 
-# create a window
+# import libraries
+from psychopy import core, visual, gui, data, event
+from psychopy.tools.filetools import fromFile, toFile
+import numpy, random
 
-mywin = visual.Window([800,600], monitor="testMonitor", units="deg")
+# dialogue box
+expInfo = {'Number': '', 'Today': data.getDateStr()}
+dlg = gui.DlgFromDict(expInfo, fixed = ['Today'], title="Rating Pilot")
+if dlg.OK == False:
+    core.quit()
 
-# create some stimuli
-grating = visual.GratingStim(win=mywin, mask="circle", size=3, pos=[-4,0], sf=3)
-fixation = visual.GratingStim(win=mywin, size=0.5, pos=[0,0], sf=0, rgb=-1)
+"""
+   ('-.  ) (`-.       _ (`-.    ('-.  _  .-')          _   .-')       ('-.       .-') _  .-') _    
+ _(  OO)  ( OO ).    ( (OO  ) _(  OO)( \( -O )        ( '.( OO )_   _(  OO)     ( OO ) )(  OO) )   
+(,------.(_/.  \_)-._.`     \(,------.,------.  ,-.-') ,--.   ,--.)(,------.,--./ ,--,' /     '._  
+ |  .---' \  `.'  /(__...--'' |  .---'|   /`. ' |  |OO)|   `.'   |  |  .---'|   \ |  |\ |'--...__) 
+ |  |      \     /\ |  /  | | |  |    |  /  | | |  |  \|         |  |  |    |    \|  | )'--.  .--' 
+(|  '--.    \   \ | |  |_.' |(|  '--. |  |_.' | |  |(_/|  |'.'|  | (|  '--. |  .     |/    |  |    
+ |  .--'   .'    \_)|  .___.' |  .--' |  .  '.',|  |_.'|  |   |  |  |  .--' |  |\    |     |  |    
+ |  `---. /  .'.  \ |  |      |  `---.|  |\  \(_|  |   |  |   |  |  |  `---.|  | \   |     |  |    
+ `------''--'   '--'`--'      `------'`--' '--' `--'   `--'   `--'  `------'`--'  `--'     `--'    
 
-# draw the stimuli and update the window
-for frameN in range(200):
-    grating.setPhase(0.05, '+')
-    grating.draw()
-    fixation.draw()
[…]

### 2020-08-08 script/psychopy/rating.py
implementing sound/rating part...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,9 +1,14 @@
+#!/usr/bin/env python
+
 """
 This script is based on tutorials on Psychopy3 website.
+https://www.psychopy.org/coder/tutorial2.html
 """
 
 # import libraries
-from psychopy import core, visual, gui, data, event
+from psychopy import core, visual, gui, data, event, prefs
+prefs.hardware['audioLib'] = ['PTB']
+from psychopy import sound
 from psychopy.tools.filetools import fromFile, toFile
 import numpy, random
 
@@ -39,11 +44,26 @@ globalClock = core.Clock()
 trialClock = core.Clock()
 
 # display instructions and wait
-mes1 = visual.TextStim(win, pos=[0, +3], font = 'Helvetica', height = 42, wrapWidth = 1400,
+mes1 = visual.TextStim(win, pos=[0, 0], font = 'Helvetica', height = 42, wrapWidth = 1400,
     text='Thank you very much for taking part in the pilot study!\n\n In this experiment, you are going to listen to a number of piano performances and asked to rate to what extent musical expressions are implemented.')
 mes1.draw()
 fixation.draw()
 win.flip()
 
 # pause until there's a keypress
+event.waitKeys(keyList = ['space'])
+
+# stimuli presentation - TBC
+
+# get response (rating scale)
+ratingScale = visual.RatingScale(win, low = 1, high = 5, markerStart = 3, leftKeys = '1', rightKeys = '2', acceptKeys = 'space', acceptPreText='move left (1) / right (2)')
+while ratingScale.noResponse:
+    mid = sound.Sound('art_1.wav')
+    mid.setVolume(1)
+    mid.play(when=win)
+    ratingScale.draw()
+    win.flip()
+rating = ratingScale.getRating()
+decisionTime = ratingScale.getRT()
+
 event.waitKeys(keyList = ['space'])

### 2020-08-10 script/psychopy/rating.py
implemented mid.play with mido
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -6,13 +6,17 @@ https://www.psychopy.org/coder/tutorial2.html
 """
 
 # import libraries
-from psychopy import core, visual, gui, data, event, prefs
-prefs.hardware['audioLib'] = ['PTB']
+from psychopy import core, visual, gui, data, event
 from psychopy import sound
 from psychopy.tools.filetools import fromFile, toFile
-import numpy, random
+import os, numpy, random, mido
 
-# dialogue box
+# open Max port
+# open max file
+os.system('open ' + "./midiplayer.maxpat")
+port = mido.open_output('to Max 1')
+
+# participant's info (only light mode)
 expInfo = {'Number': '', 'Today': data.getDateStr()}
 dlg = gui.DlgFromDict(expInfo, fixed = ['Today'], title="Rating Pilot")
 if dlg.OK == False:
@@ -36,7 +40,7 @@ dataFile = open(filename + '.txt', 'w')
 dataFile.write('stimType, likertScale, RT\n')
 
 # create window and stimuli
-win = visual.Window([800, 600], fullscr = True, monitor = 'testMonitor', units = 'pix')
+win = visual.Window([1920, 1200], fullscr = True, monitor = 'testMonitor', units = 'pix')
 fixation = visual.GratingStim(win, color = -1, colorSpace = 'rgb', tex=None, mask='circle', size=0.2)
 
 # add clocks
@@ -57,10 +61,13 @@ event.waitKeys(keyList = ['space'])
 
 # get response (rating scale)
 ratingScale = visual.RatingScale(win, low = 1, high = 5, markerStart = 3, leftKeys = '1', rightKeys = '2', acceptKeys = 'space', acceptPreText='move left (1) / right (2)')
+counter = 0
 while ratingScale.noResponse:
-    mid = sound.Sound('art_1.wav')
-    mid.setVolume(1)
-    mid.play(when=win)
+    mid = mido.MidiFile('../../material/stimuli/mid/art_1.mid')
+    if counter == 0:
+        for msg in mid.play():
+            port.send(msg)
+            counter += 1
     ratingScale.draw()
     win.flip()
 rating = ratingScale.getRating()

### 2020-08-10 script/psychopy/rating.py
implementing force quit
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -7,70 +7,98 @@ https://www.psychopy.org/coder/tutorial2.html
 
 # import libraries
 from psychopy import core, visual, gui, data, event
-from psychopy import sound
 from psychopy.tools.filetools import fromFile, toFile
 import os, numpy, random, mido
 
+### functions ###
+
+#################
+
 # open Max port
 # open max file
-os.system('open ' + "./midiplayer.maxpat")
-port = mido.open_output('to Max 1')
+os.system("open " + "./midiplayer.maxpat")
+port = mido.open_output("to Max 1")
 
-# participant's info (only light mode)
-expInfo = {'Number': '', 'Today': data.getDateStr()}
-dlg = gui.DlgFromDict(expInfo, fixed = ['Today'], title="Rating Pilot")
+# participant"s info (only works with light mode - Mac)
+expInfo = {"Number": "", "Today": data.getDateStr()}
+dlg = gui.DlgFromDict(expInfo, fixed = ["Today"], title="Rating Pilot")
 if dlg.OK == False:
     core.quit()
 
 """
-   ('-.  ) (`-.       _ (`-.    ('-.  _  .-')          _   .-')       ('-.       .-') _  .-') _    
- _(  OO)  ( OO ).    ( (OO  ) _(  OO)( \( -O )        ( '.( OO )_   _(  OO)     ( OO ) )(  OO) )   
-(,------.(_/.  \_)-._.`     \(,------.,------.  ,-.-') ,--.   ,--.)(,------.,--./ ,--,' /     '._  
- |  .---' \  `.'  /(__...--'' |  .---'|   /`. ' |  |OO)|   `.'   |  |  .---'|   \ |  |\ |'--...__) 
- |  |      \     /\ |  /  | | |  |    |  /  | | |  |  \|         |  |  |    |    \|  | )'--.  .--' 
-(|  '--.    \   \ | |  |_.' |(|  '--. |  |_.' | |  |(_/|  |'.'|  | (|  '--. |  .     |/    |  |    
- |  .--'   .'    \_)|  .___.' |  .--' |  .  '.',|  |_.'|  |   |  |  |  .--' |  |\    |     |  |    
- |  `---. /  .'.  \ |  |      |  `---.|  |\  \(_|  |   |  |   |  |  |  `---.|  | \   |     |  |    
- `------''--'   '--'`--'      `------'`--' '--' `--'   `--'   `--'  `------'`--'  `--'     `--'    
[…]

### 2020-09-04 script/psychopy/data/summary.Rmd
summary and plot for the stimuli selection experiment
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Summary for the stimuli selection experiment"
output: html_notebook
---

Participants: Anna, Thomas and me

```{r setup, include = FALSE}
# install packages
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# setting
theme_update(legend.position = "bottom")
```

## Overall rating for each category
```{r raw, echo = FALSE}
# Create a list of data file names
lf <- list.files(".", pattern = "csv")

# Create raw_data - merge all data files into one
raw_data <- data.table()
for (i in 1:length(lf)){
  data_i <- fread(file.path(".", lf[i]), header = T, sep = ",", dec = ".")
  raw_data <- rbind(raw_data, data_i)
}
# remove unnecessary strings from midFile
raw_data$midFile <- gsub("./practice/", "", raw_data$midFile)
raw_data$midFile <- gsub("./mid/", "", raw_data$midFile)
raw_data$midFile <- gsub(".mid", "", raw_data$midFile)

# remove practice trials
data <- raw_data[expMode == "experiment"]
```

```{r summary, echo = FALSE}
data$stimCategory <- "NA"
data[startsWith(midFile, "high") == TRUE]$stimCategory <- "High"
data[startsWith(midFile, "art") == TRUE]$stimCategory <- "Art_only"
data[startsWith(midFile, "dyn") == TRUE]$stimCategory <- "Dyn_only"
data[startsWith(midFile, "low") == TRUE]$stimCategory <- "Low"

# overall rating for each category
summary_data <- data[, .(N = .N/2, meanRating = mean(rating), SD = sd(rating), SEM = sd(rating)/sqrt(.N/2)), by = .(stimCategory, ratingCategory)]
summary_data[order(ratingCategory, stimCategory)]
```

## Plot
- Error bars represent SEM.
```{r summary_plot, echo = FALSE}
plot <- ggplot(summary_data, aes(x = stimCategory, y = meanRating)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin = meanRating-SEM, ymax = meanRating + SEM), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(ratingCategory ~ .)
plot
```

## For each performance
### High performance
```{r high, echo = ALSE}
data_high <- data[stimCategory == "High", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_high$subjectNumber <- factor(data_high$subjectNumber)
data_high$Rating <- factor(data_high$rating)
data_high$midFile <- gsub("high_", "", data_high$midFile)

# plot
p_high <- ggplot(data_high, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .))
p_high
```

### Articulation-only performance
```{r art, echo = FALSE}
data_art <- data[stimCategory == "Art_only", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_art$subjectNumber <- factor(data_art$subjectNumber)
data_art$Rating <- factor(data_art$rating)
data_art$midFile <- gsub("art_", "", data_art$midFile)

# plot
p_art <- ggplot(data = data_art, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_art
```

### Dynamics-only performance
```{r dyn, echo = FALSE}
data_dyn <- data[stimCategory == "Dyn_only", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_dyn$subjectNumber <- factor(data_dyn$subjectNumber)
data_dyn$Rating <- factor(data_dyn$rating)
data_dyn$midFile <- gsub("dyn_", "", data_dyn$midFile)

# plot
p_dyn <- ggplot(data = data_dyn, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_dyn
```
### Low performance
```{r low, echo = FALSE}
data_low <- data[stimCategory == "Low", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_low$subjectNumber <- factor(data_low$subjectNumber)
data_low$Rating <- factor(data_low$rating)
data_low$midFile <- gsub("low_", "", data_low$midFile)

# plot
p_low <- ggplot(data = data_low, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_low
```

## Feedback from Anna and Thomas

### 2020-09-04 script/psychopy/data/summary.Rmd
add some notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Participants: Anna, Thomas and me
+**Participants:** Anna, Thomas and me
+
+**Experiment brief overview:**
+
+- Participants were asked to listen to a number of recordings and required to rate to what extent articulation/dynamics was implemented. They rated for each expression (articulation, dynamics).
+- We used a 5-point likert scale for rating.
+- 64 stimuli, consisting of high (both articulation and dynamics implemented), articulation-only (only articulation implemented), dynamics-only (only dynamics implemeneted) and low (none of them implemented).
+
+**Technical things:**
+
+- The stimuli were presented randomly for each participant.
+- The order of rating (articulation first or dynamics first) was randomised within participants, but the half of the ratings started with articulation and the other half started with 
+- There were 3 practice trials.
+
+**Goal:**
+
+- Pick 4 performances which represent the characteristics of each category the most.
+
+**Issues:**
+
+- How to select? Need stats?
 
@@
 ## Overall rating for each category
+- Art_only (articulation only)
+- Dyn_only (dynamics only)
+- High (articulation + dynamics)
+- Low (none)
+
 ```{r raw, echo = FALSE}
@@
 ```
+**Comments:**
+
+It seems participants can easily detect articulation but not dynamics.
 
@@
 ## Feedback from Anna and Thomas
+- If expressions are not matched with notated ones but he likes the expressions, what to rate?
+    + Judge to what extent expressions are corresponded with notated ones
+- There are many changes in the first system (especially for articulation), it was easy to judge articulation and then focused on dynamics afterwards.
+When rating, forgot what he thought about articulation.
+- Staccato and piano seem to be correlated very much. Could not tell the intention of producing piano. Might be that participants played softly (piano) but just as a result of staccato, for instance.
+- The same one for the definition of expressions (should be matched or whether I like them?)
+- Definition of staccato.
+Especially when participants are not implementing articulation, all sounds like staccato or just detached (detaché).
+she did not know whether participants meant to implement staccato or not.
+Range of staccato is much smaller than that of legato, so hard to judge.
+How long should staccato be?
+- Range 1-5 is too rough as differences in each performance is so subtle.
+- **Several performances have timing issues.**
+    + I only looked at 8th not so maybe better to look at 
+- Wordings for instructions: to what extent participants follow instructions
+
+## Feedback for my performance
+- As I generated stimuli, I am not sure whether my responses are valid because I could mostly categorise stimuli for 4 types even though they were presented randomly (of course some of them, I couldn't).

### 2020-09-04 script/psychopy/data/summary.Rmd
fix typo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - We used a 5-point likert scale for rating.
-- 64 stimuli, consisting of high (both articulation and dynamics implemented), articulation-only (only articulation implemented), dynamics-only (only dynamics implemeneted) and low (none of them implemented).
+- 64 stimuli, consisting of 16 high (both articulation and dynamics implemented), 16 articulation-only (only articulation implemented), 16 dynamics-only (only dynamics implemented) and 16 low (none of them implemented).
 
@@
 How long should staccato be?
-- Range 1-5 is too rough as differences in each performance is so subtle.
+- Range 1-5 is too rough as differences in each performance are so subtle.
 - **Several performances have timing issues.**

### 2021-01-11 material/expstimuli/analysis/expression/stim_d/duration.Rmd
organised files and updated readme
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/expression/stim_a/ioi_duration.Rmd
@@
 ---
-title: 'Duration Analysis for Performing-Articulation performances'
+title: 'Duration Analysis for Performing-Dynamics performances'
 output:
@@
 
-Description: This is a rough overview of performing performance (duration) from teaching-v2.0.
+Description: This is a rough overview of performing performance (Duration) from teaching-v2.0.
 
@@
 dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
-# include only articulation
-dt_onset <- dt_onset[Condition == "performing" & Skill == "articulation"]
+# include only dynamics
+dt_onset <- dt_onset[Condition == "performing" & Skill == "dynamics"]
 
@@
 dt_offset <- dt_offset[order(SubNr, BlockNr, TrialNr, RowNr),]
-# include only articulation
-dt_offset <- dt_offset[Condition == "performing" & Skill == "articulation"]
+# include only dynamics
+dt_offset <- dt_offset[Condition == "performing" & Skill == "dynamics"]
 
@@
 # Legato
-for (phrase in 1:length(ls_legato2)){
-  for (note in 1:length(ls_legato2[[phrase]])){
-    dt_duration[RowNr == ls_legato2[[phrase]][note]]$Subcomponent <- "Legato"
+for (phrase in 1:length(ls_forte2)){
+  for (note in 1:length(ls_forte2[[phrase]])){
+    dt_duration[RowNr == ls_forte2[[phrase]][note]]$Subcomponent <- "Forte"
   }
@@
 # Staccato
-for (phrase in 1:length(ls_staccato2)){
-  for (note in 1:length(ls_staccato2[[phrase]])){
-    dt_duration[RowNr == ls_staccato2[[phrase]][note]]$Subcomponent <- "Staccato"
+for (phrase in 1:length(ls_piano2)){
+  for (note in 1:length(ls_piano2[[phrase]])){
+    dt_duration[RowNr == ls_piano2[[phrase]][note]]$Subcomponent <- "Piano"
   }
@@
 # for each individual
-duration <- dt_duration[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, Subcomponent)]
+duration <- dt_duration[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, Subcomponent)]
 print(duration)
@@
   facet_grid(Subcomponent ~ .) +
-  labs(y = "Normalised Duration", subtitle = "Duration: Performing-Articulation")
+  labs(y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics")
 p_duration
@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Articulation(All)") +
+  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics(All)") +
   scale_x_continuous(breaks=seq(1,66,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
@@
 
-#### Legato
+#### Forte
 ```{r duration_seq_leg, fig.width = 14, fig.height = 12, echo = FALSE}
@@
 dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
-duration_seq_leg <- dt_duration[Subcomponent == "Legato", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
+duration_seq_leg <- dt_duration[Subcomponent == "Forte", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
 
@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Articulation(Legato)") +
+  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics(Forte)") +
   scale_x_continuous(breaks=seq(1,66,1))
@@
 
-#### Staccato
+#### Piano
 ```{r duration_seq_sta, fig.width = 14, fig.height = 12, echo = FALSE}
@@
 dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
-duration_seq <- dt_duration[Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
+duration_seq <- dt_duration[Subcomponent == "Piano", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]
 
@@
   facet_grid(SubNr ~ .) +
-  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Articulation(Staccato)") +
+  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics(Piano)") +
   scale_x_continuous(breaks=seq(1,66,1))
@@
 
-#### Legato
-```{r duration_dis_leg, fig.width = 12, echo = FALSE}
-p_duration_dis_leg <- ggplot(dt_duration[Subcomponent == "Legato"], aes(x = normDu)) +
+#### Forte
+```{r duration_dis_for, fig.width = 12, echo = FALSE}
+p_duration_dis_for <- ggplot(dt_duration[Subcomponent == "Forte"], aes(x = normDu)) +
   geom_histogram(binwidth = .1, alpha=.5, position="identity") +
   facet_grid(TrialNr ~ SubNr) +
-  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Articulation(Legato)")
-p_duration_dis_leg
+  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Dynamics(Forte)")
+p_duration_dis_for
 ```
 
-#### Staccato
-```{r duration_dis_sta, fig.width = 12, echo = FALSE}
-p_duration_dis_sta <- ggplot(dt_duration[Subcomponent == "Staccato"], aes(x = normDu)) +
+#### Piano
+```{r duration_dis_pia, fig.width = 12, echo = FALSE}
+p_duration_dis_pia <- ggplot(dt_duration[Subcomponent == "Piano"], aes(x = normDu)) +
   geom_histogram(binwidth = .1, alpha = .5, position="identity") +
   facet_grid(TrialNr ~ SubNr) +
-  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Articulation(Staccato)")
-p_duration_dis_sta
+  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Dynamics(Piano)")
+p_duration_dis_pia
 ```
@@
 ### Valid performances (exclude outliers)
-#### Legato
+#### Forte
 ```{r outlier_leg, echo = FALSE}
-# for legato
-excluded_duration_leg <- duration[Subcomponent == "Legato" & Mean < duration_all[Subcomponent == "Legato"]$Mean+3*duration_all[Subcomponent == "Legato"]$SD & Mean > duration_all[Subcomponent == "Legato"]$Mean-3*duration_all$SD]
-print(excluded_duration_leg)
+# for forte
+excluded_duration_for <- duration[Subcomponent == "Forte" & Mean < duration_all[Subcomponent == "Forte"]$Mean+3*duration_all[Subcomponent == "Forte"]$SD & Mean > duration_all[Subcomponent == "Forte"]$Mean-3*duration_all$SD]
+print(excluded_duration_for)
 ```
-- The number of excluded performance: `r duration_all[Subcomponent == "Legato"]$N-nrow(excluded_duration_leg)`
+- The number of excluded performance: `r duration_all[Subcomponent == "Forte"]$N-nrow(excluded_duration_for)`
 
-#### Staccato
+#### Piano
 ```{r outlier_sta, echo = FALSE}
-# for staccato
-excluded_duration_sta <- duration[Subcomponent == "Staccato" & Mean < duration_all[Subcomponent == "Staccato"]$Mean+3*duration_all[Subcomponent == "Staccato"]$SD & Mean > duration_all$Mean-3*duration_all[Subcomponent == "Staccato"]$SD]
-print(excluded_duration_sta)
+# for piano
+excluded_duration_pia <- duration[Subcomponent == "Piano" & Mean < duration_all[Subcomponent == "Piano"]$Mean+3*duration_all[Subcomponent == "Piano"]$SD & Mean > duration_all$Mean-3*duration_all[Subcomponent == "Piano"]$SD]
+print(excluded_duration_pia)
 ```
-- The number of excluded performance: `r duration_all[Subcomponent == "Staccato"]$N-nrow(excluded_duration_sta)`
+- The number of excluded performance: `r duration_all[Subcomponent == "Piano"]$N-nrow(excluded_duration_pia)`
 
-```{r export, include = FALSE}
-duration_valid <- dt_duration[, .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr)]
-fwrite(duration_valid, "./duration_valid.csv")
-```
+<!-- ```{r export, include = FALSE} -->
+<!-- duration_valid <- dt_duration[, .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr)] -->
+<!-- fwrite(duration_valid, "./duration_valid.csv") -->
+<!-- ``` -->
 

### 2021-01-12 material/expstimuli/analysis/expression/stim_a/duration.Rmd
organised files and updated readme
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
changed path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/expression/stim_a/ioi_duration.Rmd
@@
 # onset
-dt_onset <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
+dt_onset_all<- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 # assign RowNr
-dt_onset$RowNr <- rep(1:72, nrow(dt_onset)/72)
+dt_onset_all$RowNr <- rep(1:72, nrow(dt_onset_all)/72)
 # sort by SubNr, BlockNr, TrialNr and NoteNr
-dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
+dt_onset_all <- dt_onset_all[order(SubNr, BlockNr, TrialNr, RowNr),]
 # include only articulation
-dt_onset <- dt_onset[Condition == "performing" & Skill == "articulation"]
+dt_onset <- dt_onset_all[Condition == "performing" & Skill == "articulation"]
 
@@
 dt_onset$TrialNr <- as.factor(dt_onset$TrialNr)
-
-# remove performances which contain NA
-dt_onset$Error <- 0
-for (error in 1:nrow(dt_error_on)){
-  dt_onset[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
-}
-dt_onset_ori <- dt_onset # save original data
-dt_onset <- dt_onset_ori[Error == 0]
+dt_onset <- dt_onset[Error == 0]
 
 # offset
-dt_offset <- fread("../filtered/data_correct_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-dt_error_off <- fread("../filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".") # read an error csv
+dt_offset_all <- fread("../preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 # assign RowNr
-dt_offset$RowNr <- rep(1:72, nrow(dt_offset)/72)
+dt_offset_all$RowNr <- rep(1:72, nrow(dt_offset_all)/72)
 # sort by SubNr, BlockNr, TrialNr and NoteNr
-dt_offset <- dt_offset[order(SubNr, BlockNr, TrialNr, RowNr),]
+dt_offset_all <- dt_offset_all[order(SubNr, BlockNr, TrialNr, RowNr),]
 # include only articulation
-dt_offset <- dt_offset[Condition == "performing" & Skill == "articulation"]
+dt_offset <- dt_offset_all[Condition == "performing" & Skill == "articulation"]
 
@@
 dt_offset$TrialNr <- as.factor(dt_offset$TrialNr)
-
-# remove performances which contain NA
-dt_offset$Error <- 0
-for (error in 1:nrow(dt_error_off)){
-  dt_offset[SubNr == dt_error_off[error]$SubNr & BlockNr == dt_error_off[error]$BlockNr & TrialNr == dt_error_off[error]$TrialNr]$Error <- 1
-}
-dt_offset_ori <- dt_offset # save original data
-dt_offset <- dt_offset_ori[Error == 0]
+dt_offset <- dt_offset[Error == 0]
 ```

### 2021-01-12 material/expstimuli/analysis/expression/stim_a/ioi_onset.Rmd
changed path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/expression/stim_a/ioi_onset.Rmd
@@
 ---
-title: 'Analysis for Performing-Articulation performances (ONSET)'
+title: 'IOI-Onset Analysis for Performing-Articulation performances'
 output:
@@
 ```{r extract, include = FALSE}
-dt_onset <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-dt_error <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
+dt_onset <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+
 # assign RowNr
@@
 dt_ioi$TrialNr <- as.factor(dt_ioi$TrialNr)
-
-# remove performances which contain NA
-dt_ioi$Error <- 0
-for (error in 1:nrow(dt_error)){
-  dt_ioi[SubNr == dt_error[error]$SubNr & BlockNr == dt_error[error]$BlockNr & TrialNr == dt_error[error]$TrialNr]$Error <- 1
-}
-dt_ioi_ori <- dt_ioi # save original data
-dt_ioi <- dt_ioi_ori[Error == 0]
+dt_ioi <- dt_ioi[Error == 0]
 ```

### 2021-01-12 material/expstimuli/analysis/expression/stim_a/kv.Rmd
changed path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-08-08 material/stimuli/analysis/expression/stim_a/kv.Rmd
@@
 # onset
-dt_kv <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
+dt_kv <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 dt_kv <- dt_kv[Condition == "performing" & Skill == "articulation"]
-
-# remove performances which contain NA
-dt_kv$Error <- 0
-for (error in 1:nrow(dt_error_on)){
-  dt_kv[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
-}
-dt_kv_ori <- dt_kv # save original data
-dt_kv <- dt_kv_ori[Error == 0]
+dt_kv <- dt_kv[Error == 0]
 ```
 
-```{r Velocitys, include = FALSE}
+```{r velocity, include = FALSE}
 # # convert bpm to ms

### 2021-01-12 material/expstimuli/analysis/expression/stim_d/duration.Rmd
changed path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # onset
-dt_onset <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
+dt_onset_all <- fread("../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+
 # assign RowNr
-dt_onset$RowNr <- rep(1:72, nrow(dt_onset)/72)
+dt_onset_all$RowNr <- rep(1:72, nrow(dt_onset_all)/72)
 # sort by SubNr, BlockNr, TrialNr and NoteNr
-dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
+dt_onset_all <- dt_onset_all[order(SubNr, BlockNr, TrialNr, RowNr),]
 # include only dynamics
-dt_onset <- dt_onset[Condition == "performing" & Skill == "dynamics"]
+dt_onset <- dt_onset_all[Condition == "performing" & Skill == "dynamics"]
 
@@
 dt_onset$TrialNr <- as.factor(dt_onset$TrialNr)
-
-# remove performances which contain NA
-dt_onset$Error <- 0
-for (error in 1:nrow(dt_error_on)){
-  dt_onset[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
-}
-dt_onset_ori <- dt_onset # save original data
-dt_onset <- dt_onset_ori[Error == 0]
+dt_onset <- dt_onset[Error == 0]
 
 # offset
-dt_offset <- fread("../filtered/data_correct_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
-dt_error_off <- fread("../filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".") # read an error csv
+dt_offset_all <- fread("../preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+
 # assign RowNr
-dt_offset$RowNr <- rep(1:72, nrow(dt_offset)/72)
+dt_offset_all$RowNr <- rep(1:72, nrow(dt_offset_all)/72)
 # sort by SubNr, BlockNr, TrialNr and NoteNr
-dt_offset <- dt_offset[order(SubNr, BlockNr, TrialNr, RowNr),]
+dt_offset_all <- dt_offset_all[order(SubNr, BlockNr, TrialNr, RowNr),]
 # include only dynamics
-dt_offset <- dt_offset[Condition == "performing" & Skill == "dynamics"]
+dt_offset <- dt_offset_all[Condition == "performing" & Skill == "dynamics"]
 

### 2021-01-12 material/expstimuli/analysis/expression/stim_d/kv.Rmd
organised files and updated readme
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
changed path
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-20 material/analysis/expression/stim_d/kv.Rmd
@@
 
-# include only articulation
+# include only dynamics
 dt_kv <- dt_kv[Condition == "performing" & Skill == "dynamics"]
-
-# remove performances which contain NA
-dt_kv$Error <- 0
-for (error in 1:nrow(dt_error_on)){
-  dt_kv[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
-}
-dt_kv_ori <- dt_kv # save original data
-dt_kv <- dt_kv_ori[Error == 0]
+dt_kv <- dt_kv[Error == 0]
 ```

### 2021-01-12 material/expstimuli/analysis/expression/preprocessor/filtering.R
add preprocessor for data with expression
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,105 @@
+#!/usr/local/bin/R
+#rm(list=ls(all=T)) - clear all in Grobal Environment
+
+####################################
+#  Documentation
+####################################
+# Created: 02/04/2020
+# This script organises raw data and removes pitch errors.
+# GitHub repo (private): https://github.com/atsukotominaga/teaching-v2.0/tree/master/script/R 
+
+####################################
+#  Requirements
+####################################
+# set working directory to file source location
+# install and load required packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# read functions
+source("./function.R")
+
+# create necessary folders if not exist
+# filtered - all of the outputs will be stored in this folder
+if (!file.exists("filtered")){
+  dir.create("filtered")
+}
+
+# read a text file for ideal performance
+df_ideal <- fread("./ideal.txt")
+colnames(df_ideal) <- "Pitch"
+df_ideal$RowNr <- c(1:nrow(df_ideal))
+
+# create a list of data file names
+lf <- list.files("./data", pattern = "txt")
+
+# create raw_data - merge all data files into one
+raw_data <- data.table()
+for (i in 1:length(lf)){
+  data_i <- read.csv(file.path("./data", lf[i]), header = F, sep = " ", dec = ".")
+  raw_data <- rbind(raw_data, data_i)
+}
+
+# add column namesls
+colnames(raw_data) <- c("NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "Tempo",
+                        "SubNr", "BlockNr", "TrialNr", "Skill", "Condition", "Image")
+
+# clean raw_data
+raw_data$NoteNr <- as.numeric(gsub(",", "", raw_data$NoteNr))
+raw_data$Image <- gsub(";", "", raw_data$Image)
+
+# sort by SubNr, BlockNr, TrialNr
+raw_data <- raw_data[order(raw_data$SubNr, raw_data$BlockNr, raw_data$TrialNr),]
+
+# separate raw_data into each skill
+df_stim_a <- raw_data[Skill == "articulation"]
+df_stim_d <- raw_data[Skill == "dynamics"]
+df_stim <- rbind(df_stim_a, df_stim_d)
+
+####################################
+# Detect pitch errors
+####################################
+# raw_data without metronome
+df_all <- df_stim[Pitch != 31 & Pitch != 34]
+
+# onset and offset
[…]

### 2021-01-12 material/expstimuli/analysis/expression/preprocessor/function.R

@@ -1,0 +1,41 @@
+##### FUNCTIONS #####
+
+# install and load required packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+### define pitch remove function (pitch_remover)
+# data: data of all trials (df_onset/df_offset)
+# ideal: df_ideal
+pitch_remover <- function(data, ideal){
+  ls_removed <- list()
+  for (subnr in unique(data$SubNr)){ #set # of participants
+    print(sprintf("---SubNr %i---", subnr))
+    for (block in unique(data[SubNr == subnr]$BlockNr)){
+      for (trial in unique(unique(data[SubNr == subnr & BlockNr == block]$TrialNr))){
+        current <- data[SubNr == subnr & BlockNr == block & TrialNr == trial]
+        if (nrow(current) != 0){ #if current data is not empty
+          if (length(ideal$Pitch) != length(current$Pitch)){ #if # of onsets/offsets is not equal to ideal performance
+            ls_removed <- c(ls_removed, list(c(subnr, block, trial, "NoteNr error")))
+            print(sprintf("NoteNr error - SubNr/BlockNr/TrialNr: %i/%i/%i", subnr, block, trial))
+          } else if (length(ideal$Pitch) == length(current$Pitch)) { #if # of onsets and offsets are correct
+            counter = 0 #set a counter so that the following loop will terminate once it detects one pitch error in a trial
+            for (note in 1:length(ideal$Pitch)){
+              # detect onset error
+              if (current$Pitch[note] != ideal$Pitch[note]){
+                while (counter == 0){
+                  ls_removed <- c(ls_removed, list(c(subnr, block, trial, "Pitch error")))
+                  print(sprintf("Pitch error - SubNr/TrialNr/NoteNr: %i//%i/%i/%i", subnr, block, trial, note))
+                  counter = counter + 1
+                }
+              }
+            }
+          }
+        } else { #if current data is emtpy"
+          ls_removed <- c(ls_removed, list(c(subnr, block, trial, "Missing")))
+          print(sprintf("Missing - SubNr/BlockNr/TrialNr: %i/%i/%i", subnr, block, trial))
+        }
+      }
+    }
+  }
+  return(ls_removed)
+}

### 2021-01-28 material/prerating/data/summary.Rmd
organised files and updated readme
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
add new plots with 3 or 2 raters
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-09-04 script/psychopy/data/summary.Rmd
@@
 ---
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
+
+# Descriptive Stats for the generated stimuli
 
@@
 ```
-**Comments:**
+**Comments**
 
@@
 - As I generated stimuli, I am not sure whether my responses are valid because I could mostly categorise stimuli for 4 types even though they were presented randomly (of course some of them, I couldn't).
+
+# Ranking for the generated stimuli
+### High performance
+- summed the rating for both articulation and dynamics and ranked the performances
+
+### with 3 raters
+```{r ranking-high, echo = FALSE}
+data_high <- data_high[order(midFile, ratingCategory)]
+data_high_sep <- data_high[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_high_both <- data_high[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_high_both[order(-Sum)]
+```
+
+### with 2 raters
+```{r ranking-high-2, echo = FALSE}
+data_high_2 <- data_high[subjectNumber != 3]
+data_high_sep_2 <- data_high_2[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_high_both_2 <- data_high_2[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_high_both_2[order(-Sum)]
+```
+
+### Articulation performance
+- summed the rating for both articulation and dynamics and ranked the performances
+
+### with 3 raters
+```{r ranking-art, echo = FALSE}
+data_art <- data_art[order(midFile, ratingCategory)]
+data_art_sep <- data_art[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_art_both <- data_art[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_art_both[order(-Sum)]
+```
+
+### Difference (not sum)
+- calculated the difference between the rating for articulation and the rating for dynamics and ranked the performances
+
+```{r ranking-art-diff, echo = FALSE}
+# or difference
+data_art_diff <- cbind(data_art[ratingCategory == "articulation"], data_art[ratingCategory == "dynamics"])
+data_art_diff$Diff <- 0
+colnames(data_art_diff)[c(5,11)] <- c("rating1", "rating2")
+data_art_diff$Diff <- data_art_diff$rating1 - data_art_diff$rating2
+data_art_diff_both <- data_art_diff[, .(N = .N, Diff = sum(Diff)), by = .(midFile)]
+
+data_art_diff_both[order(-Diff)]
+```
+
+### with 2 raters
+```{r ranking-art-2, echo = FALSE}
+data_art_2 <- data_art[subjectNumber != 3]
+data_art_sep_2 <- data_art_2[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_art_both_2 <- data_art_2[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_art_both_2[order(-Sum)]
+```
+
+### Dynamics performance
+- summed the rating for both articulation and dynamics and ranked the performances
+
+### with 3 raters
+```{r ranking-dyn, echo = FALSE}
+data_dyn <- data_dyn[order(midFile, ratingCategory)]
+data_dyn_sep <- data_dyn[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_dyn_both <- data_dyn[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_dyn_both[order(-Sum)]
+```
+
+### Difference (not sum)
+- calculated the difference between the rating for articulation and the rating for dynamics and ranked the performances
+
+```{r ranking-dyn-diff, echo = FALSE}
+# or difference
+data_dyn_diff <- cbind(data_dyn[ratingCategory == "dynamics"], data_dyn[ratingCategory == "articulation"])
+data_dyn_diff$Diff <- 0
+colnames(data_dyn_diff)[c(5,11)] <- c("rating1", "rating2")
+data_dyn_diff$Diff <- data_dyn_diff$rating1 - data_dyn_diff$rating2
+data_dyn_diff_both <- data_dyn_diff[, .(N = .N, Diff = sum(Diff)), by = .(midFile)]
+
+data_dyn_diff_both[order(-Diff)]
+```
+
+### with 2 raters
+```{r ranking-dyn-2, echo = FALSE}
+data_dyn_2 <- data_dyn[subjectNumber != 3]
+data_dyn_sep_2 <- data_dyn_2[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_dyn_both_2 <- data_dyn_2[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_dyn_both_2[order(-Sum)]
+```
+
+### Low performance
+- summed the rating for both articulation and dynamics and ranked the performances
+
+### with 3 raters
+```{r ranking-low, echo = FALSE}
+data_low <- data_low[order(midFile, ratingCategory)]
+data_low_sep <- data_low[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_low_both <- data_low[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_low_both[order(Sum)]
+```
+
+### with 2 raters
+```{r ranking-low-2, echo = FALSE}
+data_low_2 <- data_low[subjectNumber != 3]
+data_low_sep_2 <- data_low_2[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_low_both_2 <- data_low_2[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_low_both_2[order(Sum)]
+```

### 2021-01-28 material/expstimuli/analysis/expression/preprocessor/filtering.R
move files to original folder
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -6,7 +6,6 @@
 ####################################
 # Created: 02/04/2020
 # This script organises raw data and removes pitch errors.
-# GitHub repo (private): https://github.com/atsukotominaga/teaching-v2.0/tree/master/script/R 
 
 ####################################
 #  Requirements

### 2021-01-28 material/expstimuli/original/averaging/art/average.R

@@ -1,5 +1,5 @@
 # /usr/local/bin/r
-# Average key velocity for averaged expressive performance
+# Average duration for averaged expressive performance
 # Created: 06/05/2020
 
 # packages
@@ -14,54 +14,60 @@ foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current tim
 dir.create(foldername)
 
 # file location
-# 1. data from stim_n (tempo, articulation)
+# 1. data from stim_n (tempo, dynamics)
 filename_ioi = "../low/1596207520-310720/dt_ioi_instance.txt"
-filename_du = "../low/1596207520-310720/dt_du_instance.txt"
-# 2. data from stim_d (dynamics)
-filename_onset = "../../analysis/expression/filtered/data_onset.csv"
-filename_offset = "../../analysis/expression/filtered/data_offset.csv"
-filename_valid_kv = "../../analysis/expression/stim_d/kv_valid.csv"
+filename_kv = "../low/1596207520-310720/dt_kv_instance.txt"
+# 2. data from stim_a (articulation)
+filename_onset = "../../../analysis/expression/filtered/data_onset.csv"
+filename_offset = "../../../analysis/expression/filtered/data_offset.csv"
+filename_valid_du = "../../../analysis/expression/stim_a/duration_valid.csv"
 
 # read csv/txt
-# 1. use normative performance for ioi, duration
+# 1. use normative performance for ioi, kv
 dt_ioi_instance <- fread(filename_ioi, header = T, sep = ",", dec = ".")
-dt_du_instance <- fread(filename_du, header = T, sep = ",", dec = ".")
-# 2. use performing performance for kv
-kv_valid <- fread(filename_valid_kv, header = T, sep = ",", dec = ".")
+dt_kv_instance <- fread(filename_kv, header = T, sep = ",", dec = ".")
+# 2. use performing performance for duration
 dt_onset <- fread(filename_onset, header = T, sep = ",", dec = ".")
 dt_offset <- fread(filename_offset, header = T, sep = ",", dec = ".")
-# only for performing/dynamics
-dt_onset_kv <- dt_onset[Condition == "performing" & Skill == "dynamics"]
+# only for performing/articulation
+dt_onset_du <- dt_onset[Condition == "performing" & Skill == "articulation"]
[…]

### 2021-01-28 material/expstimuli/original/averaging/dyn/average.R

@@ -1,5 +1,5 @@
 # /usr/local/bin/r
-# Average duration for averaged expressive performance
+# Average key velocity for averaged expressive performance
 # Created: 06/05/2020
 
 # packages
@@ -14,60 +14,54 @@ foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current tim
 dir.create(foldername)
 
 # file location
-# 1. data from stim_n (tempo, dynamics)
+# 1. data from stim_n (tempo, articulation)
 filename_ioi = "../low/1596207520-310720/dt_ioi_instance.txt"
-filename_kv = "../low/1596207520-310720/dt_kv_instance.txt"
-# 2. data from stim_a (articulation)
-filename_onset = "../../analysis/expression/filtered/data_onset.csv"
-filename_offset = "../../analysis/expression/filtered/data_offset.csv"
-filename_valid_du = "../../analysis/expression/stim_a/duration_valid.csv"
+filename_du = "../low/1596207520-310720/dt_du_instance.txt"
+# 2. data from stim_d (dynamics)
+filename_onset = "../../../analysis/expression/filtered/data_onset.csv"
+filename_offset = "../../../analysis/expression/filtered/data_offset.csv"
+filename_valid_kv = "../../../analysis/expression/stim_d/kv_valid.csv"
 
 # read csv/txt
-# 1. use normative performance for ioi, kv
+# 1. use normative performance for ioi, duration
 dt_ioi_instance <- fread(filename_ioi, header = T, sep = ",", dec = ".")
-dt_kv_instance <- fread(filename_kv, header = T, sep = ",", dec = ".")
-# 2. use performing performance for duration
+dt_du_instance <- fread(filename_du, header = T, sep = ",", dec = ".")
+# 2. use performing performance for kv
+kv_valid <- fread(filename_valid_kv, header = T, sep = ",", dec = ".")
 dt_onset <- fread(filename_onset, header = T, sep = ",", dec = ".")
 dt_offset <- fread(filename_offset, header = T, sep = ",", dec = ".")
-# only for performing/articulation
-dt_onset_du <- dt_onset[Condition == "performing" & Skill == "articulation"]
-dt_offset_du <- dt_offset[Condition == "performing" & Skill == "articulation"]
[…]

### 2021-01-28 material/expstimuli/original/averaging/low/average.R

@@ -14,12 +14,12 @@ foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current tim
 dir.create(foldername)
 
 # file location
-filename_onset = "../../analysis/stim_n/preprocessor/filtered/data_onset.csv"
-filename_offset = "../../analysis/stim_n/preprocessor/filtered/data_offset.csv"
-filename_valid_on = "../../analysis/stim_n/onset_valid.csv"
-filename_valid_off = "../../analysis/stim_n/offset_valid.csv"
-filename_valid_du = "../../analysis/stim_n/duration_valid.csv"
-filename_valid_kv = "../../analysis/stim_n/kv_valid.csv"
+filename_onset = "../../../analysis/stim_n/preprocessor/filtered/data_onset.csv"
+filename_offset = "../../../analysis/stim_n/preprocessor/filtered/data_offset.csv"
+filename_valid_on = "../../../analysis/stim_n/onset_valid.csv"
+filename_valid_off = "../../../analysis/stim_n/offset_valid.csv"
+filename_valid_du = "../../../analysis/stim_n/duration_valid.csv"
+filename_valid_kv = "../../../analysis/stim_n/kv_valid.csv"
 
 # read csv/txt
 dt_onset <- fread(filename_onset, header = T, sep = ",", dec = ".")

### 2021-02-01 material/expstimuli/analysis/stim_n/duration.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/stim_n/duration.Rmd
@@
 
-Description: This is a rough overview of non-expressive performance (Duration) from teaching-v2.0.
+Description: This is a rough overview of non-expressive performance (Duration) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Non-expressive performances: data from practice trials
 
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude Duration outliers)
+- Exclude mean Duration outside 3SD
+
 ```{r outlier, echo = FALSE}
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude CVs outliers)
+- Exclude mean CVs outside 3SD
+
 ```{r cv_outlier, echo = FALSE}

### 2021-02-01 material/expstimuli/analysis/stim_n/ioi_offset.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of non-expressive performance (IOI-offset) from teaching-v2.0.
+Description: This is a rough overview of non-expressive performance (IOI-Offset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Non-expressive performances: data from practice trials
 
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude outliers)
+- Exclude mean IOIs outside 3SD
+
 ```{r outlier, echo = FALSE}
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude outliers)
+- Exclude mean CVs outside 3SD
+
 ```{r cv_outlier, echo = FALSE}

### 2021-02-01 material/expstimuli/analysis/stim_n/ioi_onset.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/stim_n/ioi_onset.Rmd
@@
 
-Description: This is a rough overview of non-expressive performance (IOI-Onset) from teaching-v2.0.
+Description: This is a rough overview of non-expressive performance (IOI-Onset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Non-expressive performances: data from practice trials
 
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude IOIs outliers)
+- Exclude mean IOIs outside 3SD
+
 ```{r outlier, echo = FALSE}

### 2021-02-01 material/expstimuli/analysis/stim_n/kv.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-05-19 material/analysis/stim_n/kv.Rmd
@@
 
-Description: This is a rough overview of non-expressive performance (KV) from teaching-v2.0.
+Description: This is a rough overview of non-expressive performance (KV) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Non-expressive performances: data from practice trials
 
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude KVs outliers)
+- Exclude mean KVs outside 3SD
+
 ```{r outlier, echo = FALSE}

### 2021-02-01 material/expstimuli/analysis/expression/stim_a/duration.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance (Duration) from teaching-v2.0.
+***Duration values from this analysis were used.***
+
+Description: This is a rough overview of performing performance (Duration) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Performing-Articulation performances: data from the performing condition (articulation)
 
@@
 
-### Valid performances (exclude outliers)
+# Valid performances (exclude Duration outliers)
+- Exclude mean Duration outside 3SD
 #### Legato
@@
 ```
-
-<!-- # Duration Variability -->
-<!-- ```{r cv, echo = FALSE} -->
-<!-- # for each participant -->
-<!-- cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)] -->
-<!-- print(cv) -->
-<!-- ``` -->
-
-<!-- ```{r cv_box, echo = FALSE} -->
-<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
-<!--   geom_boxplot() + -->
-<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
-<!-- p_cv_box -->
-<!-- ``` -->
-
-<!-- ### Exclude outlier performacnes -->
-<!-- ```{r cv_outlier, echo = FALSE} -->
-<!-- # for all -->
-<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
-<!-- print(cv_all) -->
-
-<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
-<!-- print(excluded_cv) -->
-<!-- ``` -->

### 2021-02-01 material/expstimuli/analysis/expression/stim_a/ioi_onset.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance (IOI-Onset) from teaching-v2.0.
+Description: This is a rough overview of performing performance (IOI-Onset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Performing-Articulation performances: data from the performing condition (articulation)
 
@@
 
-### Valid performances (exclude outliers)
-```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
-print(excluded_ioi)
-```
-- The number of excluded performance: `r ioi_all$N-nrow(excluded_ioi)`
-
 # IOI Variability
@@
 ```
-
-### Valid performances (exclude outliers)
-```{r cv_outlier, echo = FALSE}
-excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
-print(excluded_cv)
-```
-- The number of excluded performance: `r cv_all$N-nrow(excluded_cv)`

### 2021-02-01 material/expstimuli/analysis/expression/stim_a/kv.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance (KV) from teaching-v2.0.
+Description: This is a rough overview of performing performance (KV) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Performing-Articulation performances: data from the performing condition (articulation)
 
@@
 ```{r velocity, include = FALSE}
-# # convert bpm to ms
-# dt_kv[Tempo == 120]$Tempo <- 250
-# dt_kv[Tempo == 110]$Tempo <- 273
-# dt_kv[Tempo == 100]$Tempo <- 300
-
 # define subcomponents
@@
 ```
-
-### Valid performances (exclude outliers)
-#### Legato
-```{r outlier_leg, echo = FALSE}
-# for legato
-excluded_kv_leg <- kv[Subcomponent == "Legato" & Mean < kv_all[Subcomponent == "Legato"]$Mean+3*kv_all[Subcomponent == "Legato"]$SD & Mean > kv_all[Subcomponent == "Legato"]$Mean-3*kv_all$SD]
-print(excluded_kv_leg)
-```
-- The number of excluded performance: `r kv_all[Subcomponent == "Legato"]$N-nrow(excluded_kv_leg)`
-
-#### Staccato
-```{r outlier_sta, echo = FALSE}
-# for staccato
-excluded_kv_sta <- kv[Subcomponent == "Staccato" & Mean < kv_all[Subcomponent == "Staccato"]$Mean+3*kv_all[Subcomponent == "Staccato"]$SD & Mean > kv_all[Subcomponent == "Staccato"]$Mean-3*kv_all$SD]
-print(excluded_kv_sta)
-```
-- The number of excluded performance: `r kv_all[Subcomponent == "Staccato"]$N-nrow(excluded_kv_sta)`
-
-<!-- ```{r export, include = FALSE} -->
-<!-- kv_valid <- dt_kv[, .(N = length(Velocity), Mean = mean(Velocity)), .(SubNr, TrialNr)] -->
-<!-- fwrite(kv_valid, "./kv_valid.csv") -->
-<!-- ``` -->
-
-<!-- # KV Variability -->
-<!-- ```{r cv, echo = FALSE} -->
-<!-- # for each participant -->
-<!-- cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)] -->
-<!-- cv -->
-<!-- ``` -->
-
-<!-- ```{r cv_box, echo = FALSE} -->
-<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
-<!--   geom_boxplot() + -->
-<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
-<!-- p_cv_box -->
-<!-- ``` -->
-
-<!-- ### Exclude outlier performances -->
-<!-- ```{r cv_outlier, echo = FALSE} -->
-<!-- # for all -->
-<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
-<!-- print(cv_all) -->
-
-<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
-<!-- print(excluded_cv) -->
-<!-- ``` -->

### 2021-02-01 material/expstimuli/analysis/expression/stim_a/duration.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 - Exclude mean Duration outside 3SD
-#### Legato
+
+## Legato
 ```{r outlier_leg, echo = FALSE}
@@
 
-#### Staccato
+## Staccato
 ```{r outlier_sta, echo = FALSE}

### 2021-02-01 material/expstimuli/analysis/expression/stim_d/duration.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance (Duration) from teaching-v2.0.
+Description: This is a rough overview of performing performance (Duration) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Performing-Dynamics performances: data from the performing condition (dynamics)
 
@@
 ```
-
-### Valid performances (exclude outliers)
-#### Forte
-```{r outlier_leg, echo = FALSE}
-# for forte
-excluded_duration_for <- duration[Subcomponent == "Forte" & Mean < duration_all[Subcomponent == "Forte"]$Mean+3*duration_all[Subcomponent == "Forte"]$SD & Mean > duration_all[Subcomponent == "Forte"]$Mean-3*duration_all$SD]
-print(excluded_duration_for)
-```
-- The number of excluded performance: `r duration_all[Subcomponent == "Forte"]$N-nrow(excluded_duration_for)`
-
-#### Piano
-```{r outlier_sta, echo = FALSE}
-# for piano
-excluded_duration_pia <- duration[Subcomponent == "Piano" & Mean < duration_all[Subcomponent == "Piano"]$Mean+3*duration_all[Subcomponent == "Piano"]$SD & Mean > duration_all$Mean-3*duration_all[Subcomponent == "Piano"]$SD]
-print(excluded_duration_pia)
-```
-- The number of excluded performance: `r duration_all[Subcomponent == "Piano"]$N-nrow(excluded_duration_pia)`
-
-<!-- ```{r export, include = FALSE} -->
-<!-- duration_valid <- dt_duration[, .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr)] -->
-<!-- fwrite(duration_valid, "./duration_valid.csv") -->
-<!-- ``` -->
-
-<!-- # Duration Variability -->
-<!-- ```{r cv, echo = FALSE} -->
-<!-- # for each participant -->
-<!-- cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)] -->
-<!-- print(cv) -->
-<!-- ``` -->
-
-<!-- ```{r cv_box, echo = FALSE} -->
-<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
-<!--   geom_boxplot() + -->
-<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
-<!-- p_cv_box -->
-<!-- ``` -->
-
-<!-- ### Exclude outlier performacnes -->
-<!-- ```{r cv_outlier, echo = FALSE} -->
-<!-- # for all -->
-<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
-<!-- print(cv_all) -->
-
-<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
-<!-- print(excluded_cv) -->
-<!-- ``` -->

### 2021-02-01 material/expstimuli/analysis/expression/stim_d/ioi_onset.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance (IOI-Onset) from teaching-v2.0.
+Description: This is a rough overview of performing performance (IOI-Onset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Performing-Dynamics performances: data from the performing condition (dynamics)
 

### 2021-02-01 material/expstimuli/analysis/expression/stim_d/kv.Rmd
update comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a rough overview of performing performance (KV) from teaching-v2.0.
+***KV values from this analysis were used.***
+
+Description: This is a rough overview of performing performance (KV) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+
+Performing-Dynamics performances: data from the performing condition (dynamics)
 
@@
 ```{r Velocitys, include = FALSE}
-# # convert bpm to ms
-# dt_kv[Tempo == 120]$Tempo <- 250
-# dt_kv[Tempo == 110]$Tempo <- 273
-# dt_kv[Tempo == 100]$Tempo <- 300
-
 # define subcomponents
@@
 
-### Valid performances (exclude outliers)
-#### Forte
+# Valid performances (exclude KVs outliers)
+## Forte
 ```{r outlier_for, echo = FALSE}
@@
 
-#### Piano
+## Piano
 ```{r outlier_pia, echo = FALSE}
@@
 ```
-
-<!-- # KV Variability -->
-<!-- ```{r cv, echo = FALSE} -->
-<!-- # for each participant -->
-<!-- cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)] -->
-<!-- cv -->
-<!-- ``` -->
-
-<!-- ```{r cv_box, echo = FALSE} -->
-<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
-<!--   geom_boxplot() + -->
-<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
-<!-- p_cv_box -->
-<!-- ``` -->
-
-<!-- ### Exclude outlier performances -->
-<!-- ```{r cv_outlier, echo = FALSE} -->
-<!-- # for all -->
-<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
-<!-- print(cv_all) -->
-
-<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
-<!-- print(excluded_cv) -->
-<!-- ``` -->

### 2021-02-01 material/expstimuli/analysis/stim_n/duration.Rmd
summarise data by subnr, trial
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r outlier, echo = FALSE}
-excluded_duration <- duration[Mean < duration_all$Mean+3*duration_all$SD & Mean > duration_all$Mean-3*duration_all$SD]
+duration_trial <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr)]
+
+excluded_duration <- duration_trial[Mean < duration_all$Mean+3*duration_all$SD & Mean > duration_all$Mean-3*duration_all$SD]
 print(excluded_duration)
 ```
-- The number of excluded performance: `r duration_all$N-nrow(excluded_duration)`
+- The number of excluded performance: `r nrow(duration_trial)-nrow(excluded_duration)`
 

### 2021-02-01 material/expstimuli/analysis/stim_n/ioi_offset.Rmd
summarise data by subnr, trial
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
+ioi_trial <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, TrialNr)]
+
+excluded_ioi <- ioi_trial[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
 print(excluded_ioi)
 ```
-- The number of excluded performance: `r ioi_all$N-nrow(excluded_ioi)`
+- The number of excluded performance: `r nrow(ioi_trial)-nrow(excluded_ioi)`
 

### 2021-02-01 material/expstimuli/analysis/stim_n/kv.Rmd
summarise data by subnr, trial
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r outlier, echo = FALSE}
-excluded_kv <- kv[Mean < kv_all$Mean+3*kv_all$SD & Mean > kv_all$Mean-3*kv_all$SD]
+kv_trial <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr)]
+
+excluded_kv <- kv_trial[Mean < kv_all$Mean+3*kv_all$SD & Mean > kv_all$Mean-3*kv_all$SD]
 print(excluded_kv)
 ```
-- The number of excluded performance: `r kv_all$N-nrow(excluded_kv)`
+- The number of excluded performance: `r nrow(kv_trial)-nrow(excluded_kv)`
 

### 2021-02-01 material/expstimuli/analysis/expression/stim_a/duration.Rmd
correct comments and show data by trial
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Valid performances (exclude Duration outliers)
-- Exclude mean Duration outside 3SD
-
+# Included performances for averaging
 ## Legato
@@
 # for legato
-excluded_duration_leg <- duration[Subcomponent == "Legato" & Mean < duration_all[Subcomponent == "Legato"]$Mean+3*duration_all[Subcomponent == "Legato"]$SD & Mean > duration_all[Subcomponent == "Legato"]$Mean-3*duration_all$SD]
-print(excluded_duration_leg)
+duration_trial_leg <- dt_duration[Subcomponent == "Legato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr)]
+duration_trial_leg
 ```
-- The number of excluded performance: `r duration_all[Subcomponent == "Legato"]$N-nrow(excluded_duration_leg)`
 
@@
 # for staccato
-excluded_duration_sta <- duration[Subcomponent == "Staccato" & Mean < duration_all[Subcomponent == "Staccato"]$Mean+3*duration_all[Subcomponent == "Staccato"]$SD & Mean > duration_all$Mean-3*duration_all[Subcomponent == "Staccato"]$SD]
-print(excluded_duration_sta)
+duration_trial_sta <- dt_duration[Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr, Subcomponent)]
+duration_trial_sta
 ```
-- The number of excluded performance: `r duration_all[Subcomponent == "Staccato"]$N-nrow(excluded_duration_sta)`
 

### 2021-02-01 material/expstimuli/analysis/expression/stim_d/kv.Rmd
correct comments and show data by trial
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-# Valid performances (exclude KVs outliers)
+# Included performances for averaging
 ## Forte
@@
 # for forte
-excluded_kv_for <- kv[Subcomponent == "Forte" & Mean < kv_all[Subcomponent == "Forte"]$Mean+3*kv_all[Subcomponent == "Forte"]$SD & Mean > kv_all[Subcomponent == "Forte"]$Mean-3*kv_all$SD]
-print(excluded_kv_for)
+kv_trial_for <- dt_kv[Subcomponent == "Forte", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr)]
+kv_trial_for
 ```
-- The number of excluded performance: `r kv_all[Subcomponent == "Forte"]$N-nrow(excluded_kv_for)`
 
@@
 # for piano
-excluded_kv_pia <- kv[Subcomponent == "Piano" & Mean < kv_all[Subcomponent == "Piano"]$Mean+3*kv_all[Subcomponent == "Piano"]$SD & Mean > kv_all[Subcomponent == "Piano"]$Mean-3*kv_all$SD]
-print(excluded_kv_pia)
+kv_trial_pia <- dt_kv[Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr)]
+kv_trial_pia
 ```
-- The number of excluded performance: `r kv_all[Subcomponent == "Piano"]$N-nrow(excluded_kv_pia)`
 

### 2021-02-02 material/expstimuli/selected/IOI/mid_export.py
select 4 performances for IOIs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,52 @@
+#!/usr/local/bin/python
+####################
+# Created: 20/04/2020
+# This script converts txt files to mid files.
+
+#%% import packages
+import os, csv, mido
+
+#%% create mid folder if not existed
+if not os.path.exists("mid"):
+    os.mkdir("mid")
+
+#%% folder names
+folder_ioi = "./1612276046-020221/"
+folders = [folder_ioi]
+
+#%% mid export
+for folder in folders:
+    for instance in range(1, 17):
+        filename = folder + str(instance) + "_instance.txt"
+        with open(filename) as csvfile:
+            # create mid file
+            mid = mido.MidiFile()
+            track = mido.MidiTrack()
+            mid.tracks.append(track)
+            track.append(mido.Message('program_change', program=0, time=0)) # program 0 = Acoustic Grand Piano
+            # name for mid file
+            if folder == "./1612276046-020221/":
+                midname = "./mid/ioi_" + str(instance) + ".mid"
+            # read current track data (txt file)   
+            current = csv.reader(csvfile, delimiter = ",")
+            next(current) # skip first row
+            counter = 0
+            for row in current:
+                if counter == 0:
+                    previousTime = 0
+                    counter += 1
+                else:
+                    previousTime = currentTime
+                currentTime = int(mido.second2tick(int(row[3])*0.001+3, 480, 500000))
+                currentPitch = int(row[5])
+                currentVelocity = int(row[4])
+                currentOnOff = int(row[2])
+                # assign midi values
+                if currentOnOff == 1:
+                    track.append(mido.Message('note_on', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
+                elif currentOnOff == 0:
+                    track.append(mido.Message('note_off', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
+            mid.save(midname) # save track
+
+print("Done :D")
+# %%

### 2021-02-02 material/prerating/data/summary.Rmd
select 4 stimuli from each category
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2020-09-04 script/psychopy/data/summary.Rmd
@@
 ---
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
 
-**Participants:** Anna, Thomas and me
+# Descriptive Stats for the generated stimuli
+
+**Participants:** 3 including myself (author)
 
@@
 - The stimuli were presented randomly for each participant.
-- The order of rating (articulation first or dynamics first) was randomised within participants, but the half of the ratings started with articulation and the other half started with 
+- The order of rating (articulation first or dynamics first) was randomised within participants, but the half of the ratings started with articulation and the other half started with dynamics.
 - There were 3 practice trials.
@@
 - Pick 4 performances which represent the characteristics of each category the most.
-
-**Issues:**
-
-- How to select? Need stats?
 
@@
 ```{r raw, echo = FALSE}
-# Create a list of data file names
-lf <- list.files(".", pattern = "csv")
+# create raw_data - merge all data files into one
+data_ls <- list.files(".", pattern = "csv")
+combined <- lapply(data_ls, function(f){
+  fread(paste("./", f, sep = ""), sep = ",")
+})
+raw_data <- do.call(rbind.data.frame, combined)
 
-# Create raw_data - merge all data files into one
-raw_data <- data.table()
-for (i in 1:length(lf)){
-  data_i <- fread(file.path(".", lf[i]), header = T, sep = ",", dec = ".")
-  raw_data <- rbind(raw_data, data_i)
-}
 # remove unnecessary strings from midFile
@@
 # overall rating for each category
-summary_data <- data[, .(N = .N/2, meanRating = mean(rating), SD = sd(rating), SEM = sd(rating)/sqrt(.N/2)), by = .(stimCategory, ratingCategory)]
+summary_data <- data[, .(N = .N, meanRating = mean(rating), SD = sd(rating), SEM = sd(rating)/sqrt(.N)), by = .(stimCategory, ratingCategory)]
 summary_data[order(ratingCategory, stimCategory)]
@@
 ```
-**Comments:**
-
-It seems participants can easily detect articulation but not dynamics.
 
@@
 
-## Feedback from Anna and Thomas
-- If expressions are not matched with notated ones but he likes the expressions, what to rate?
-    + Judge to what extent expressions are corresponded with notated ones
-- There are many changes in the first system (especially for articulation), it was easy to judge articulation and then focused on dynamics afterwards.
-When rating, forgot what he thought about articulation.
-- Staccato and piano seem to be correlated very much. Could not tell the intention of producing piano. Might be that participants played softly (piano) but just as a result of staccato, for instance.
-- The same one for the definition of expressions (should be matched or whether I like them?)
-- Definition of staccato.
-Especially when participants are not implementing articulation, all sounds like staccato or just detached (detaché).
-she did not know whether participants meant to implement staccato or not.
-Range of staccato is much smaller than that of legato, so hard to judge.
-How long should staccato be?
-- Range 1-5 is too rough as differences in each performance are so subtle.
-- **Several performances have timing issues.**
-    + I only looked at 8th not so maybe better to look at 
-- Wordings for instructions: to what extent participants follow instructions
+# Selecting stimuli
+## High performance
+- sum the ratings for both articulation and dynamics and rank the performances by Sum
 
-## Feedback for my performance
-- As I generated stimuli, I am not sure whether my responses are valid because I could mostly categorise stimuli for 4 types even though they were presented randomly (of course some of them, I couldn't).
+```{r ranking-high, echo = FALSE}
+data_high <- data_high[order(midFile, ratingCategory)]
+data_high_sep <- data_high[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_high_both <- data_high[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_high_both[order(-Sum)]
+```
+
+### Selected 4 performances
+```{r, echo = FALSE}
+select_high <- data_high_both[order(-Sum)]
+select_high$Rank <- 1:nrow(select_high)
+select_high$Category <- "High"
+
+# dummy
+select_high$Sample <- 0
+select_high$Diff <- 0
+select_high$Total <- 0
+
+select_high[Rank < 5]
+```
+
+## Articulation performance
+- sum the ratings for articulation (Sum) and calculate the difference between articulation and dynamics ratings (Diff)
+- find bigger articulation ratings and bigger differences between articulation and dynamics ratings
+- rank the performances by Total
+
+```{r ranking-art, echo = FALSE}
+data_art <- data_art[order(midFile, ratingCategory)]
+data_art_sep <- data_art[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_art_art <- data_art[ratingCategory == "articulation", .(N = .N, Sum = sum(rating)), by = .(midFile)]
+data_art_art$Diff <- data_art_sep[ratingCategory == "articulation"]$Sum - data_art_sep[ratingCategory == "dynamics"]$Sum
+
+# articulation ratings + difference between articulation and dynamics ratings (bigger values are good stimuli)
+data_art_art$Total <- data_art_art$Sum + data_art_art$Diff
+data_art_art[order(-Total)]
+```
+
+### Selected 4 performances
+```{r, echo = FALSE}
+select_art <- data_art_art[order(-Total)]
+select_art$Rank <- 1:nrow(select_art)
+select_art$Category <- "Art_only"
+
+# randomly select 3 from Total == 22
+select_art$Sample <- 0
+select_art[Rank == 1]$Sample <- 1
+select_art[Total == 22]$Sample <- sample(c(1:nrow(select_art[Total == 22])), replace = FALSE)
+select_art[Sample != 0 & Sample < 4]
+```
+
+## Dynamics performance
+- sum the ratings for dynamics (Sum) and calculate the difference between dynamics and articulation ratings (Diff)
+- find bigger dynamics ratings and bigger differences between dynamics and articulation ratings
+- rank the performances by Total
+
+```{r ranking-dyn, echo = FALSE}
+data_dyn <- data_dyn[order(midFile, ratingCategory)]
+data_dyn_sep <- data_dyn[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_dyn_dyn <- data_dyn[ratingCategory == "dynamics", .(N = .N, Sum = sum(rating)), by = .(midFile)]
+data_dyn_dyn$Diff <- data_dyn_sep[ratingCategory == "dynamics"]$Sum - data_dyn_sep[ratingCategory == "articulation"]$Sum
+
+# dynamics ratings + difference between dynamics and articulation ratings (bigger values are good stimuli)
+data_dyn_dyn$Total <- data_dyn_dyn$Sum + data_dyn_dyn$Diff
+data_dyn_dyn[order(-Total)]
+```
+
+### Selected 4 performances
+```{r, echo = FALSE}
+select_dyn <- data_dyn_dyn[order(-Total)]
+select_dyn$Rank <- 1:nrow(select_dyn)
+select_dyn$Category <- "Dyn_only"
+
+# dummy
+select_dyn$Sample <- 0
+
+select_dyn[Total >= 11]
+```
+
+## Low performance
+- sum the ratings for both articulation and dynamics and rank the performances by Sum
+
+```{r ranking-low, echo = FALSE}
+data_low <- data_low[order(midFile, ratingCategory)]
+data_low_sep <- data_low[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
+data_low_both <- data_low[, .(N = .N, Sum = sum(rating)), by = .(midFile)]
+
+data_low_both[order(Sum)]
+```
+
+### Selected 4 performances
+```{r, echo = FALSE}
+select_low <- data_low_both[order(Sum)]
+select_low$Rank <- 1:nrow(select_low)
+select_low$Category <- "Low"
+select_low <- select_low[Sum == min(select_low$Sum)]
+
+# randomly select 4
+select_low$Sample <- sample(c(1:nrow(select_low)), replace = FALSE)
+
+# dummy
+select_low$Diff <- 0
+select_low$Total <- 0
+
+select_low[Sample < 5]
+```
+# Export selected stimuli
+```{r, echo = FALSE}
+selected <- rbind(select_high[Rank < 5], select_art[Sample != 0 & Sample < 4], select_dyn[Total >= 11], select_low[Sample < 5])
+selected
+
+# export txt file
+filename = paste("./selected_", format(Sys.time(), "%s-%d%m%y"), ".txt", sep = "") # current time
+fwrite(selected, filename)
+```

### 2021-02-02 material/expstimuli/selected/IOI/mid_export.py

@@ -1,6 +1,5 @@
 #!/usr/local/bin/python
 ####################
-# Created: 20/04/2020
 # This script converts txt files to mid files.
 
 #%% import packages

### 2021-02-03 material/expstimuli/selected/combine.R
combine selected performances and export exp stimuli
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,192 @@
+# /usr/local/bin/r
+# Combine selected performances and create experimental stimuli
+# Created: 03/02/2021
+
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+
+# read base tempo
+data_ls <- list.files("./IOI/tempo/", pattern = "txt")
+combined <- lapply(data_ls, function(f){
+  fread(paste("./IOI/tempo/", f, sep = ""), sep = ",")
+})
+all_data_onset <- do.call(rbind.data.frame, combined)
+all_data_onset$KeyOnOff <- 1 # define onsets
+
+# assign stimuli number
+all_data_onset$Stimuli <- 0
+counter = 1
+for (i in unique(all_data_onset$Instance)){
+  all_data_onset[Instance == i]$Stimuli <- counter
+  counter = counter + 1
+}
+all_data_onset[, c("Mean","SD", "N")] <- NULL # remove irrelevant columns
+
+# read info about selected performances
+selected <- fread("../../prerating/data/selected_1612309984-030221.txt")
+
+# create TimeStamp
+all_data_onset$TimeStamp <- 0
+for (i in 1:nrow(all_data_onset)){
+  if (all_data_onset$RowNr[i] == 1){
+    all_data_onset$TimeStamp[i] <- 0
+  } else {
+    all_data_onset$TimeStamp[i] <- round(all_data_onset$TimeStamp[i-1]+all_data_onset$IOI[i])
+  }
+}
+
+# create Pitch
+dt_ideal <- fread("./ideal.txt")
+all_data_onset$Pitch <- rep(dt_ideal$V1, 4)
+
+# 1. High performances
+high_art <- fread("../original/averaging/art/1596379937-020820/dt_du_instance.txt")
+high_dyn <- fread("../original/averaging/dyn/1596379899-020820/dt_kv_instance.txt")
+
+# offsets
+du_high <- data.table()
+for (i in unique(selected[Category == "High"]$midFile)){
+  current <- high_art[Instance == i]
+  du_high <- rbind(du_high, current)
+}
+
+all_data_offset_high <- all_data_onset[, c("Instance" ,"Stimuli", "RowNr")]
+all_data_offset_high$KeyOnOff <- 0
+all_data_offset_high$Instance <- du_high$Instance
[…]

### 2021-02-03 material/expstimuli/selected/mid_export.py

@@ -1,0 +1,49 @@
+#!/usr/local/bin/python
+####################
+# Created: 03/02/2021
+# This script converts txt files to mid files.
+
+#%% import packages
+import os, csv, mido
+
+#%% create mid folder if not existed
+if not os.path.exists("mid"):
+    os.mkdir("mid")
+
+#%% folder names
+folder = "./stimuli/"
+
+#%% mid export
+for file in os.listdir(folder):
+    filename = folder + file
+    with open(filename) as csvfile:
+        # create mid file
+        mid = mido.MidiFile()
+        track = mido.MidiTrack()
+        mid.tracks.append(track)
+        track.append(mido.Message('program_change', program=0, time=0)) # program 0 = Acoustic Grand Piano
+        # name for mid file
+        midname = "./mid/" + file.replace('.txt', '') + ".mid"
+        # read current track data (txt file)   
+        current = csv.reader(csvfile, delimiter = ",")
+        next(current) # skip first row
+        counter = 0
+        for row in current:
+            if counter == 0:
+                previousTime = 0
+                counter += 1
+            else:
+                previousTime = currentTime
+            currentTime = int(mido.second2tick(int(row[4])*0.001+3, 480, 500000))
+            currentPitch = int(row[5])
+            currentVelocity = int(row[6])
+            currentOnOff = int(row[2])
+            # assign midi values
+            if currentOnOff == 1:
+                track.append(mido.Message('note_on', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
+            elif currentOnOff == 0:
+                track.append(mido.Message('note_off', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
+        mid.save(midname) # save track
+
+print("Done :D")
+# %%

### 2021-02-03 material/expstimuli/selected/combine.R
remove unnecessary part
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -172,21 +172,7 @@ dir.create("./stimuli")
 for (category in unique(all_data$Category)){
   for (i in 1:4){
     current <- all_data[Category == category & Stimuli == i]
-    filename = paste("./stimuli/", tolower(category), )
+    filename = paste("./stimuli/", tolower(category), "_", i, ".txt", sep = "")
     fwrite(current, filename)
   }
-}
-
-for (i in 1:16){
-  onset <- dt_playback_onset[Instance == i]
-  offset <- dt_playback_offset[Instance == i]
-  onset$Pitch <- dt_ideal$V1
-  offset$Pitch <- dt_ideal$V1
-  instance <- rbind(onset, offset)
-  # sort by TimeStamp
-  instance <- instance[order(TimeStamp)]
-  
-  # export txt
-  filename = paste(foldername, i, "_instance.txt", sep = "")
-  fwrite(instance, filename)
 }

### 2021-02-15 material/expstimuli/practice/random.R
correct paths and adjust tempo
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -21,8 +21,8 @@ filename_offset1 = "../analysis/stim_n/preprocessor/filtered/data_offset.csv"
 filename_valid_du1 = "../analysis/stim_n/duration_valid.csv"
 filename_valid_kv1 = "../analysis/stim_n/kv_valid.csv"
 # 2. data for raw expressive performances
-filename_onset2 = "../analysis/expression/filtered/data_onset.csv"
-filename_offset2 = "../analysis/expression/filtered/data_offset.csv"
+filename_onset2 = "../analysis/expression/preprocessor/filtered/data_onset.csv"
+filename_offset2 = "../analysis/expression/preprocessor/filtered/data_offset.csv"
 filename_valid_du2 = "../analysis/expression/stim_a/duration_valid.csv"
 filename_valid_kv2 = "../analysis/expression/stim_d/kv_valid.csv"
 
@@ -50,9 +50,6 @@ valid1$Skill <- "non-expression"
 # 2.expressive performances
 dt_onset2 <- fread(filename_onset2, header = T, sep = ",", dec = ".")
 dt_offset2 <- fread(filename_offset2, header = T, sep = ",", dec = ".")
-# remove unnecessary rows/columns
-dt_onset2$RowNr <- NULL
-dt_offset2$RowNr <- NULL
 dt_onset2 <- dt_onset2[Condition == "performing"]
 dt_offset2 <- dt_offset2[Condition == "performing"]
 # valid performances for each parameter
@@ -102,9 +99,9 @@ for (subnr in unique(dt_demo$SubNr)){
   current <- dt_demo[SubNr == subnr]
   current$TimeStamp <- current$TimeStamp-current$TimeStamp[1]
   current <- current[order(TimeStamp),]
-  # if (unique(current$Tempo) != 100){ # NEED TO BE FIXED!
-  #   current$TimeStamp <- round((current$TimeStamp)/unique(current$Tempo)*100, 0)
-  # }
+  if (unique(current$Tempo) != 100){
+    current$TimeStamp <- round((current$TimeStamp)/unique(current$Tempo)*100, 0)
+  }
   if (length(unique(current$Skill)) > 1){
     print("It seems the same participant was used more than twice. Run the script again.")
   } else if (unique(current$Skill == "non-expression")){

### 2021-02-15 material/expstimuli/practice/mid_export.py
correct comments and remove unnecessary parts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -39,12 +39,6 @@ for folder in folders:
                     counter += 1
                 else:
                     previousTime = currentTime
-                # tempo always 100bpm
-                if int(row[6]) == 120:
-                    currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, round(500000*(500/600), 0))) # adjust tempo
-                elif int(row[6]) == 110:
-                    currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, round(500000*(545/600), 0))) # adjust tempo
-                elif int(row[6]) == 100:
                     currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, round(500000*(600/600), 0)))
                 currentPitch = int(row[2])
                 currentVelocity = int(row[3])

### 2021-02-15 material/expstimuli/practice/random.R

@@ -72,7 +72,7 @@ stim_a <- min(valid[Skill == "articulation"]$Sample)
 stim_d <- min(valid[Skill == "dynamics"]$Sample)
 writeLines(paste("stim_n: ", stim_n, "stim_a: ", stim_a, "stim_d: ", stim_d), paste(foldername, "demo.txt", sep = ""))
 
-# remove first silence and adjust tempo for 100 bpm
+# create demo data
 dt_demo <- data.table()
 counter = 0
 for (i in c(stim_n, stim_a, stim_d)){

### 2021-02-15 material/expstimuli/tweaked/analysis.Rmd
summary for selected stimuli and tweak problematic KOTs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Adaptation: Analysis for selected stimuli"
output: html_notebook
---

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r midi, include = FALSE}
# create a list of data file names
lf <- list.files("../selected/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../selected/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:144, nrow(data)/144)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$SubArt <- "NA"
data$SubDyn <- "NA"
data$MidFileShort <- gsub("art_only", "a", data$MidFile)
data$MidFileShort <- gsub("dyn_only", "d", data$MidFileShort)
data$MidFileShort <- gsub("high", "h", data$MidFileShort)
data$MidFileShort <- gsub("low", "l", data$MidFileShort)

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (velocity)
ls_legato_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

# IOI
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Interval == number]$SubArt <- "LtoS"
 dt_ioi[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_ioi[Interval == number]$SubArt <- "StoL"
 dt_ioi[Interval == number]$SubDyn <- "PtoF"
}
```

- There are 4 base tempi

## Histogram
```{r ioi-distribution, echo = FALSE}
gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean
```{r ioi-mean, echo = FALSE}
# mean
dt_ioi[SubArt != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), "MidFile"]

ggboxplot(dt_ioi[SubArt != "NA"], x = "MidFileShort", y = "IOI", add = "jitter", width = 0.8)
```

# KOT

```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time[row] - data_onset$time[row+1] # offset(n) - onset(n+1)
   }
}

dt_kot <- data_onset[NoteNr != 1]
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_kot[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_kot[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_kot[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_kot[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_kot[Interval == number]$SubArt <- "LtoS"
 dt_kot[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_kot[Interval == number]$SubArt <- "StoL"
 dt_kot[Interval == number]$SubDyn <- "PtoF"
}
```

## Histogram
```{r kot-distribution, echo = FALSE}
gghistogram(dt_kot[SubArt != "NA"], x = "KOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean
```{r kot-mean, echo = FALSE}
# mean
dt_kot[SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, SubArt)]

ggboxplot(dt_kot[SubArt != "NA"], x = "MidFileShort", y = "KOT", color = "SubArt", add = "jitter", width = 0.8)
```

## Tweak
```{r kot-tweak, echo = FALSE}
# grandMean of art_only & high stimuli
kot_grandMean <- mean(dt_kot[grepl("art_only", MidFile) | grepl("high", MidFile) & SubArt != "NA"]$KOT)

# calculate the diff between each KOT mean and grandMean
tweak_data <- dt_kot[grepl("dyn_only", MidFile) | grepl("low", MidFile) & SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]
tweak_data$Diff <- abs(tweak_data$Mean-kot_grandMean)
tweak_data

# add each fixed Diff value to TimeStamp
dt_kot$FixedKOT <- dt_kot$KOT
for (midfile in unique(tweak_data$MidFile)){
   dt_kot[MidFile == midfile]$FixedKOT <- dt_kot[MidFile == midfile]$KOT+tweak_data[MidFile == midfile]$Diff
}
```

## Histogram (Fixed KOT)
```{r kot-distribution-fixed, echo = FALSE}
gghistogram(dt_kot[SubArt != "NA"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean (Fixed KOT)
```{r kot-mean-fixed, echo = FALSE}
# mean
dt_kot[SubArt != "NA", .(N = .N, Mean = mean(FixedKOT), SD = sd(FixedKOT)), by = .(MidFile, SubArt)]

ggboxplot(dt_kot[SubArt != "NA"], x = "MidFileShort", y = "FixedKOT", color = "SubArt", add = "jitter", width = 0.8)
```

# KV

```{r vel, include = FALSE}
dt_vel <- data_onset
dt_vel$NoteOnsetNr <- rep(1:72, nrow(dt_vel)/72)
# assign Subcomponents
# for each note
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_legato_2[[phrase]][note]]$SubArt <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_staccato_2[[phrase]][note]]$SubArt <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_forte_2[[phrase]][note]]$SubDyn <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_piano_2[[phrase]][note]]$SubDyn <- "Piano"
   }
}
```

## Histogram
```{r vel-distribution, echo = FALSE}
gghistogram(dt_vel[SubArt != "NA"], x = "Velocity", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
```

## Mean
```{r vel-mean, echo = FALSE}
# mean
dt_vel[SubDyn != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(MidFile, SubDyn)]

ggboxplot(dt_vel[SubArt != "NA"], x = "MidFileShort", y = "Velocity", color = "SubDyn", add = "jitter", width = 0.8)
```

# KV Diff
```{r vel-diff, include = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_vel_diff[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_vel_diff[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_vel_diff[Interval == number]$SubArt <- "LtoS"
 dt_vel_diff[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_vel_diff[Interval == number]$SubArt <- "StoL"
 dt_vel_diff[Interval == number]$SubDyn <- "PtoF"
}
```

## Histogram
```{r vel-diff-distribution, echo = FALSE}
gghistogram(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "Diff", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
```

## Mean
```{r vel-diff-mean, echo = FALSE}
# mean
dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(MidFile, SubDyn)]

ggboxplot(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "MidFileShort", y = "Diff", color = "SubDyn", add = "jitter", width = 0.8)
```

### 2021-02-16 material/expstimuli/tweaked/mid_export.py
add py script to export midi files
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,0 +1,47 @@
+#!/usr/local/bin/python
+####################
+# This script converts txt files to mid files.
+
+#%% import packages
+import os, csv, mido
+
+#%% create mid folder if not existed
+if not os.path.exists("mid"):
+    os.mkdir("mid")
+
+#%% list exported instances
+files = [f for f in os.listdir("./") if f.endswith('.txt')]
+
+#%% mid export
+for file in files:
+    with open(file) as csvfile:
+        # create mid file
+        mid = mido.MidiFile()
+        track = mido.MidiTrack()
+        mid.tracks.append(track)
+        track.append(mido.Message('program_change', program=0, time=0)) # program 0 = Acoustic Grand Piano
+        # name for mid file
+        midname = "./mid/" + file.replace(".txt",".mid")
+        # read current track data (txt file)   
+        current = csv.reader(csvfile, delimiter = ",")
+        next(current) # skip first row
+        counter = 0
+        for row in current:
+            if counter == 0:
+                previousTime = 0
+                counter += 1
+            else:
+                previousTime = currentTime
+            currentTime = int(mido.second2tick(int(row[9])*0.001, 480, 500000))
+            currentPitch = int(row[2])
+            currentVelocity = int(row[3])
+            currentOnOff = int(row[10])
+            # assign midi values
+            if currentOnOff == 1:
+                track.append(mido.Message('note_on', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
+            elif currentOnOff == 0:
+                track.append(mido.Message('note_off', note=currentPitch, velocity=currentVelocity, time=currentTime-previousTime))
+        mid.save(midname) # save track
+
+print("Done :D")
+# %%

### 2021-02-16 material/expstimuli/tweaked/analysis.Rmd
add hline
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
calculate new time (timeFixed for offsets) based on fixedKOT
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- There are 4 base tempi
+- There are 4 base tempi.
 
@@
 ## Mean
+- The red line indicates the ideal tempo (100bpm - IOI: 300 ms)
+
 ```{r ioi-mean, echo = FALSE}
@@
 
-ggboxplot(dt_ioi[SubArt != "NA"], x = "MidFileShort", y = "IOI", add = "jitter", width = 0.8)
+ggboxplot(dt_ioi[SubArt != "NA"], x = "MidFileShort", y = "IOI", add = "jitter", width = 0.8) + geom_hline(yintercept = 300, linetype = "dashed", color = "red")
 ```
@@
 # KOT
+- For art_only and high stimuli, there should be a bimodal distribution to represent legato and staccato.
+
+- For dyn_only and low stimuli, there should be a normal distribution.
 
@@
 ## Tweak
+- For dyn_only and low stimuli, it seems the KOT values are too close to the staccato range.
+We calculated the difference between the mean of each stimulus and the grand mean of art_only and high stimuli and add that difference to all KOT values for dyn_only and low, so that the KOT values should shift closer to the grand mean of art_only and high stimuli.
+
 ```{r kot-tweak, echo = FALSE}
@@
 ```
+
+# Export stimuli with fixed KOT as mid files
+```{r export, include = FALSE}
+# fix time for onset and offset data
+data_offset$timeFixed <- data_onset$time
+data_offset$timeFixed <- data_offset$time
+for (midfile in unique(tweak_data$MidFile)){
+   # offset[i] = onset[i+1]+KOT[i]
+   for (i in 1:nrow(data_onset[MidFile == midfile])-1){
+      data_offset[MidFile == midfile][i]$timeFixed <- data_onset[MidFile == midfile][i+1]$time+dt_kot[MidFile == midfile][i]$FixedKOT
+   }
+}
+
+for (stim in unique(data_onset$MidFile)){
+   onset <- data_onset[MidFile == stim]
+   offset <- data_offset[MidFile == stim]
+   instance <- rbind(onset, offset)
+   # round timeFixed
+   instance$timeFixed <- round(instance$timeFixed, 0)
+   # Note_On/Off >> 1 or 0
+   instance$KeyOnOff <- 0
+   instance[event == "Note On"]$KeyOnOff <- 1
+   # sort by timeFixed
+   instance <- instance[order(timeFixed)]
+
+   filename = paste(stim, ".txt", sep = "")
+   fwrite(instance, filename)
+}
+```
+
+```{bash, echo = TRUE}
+# python
+source /Users/atsukotominaga/Pyenv/bin/activate
+python mid_export.py
+```

### 2021-07-05 material/expstimuli/practice/mid_export.py
tempo change
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -39,7 +39,12 @@ for folder in folders:
                     counter += 1
                 else:
                     previousTime = currentTime
-                    currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, round(500000*(600/600), 0)))
+                if int(row[6]) == 100:
+                    currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, 500000*(500/500)))
+                elif int(row[6]) == 110:
+                    currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, 500000*(500/545)))
+                elif int(row[6]) == 120:
+                    currentTime = int(mido.second2tick(int(row[1])*0.001+3, 480, 500000*(500/600)))
                 currentPitch = int(row[2])
                 currentVelocity = int(row[3])
                 currentOnOff = int(row[4])

### 2021-07-05 material/expstimuli/practice/random.R

@@ -99,9 +99,6 @@ for (subnr in unique(dt_demo$SubNr)){
   current <- dt_demo[SubNr == subnr]
   current$TimeStamp <- current$TimeStamp-current$TimeStamp[1]
   current <- current[order(TimeStamp),]
-  if (unique(current$Tempo) != 100){
-    current$TimeStamp <- round((current$TimeStamp)/unique(current$Tempo)*100, 0)
-  }
   if (length(unique(current$Skill)) > 1){
     print("It seems the same participant was used more than twice. Run the script again.")
   } else if (unique(current$Skill == "non-expression")){

### 2021-07-05 material/expstimuli/analysis/stim_n/preprocessor/filtering.R
df_ > dt_
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -28,9 +28,9 @@ if (!file.exists("filtered")){
 }
 
 # read a text file for ideal performance
-df_ideal <- fread("./ideal.txt")
-colnames(df_ideal) <- "Pitch"
-df_ideal$RowNr <- c(1:nrow(df_ideal))
+dt_ideal <- fread("./ideal.txt")
+colnames(dt_ideal) <- "Pitch"
+dt_ideal$RowNr <- c(1:nrow(dt_ideal))
 
 # create a list of data file names
 lf <- list.files("./practice", pattern = "txt")
@@ -54,55 +54,55 @@ raw_data$Image <- gsub(";", "", raw_data$Image)
 raw_data <- raw_data[order(raw_data$SubNr, raw_data$BlockNr, raw_data$TrialNr),]
 
 # separate raw_data into each skill
-df_stim_n <- raw_data[BlockNr == 0] # non-expressive sheet music
-df_stim_a <- raw_data[Skill == "articulation"]
-df_stim_d <- raw_data[Skill == "dynamics"]
+dt_stim_n <- raw_data[BlockNr == 0] # non-expressive sheet music
+dt_stim_a <- raw_data[Skill == "articulation"]
+dt_stim_d <- raw_data[Skill == "dynamics"]
 
 ####################################
 # Detect pitch errors
 ####################################
 # raw_data without metronome
-df_all <- df_stim_n[Pitch != 31 & Pitch != 34]
+dt_all <- dt_stim_n[Pitch != 31 & Pitch != 34]
 
 # onset and offset
-df_onset <- df_all[Key_OnOff == 1]
-df_offset <- df_all[Key_OnOff == 0]
+dt_onset <- dt_all[Key_OnOff == 1]
+dt_offset <- dt_all[Key_OnOff == 0]
 
 ####################################
 # ONSET
 ####################################
 # detect pitch errors
-ls_error_onset  <- pitch_remover(df_onset, df_ideal)
+ls_error_onset  <- pitch_remover(dt_onset, dt_ideal)
 
 # mark trials with errors by the first filtering
-df_onset$Error <- 0
+dt_onset$Error <- 0
 for (error in 1:length(ls_error_onset)){
-  df_onset[SubNr == ls_error_onset[[error]][1] & TrialNr == ls_error_onset[[error]][2]]$Error <- 1
[…]

### 2021-07-05 material/expstimuli/analysis/stim_n/preprocessor/function.R

@@ -4,8 +4,8 @@
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 
 ### define pitch remove function (pitch_remover)
-# data: data of all trials (df_onset/df_offset)
-# ideal: df_ideal
+# data: data of all trials (dt_onset/dt_offset)
+# ideal: dt_ideal
 pitch_remover <- function(data, ideal){
   ls_removed <- list()
   for (subnr in unique(data$SubNr)){ #set # of participants
@@ -23,7 +23,7 @@ pitch_remover <- function(data, ideal){
             if (current$Pitch[note] != ideal$Pitch[note]){
               while (counter == 0){
                 ls_removed <- c(ls_removed, list(c(subnr, trial, "Pitch error")))
-                print(sprintf("Pitch error - SubNr/TrialNr/NoteNr: %i//%i/%i", subnr, trial, note))
+                print(sprintf("Pitch error - SubNr/TrialNr/NoteNr: %i/%i/%i", subnr, trial, note))
                 counter = counter + 1
               }
             }

### 2021-07-05 material/expstimuli/analysis/stim_n/preprocessor/function.R
update comments in the code
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -16,7 +16,7 @@ pitch_remover <- function(data, ideal){
         if (length(ideal$Pitch) != length(current$Pitch)){ #if # of onsets/offsets is not equal to ideal performance
           ls_removed <- c(ls_removed, list(c(subnr, trial, "NoteNr error")))
           print(sprintf("NoteNr error - SubNr/TrialNr: %i/%i", subnr, trial))
-        } else if (length(ideal$Pitch) == length(current$Pitch)) { #if # of onsets and offsets are correct
+        } else if (length(ideal$Pitch) == length(current$Pitch)) { #if # of onsets/offsets is correct
           counter = 0 #set a counter so that the following loop will terminate once it detects one pitch error in a trial
           for (note in 1:length(ideal$Pitch)){
             # detect onset error
@@ -29,7 +29,7 @@ pitch_remover <- function(data, ideal){
             }
           }
         }
-      } else { #if current data is emtpy"
+      } else { #if current data is empty"
         ls_removed <- c(ls_removed, list(c(subnr, trial, "Missing")))
         print(sprintf("Missing - SubNr/TrialNr: %i/%i", subnr, trial))
       }

### 2021-07-05 material/expstimuli/analysis/stim_n/preprocessor/filtering.R
write.csv > fwrite
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -100,8 +100,8 @@ dt_correct_offset <- dt_offset[Error == 0]
 # Export csv files
 ####################################
 # Export corrected onset/offset for stim_n
-write.csv(dt_correct_onset, file = "./filtered/data_onset.csv", row.names = F)
-write.csv(dt_correct_offset, file = "./filtered/data_offset.csv", row.names = F)
+fwrite(dt_correct_onset, file = "./filtered/data_onset.csv", row.names = F)
+fwrite(dt_correct_offset, file = "./filtered/data_offset.csv", row.names = F)
 
 # check whether there is still error
 pitch_remover(dt_correct_onset, dt_ideal)

### 2021-07-05 material/expstimuli/analysis/expression/preprocessor/filtering.R
write.csv > fwrite, df_ > dt_
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -28,9 +28,9 @@ if (!file.exists("filtered")){
 }
 
 # read a text file for ideal performance
-df_ideal <- fread("./ideal.txt")
-colnames(df_ideal) <- "Pitch"
-df_ideal$RowNr <- c(1:nrow(df_ideal))
+dt_ideal <- fread("./ideal.txt")
+colnames(dt_ideal) <- "Pitch"
+dt_ideal$RowNr <- c(1:nrow(dt_ideal))
 
 # create a list of data file names
 lf <- list.files("./data", pattern = "txt")
@@ -54,55 +54,55 @@ raw_data$Image <- gsub(";", "", raw_data$Image)
 raw_data <- raw_data[order(raw_data$SubNr, raw_data$BlockNr, raw_data$TrialNr),]
 
 # separate raw_data into each skill
-df_stim_a <- raw_data[Skill == "articulation"]
-df_stim_d <- raw_data[Skill == "dynamics"]
-df_stim <- rbind(df_stim_a, df_stim_d)
+dt_stim_a <- raw_data[Skill == "articulation"]
+dt_stim_d <- raw_data[Skill == "dynamics"]
+dt_stim <- rbind(dt_stim_a, dt_stim_d)
 
 ####################################
 # Detect pitch errors
 ####################################
 # raw_data without metronome
-df_all <- df_stim[Pitch != 31 & Pitch != 34]
+dt_all <- dt_stim[Pitch != 31 & Pitch != 34]
 
 # onset and offset
-df_onset <- df_all[Key_OnOff == 1]
-df_offset <- df_all[Key_OnOff == 0]
+dt_onset <- dt_all[Key_OnOff == 1]
+dt_offset <- dt_all[Key_OnOff == 0]
 
 ####################################
 # ONSET
 ####################################
 # detect pitch errors
-ls_error_onset  <- pitch_remover(df_onset, df_ideal)
+ls_error_onset  <- pitch_remover(dt_onset, dt_ideal)
 
 # mark trials with errors by the first filtering
-df_onset$Error <- 0
+dt_onset$Error <- 0
 for (error in 1:length(ls_error_onset)){
-  df_onset[SubNr == ls_error_onset[[error]][1] & BlockNr == ls_error_onset[[error]][2] & TrialNr == ls_error_onset[[error]][3]]$Error <- 1
[…]

### 2021-07-05 material/expstimuli/analysis/expression/preprocessor/function.R

@@ -4,8 +4,8 @@
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 
 ### define pitch remove function (pitch_remover)
-# data: data of all trials (df_onset/df_offset)
-# ideal: df_ideal
+# data: data of all trials (dt_onset/dt_offset)
+# ideal: dt_ideal
 pitch_remover <- function(data, ideal){
   ls_removed <- list()
   for (subnr in unique(data$SubNr)){ #set # of participants
@@ -17,7 +17,7 @@ pitch_remover <- function(data, ideal){
           if (length(ideal$Pitch) != length(current$Pitch)){ #if # of onsets/offsets is not equal to ideal performance
             ls_removed <- c(ls_removed, list(c(subnr, block, trial, "NoteNr error")))
             print(sprintf("NoteNr error - SubNr/BlockNr/TrialNr: %i/%i/%i", subnr, block, trial))
-          } else if (length(ideal$Pitch) == length(current$Pitch)) { #if # of onsets and offsets are correct
+          } else if (length(ideal$Pitch) == length(current$Pitch)) { #if # of onsets/offsets is correct
             counter = 0 #set a counter so that the following loop will terminate once it detects one pitch error in a trial
             for (note in 1:length(ideal$Pitch)){
               # detect onset error
@@ -30,7 +30,7 @@ pitch_remover <- function(data, ideal){
               }
             }
           }
-        } else { #if current data is emtpy"
+        } else { #if current data is empty"
           ls_removed <- c(ls_removed, list(c(subnr, block, trial, "Missing")))
           print(sprintf("Missing - SubNr/BlockNr/TrialNr: %i/%i/%i", subnr, block, trial))
         }

### 2021-07-05 material/expstimuli/analysis/stim_n/duration.Rmd
bar plots >> box plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/stim_n/duration.Rmd
@@
 ---
+
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
 
@@
 theme_set(theme_classic())
-theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
+theme_update(legend.position = "bottom")
 ```
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
 ```{r duration_bar,  echo = FALSE}
-p_duration <- ggplot(data = duration, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+duration_trial <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr)]
+
+p_duration <- ggplot(data = duration_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   labs(y = "Normalised Duration", subtitle = "Duration: Non-Expression")
@@
 ```{r outlier, echo = FALSE}
-excluded_duration <- duration[Mean < duration_all$Mean+3*duration_all$SD & Mean > duration_all$Mean-3*duration_all$SD]
+excluded_duration <- duration_trial[Mean < duration_all$Mean+3*duration_all$SD & Mean > duration_all$Mean-3*duration_all$SD]
 print(excluded_duration)
 ```
-- The number of excluded performance: `r duration_all$N-nrow(excluded_duration)`
+- The number of excluded performance: `r nrow(duration_trial)-nrow(excluded_duration)`
 
@@
 # for each participant
-cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
+cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr)]
 print(cv)
@@
 ```{r cv_box, echo = FALSE}
-p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+cv_trial <- dt_duration[EighthNote == "Yes", .(N = length(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)]
+
+p_cv_box <- ggplot(cv_trial, aes(x = SubNr, y = CV)) +
   geom_boxplot() +
@@
 ```{r cv_outlier, echo = FALSE}
-excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
+excluded_cv <- cv_trial[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD]
 print(excluded_cv)

### 2021-07-05 material/expstimuli/analysis/stim_n/kv.Rmd
bar plots >> box plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/stim_n/kv.Rmd
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
 ```{r kv_bar,  echo = FALSE}
-p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+kv_trial <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr)]
+
+p_kv <- ggplot(data = kv_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   labs(y = "Key Velocity", subtitle = "KV: Non-Expression")
@@
 ```{r outlier, echo = FALSE}
-excluded_kv <- kv[Mean < kv_all$Mean+3*kv_all$SD & Mean > kv_all$Mean-3*kv_all$SD]
+excluded_kv <- kv_trial[Mean < kv_all$Mean+3*kv_all$SD & Mean > kv_all$Mean-3*kv_all$SD]
 print(excluded_kv)
 ```
-- The number of excluded performance: `r kv_all$N-nrow(excluded_kv)`
+- The number of excluded performance: `r nrow(kv_trial)-nrow(excluded_kv)`
 
@@
 # for each participant
-cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)]
+cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr)]
 cv
@@
 
+#### Box plot
 ```{r cv_box, echo = FALSE}
-p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+cv_trial <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)]
+
+p_cv_box <- ggplot(cv_trial, aes(x = SubNr, y = CV)) +
   geom_boxplot() +

### 2021-07-05 material/expstimuli/analysis/expression/stim_a/duration.Rmd
bar plots >> box plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ***Duration values from this analysis were used.***
+
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
 
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
 ```{r duration_bar,  echo = FALSE}
-p_duration <- ggplot(data = duration, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+duration_trial <- dt_duration[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr, Subcomponent)]
+
+p_duration <- ggplot(data = duration_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   facet_grid(Subcomponent ~ .) +

### 2021-07-05 material/expstimuli/analysis/expression/stim_d/duration.Rmd
bar plots >> box plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
 ```{r duration_bar,  echo = FALSE}
-p_duration <- ggplot(data = duration, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+duration_trial <- dt_duration[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr, Subcomponent)]
+
+p_duration <- ggplot(data = duration_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   facet_grid(Subcomponent ~ .) +
@@
 ```
+
+# Included performances for averaging
+## Legato
+```{r outlier_leg, echo = FALSE}
+# for legato
+duration_trial_leg <- dt_duration[Subcomponent == "Legato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr)]
+duration_trial_leg
+```
+
+## Staccato
+```{r outlier_sta, echo = FALSE}
+# for staccato
+duration_trial_sta <- dt_duration[Subcomponent == "Staccato", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, TrialNr, Subcomponent)]
+duration_trial_sta
+```

### 2021-07-06 material/expstimuli/analysis/stim_n/duration.Rmd
add comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+#### Box plot
 ```{r cv_box, echo = FALSE}

### 2021-09-16 material/expstimuli/audio/analysis/expression/stim_a/ioi_onset.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/expression/stim_a/ioi_onset.Rmd
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 - A dashed line represents the ideal tempo
@@
 ```{r ioi_bar,  echo = FALSE}
-p_ioi <- ggplot(data = ioi, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+ioi_trial <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, TrialNr)]
+
+p_ioi <- ggplot(data = ioi_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
-  labs(y = "Normalised IOI", subtitle = "IOI: Performing-Articulation")
+  labs(y = "Normalised IOI", subtitle = "IOI: Non-Expression")
 p_ioi
@@
 # for each participant
-cv <- dt_ioi[, .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+cv <- dt_ioi[, .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr)]
 cv
@@
 
+#### Box plot
 ```{r cv_box, echo = FALSE}
-p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+cv_trial <- dt_ioi[, .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+
+p_cv_box <- ggplot(cv_trial, aes(x = SubNr, y = CV)) +
   geom_boxplot() +

### 2021-09-16 material/expstimuli/audio/analysis/expression/stim_a/kv.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/expression/stim_a/kv.Rmd
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
 ```{r kv_bar,  echo = FALSE}
-p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+kv_trial <- dt_kv[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Subcomponent)]
+
+p_kv <- ggplot(data = kv_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   facet_grid(Subcomponent ~ .) +

### 2021-09-16 material/expstimuli/audio/analysis/expression/stim_d/ioi_onset.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/expression/stim_d/ioi_onset.Rmd
@@
 ```{r ioi_bar,  echo = FALSE}
-p_ioi <- ggplot(data = ioi, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+ioi_trial <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr, TrialNr)]
+
+p_ioi <- ggplot(data = ioi_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
@@
 ```{r cv_box, echo = FALSE}
-p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+cv_trial <- dt_ioi[, .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+
+p_cv_box <- ggplot(cv_trial, aes(x = SubNr, y = CV)) +
   geom_boxplot() +

### 2021-09-16 material/expstimuli/audio/analysis/expression/stim_d/kv.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/expression/stim_d/kv.Rmd
@@
 ***KV values from this analysis were used.***
+
+- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
 
@@
 ### Plot
-#### Bar chart
-- Error bars show SD of the mean
+#### Box plot
 
 ```{r kv_bar,  echo = FALSE}
-p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
-                width = .2, position = position_dodge(.9)) +
+kv_trial <- dt_kv[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Subcomponent)]
+
+p_kv <- ggplot(data = kv_trial, aes(x = SubNr, y = Mean)) +
+  geom_boxplot() +
+  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) +
   facet_grid(Subcomponent ~ .) +

### 2021-09-16 material/expstimuli/audio/analysis/stim_n/ioi_offset.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-01 material/expstimuli/analysis/stim_n/ioi_offset.Rmd
@@
 ```{r outlier, echo = FALSE}
-excluded_ioi <- ioi[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
+excluded_ioi <- ioi_trial[Mean < ioi_all$Mean+3*ioi_all$SD & Mean > ioi_all$Mean-3*ioi_all$SD]
 print(excluded_ioi)
 ```
-- The number of excluded performance: `r ioi_all$N-nrow(excluded_ioi)`
+- The number of excluded performance: `r nrow(ioi_trial)-nrow(excluded_ioi)`
 
@@
 # for each participant
-cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+cv <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr)]
 cv
@@
 
+#### Box plot
 ```{r cv_box, echo = FALSE}
-p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) +
+cv_trial <- dt_ioi[EighthNote == "Yes", .(N = length(normIOI), CV = sd(normIOI)/mean(normIOI)) , by = .(SubNr, TrialNr)]
+
+p_cv_box <- ggplot(cv_trial, aes(x = SubNr, y = CV)) +
   geom_boxplot() +

### 2021-09-16 material/expstimuli/audio/analysis/stim_n/ioi_onset.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-09-16 material/expstimuli/audio/analysis/stim_n/ioi_offset.Rmd
@@
 ---
-title: 'IOI-Offset Analysis for Non-Expressive performances'
+title: 'IOI-Onset Analysis for Non-Expressive performances'
 output:
@@
 
-Description: This is a rough overview of non-expressive performance (IOI-Offset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
+Description: This is a rough overview of non-expressive performance (IOI-Onset) from [music-teaching (experiment2)](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2).
 
@@
 ```{r extract, include = FALSE}
-dt_ioi <- fread("./preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
+dt_ioi <- fread("./preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
 
@@
 
-# Valid performances (exclude outliers)
+# Valid performances (exclude IOIs outliers)
 - Exclude mean IOIs outside 3SD
@@
 ```
-
 ## 2. All - group
@@
 
-# Valid performances (exclude outliers)
+# Valid performances (exclude CVs outliers)
 - Exclude mean CVs outside 3SD
@@
 ```{r export, include = FALSE}
-fwrite(excluded_cv, "./offset_valid.csv")
+fwrite(excluded_cv, "./onset_valid.csv")
 ```

### 2021-09-16 material/expstimuli/audio/selected/IOI/ioi_onset.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'IOI-Onset Selection'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: In order to select 4 performances to determine a tempo of each performance, we look at previously generated 16 instances and randomly selected from them.

Original stimuli: `material/expstimuli/original/averaging/low/1596207520-310720/dt_ioi_instance.txt`

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
dt_ioi <- fread("../../original/averaging/low/1596207520-310720/dt_ioi_instance.txt")
dt_ioi$IOI <- dt_ioi$Mean*300
```

# IOIs plot
```{r plot-ioi, echo = FALSE, fig.height = 6}
ggplot(data = dt_ioi, aes(x = RowNr, y = IOI)) +
  geom_line() +
  geom_point() +
  facet_grid(Instance ~ .) +
  labs(x = "RowNr", y = "IOIs (100bpm)", subtitle = "IOI: 16 instances") +
  scale_x_continuous(breaks=seq(1,72,1))
```
- Looks like 13 may have a delay at Row 24?

```{r playback, include = FALSE}
# create a folder to store stimuli
foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
dir.create(foldername)

# 1. determine onsets/offsets
# onsets
dt_playback_onset <- dt_ioi[, c("Instance", "RowNr")]
dt_playback_onset$Key_OnOff <- 1
for (i in 1:nrow(dt_playback_onset)){
  if (dt_playback_onset$RowNr[i] == 1){
    dt_playback_onset$TimeStamp[i] <- 0
  } else {
    dt_playback_onset$TimeStamp[i] <- dt_playback_onset$TimeStamp[i-1]+dt_ioi$Mean[i]
  }
}
# Tempo 100 bpm (IOI - 300ms (8th notes))
dt_playback_onset$TimeStamp <- round(dt_playback_onset$TimeStamp*300)

# offsets
dt_playback_offset <- dt_ioi[, c("Instance", "RowNr")]
dt_playback_offset$Key_OnOff <- 0
# Tempo 100 bpm (IOI - 300ms (8th notes))
dt_playback_offset$TimeStamp <- round(dt_playback_onset$TimeStamp+0.5*300)

# kv
dt_playback_onset$Velocity <- 80
dt_playback_offset$Velocity <- 64

# create txt for each instance
for (i in 1:16){
  onset <- dt_playback_onset[Instance == i]
  offset <- dt_playback_offset[Instance == i]
  onset$Pitch <- 60
  offset$Pitch <- 60
  instance <- rbind(onset, offset)
  # sort by TimeStamp
  instance <- instance[order(TimeStamp)]

  # export txt
  filename = paste(foldername, i, "_instance.txt", sep = "")
  fwrite(instance, filename)
}
```

# Export as mid files
```{bash, echo = TRUE}
# python
source /Users/atsukotominaga/Pyenv/bin/activate
python mid_export.py
```
- Listened to each performance (mid files between 1-16) and `ioi_13.mid` was excluded because there was one tiny delay in a performance.
- We select four of these performances (15 mid files) to determine a tempo of each stimulus.

# Selected 4 performances (IOIs)
```{r, echo = FALSE}
select_ioi <- dt_ioi[Instance != 13, .N, .(Instance)]
select_ioi$Sample <- sample(c(1:nrow(select_ioi)), replace = FALSE)
select_ioi[Sample < 5]

# export
dir.create("./tempo/")
for (i in 1:4){
  current <- dt_ioi[Instance == select_ioi[Sample == i]$Instance]
  filename = paste("./tempo/", i, ".txt", sep = "")
  fwrite(current, filename)
}
```

### 2021-09-16 material/expstimuli/audio/tweaked/analysis.Rmd
moved to audio file
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-16 material/expstimuli/tweaked/analysis.Rmd
@@
 tweak_data <- dt_kot[grepl("dyn_only", MidFile) | grepl("low", MidFile) & SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]
-tweak_data$Diff <- abs(tweak_data$Mean-kot_grandMean)
+tweak_data$Diff <- round(abs(tweak_data$Mean-kot_grandMean), 0)
 tweak_data
@@
 
-## Histogram (Fixed KOT)
+d KOT)
 ```{r kot-distribution-fixed, echo = FALSE}
@@
 ```{r export, include = FALSE}
+# remove unnecessary columns
+data_onset$IOI <- NULL
+data_onset$KOT <- NULL
+data_onset$Diff <- NULL
 # fix time for onset and offset data
-data_offset$timeFixed <- data_onset$time
+data_onset$timeFixed <- data_onset$time
 data_offset$timeFixed <- data_offset$time
@@
 
+# export each instance as txt file
 for (stim in unique(data_onset$MidFile)){
@@
    instance <- rbind(onset, offset)
-   # round timeFixed
-   instance$timeFixed <- round(instance$timeFixed, 0)
    # Note_On/Off >> 1 or 0

### 2021-10-10 analysis/filtering.Rmd
add basic analysis scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Preprocessing: filtering (experiment)"
output: html_notebook
---

```{r setup, include = FALSE}
# set working directory
if (!require("here")) {install.packages("here"); require("here")}
here::i_am("filtering.Rmd")

# create a folder if not exists
if (!file.exists(here("filtered"))){
  dir.create(here("filtered"))
}

# read functions
source(here("function.R"))

# read a text file for ideal performance
dt_ideal <- read.table(here("ideal.txt"))
colnames(dt_ideal) <- "Pitch"
dt_ideal$RowNr <- 1:nrow(dt_ideal)
setcolorder(dt_ideal, c(2, 1))

# create a list of data file names
lf <- list.files(here("raw_data"), pattern = "txt")

# create raw_data - merge all data files into one
raw_data <- data.table()
for (i in 1:length(lf)){
  data_i <- fread(file.path(here("raw_data"), lf[i]), header = F, sep = " ", dec = ".")
  raw_data <- rbind(raw_data, data_i)
}

# add column namesls
colnames(raw_data) <- c("NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "SubNr", "TrialNr", "Stimuli", "Session")

# clean raw_data
raw_data$NoteNr <- as.numeric(gsub(",", "", raw_data$NoteNr))
raw_data$Session <- gsub(";", "", raw_data$Session)

# sort by SubNr, TrialNr
raw_data <- raw_data[order(raw_data$SubNr, raw_data$TrialNr),]

# raw_data without metronome
dt_all <- raw_data[Pitch != 31 & Pitch != 34]

# onset and offset for experimental session
dt_onset <- dt_all[Key_OnOff == 1 & Session == "experiment"]
dt_offset <- dt_all[Key_OnOff == 0 & Session == "experiment"]
```

# Onset
## Detect erroneous trials
```{r, echo = FALSE}
dt_error_onset <- checker(dt_onset, dt_ideal)

# mark imperfect performances
dt_onset$Error <- 0
for (error in 1:nrow(dt_error_onset)){
  dt_onset[SubNr == dt_error_onset$SubNr[error] & TrialNr == dt_error_onset$TrialNr[error]]$Error <- 1
}

dt_correct_onset_1 <- dt_onset[Error == 0]
dt_correct_onset_1$RowNr <- rep(1:72, nrow(dt_correct_onset_1)/72)
setcolorder(dt_correct_onset_1, c(12, 1:11))

# export csv
fwrite(dt_correct_onset_1, file = here("filtered", "dt_correct_onset_1.txt"))

# print the result
dt_error_onset
```

# Offset

## Detect erroneous trials
```{r, echo = FALSE}
dt_error_offset <- checker(dt_offset, dt_ideal)

# mark imperfect performances
dt_offset$Error <- 0
for (error in 1:nrow(dt_error_offset)){
  dt_offset[SubNr == dt_error_offset$SubNr[error] & TrialNr == dt_error_offset$TrialNr[error]]$Error <- 1
}

dt_correct_offset_1 <- dt_offset[Error == 0]
dt_correct_offset_1$RowNr <- rep(1:72, nrow(dt_correct_offset_1)/72)
setcolorder(dt_correct_offset_1, c(12, 1:11))

# export csv
fwrite(dt_correct_offset_1, file = here("filtered", "dt_correct_offset_1.txt"))

# print the result
dt_error_offset
```

# Checking

### 2021-10-10 analysis/function.R

@@ -1,0 +1,106 @@
+# install and load required packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+if (!require("editData")) {install.packages("editData"); require("editData")}
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+
+# ggplot settings
+theme_set(theme_classic())
+
+### check whether a trial contains pitch errors
+# data: data of the current trial
+# ideal: dt_ideal
+checker <- function(data, ideal){
+  dt_errors <- data.table() # return data of erroneous trials - SubNr/BlockNr/TrialNr/Reason
+  for (subject in unique(data$SubNr)){
+    for (trial in c(1:16)){
+      current <- data[SubNr == subject & TrialNr == trial]
+      if (nrow(current) != 0){ # current data not empty
+        if (nrow(current) != nrow(ideal)){ # extra/missing note
+          if (nrow(current) > nrow(ideal)){
+            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Extra Notes")) 
+          } else if (nrow(current) < nrow(ideal)){
+            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Missing Notes")) 
+          }
+        } else if (nrow(current) == nrow(ideal)){ # substituted note
+          for (note in 1:nrow(ideal)){
+            if (current$Pitch[note] != ideal$Pitch[note]){
+              dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Substituted Notes - RowNr ", as.character(note), sep = "")))
+              break
+            }
+          }
+        }
+      } else if (nrow(current) == 0){ # current data empty
+        dt_errors <- rbind(dt_errors, data.table(subject, block, trial, paste("Missing Trial")))
+      }
+    }
+  }
+  colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
+  return(dt_errors)
+}
+
+### edit data
+# data: data of the current trial
+# ideal: dt_ideal
+edit <- function(data, ideal){
+  data$RowNr <- c(1:nrow(data))
+  length_diff <- abs(nrow(data) - nrow(ideal))
+  data$Ideal <- c(ideal$Pitch, rep(NA, length_diff))
+  data$Diff <- "NA"
+  data[Pitch != Ideal]$Diff <- "DIFFERENT"
[…]

### 2021-12-01 analysis/primary.Rmd
upload primary analysis for ioi, kv, kv-diff
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Primary analysis'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of primary analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline
- IOIs
- KOT
- KV
- KV-Diff

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r file, include = FALSE}
filename = "filtered/dt_correct_onset_1.txt"
```

```{r extract, include = FALSE}
data <- fread(filename, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
data$SubNr <- as.factor(data$SubNr)

# assign stimuli categories
data$Stimuli <- as.numeric(gsub(".mid", "", data$Stimuli))
data$Category <- "NA"
data[Stimuli < 5]$Category <- "art_only"
data[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
data[Stimuli > 8 & Stimuli < 13]$Category <- "high"
data[Stimuli > 12]$Category <- "low"

# assign articulation (yes/no) & dynamics (yes/no)
data$Articulation <- "NA"
data$Dynamics <- "NA"
data[Category == "art_only"]$Articulation <- "Present"
data[Category == "art_only"]$Dynamics <- "Absent"
data[Category == "dyn_only"]$Articulation <- "Absent"
data[Category == "dyn_only"]$Dynamics <- "Present"
data[Category == "high"]$Articulation <- "Present"
data[Category == "high"]$Dynamics <- "Present"
data[Category == "low"]$Articulation <- "Absent"
data[Category == "low"]$Dynamics <- "Absent"

# IOIs
dt_ioi <- data
dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))

# remove the first note
dt_ioi <- dt_ioi[RowNr != 1]

# assign Interval
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# define intervals
ls_l_f <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_s_p <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
change_1 <- c(5, 17, 47) # legato to staccato / forte to piano
change_2 <- c(8, 20, 39) # staccato to legato / piano to forte

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_ioi[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_ioi[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_ioi[Interval == number]$Subcomponent <- "LFtoPS"
}
for (number in change_2){
  dt_ioi[Interval == number]$Subcomponent <- "SPtoLF"
}

# KOT

# KV
dt_vel <- data
dt_vel$Diff <- diff(c(0, dt_vel$Velocity))

# remove the first note
dt_vel_diff <- dt_vel[RowNr != 1]
dt_vel$Diff <- NULL # remove Diff from dt_vel

# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# for each note (velocity)
ls_l_f_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_s_p_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# assign Subcomponents
# for each note
dt_vel$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f_2)){
  for (note in 1:length(ls_l_f_2[[phrase]])){
    dt_vel[RowNr == ls_l_f_2[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p_2)){
  for (note in 1:length(ls_s_p_2[[phrase]])){
    dt_vel[RowNr == ls_s_p_2[[phrase]][note]]$Subcomponent <- "S-P"
  }
}

dt_vel_diff$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_vel_diff[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_vel_diff[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_vel_diff[Interval == number]$Subcomponent <- "LFtoSP"
}
for (number in change_2){
  dt_vel_diff[Interval == number]$Subcomponent <- "SPtoLF"
}
```

# IOI 
## 1. Individual

```{r ioi, echo = FALSE}
# for each individual
ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Stimuli, Articulation, Dynamics)]
ioi <- ioi[order(SubNr, Stimuli, Articulation, Dynamics)]
ioi
```

### Plot
#### Boxplot

```{r ioi-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

## 2. All - group

```{r ioi-all, echo = FALSE}
# group mean
ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
ioi_all <- ioi_all[order(Articulation, Dynamics)]
ioi_all
```

### Plot
#### Box plot

```{r ioi-all-plot, echo = FALSE}
ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

# Dynamics
## 1. Individual
```{r vel, echo = FALSE}
# for each individual
vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel <- vel[order(SubNr, Stimuli, Articulation, Dynamics)]
vel
```

### Plot
#### Boxplot

```{r vel-plot,  echo = FALSE, fig.width = 4.5}
ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
```

## 2. All - group
```{r vel-all, echo = FALSE}
# group mean
vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
vel_all <- vel_all[order(Articulation, Dynamics, Subcomponent)]
vel_all
```

### Plot
#### Boxplot

```{r vel-all-plot,  echo = FALSE}
ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

# KV-Diff
## 1. Individual

```{r vel-diff, echo = FALSE}
# for each individual
vel_diff <- dt_vel_diff[Subcomponent == "LFtoSP" | Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel_diff <- vel_diff[order(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel_diff
```

### Plot
#### Boxplot

```{r vel-diff-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (LFtoSP)")
ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (SPtoLF)")
```

## 2. All - group
```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
vel_diff_all <- vel_diff_all[order(Articulation, Dynamics, Subcomponent)]
vel_diff_all
```

### Plot
#### Boxplot

```{r vel-diff-all-plot,  echo = FALSE}
ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (LF to SP)")
ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (SP to LF)")
```

### 2021-12-01 analysis/filtering.Rmd
upload filtering
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ## Detect erroneous trials
-```{r, echo = FALSE}
+```{r onset, echo = FALSE}
 dt_error_onset <- checker(dt_onset, dt_ideal)
@@
 ## Detect erroneous trials
-```{r, echo = FALSE}
+```{r offset, echo = FALSE}
 dt_error_offset <- checker(dt_offset, dt_ideal)

### 2021-12-01 analysis/function.R

@@ -10,7 +10,7 @@ theme_set(theme_classic())
 # data: data of the current trial
 # ideal: dt_ideal
 checker <- function(data, ideal){
-  dt_errors <- data.table() # return data of erroneous trials - SubNr/BlockNr/TrialNr/Reason
+  dt_errors <- data.table() # return data of erroneous trials - SubNr/TrialNr/Reason
   for (subject in unique(data$SubNr)){
     for (trial in c(1:16)){
       current <- data[SubNr == subject & TrialNr == trial]
@@ -30,7 +30,7 @@ checker <- function(data, ideal){
           }
         }
       } else if (nrow(current) == 0){ # current data empty
-        dt_errors <- rbind(dt_errors, data.table(subject, block, trial, paste("Missing Trial")))
+        dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Missing Trial")))
       }
     }
   }
@@ -48,7 +48,7 @@ edit <- function(data, ideal){
   data$Diff <- "NA"
   data[Pitch != Ideal]$Diff <- "DIFFERENT"
   # sort the order of columns
-  setcolorder(data, c("RowNr", "NoteNr", "TimeStamp", "Pitch", "Ideal", "Diff", "Velocity", "Key_OnOff", "Device", "SubNr", "BlockNr", "TrialNr", "Skill", "Condition", "Image", "Error"))
+  setcolorder(data, c("RowNr", "NoteNr", "TimeStamp", "Pitch", "Ideal", "Diff", "Velocity", "Key_OnOff", "Device", "SubNr", "TrialNr", "Stimuli", "Session", "Error"))
   graph <- ggplot() +
     geom_line(data = data, aes(x = RowNr, y = Pitch), colour = "#F8766D") +
     geom_line(data = data, aes(x = RowNr, y = Ideal), colour = "#00BFC4") +
@@ -56,7 +56,7 @@ edit <- function(data, ideal){
     geom_point(data = data, aes(x = RowNr, y = Ideal), colour = "#00BFC4") +
     scale_x_continuous("RowNr", data$RowNr) +
     coord_fixed(ratio = 1/4) +
-    labs(title = sprintf("SubNr: %s, BlockNr: %s, TrialNr: %s", unique(data$SubNr), unique(data$BlockNr), unique(data$TrialNr)), y = "Pitch")
[…]

### 2021-12-02 analysis/primary.Rmd
add kot analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r file, include = FALSE}
-filename = "filtered/dt_correct_onset_1.txt"
+onset = "filtered/dt_correct_onset_1.txt"
+offset = "filtered/dt_correct_offset_1.txt"
 ```
 
-```{r extract, include = FALSE}
-data <- fread(filename, header = T, sep = ",", dec = ".") # read a trimmed csv
+```{r prep, include = FALSE}
+data_on <- fread(onset, header = T, sep = ",", dec = ".") # read a trimmed csv
+data_off <- fread(offset, header = T, sep = ",", dec = ".") # read a trimmed csv
 
 # SubNr as a factor
-data$SubNr <- as.factor(data$SubNr)
+data_on$SubNr <- as.factor(data_on$SubNr)
 
 # assign stimuli categories
-data$Stimuli <- as.numeric(gsub(".mid", "", data$Stimuli))
-data$Category <- "NA"
-data[Stimuli < 5]$Category <- "art_only"
-data[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
-data[Stimuli > 8 & Stimuli < 13]$Category <- "high"
-data[Stimuli > 12]$Category <- "low"
+data_on$Stimuli <- as.numeric(gsub(".mid", "", data_on$Stimuli))
+data_on$Category <- "NA"
+data_on[Stimuli < 5]$Category <- "art_only"
+data_on[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
+data_on[Stimuli > 8 & Stimuli < 13]$Category <- "high"
+data_on[Stimuli > 12]$Category <- "low"
 
 # assign articulation (yes/no) & dynamics (yes/no)
-data$Articulation <- "NA"
-data$Dynamics <- "NA"
-data[Category == "art_only"]$Articulation <- "Present"
-data[Category == "art_only"]$Dynamics <- "Absent"
-data[Category == "dyn_only"]$Articulation <- "Absent"
-data[Category == "dyn_only"]$Dynamics <- "Present"
-data[Category == "high"]$Articulation <- "Present"
-data[Category == "high"]$Dynamics <- "Present"
-data[Category == "low"]$Articulation <- "Absent"
-data[Category == "low"]$Dynamics <- "Absent"
+data_on$Articulation <- "NA"
+data_on$Dynamics <- "NA"
+data_on[Category == "art_only"]$Articulation <- "Present"
+data_on[Category == "art_only"]$Dynamics <- "Absent"
+data_on[Category == "dyn_only"]$Articulation <- "Absent"
+data_on[Category == "dyn_only"]$Dynamics <- "Present"
+data_on[Category == "high"]$Articulation <- "Present"
+data_on[Category == "high"]$Dynamics <- "Present"
+data_on[Category == "low"]$Articulation <- "Absent"
+data_on[Category == "low"]$Dynamics <- "Absent"
 
 # IOIs
-dt_ioi <- data
+dt_ioi <- data_on
 dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))
@@
 
-# define intervals
+# define Subcomponents
+# for each interval
 ls_l_f <- list(c(1:4), c(9:16), c(21:24), c(40:46))
@@
 # KOT
+onset_trials <- data_on[, .(N = .N), by = .(SubNr, TrialNr)]
+offset_trials <- data_off[, .(N = .N), by = .(SubNr, TrialNr)]
+valid_trials <- rbind(onset_trials, offset_trials)
+valid_trials$Duplicate <- duplicated(valid_trials)
+valid_trials_included <- valid_trials[Duplicate == TRUE]
 
+dt_kot_on <- data.table()
+for (row in 1:nrow(valid_trials_included)){
+  current <- data_on[SubNr == valid_trials_included$SubNr[row] & TrialNr == valid_trials_included$TrialNr[row]]
+  dt_kot_on <- rbind(dt_kot_on, current)
+}
+
+dt_kot_off <- data.table()
+for (row in 1:nrow(valid_trials_included)){
+  current <- data_off[SubNr == valid_trials_included$SubNr[row] & TrialNr == valid_trials_included$TrialNr[row]]
+  dt_kot_off <- rbind(dt_kot_off, current)
+}
+
+# sort by SubNr, TrialNr and RowNr
+dt_kot_on <- dt_kot_on[order(SubNr, TrialNr, RowNr)]
+dt_kot_off <- dt_kot_off[order(SubNr, TrialNr, RowNr)]
+
+# Offset 1 - Onset 2
+dt_kot_on$KOT <- NA
+for (row in 1:nrow(dt_kot_on)){
+  if (row < nrow(dt_kot_on)){
+    dt_kot_on$KOT[row+1] <- dt_kot_off$TimeStamp[row] - dt_kot_on$TimeStamp[row+1] # offset(n) - onset(n+1)
+  }
+}
+
+# remove the first note
+dt_kot_on <- dt_kot_on[RowNr != 1]
+
+# assign Interval
+dt_kot_on$Interval <- rep(1:71, nrow(dt_kot_on)/71)
+
+# assign Subcomponents
+dt_kot_on$Subcomponent <- "NA"
+# legato & forte
+for (phrase in 1:length(ls_l_f)){
+  for (note in 1:length(ls_l_f[[phrase]])){
+    dt_kot_on[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
+  }
+}
+# staccato & piano
+for (phrase in 1:length(ls_s_p)){
+  for (note in 1:length(ls_s_p[[phrase]])){
+    dt_kot_on[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
+  }
+}
+for (number in change_1){
+  dt_kot_on[Interval == number]$Subcomponent <- "LFtoPS"
+}
+for (number in change_2){
+  dt_kot_on[Interval == number]$Subcomponent <- "SPtoLF"
+}
 # KV
-dt_vel <- data
+dt_vel <- data_on
 dt_vel$Diff <- diff(c(0, dt_vel$Velocity))
@@
 ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Stimuli, Articulation, Dynamics)]
-ioi <- ioi[order(SubNr, Stimuli, Articulation, Dynamics)]
-ioi
+ioi_ind <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)]
+ioi_ind
 ```
@@
 # group mean
-ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
-ioi_all <- ioi_all[order(Articulation, Dynamics)]
+ioi_all <- ioi_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
 ioi_all
@@
 ```{r ioi-all-plot, echo = FALSE}
-ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+ggboxplot(ioi_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+```
+
+# Articulation
+## 1. Interval
+```{r kot, echo = FALSE}
+# for each interval
+kot <- dt_kot_on[Subcomponent == "L-F" | Subcomponent == "S-P", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
+kot_ind <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Subcomponent)]
+kot_ind
+```
+
+### Plot
+#### Boxplot
+
+```{r kot-plot,  echo = FALSE, fig.width = 4.5}
+ggboxplot(kot[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+ggboxplot(kot[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
+```
+
+## 2. All - group
+```{r kot-all, echo = FALSE}
+# group mean
+kot_all <- kot_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
+kot_all
+```
+
+### Plot
+#### Boxplot
+
+```{r kot-all-plot, echo = FALSE}
+ggboxplot(kot_ind[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+ggboxplot(kot_ind[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
 ```
@@
 ```{r vel, echo = FALSE}
-# for each individual
+# for each note
 vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
-vel <- vel[order(SubNr, Stimuli, Articulation, Dynamics)]
-vel
+vel_ind <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Subcomponent)]
+vel_ind
 ```
@@
 # group mean
-vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
-vel_all <- vel_all[order(Articulation, Dynamics, Subcomponent)]
+vel_all <- vel_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
 vel_all
@@
 ```{r vel-all-plot,  echo = FALSE}
-ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
-ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
+ggboxplot(vel_ind[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
+ggboxplot(vel_ind[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
 ```
@@
 vel_diff <- dt_vel_diff[Subcomponent == "LFtoSP" | Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
-vel_diff <- vel_diff[order(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
-vel_diff
+vel_diff_ind <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Subcomponent)]
+vel_diff_ind
 ```
@@
 # group mean
-vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
-vel_diff_all <- vel_diff_all[order(Articulation, Dynamics, Subcomponent)]
+vel_diff_all <- vel_diff_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
 vel_diff_all
@@
 ```{r vel-diff-all-plot,  echo = FALSE}
-ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (LF to SP)")
-ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (SP to LF)")
+ggboxplot(vel_diff_ind[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (LF to SP)")
+ggboxplot(vel_diff_ind[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (SP to LF)")
 ```

### 2022-08-11 analysis/filtering.Rmd
add funcsions to remove extra notes
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-10-10 analysis/filtering.Rmd
@@
 ---
-title: "Preprocessing: filtering (experiment)"
+title: "Preprocessing: filtering"
 output: html_notebook
@@
 ```{r setup, include = FALSE}
+# set chunk option
+knitr::opts_chunk$set(echo = FALSE)
+
 # set working directory
@@
 
-# Onset
-## Detect erroneous trials
-```{r, echo = FALSE}
+# Detect erroneous trials
+
+1) Onset
+
+```{r onset}
 dt_error_onset <- checker(dt_onset, dt_ideal)
@@
 
-# Offset
+2) Offset
 
-## Detect erroneous trials
-```{r, echo = FALSE}
+```{r offset, echo = FALSE}
 dt_error_offset <- checker(dt_offset, dt_ideal)
@@
 
-# Checking
+# Manual pitch removal
+
+## Extra notes
+1) Onset
+
+```{r extra, include = FALSE}
+dt_error_onset$CorrectionNr <- NA
+
+error_extra_onset <- dt_error_onset[Reason == "Extra Notes"]
+dt_correct_onset_2 <- data.table()
+for (row in 1:nrow(error_extra_onset)){
+  current <- dt_onset[SubNr == error_extra_onset$SubNr[row] & TrialNr == error_extra_onset$TrialNr[row]]
+  decision = 2 
+  correction = 0 # # of correction for reporting stats
+  while (decision == 2){
+    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
+    print("----- First check -----")
+    current <- edit(current, dt_ideal)
+    print("----- Correction check -----")
+    edit(current, dt_ideal)
+    decision <- menu(c("y", "n", "other"), title = "Save the current data? (to continue, enter 'n')")
+    if (decision == 1){
+      correction = correction + 1
+      error_extra_onset$CorrectionNr[row] <- correction
+      dt_correct_onset_2 <- rbind(dt_correct_onset_2, current[, -c(5:6)])
+    } else if (decision == 3){
+      error_extra_onset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
+    } else if (decision == 2){
+      correction = correction + 1
+      print("----- Continue correction -----")
+    }
+  }
+}
+
+# export csv
+fwrite(dt_correct_onset_2, file = here("filtered", "dt_correct_onset_2.txt"))
+```
+
+2) Offset
+
+```{r extra, include = FALSE}
+dt_error_offset$CorrectionNr <- NA
+
+error_extra_offset <- dt_error_offset[Reason == "Extra Notes"]
+dt_correct_offset_2 <- data.table()
+for (row in 1:nrow(error_extra_offset)){
+  current <- dt_offset[SubNr == error_extra_offset$SubNr[row] & TrialNr == error_extra_offset$TrialNr[row]]
+  decision = 2 
+  correction = 0 # # of correction for reporting stats
+  while (decision == 2){
+    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
+    print("----- First check -----")
+    current <- edit(current, dt_ideal)
+    print("----- Correction check -----")
+    edit(current, dt_ideal)
+    decision <- menu(c("y", "n", "other"), title = "Save the current data? (to continue, enter 'n')")
+    if (decision == 1){
+      correction = correction + 1
+      error_extra_offset$CorrectionNr[row] <- correction
+      dt_correct_offset_2 <- rbind(dt_correct_offset_2, current[, -c(5:6)])
+    } else if (decision == 3){
+      error_extra_offset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
+    } else if (decision == 2){
+      correction = correction + 1
+      print("----- Continue correction -----")
+    }
+  }
+}
+
+# export csv
+fwrite(dt_correct_offset_2, file = here("filtered", "dt_correct_offset_2.txt"))
+```

### 2022-08-11 analysis/function.R

@@ -49,16 +49,7 @@ edit <- function(data, ideal){
   data[Pitch != Ideal]$Diff <- "DIFFERENT"
   # sort the order of columns
   setcolorder(data, c("RowNr", "NoteNr", "TimeStamp", "Pitch", "Ideal", "Diff", "Velocity", "Key_OnOff", "Device", "SubNr", "TrialNr", "Stimuli", "Session", "Error"))
-  graph <- ggplot() +
-    geom_line(data = data, aes(x = RowNr, y = Pitch), colour = "#F8766D") +
-    geom_line(data = data, aes(x = RowNr, y = Ideal), colour = "#00BFC4") +
-    geom_point(data = data, aes(x = RowNr, y = Pitch), colour = "#F8766D") +
-    geom_point(data = data, aes(x = RowNr, y = Ideal), colour = "#00BFC4") +
-    scale_x_continuous("RowNr", data$RowNr) +
-    coord_fixed(ratio = 1/4) +
-    labs(title = sprintf("SubNr: %s, TrialNr: %s", unique(data$SubNr), unique(data$TrialNr)), y = "Pitch")
-  print(graph)
-  corrected <- editData(data, viewer = "pane")
+  corrected <- editData(data)
   return(data.table(corrected)) # convert to data.table
 }
 

### 2022-08-14 analysis/filtering.Rmd
add missing notes removal
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r extra, include = FALSE}
+```{r extra-onset, include = FALSE, eval = FALSE}
 dt_error_onset$CorrectionNr <- NA
@@
 fwrite(dt_correct_onset_2, file = here("filtered", "dt_correct_onset_2.txt"))
+fwrite(error_extra_onset, file = here("filtered", "error_extra_onset.txt"))
+```
+
+```{r}
+error_extra_onset <- fread(here("filtered/error_extra_onset.txt"))
+error_extra_onset
 ```
@@
 
-```{r extra, include = FALSE}
+```{r extra-offset, include = FALSE, eval = FALSE}
 dt_error_offset$CorrectionNr <- NA
@@
 fwrite(dt_correct_offset_2, file = here("filtered", "dt_correct_offset_2.txt"))
+fwrite(error_extra_offset, file = here("filtered", "error_extra_offset.txt"))
 ```
+
+```{r}
+error_extra_offset <- fread(here("filtered/error_extra_offset.txt"))
+error_extra_offset
+```
+
+## Missing notes
+1) Onset
+
+```{r missing-onset, include = FALSE, eval = FALSE}
+error_missing_onset <- dt_error_onset[Reason == "Missing Notes"]
+
+dt_correct_onset_3 <- data.table()
+for (row in 1:nrow(error_missing_onset)){
+  current <- dt_onset[SubNr == error_missing_onset$SubNr[row] & TrialNr == error_missing_onset$TrialNr[row]]
+  diff <- abs(nrow(dt_ideal) - nrow(current))
+  if (diff < 5){
+    # insert NA row
+    current <- insert_na(current, dt_ideal)
+    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
+    print("----- Correction check -----")
+    current <- edit(current, dt_ideal)
+    decision <- menu(c("y", "other"), title = "Save the current data?")
+    if (decision == 1){
+      error_missing_onset$CorrectionNr[row] <- diff
+      dt_correct_onset_3 <- rbind(dt_correct_onset_3, current[, -c(5:6)])
+    } else if (decision == 2){
+      error_missing_onset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
+    }
+  } else {
+    error_missing_onset$CorrectionNr[row] <- "Check individually"
+  }
+}
+
+# export csv
+fwrite(dt_correct_onset_3, file = here("filtered", "dt_correct_onset_3.txt"))
+fwrite(error_missing_onset, file = here("filtered", "error_missing_onset.txt"))
+```
+
+```{r}
+error_missing_onset <- fread(here("filtered/error_missing_onset.txt"))
+error_missing_onset
+```
+
+2) Offset
+
+```{r missing-offset, include = FALSE}
+error_missing_offset <- dt_error_offset[Reason == "Missing Notes"]
+
+dt_correct_offset_3 <- data.table()
+for (row in 1:nrow(error_missing_offset)){
+  current <- dt_offset[SubNr == error_missing_offset$SubNr[row] & TrialNr == error_missing_offset$TrialNr[row]]
+  diff <- abs(nrow(dt_ideal) - nrow(current))
+  if (diff < 5){
+    # insert NA row
+    current <- insert_na(current, dt_ideal)
+    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
+    print("----- Correction check -----")
+    current <- edit(current, dt_ideal)
+    decision <- menu(c("y", "other"), title = "Save the current data?")
+    if (decision == 1){
+      error_missing_offset$CorrectionNr[row] <- diff
+      dt_correct_offset_3 <- rbind(dt_correct_offset_3, current[, -c(5:6)])
+    } else if (decision == 2){
+      error_missing_offset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
+    }
+  } else {
+    error_missing_offset$CorrectionNr[row] <- "Check individually"
+  }
+}
+
+# export csv
+fwrite(dt_correct_offset_3, file = here("filtered", "dt_correct_offset_3.txt"))
+fwrite(error_missing_offset, file = here("filtered", "error_missing_offset.txt"))
+```
+
+```{r}
+error_missing_offset <- fread(here("filtered/error_missing_offset.txt"))
+error_missing_offset
+```

### 2022-08-17 analysis/filtering.Rmd
update filtering (up to subsituted notes corrections)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 dt_error_onset
+
+# add CorrectionNr column
+dt_error_onset$CorrectionNr <- NA
 ```
@@
 dt_error_offset
+
+# add CorrectionNr column
+dt_error_offset$CorrectionNr <- NA
 ```
@@
 ```{r extra-onset, include = FALSE, eval = FALSE}
-dt_error_onset$CorrectionNr <- NA
-
 error_extra_onset <- dt_error_onset[Reason == "Extra Notes"]
@@
 ```{r extra-offset, include = FALSE, eval = FALSE}
-dt_error_offset$CorrectionNr <- NA
-
 error_extra_offset <- dt_error_offset[Reason == "Extra Notes"]
@@
 
-```{r missing-offset, include = FALSE}
+```{r missing-offset, include = FALSE, eval = FALSE}
 error_missing_offset <- dt_error_offset[Reason == "Missing Notes"]
@@
 ```
+
+## Substituted notes
+1) Onset
+
+```{r sub-onset, include = FALSE, eval = FALSE}
+error_sub_onset <- dt_error_onset[startsWith(Reason, "Substituted")]
+
+dt_correct_onset_4 <- data.table()
+for (row in 1:nrow(error_sub_onset)){
+  current <- dt_onset[SubNr == error_sub_onset$SubNr[row] & TrialNr == error_sub_onset$TrialNr[row]]
+  decision = 2 
+  correction = 0 # # of correction for reporting stats
+  while (decision == 2){
+    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
+    print("----- First check -----")
+    current <- edit(current, dt_ideal)
+    print("----- Correction check -----")
+    edit(current, dt_ideal)
+    decision <- menu(c("y", "n", "other"), title = "Save the current data? (to continue, enter 'n')")
+    if (decision == 1){
+      correction = correction + 1
+      error_sub_onset$CorrectionNr[row] <- correction
+      dt_correct_onset_4 <- rbind(dt_correct_onset_4, current[, -c(5:6)])
+    } else if (decision == 3) {
+      error_sub_onset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
+    } else if (decision == 2){
+      correction = correction + 1
+      print("----- Continue correction -----")
+    }
+  }
+}
+
+# export csv
+fwrite(dt_correct_onset_4, file = here("filtered", "dt_correct_onset_4.txt"))
+fwrite(error_sub_onset, file = here("filtered", "error_sub_onset.txt"))
+```
+
+```{r}
+error_sub_onset <- fread(here("filtered/error_sub_onset.txt"))
+error_sub_onset
+```
+
+2) Offset
+
+```{r sub-offset, include = FALSE, eval = FALSE}
+error_sub_offset <- dt_error_offset[startsWith(Reason, "Substituted")]
+
+dt_correct_offset_4 <- data.table()
+for (row in 1:nrow(error_sub_offset)){
+  current <- dt_offset[SubNr == error_sub_offset$SubNr[row] & TrialNr == error_sub_offset$TrialNr[row]]
+  decision = 2 
+  correction = 0 # # of correction for reporting stats
+  while (decision == 2){
+    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
+    print("----- First check -----")
+    current <- edit(current, dt_ideal)
+    print("----- Correction check -----")
+    edit(current, dt_ideal)
+    decision <- menu(c("y", "n", "other"), title = "Save the current data? (to continue, enter 'n')")
+    if (decision == 1){
+      correction = correction + 1
+      error_sub_offset$CorrectionNr[row] <- correction
+      dt_correct_offset_4 <- rbind(dt_correct_offset_4, current[, -c(5:6)])
+    } else if (decision == 3) {
+      error_sub_offset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
+    } else if (decision == 2){
+      correction = correction + 1
+      print("----- Continue correction -----")
+    }
+  }
+}
+
+# export csv
+fwrite(dt_correct_offset_4, file = here("filtered", "dt_correct_offset_4.txt"))
+fwrite(error_sub_offset, file = here("filtered", "error_sub_offset.txt"))
+```
+
+```{r}
+error_sub_offset <- fread(here("filtered/error_sub_offset.txt"))
+error_sub_offset
+```

### 2022-08-17 analysis/function.R

@@ -1,5 +1,6 @@
 # install and load required packages
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+if (!require("tibble")) {install.packages("tibble"); require("tibble")}
 if (!require("editData")) {install.packages("editData"); require("editData")}
 if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
 
@@ -60,7 +61,7 @@ insert_na <- function(data, ideal){
   # insert NA row
   while (nrow(data) < nrow(ideal)){
     for (note in 1:nrow(data)){
-      if (note != 66){
+      if (note != 71){
         if (data$Pitch[note] != ideal$Pitch[note]){
           data <- add_row(data, .before = note)
           data[note] <- data[note-1]
@@ -69,20 +70,20 @@ insert_na <- function(data, ideal){
           data$Pitch[note] <- ideal$Pitch[note]
           break
         }
-      } else if (note == 66){
+      } else if (note == 71){
         if (data$Pitch[note] != ideal$Pitch[note]){
           data <- add_row(data, .after = note)
           data[note+1] <- data[note]
-          # NA66
+          # NA71
           data$TimeStamp[note] <- NA
           data$Velocity[note] <- NA
           data$Pitch[note] <- ideal$Pitch[note]
-          # NA67
+          # NA72
           data$TimeStamp[note+1] <- NA
           data$Velocity[note+1] <- NA
           data$Pitch[note+1] <- ideal$Pitch[note+1]
           break
-        } else if (data$Pitch[note] == ideal$Pitch[note]){
+        } else if (data$Pitch[note] == ideal$Pitch[note]){ # if the very last note is missing
           data <- add_row(data, .after = note)
           data[note+1] <- data[note]
           data$TimeStamp[note+1] <- NA

### 2022-08-19 analysis/filtering.Rmd
update filtering (trying to implement invidual manipulations
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # Detect erroneous trials
+Firstly, separate non-erroneous performances from erroneous performances.
 
@@
 # Manual pitch removal
+Second, manually correct some pitch errors if the performance has only one type of the following errors:
+
+- Extra notes
+- Missing notes
+- Substituted notes
 
@@
 ```
+
+# Individual investigation
+If the performance has complicated pitch errors, investigate individually.
+
+1) Onset
+
+```{r}
+# combine error
+error_onset <- rbind(error_extra_onset, error_missing_onset, error_sub_onset)
+
+error_ind_onset <- error_onset[startsWith(CorrectionNr, "Check")]
+error_ind_onset
+```
+
+- SubNr: 12, TrialNr: 13
+
+```{r}
+current_1 <- dt_onset[SubNr == error_ind_onset$SubNr[1] & TrialNr == error_ind_onset$TrialNr[1]]
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# remove the first error
+current_1 <- edit(current_1, dt_ideal)
+graph(current_1)
+
+# corrected the second error (substituted note)
+current_1 <- edit(current_1, dt_ideal)
+graph(current_1)
+
+error_ind_onset$CorrectionNr[1] <- 2
+```
+
+- SubNr: 10, TrialNr: 10
+
+```{r}
+current_2 <- dt_onset[SubNr == error_ind_onset$SubNr[2] & TrialNr == error_ind_onset$TrialNr[2]]
+current_2 <- insert_na(current_2, dt_ideal)
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+# insert one skipped note
+current_2 <- edit(current_2, dt_ideal)
+graph(current_2)
+
+# remove an extra note
+current_2 <- edit(current_2, dt_ideal)
+graph(current_2)
+
+error_ind_onset$CorrectionNr[2] <- 2
+```
+
+- SubNr
+
+```{r}
+
+```
+
+2) Offset
+
+```{r}
+# combine error
+error_offset <- rbind(error_extra_offset, error_missing_offset, error_sub_offset)
+
+error_ind_offset <- error_offset[startsWith(CorrectionNr, "Check")]
+error_ind_offset
+```

### 2022-08-19 analysis/function.R

@@ -96,3 +96,17 @@ insert_na <- function(data, ideal){
   }
   return(data)
 }
+
+# make and print a graph
+graph <- function(data){
+  graph <- ggplot() +
+    geom_line(data = data, aes(x = RowNr, y = Pitch), colour = "#F8766D") +
+    geom_line(data = data, aes(x = RowNr, y = Ideal), colour = "#00BFC4") +
+    geom_point(data = data, aes(x = RowNr, y = Pitch), colour = "#F8766D") +
+    geom_point(data = data, aes(x = RowNr, y = Ideal), colour = "#00BFC4") +
+    scale_x_continuous("RowNr", current_1$RowNr) +
+    coord_fixed(ratio = 1/4) +
+    labs(title = sprintf("SubNr: %s, TrialNr: %s", unique(data$SubNr), unique(data$TrialNr)), y = "Pitch")
+  print(graph)
+}
+  

### 2022-09-14 analysis/filtering.Rmd
finished onset corrections
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-08-17 analysis/filtering.Rmd
@@
 ```
+
+# Individual investigation
+If the performance has complicated pitch errors, investigate individually.
+
+1) Onset
+
+```{r}
+# combine error
+error_onset <- rbind(error_extra_onset, error_missing_onset, error_sub_onset)
+
+error_ind_onset <- error_onset[startsWith(CorrectionNr, "Check")]
+error_ind_onset
+```
+
+1) SubNr: 12, TrialNr: 13
+
+```{r, include = FALSE}
+current_1 <- dt_onset[SubNr == error_ind_onset$SubNr[1] & TrialNr == error_ind_onset$TrialNr[1]]
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# remove the first error
+current_1 <- edit(current_1, dt_ideal)
+current_1$Ideal <- dt_ideal$Pitch
+graph(current_1)
+
+# corrected the second error (substituted note)
+current_1 <- edit(current_1, dt_ideal)
+graph(current_1)
+
+error_ind_onset$CorrectionNr[1] <- 2
+```
+
+2) SubNr: 10, TrialNr: 10
+
+```{r, include = FALSE}
+current_2 <- dt_onset[SubNr == error_ind_onset$SubNr[2] & TrialNr == error_ind_onset$TrialNr[2]]
+current_2 <- insert_na(current_2, dt_ideal)
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+# insert one skipped note
+current_2 <- edit(current_2, dt_ideal)
+current_2$Ideal <- dt_ideal$Pitch
+current_2$RowNr <- 1:nrow(current_2)
+graph(current_2)
+
+# remove an extra note
+current_2 <- edit(current_2, dt_ideal)
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+error_ind_onset$CorrectionNr[2] <- 3
+```
+
+3) SubNr
+
+```{r, include = FALSE}
+current_3 <- dt_onset[SubNr == error_ind_onset$SubNr[3] & TrialNr == error_ind_onset$TrialNr[3]]
+
+```
+
+2) Offset
+
+```{r, include = FALSE}
+# combine error
+error_offset <- rbind(error_extra_offset, error_missing_offset, error_sub_offset)
+
+error_ind_offset <- error_offset[startsWith(CorrectionNr, "Check")]
+error_ind_offset
+```

### 2022-09-14 analysis/filtering.Rmd
finished offset corrections / implement checking
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-08-17 analysis/filtering.Rmd
@@
 ```{r}
-error_extra_offset <- fread(here("filtered/error_extra_offset.txt"))
+error_extra_offset <- fread(here("filtered", "error_extra_offset.txt"))
 error_extra_offset
@@
 ```{r}
-error_missing_onset <- fread(here("filtered/error_missing_onset.txt"))
+error_missing_onset <- fread(here("filtered", "error_missing_onset.txt"))
 error_missing_onset
@@
 ```{r}
-error_missing_offset <- fread(here("filtered/error_missing_offset.txt"))
+error_missing_offset <- fread(here("filtered", "error_missing_offset.txt"))
 error_missing_offset
@@
 ```{r}
-error_sub_onset <- fread(here("filtered/error_sub_onset.txt"))
+error_sub_onset <- fread(here("filtered", "error_sub_onset.txt"))
 error_sub_onset
@@
 ```{r}
-error_sub_offset <- fread(here("filtered/error_sub_offset.txt"))
+error_sub_offset <- fread(here("filtered", "error_sub_offset.txt"))
 error_sub_offset
 ```
+
+# Individual investigation
+If the performance has complicated pitch errors, investigate individually.
+
+1) Onset
+
+```{r, include = FALSE}
+# combine error
+error_onset <- rbind(error_extra_onset, error_missing_onset, error_sub_onset)
+
+error_ind_onset <- error_onset[startsWith(CorrectionNr, "Check")]
+```
+
+- SubNr: 12, TrialNr: 13
+
+```{r, include = FALSE, eval = FALSE}
+current_1 <- dt_onset[SubNr == error_ind_onset$SubNr[1] & TrialNr == error_ind_onset$TrialNr[1]]
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# remove the first error
+current_1 <- edit(current_1, dt_ideal)
+current_1$Ideal <- dt_ideal$Pitch
+graph(current_1)
+
+# corrected the second error (substituted note)
+current_1 <- edit(current_1, dt_ideal)
+graph(current_1)
+
+error_ind_onset$CorrectionNr[1] <- 2
+```
+
+- SubNr: 10, TrialNr: 10
+
+```{r, include = FALSE, eval = FALSE}
+current_2 <- dt_onset[SubNr == error_ind_onset$SubNr[2] & TrialNr == error_ind_onset$TrialNr[2]]
+
+# insert one note
+current_2 <- insert_na(current_2, dt_ideal)
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+# insert one note
+current_2 <- edit(current_2, dt_ideal)
+current_2$RowNr <- 1:nrow(current_2)
+current_2$Ideal <- cc(dt_ideal$Pitch, rep(NA, abs(nrow(current_2) - nrow(dt_ideal))))
+graph(current_2)
+
+# remove one extra note
+current_2 <- edit(current_2, dt_ideal)
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+error_ind_onset$CorrectionNr[2] <- 3
+```
+
+- SubNr: 6, TrialNr: 2
+
+```{r, include = FALSE, eval = FALSE}
+current_3 <- dt_onset[SubNr == error_ind_onset$SubNr[3] & TrialNr == error_ind_onset$TrialNr[3]]
+current_3$RowNr <- c(1:nrow(current_3))
+current_3$Ideal <- dt_ideal$Pitch
+graph(current_3)
+
+# remove the first error
+current_3 <- edit(current_3, dt_ideal)
+
+# insert one note
+current_3 <- insert_na(current_3, dt_ideal)
+current_3$Ideal <- dt_ideal$Pitch
+graph(current_3)
+
+error_ind_onset$CorrectionNr[3] <- 2
+```
+
+- SubNr: 6, TrialNr: 12
+
+```{r, include = FALSE, eval = FALSE}
+current_4 <- dt_onset[SubNr == error_ind_onset$SubNr[4] & TrialNr == error_ind_onset$TrialNr[4]]
+current_4$RowNr <- c(1:nrow(current_4))
+current_4$Ideal <- dt_ideal$Pitch
+graph(current_4)
+
+# insert one skipped note
+current_4 <- edit(current_4, dt_ideal)
+current_4$RowNr <- 1:nrow(current_4)
+current_4$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_4) - nrow(dt_ideal))))
+graph(current_4)
+
+# remove one extra note
+current_4 <- edit(current_4, dt_ideal)
+current_4$RowNr <- c(1:nrow(current_4))
+current_4$Ideal <- dt_ideal$Pitch
+graph(current_4)
+
+error_ind_onset$CorrectionNr[4] <- 2
+```
+
+- SubNr: 8, TrialNr: 6
+
+```{r, include = FALSE, eval = FALSE}
+current_5 <- dt_onset[SubNr == error_ind_onset$SubNr[5] & TrialNr == error_ind_onset$TrialNr[5]]
+current_5$RowNr <- c(1:nrow(current_5))
+current_5$Ideal <- dt_ideal$Pitch
+graph(current_5)
+
+# remove one wrong note
+current_5 <- edit(current_5, dt_ideal)
+
+# insert one note
+current_5 <- insert_na(current_5, dt_ideal)
+current_5$RowNr <- c(1:nrow(current_5))
+current_5$Ideal <- dt_ideal$Pitch
+graph(current_5)
+
+error_ind_onset$CorrectionNr[5] <- 2
+```
+
+- SubNr: 8, TrialNr: 10
+
+```{r, include = FALSE, eval = FALSE}
+current_6 <- dt_onset[SubNr == error_ind_onset$SubNr[6] & TrialNr == error_ind_onset$TrialNr[6]]
+current_6$RowNr <- c(1:nrow(current_6))
+current_6$Ideal <- dt_ideal$Pitch
+graph(current_6)
+
+# remove one extra note
+current_6 <- edit(current_6, dt_ideal)
+
+# insert one note
+current_6 <- insert_na(current_6, dt_ideal)
+current_6$RowNr <- c(1:nrow(current_6))
+current_6$Ideal <- dt_ideal$Pitch
+graph(current_6)
+
+error_ind_onset$CorrectionNr[6] <- 2
+```
+
+- SubNr: 13, TrialNr: 1
+
+```{r, include = FALSE, eval = FALSE}
+current_7 <- dt_onset[SubNr == error_ind_onset$SubNr[7] & TrialNr == error_ind_onset$TrialNr[7]]
+current_7$RowNr <- c(1:nrow(current_7))
+current_7$Ideal <- dt_ideal$Pitch
+graph(current_7)
+
+# insert one note
+current_7 <- edit(current_7, dt_ideal)
+current_7$RowNr <- 1:nrow(current_7)
+current_7$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_7) - nrow(dt_ideal))))
+graph(current_7)
+
+# remove one note
+current_7 <- edit(current_7, dt_ideal)
+current_7$RowNr <- c(1:nrow(current_7))
+current_7$Ideal <- dt_ideal$Pitch
+graph(current_7)
+
+error_ind_onset$CorrectionNr[7] <- 2
+```
+
+```{r, include = FALSE, eval = FALSE}
+dt_correct_onset_5 <- rbind(current_1, current_2, current_3, current_4, current_5, current_6, current_7)
+dt_correct_onset_5 <- dt_correct_onset_5[, -c(5:6)]
+
+# export csv
+fwrite(dt_correct_onset_5, file = here("filtered", "dt_correct_onset_5.txt"))
+fwrite(error_ind_onset, file = here("filtered", "error_ind_onset.txt"))
+```
+
+```{r}
+error_ind_onset <- fread(here("filtered", "error_ind_onset.txt"))
+error_ind_onset
+```
+
+2) Offset
+
+```{r}
+# combine error
+error_offset <- rbind(error_extra_offset, error_missing_offset, error_sub_offset)
+
+error_ind_offset <- error_offset[startsWith(CorrectionNr, "Check")]
+```
+
+- SubNr: 12, TrialNr: 13
+
+```{r, include = FALSE, eval = FALSE}
+current_1 <- dt_onset[SubNr == error_ind_offset$SubNr[1] & TrialNr == error_ind_offset$TrialNr[1]]
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# remove one note
+current_1 <- edit(current_1, dt_ideal)
+current_1$Ideal <- dt_ideal$Pitch
+graph(current_1)
+
+# corrected the second error (substituted note)
+current_1 <- edit(current_1, dt_ideal)
+graph(current_1)
+
+error_ind_offset$CorrectionNr[1] <- 2
+```
+
+- SubNr: 17, TrialNr: 7
+
+```{r, include = FALSE, eval = FALSE}
+current_2 <- dt_offset[SubNr == error_ind_offset$SubNr[2] & TrialNr == error_ind_offset$TrialNr[2]]
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_2) - nrow(dt_ideal))))
+graph(current_2)
+
+# correct two notes
+current_2 <- edit(current_2, dt_ideal)
+graph(current_2)
+
+# remove one note
+current_2 <- edit(current_2, dt_ideal)
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+error_ind_offset$CorrectionNr[2] <- 3
+```
+
+- SubNr: 4, TrialNr: 4
+
+```{r, include = FALSE, eval = FALSE}
+current_3 <- dt_offset[SubNr == error_ind_offset$SubNr[3] & TrialNr == error_ind_offset$TrialNr[3]]
+current_3$RowNr <- c(1:nrow(current_3))
+current_3$Ideal <- dt_ideal$Pitch[-c(71, 72)]
+graph(current_3)
+
+# correct two notes
+current_3 <- edit(current_3, dt_ideal[-c(71, 72),])
+current_3$RowNr <- c(1:nrow(current_3))
+current_3$Ideal <- dt_ideal$Pitch[-c(71, 72)]
+graph(current_3)
+
+# insert one notes
+current_3 <- edit(current_3, dt_ideal[-c(71, 72),])
+current_3$RowNr <- c(1:nrow(current_3))
+current_3$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_3)
+
+# correct two notes
+current_3 <- edit(current_3, dt_ideal[-c(72),])
+current_3$RowNr <- c(1:nrow(current_3))
+current_3$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_3)
+
+# insert one note
+current_3 <- insert_na(current_3, dt_ideal)
+current_3$RowNr <- c(1:nrow(current_3))
+current_3$Ideal <- dt_ideal$Pitch
+graph(current_3)
+
+error_ind_offset$CorrectionNr[3] <- 6
+```
+
+- SubNr: 4, TrialNr: 4
+
+```{r, include = FALSE, eval = FALSE}
+current_4 <- dt_offset[SubNr == error_ind_offset$SubNr[4] & TrialNr == error_ind_offset$TrialNr[4]]
+current_4$RowNr <- c(1:nrow(current_4))
+current_4$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_4)
+
+# correct two notes
+current_4 <- edit(current_4, dt_ideal[-c(72),])
+current_4$RowNr <- c(1:nrow(current_4))
+current_4$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_4)
+
+# insert one note
+current_4 <- insert_na(current_4, dt_ideal)
+current_4$RowNr <- c(1:nrow(current_4))
+current_4$Ideal <- dt_ideal$Pitch
+graph(current_4)
+
+error_ind_offset$CorrectionNr[4] <- 3
+```
+
+- SubNr: 7, TrialNr: 11
+
+```{r, include = FALSE, eval = FALSE}
+current_5 <- dt_offset[SubNr == error_ind_offset$SubNr[5] & TrialNr == error_ind_offset$TrialNr[5]]
+current_5$RowNr <- c(1:nrow(current_5))
+current_5$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_5)
+
+# correct two notes
+current_5 <- edit(current_5, dt_ideal[-c(72),])
+current_5$RowNr <- c(1:nrow(current_5))
+current_5$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_5)
+
+# insert one note
+current_5 <- insert_na(current_5, dt_ideal)
+current_5$RowNr <- c(1:nrow(current_5))
+current_5$Ideal <- dt_ideal$Pitch
+graph(current_5)
+
+error_ind_offset$CorrectionNr[5] <- 3
+```
+
+- SubNr: 10, TrialNr: 10
+
+```{r, include = FALSE, eval = FALSE}
+current_6 <- dt_offset[SubNr == error_ind_offset$SubNr[6] & TrialNr == error_ind_offset$TrialNr[6]]
+current_6$RowNr <- c(1:nrow(current_6))
+current_6$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_6)
+
+# insert one note
+current_6 <- insert_na(current_6, dt_ideal)
+current_6$RowNr <- c(1:nrow(current_6))
+current_6$Ideal <- dt_ideal$Pitch
+graph(current_6)
+
+# insert one note
+current_6 <- edit(current_6, dt_ideal)
+current_6$RowNr <- c(1:nrow(current_6))
+current_6$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_6) - nrow(dt_ideal))))
+graph(current_6)
+
+# remove one note
+current_6 <- edit(current_6, dt_ideal)
+current_6$RowNr <- c(1:nrow(current_6))
+current_6$Ideal <- dt_ideal$Pitch
+graph(current_6)
+
+error_ind_offset$CorrectionNr[6] <- 3
+```
+
+- SubNr: 17, TrialNr: 1
+
+```{r, include = FALSE, eval = FALSE}
+current_7 <- dt_offset[SubNr == error_ind_offset$SubNr[7] & TrialNr == error_ind_offset$TrialNr[7]]
+current_7$RowNr <- c(1:nrow(current_7))
+current_7$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_7)
+
+# insert one note
+current_7 <- insert_na(current_7, dt_ideal)
+current_7$RowNr <- c(1:nrow(current_7))
+current_7$Ideal <- dt_ideal$Pitch
+graph(current_7)
+
+# correct four notes
+current_7 <- edit(current_7, dt_ideal)
+current_7$RowNr <- c(1:nrow(current_7))
+current_7$Ideal <- dt_ideal$Pitch
+graph(current_7)
+
+error_ind_offset$CorrectionNr[7] <- 5
+```
+
+- SubNr: 17, TrialNr: 6
+
+```{r, include = FALSE, eval = FALSE}
+current_8 <- dt_offset[SubNr == error_ind_offset$SubNr[8] & TrialNr == error_ind_offset$TrialNr[8]]
+current_8$RowNr <- c(1:nrow(current_8))
+current_8$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_8)
+
+# correct two notes
+current_8 <- edit(current_8, dt_ideal[-c(72),])
+current_8$RowNr <- c(1:nrow(current_8))
+current_8$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_8)
+
+# insert one note
+current_8 <- insert_na(current_8, dt_ideal)
+current_8$RowNr <- c(1:nrow(current_8))
+current_8$Ideal <- dt_ideal$Pitch
+graph(current_8)
+
+error_ind_offset$CorrectionNr[8] <- 3
+```
+
+- SubNr: 17, TrialNr: 14
+
+```{r, include = FALSE, eval = FALSE}
+current_9 <- dt_offset[SubNr == error_ind_offset$SubNr[9] & TrialNr == error_ind_offset$TrialNr[9]]
+current_9$RowNr <- c(1:nrow(current_9))
+current_9$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_9)
+
+# correct two notes
+current_9 <- edit(current_9, dt_ideal[-c(72),])
+current_9$RowNr <- c(1:nrow(current_9))
+current_9$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_9)
+
+# insert one note
+current_9 <- insert_na(current_9, dt_ideal)
+current_9$RowNr <- c(1:nrow(current_9))
+current_9$Ideal <- dt_ideal$Pitch
+graph(current_9)
+
+error_ind_offset$CorrectionNr[9] <- 3
+```
+
+- SubNr: 20, TrialNr: 1
+
+```{r, include = FALSE, eval = FALSE}
+current_10 <- dt_offset[SubNr == error_ind_offset$SubNr[10] & TrialNr == error_ind_offset$TrialNr[10]]
+current_10$RowNr <- c(1:nrow(current_10))
+current_10$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_10)
+
+# correct two notes
+current_10 <- edit(current_10, dt_ideal[-c(72),])
+current_10$RowNr <- c(1:nrow(current_10))
+current_10$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_10)
+
+# insert one note
+current_10 <- insert_na(current_10, dt_ideal)
+current_10$RowNr <- c(1:nrow(current_10))
+current_10$Ideal <- dt_ideal$Pitch
+graph(current_10)
+
+error_ind_offset$CorrectionNr[10] <- 3
+```
+
+- SubNr: 6, TrialNr: 2
+
+```{r, include = FALSE, eval = FALSE}
+current_11 <- dt_offset[SubNr == error_ind_offset$SubNr[11] & TrialNr == error_ind_offset$TrialNr[11]]
+current_11$RowNr <- c(1:nrow(current_11))
+current_11$Ideal <- dt_ideal$Pitch
+graph(current_11)
+
+# remove one note
+current_11 <- edit(current_11, dt_ideal)
+current_11$RowNr <- c(1:nrow(current_11))
+current_11$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_11)
+
+# insert one note
+current_11 <- insert_na(current_11, dt_ideal)
+current_11$RowNr <- c(1:nrow(current_11))
+current_11$Ideal <- dt_ideal$Pitch
+graph(current_11)
+
+error_ind_offset$CorrectionNr[11] <- 2
+```
+
+- SubNr: 6, TrialNr: 12
+
+```{r, include = FALSE, eval = FALSE}
+current_12 <- dt_offset[SubNr == error_ind_offset$SubNr[12] & TrialNr == error_ind_offset$TrialNr[12]]
+current_12$RowNr <- c(1:nrow(current_12))
+current_12$Ideal <- dt_ideal$Pitch
+graph(current_12)
+
+# insert one note
+current_12 <- edit(current_12, dt_ideal)
+current_12$RowNr <- c(1:nrow(current_12))
+current_12$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_12) - nrow(dt_ideal))))
+graph(current_12)
+
+# remove one note
+current_12 <- edit(current_12, dt_ideal)
+current_12$RowNr <- c(1:nrow(current_12))
+current_12$Ideal <- dt_ideal$Pitch
+graph(current_12)
+
+error_ind_offset$CorrectionNr[12] <- 2
+```
+
+- SubNr: 8, TrialNr: 6
+
+```{r, include = FALSE, eval = FALSE}
+current_13 <- dt_offset[SubNr == error_ind_offset$SubNr[13] & TrialNr == error_ind_offset$TrialNr[13]]
+current_13$RowNr <- c(1:nrow(current_13))
+current_13$Ideal <- dt_ideal$Pitch
+graph(current_13)
+
+# remove one note
+current_13 <- edit(current_13, dt_ideal)
+current_13$RowNr <- c(1:nrow(current_13))
+current_13$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_13)
+
+# insert one note
+current_13 <- insert_na(current_13, dt_ideal)
+current_13$RowNr <- c(1:nrow(current_13))
+current_13$Ideal <- dt_ideal$Pitch
+graph(current_13)
+
+error_ind_offset$CorrectionNr[13] <- 2
+```
+
+- SubNr: 8, TrialNr: 10
+
+```{r, include = FALSE, eval = FALSE}
+current_14 <- dt_offset[SubNr == error_ind_offset$SubNr[14] & TrialNr == error_ind_offset$TrialNr[14]]
+current_14$RowNr <- c(1:nrow(current_14))
+current_14$Ideal <- dt_ideal$Pitch
+graph(current_14)
+
+# remove one note
+current_14 <- edit(current_14, dt_ideal)
+current_14$RowNr <- c(1:nrow(current_14))
+current_14$Ideal <- dt_ideal$Pitch[-c(72)]
+graph(current_14)
+
+# insert one note
+current_14 <- insert_na(current_14, dt_ideal)
+current_14$RowNr <- c(1:nrow(current_14))
+current_14$Ideal <- dt_ideal$Pitch
+graph(current_14)
+
+error_ind_offset$CorrectionNr[14] <- 2
+```
+
+- SubNr: 13, TrialNr: 1
+
+```{r, include = FALSE, eval = FALSE}
+current_15 <- dt_offset[SubNr == error_ind_offset$SubNr[15] & TrialNr == error_ind_offset$TrialNr[15]]
+current_15$RowNr <- c(1:nrow(current_15))
+current_15$Ideal <- dt_ideal$Pitch
+graph(current_15)
+
+# insert one note
+current_15 <- edit(current_15, dt_ideal)
+current_15$RowNr <- c(1:nrow(current_15))
+current_15$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_15) - nrow(dt_ideal))))
+graph(current_15)
+
+# remove one note
+current_15 <- edit(current_15, dt_ideal)
+current_15$RowNr <- c(1:nrow(current_15))
+current_15$Ideal <- dt_ideal$Pitch
+graph(current_15)
+
+error_ind_offset$CorrectionNr[15] <- 2
+```
+
+```{r, include = FALSE, eval = FALSE}
+dt_correct_offset_5 <- rbind(current_1, current_2, current_3, current_4, current_5, current_6, current_7, current_8, current_9, current_10, current_11, current_12, current_13, current_14, current_15)
+dt_correct_offset_5 <- dt_correct_offset_5[, -c(5:6)]
+
+# export csv
+fwrite(dt_correct_offset_5, file = here("filtered", "dt_correct_offset_5.txt"))
+fwrite(error_ind_offset, file = here("filtered", "error_ind_offset.txt"))
+```
+
+```{r}
+error_ind_offset <- fread(here("filtered", "error_ind_offset.txt"))
+error_ind_offset
+```
+
+```{r, include = FALSE, eval = FALSE}
+dt_correct_onset_1 <- fread(here("filtered", "dt_correct_onset_1.txt"))
+dt_correct_onset_2 <- fread(here("filtered", "dt_correct_onset_2.txt"))
+dt_correct_onset_3 <- fread(here("filtered", "dt_correct_onset_3.txt"))
+dt_correct_onset_4 <- fread(here("filtered", "dt_correct_onset_4.txt"))
+dt_correct_onset_5 <- fread(here("filtered", "dt_correct_onset_5.txt"))
+
+# combine all
+dt_correct_onset <- rbind(dt_correct_onset_1, dt_correct_onset_2, dt_correct_onset_3, dt_correct_onset_4, dt_correct_onset_5)
+
+# export csv
+fwrite(dt_correct_onset, file = here("filtered", "dt_correct_onset.txt"))
+
+dt_correct_offset_1 <- fread(here("filtered", "dt_correct_offset_1.txt"))
+dt_correct_offset_2 <- fread(here("filtered", "dt_correct_offset_2.txt"))
+dt_correct_offset_3 <- fread(here("filtered", "dt_correct_offset_3.txt"))
+dt_correct_offset_4 <- fread(here("filtered", "dt_correct_offset_4.txt"))
+dt_correct_offset_5 <- fread(here("filtered", "dt_correct_offset_5.txt"))
+
+# combine all
+dt_correct_offset <- rbind(dt_correct_offset_1, dt_correct_offset_2, dt_correct_offset_3, dt_correct_offset_4, dt_correct_offset_5)
+
+# export csv
+fwrite(dt_correct_offset, file = here("filtered", "dt_correct_offset.txt"))
+```
+
+# Checking
+Check whether data look okay.
+
+```{r, include = FALSE}
+rm(list=ls(all=T)) # clear all
+
+# set working directory
+if (!require("here")) {install.packages("here"); require("here")}
+here::i_am("filtering.Rmd")
+
+# read functions
+source(here("function.R"))
+
+# read a text file for ideal performance
+dt_ideal <- read.table(here("ideal.txt"))
+colnames(dt_ideal) <- "Pitch"
+dt_ideal$RowNr <- 1:nrow(dt_ideal)
+setcolorder(dt_ideal, c(2, 1))
+
+# read txt files
+dt_correct_onset <- fread(here("filtered", "dt_correct_onset.txt"))
+dt_correct_offset <- fread(here("filtered", "dt_correct_offset.txt"))
+```
+
+1) Onset
+```{r}
+print(checker(dt_correct_onset, dt_ideal))
+```
+
+2) Offset
+```{r}
+print(checker(dt_correct_offset, dt_ideal))
+```

### 2022-09-14 analysis/function.R
return no errors if there is no errors/missing trials
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -35,8 +35,12 @@ checker <- function(data, ideal){
       }
     }
   }
-  colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
-  return(dt_errors)
+  if (nrow(dt_errors) != 0){
+    colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
+    return(dt_errors)
+  } else if (nrow(dt_errors) == 0){
+    return("No errors!")
+  }
 }
 
 ### edit data

### 2022-09-14 analysis/trimming.Rmd
removed outliers for IOIs
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Preprocessing: trimming"
output: html_notebook
---

```{r setup, include = FALSE}
# set chunk option
knitr::opts_chunk$set(echo = FALSE)

# set working directory
if (!require("here")) {install.packages("here"); require("here")}
here::i_am("trimming.Rmd")

# packages
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}

# create a folder if not exists
if (!file.exists(here("trimmed"))){
  dir.create(here("trimmed"))
}

# read txt files
dt_onset <- fread(here("filtered", "dt_correct_onset.txt"))
dt_offset <- fread(here("filtered", "dt_correct_offset.txt"))

# sort data
setorder(dt_onset, cols = "SubNr", "TrialNr", "RowNr")
setorder(dt_offset, cols = "SubNr", "TrialNr", "RowNr")
```

```{r, include = FALSE}
# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

# IOIs
```{r, include = FALSE}
# calculate IOIs
dt_ioi <- dt_onset
dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))

# remove the first note
dt_ioi <- dt_ioi[RowNr != 1]

# assign Interval
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# assign Articulation/Dynamics
dt_ioi$Articulation <- "NA"
dt_ioi$Dynamics <- "NA"
# Legato/Forte
for (phrase in 1:length(ls_legato)){
  for (note in 1:length(ls_legato[[phrase]])){
    dt_ioi[Interval == ls_legato[[phrase]][note]]$Articulation <- "Legato"
    dt_ioi[Interval == ls_legato[[phrase]][note]]$Dynamics <- "Forte"
  }
}
# Staccato/Piano
for (phrase in 1:length(ls_staccato)){
  for (note in 1:length(ls_staccato[[phrase]])){
    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Articulation <- "Staccato"
    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Dynamics <- "Piano"
  }
}

# assign Subcomponent Change
for (number in change_1){
  dt_ioi[Interval == number]$Articulation <- "LtoS"
  dt_ioi[ Interval == number]$Dynamics <- "FtoP"
}
for (number in change_2){
  dt_ioi[Interval == number]$Articulation <- "StoL"
  dt_ioi[Interval == number]$Dynamics <- "PtoF"
}
```

## Histogram
### Before removing outliers
```{r}
dt_ioi_subset_all <- dt_ioi[Articulation != "NA" & !is.na(IOI)]

gghistogram(dt_ioi_subset_all, x = "IOI", color = "Articulation", bins = 150)
```

### After removing outliers
Remove outliers, which are outside of +/- 3SD separately for whether responses on boundary (transition points between Legato/Forte and Staccato/Piano) because IOIs are likely to be slower in such boundaries.

```{r}
ioi_summary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr)]

# exclude tempo deviated participants
ioi_summary$Include <- "No"
ioi_summary[Mean < mean(ioi_summary$Mean)+3*sd(ioi_summary$Mean) & Mean > mean(ioi_summary$Mean)-3*sd(ioi_summary$Mean)]$Include <- "Yes"

# evaluate if it's on boundary
dt_ioi_subset_all$Boundary <- "No"
dt_ioi_subset_all[Articulation == "LtoS" | Articulation == "StoL"]$Boundary <- "Yes"

# exclude ioi > +- 3SD (separately for each Boundary)
ioi_boundary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = Boundary]

dt_ioi_trimmed <- data.table()
for (boundary in unique(ioi_boundary$Boundary)){
  upper_ioi <- ioi_boundary[Boundary == boundary]$Mean+3*ioi_boundary[Boundary == boundary]$SD
  lower_ioi <- ioi_boundary[Boundary == boundary]$Mean-3*ioi_boundary[Boundary == boundary]$SD
  dt_current <- dt_ioi_subset_all[Boundary == boundary & IOI < upper_ioi & IOI > lower_ioi]
  dt_ioi_trimmed <- rbind(dt_ioi_trimmed, dt_current)
}

gghistogram(dt_ioi_trimmed, color = "Articulation", x = "IOI", bins = 150)
```

Remove `r nrow(dt_ioi_subset_all)-nrow(dt_ioi_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_ioi_subset_all)-nrow(dt_ioi_trimmed))/nrow(dt_ioi_subset_all))` % of total responses

# KOT

# KV

# KV Difference

### 2022-09-15 analysis/trimming.Rmd
removed KOT outliers
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # define subcomponents
-# for intervals
+# for intervals (Legato/Forte, Staccato/Piano)
 ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
@@
 
-# define Component Change (LtoS, FtoP)
+# define subcomponent change (LtoS, FtoP)
 change_1 <- c(5, 17, 47)
-# define Component Change (StoL, PtoF)
+# define subcomponent change (StoL, PtoF)
 change_2 <- c(8, 20, 39)
@@
 ```{r}
-dt_ioi_subset_all <- dt_ioi[Articulation != "NA" & !is.na(IOI)]
+dt_ioi_subset <- dt_ioi[Articulation != "NA" & !is.na(IOI)]
 
-gghistogram(dt_ioi_subset_all, x = "IOI", color = "Articulation", bins = 150)
+gghistogram(dt_ioi_subset, x = "IOI", color = "Articulation", bins = 150)
 ```
@@
 ### After removing outliers
-Remove outliers, which are outside of +/- 3SD separately for whether responses on boundary (transition points between Legato/Forte and Staccato/Piano) because IOIs are likely to be slower in such boundaries.
+Remove outliers, which are outside of +/- 3SD separately depending on whether responses are on boundaries (transition points between Legato/Forte and Staccato/Piano) because IOIs are likely to be slower in the transition points.
 
 ```{r}
-ioi_summary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr)]
+ioi_summary <- dt_ioi_subset[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr)]
 
@@
 # evaluate if it's on boundary
-dt_ioi_subset_all$Boundary <- "No"
-dt_ioi_subset_all[Articulation == "LtoS" | Articulation == "StoL"]$Boundary <- "Yes"
+dt_ioi_subset$Boundary <- "No"
+dt_ioi_subset[Articulation == "LtoS" | Articulation == "StoL"]$Boundary <- "Yes"
 
 # exclude ioi > +- 3SD (separately for each Boundary)
-ioi_boundary <- dt_ioi_subset_all[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = Boundary]
+ioi_boundary <- dt_ioi_subset[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = Boundary]
 
@@
   lower_ioi <- ioi_boundary[Boundary == boundary]$Mean-3*ioi_boundary[Boundary == boundary]$SD
-  dt_current <- dt_ioi_subset_all[Boundary == boundary & IOI < upper_ioi & IOI > lower_ioi]
+  dt_current <- dt_ioi_subset[Boundary == boundary & IOI < upper_ioi & IOI > lower_ioi]
   dt_ioi_trimmed <- rbind(dt_ioi_trimmed, dt_current)
@@
 
-Remove `r nrow(dt_ioi_subset_all)-nrow(dt_ioi_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_ioi_subset_all)-nrow(dt_ioi_trimmed))/nrow(dt_ioi_subset_all))` % of total responses
+Remove `r nrow(dt_ioi_subset)-nrow(dt_ioi_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_ioi_subset)-nrow(dt_ioi_trimmed))/nrow(dt_ioi_subset))` % of total responses
+
+```{r, include = FALSE}
+fwrite(dt_ioi_trimmed, file = here("trimmed", "dt_ioi_trimmed.txt"), row.names = F)
+```
 
 # KOT
+```{r}
+dt_kot <- dt_onset
+
+# Offset 1 - Onset 2
+dt_kot$KOT <- NA
+for (row in 1:nrow(dt_kot)){
+  if (row < nrow(dt_kot)){
+    dt_kot$KOT[row+1] <- dt_offset$TimeStamp[row] - dt_onset$TimeStamp[row+1] # offset(n) - onset(n+1)
+  }
+}
+
+# remove the first note
+dt_kot <- dt_kot[RowNr != 1]
+
+# assign a sequence number for each tone
+dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)
+
+# assign Articulation/Dynamics
+dt_kot$Articulation <- "NA"
+dt_kot$Dynamics <- "NA"
+# Legato/Forte
+for (phrase in 1:length(ls_legato)){
+  for (note in 1:length(ls_legato[[phrase]])){
+    dt_kot[Interval == ls_legato[[phrase]][note]]$Articulation <- "Legato"
+    dt_kot[Interval == ls_legato[[phrase]][note]]$Dynamics <- "Forte"
+  }
+}
+# Staccato/Piano
+for (phrase in 1:length(ls_staccato)){
+  for (note in 1:length(ls_staccato[[phrase]])){
+    dt_kot[Interval == ls_staccato[[phrase]][note]]$Articulation <- "Staccato"
+    dt_kot[Interval == ls_staccato[[phrase]][note]]$Dynamics <- "Piano"
+  }
+}
+
+# assign Subcomponent Change
+for (number in change_1){
+  dt_kot[Interval == number]$Articulation <- "LtoS"
+  dt_kot[ Interval == number]$Dynamics <- "FtoP"
+}
+for (number in change_2){
+  dt_kot[Interval == number]$Articulation <- "StoL"
+  dt_kot[Interval == number]$Dynamics <- "PtoF"
+}
+```
+
+## Histogram
+### Before removing outliers
+
+```{r}
+dt_kot_subset <- dt_kot[Articulation != "NA" & !is.na(KOT)]
+
+gghistogram(dt_kot_subset, x = "KOT", color = "Articulation", bins = 150)
+```
+
+### After removing outliers
+```{r}
+kot_subcomponent <- dt_kot_subset[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = Articulation]
+dt_kot_trimmed <- data.table()
+for (subcomponent in unique(dt_kot_subset$Articulation)){
+  upper <- kot_subcomponent[Articulation == subcomponent]$Mean+3*kot_subcomponent[Articulation == subcomponent]$SD
+  lower <- kot_subcomponent[Articulation == subcomponent]$Mean-3*kot_subcomponent[Articulation == subcomponent]$SD
+  dt_current <- dt_kot_subset[Articulation == subcomponent & KOT < upper & KOT > lower]
+  dt_kot_trimmed <- rbind(dt_kot_trimmed, dt_current)
+}
+
+gghistogram(dt_kot_trimmed, x = "KOT", color = "Articulation", bins = 150)
+```
+
+```{r, include = FALSE}
+fwrite(dt_kot_trimmed, file = here("trimmed", "dt_kot_trimmed.txt"), row.names = F)
+```
 

### 2022-09-15 analysis/preprosessor/trimming.Rmd
finished trimming / moved files to preprosessor
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-09-15 analysis/trimming.Rmd
@@
 # KOT
-```{r}
+
+```{r, include = FALSE}
 dt_kot <- dt_onset
 
-# Offset 1 - Onset 2
+# calculate KOT (Offset 1 - Onset 2)
 dt_kot$KOT <- NA
@@
 
+Remove `r nrow(dt_kot_subset)-nrow(dt_kot_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_kot_subset)-nrow(dt_kot_trimmed))/nrow(dt_kot_subset))` % of total responses
+
 ```{r, include = FALSE}
@@
 # KV
+```{r, include = FALSE}
+# calculate Diff (acceleration - velocity difference between notes)
+dt_vel <- dt_onset
+dt_vel$Diff <- diff(c(0, dt_vel$Velocity))
+
+# remove the first note
+dt_vel_diff <- dt_vel[RowNr != 1]
+dt_vel$Diff <- NULL # remove Diff from dt_vel
+
+# assign Subcomponents
+# for each note
+dt_vel$Articulation <- "NA"
+dt_vel$Dynamics <- "NA"
+# Legato/Forte
+for (phrase in 1:length(ls_legato2)){
+  for (note in 1:length(ls_legato2[[phrase]])){
+    dt_vel[RowNr == ls_legato2[[phrase]][note]]$Articulation <- "Legato"
+    dt_vel[RowNr  == ls_legato2[[phrase]][note]]$Dynamics <- "Forte"
+  }
+}
+# Staccato/Piano
+for (phrase in 1:length(ls_staccato2)){
+  for (note in 1:length(ls_staccato2[[phrase]])){
+    dt_vel[RowNr  == ls_staccato2[[phrase]][note]]$Articulation <- "Staccato"
+    dt_vel[RowNr  == ls_staccato2[[phrase]][note]]$Dynamics <- "Piano"
+  }
+}
+```
+
+## Histogram
+### Before removing outliers
+```{r}
+dt_vel_subset <- dt_vel[Articulation != "NA" & !is.na(Velocity)]
+
+gghistogram(dt_vel_subset, x = "Velocity", color = "Articulation", bins = 150)
+```
+
+### After removing outliers
+```{r}
+vel_subcomponent <- dt_vel_subset[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = Articulation]
+dt_vel_trimmed <- data.table()
+for (subcomponent in unique(dt_vel_subset$Articulation)){
+  upper <- vel_subcomponent[Articulation == subcomponent]$Mean+3*vel_subcomponent[Articulation == subcomponent]$SD
+  lower <- vel_subcomponent[Articulation == subcomponent]$Mean-3*vel_subcomponent[Articulation == subcomponent]$SD
+  dt_current <- dt_vel_subset[Articulation == subcomponent & Velocity < upper & Velocity > lower]
+  dt_vel_trimmed <- rbind(dt_vel_trimmed, dt_current)
+}
+
+gghistogram(dt_vel_trimmed, x = "Velocity", color = "Articulation", bins = 150)
+```
+
+Remove `r nrow(dt_vel_subset)-nrow(dt_vel_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_vel_subset)-nrow(dt_vel_trimmed))/nrow(dt_vel_subset))` % of total responses
+
+```{r, include = FALSE}
+fwrite(dt_vel_trimmed, file = here("trimmed", "dt_vel_trimmed.txt"), row.names = F)
+```
 
 # KV Difference
+```{r, include = FALSE}
+# assign Interval
+dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)
+
+# assign Subcomponents
+# for intervals
+dt_vel_diff$Articulation <- "NA"
+dt_vel_diff$Dynamics <- "NA"
+# Legato/Forte
+for (phrase in 1:length(ls_legato)){
+  for (note in 1:length(ls_legato[[phrase]])){
+    dt_vel_diff[Interval == ls_legato[[phrase]][note]]$Articulation <- "Legato"
+    dt_vel_diff[Interval == ls_legato[[phrase]][note]]$Dynamics <- "Forte"
+  }
+}
+# Staccato/Piano
+for (phrase in 1:length(ls_staccato)){
+  for (note in 1:length(ls_staccato[[phrase]])){
+    dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$Articulation <- "Staccato"
+    dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$Dynamics <- "Piano"
+  }
+}
+
+# assign Subcomponent Change
+for (number in change_1){
+  dt_vel_diff[Interval == number]$Articulation <- "LtoS"
+  dt_vel_diff[ Interval == number]$Dynamics <- "FtoP"
+}
+for (number in change_2){
+  dt_vel_diff[Interval == number]$Articulation <- "StoL"
+  dt_vel_diff[Interval == number]$Dynamics <- "PtoF"
+}
+
+```
+
+## Histogram
+### Before removing outliers
+```{r}
+dt_vel_diff_subset <- dt_vel_diff[Articulation != "NA" & !is.na(Diff)]
+
+gghistogram(dt_vel_diff_subset, x = "Diff", color = "Articulation", bins = 150)
+```
+
+### After removing outliers
+```{r}
+vel_diff_subcomponent <- dt_vel_diff_subset[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = Articulation]
+dt_vel_diff_trimmed <- data.table()
+for (subcomponent in unique(dt_vel_diff_subset$Articulation)){
+  upper <- vel_diff_subcomponent[Articulation == subcomponent]$Mean+3*vel_diff_subcomponent[Articulation == subcomponent]$SD
+  lower <- vel_diff_subcomponent[Articulation == subcomponent]$Mean-3*vel_diff_subcomponent[Articulation == subcomponent]$SD
+  dt_current <- dt_vel_diff_subset[Articulation == subcomponent & Diff < upper & Diff > lower]
+  dt_vel_diff_trimmed <- rbind(dt_vel_diff_trimmed, dt_current)
+}
+
+gghistogram(dt_vel_diff_trimmed, x = "Diff", color = "Articulation", bins = 150)
+```
+
+Remove `r nrow(dt_vel_diff_subset)-nrow(dt_vel_diff_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_vel_diff_subset)-nrow(dt_vel_diff_trimmed))/nrow(dt_vel_diff_subset))` % of total responses
+
+```{r, include = FALSE}
+fwrite(dt_vel_diff_trimmed, file = here("trimmed", "dt_vel_diff_trimmed.txt"), row.names = F)
+```

### 2022-09-15 analysis/preprosessor/trimming.Rmd
remove ggplot2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}

### 2022-09-15 analysis/primary.Rmd
update primary analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-12-01 analysis/primary.Rmd
@@
 
-# Outline
-- IOIs
-- KOT
-- KV
-- KV-Diff
-
 ```{r setup, include = FALSE}
@@
 ```{r file, include = FALSE}
-filename = "filtered/dt_correct_onset_1.txt"
+data_ioi = "preprosessor/trimmed/dt_ioi_trimmed.txt"
+data_kot = "preprosessor/trimmed/dt_kot_trimmed.txt"
+data_vel = "preprosessor/trimmed/dt_vel_trimmed.txt"
+data_vel_diff = "preprosessor/trimmed/dt_vel_diff_trimmed.txt"
 ```
 
-```{r extract, include = FALSE}
-data <- fread(filename, header = T, sep = ",", dec = ".") # read a trimmed csv
+```{r prep, include = FALSE}
+dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
+dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
+dt_vel <- fread(data_vel, header = T, sep = ",", dec = ".")
+dt_vel_diff <- fread(data_vel_diff, header = T, sep = ",", dec = ".")
 
 # SubNr as a factor
-data$SubNr <- as.factor(data$SubNr)
+dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
 
 # assign stimuli categories
-data$Stimuli <- as.numeric(gsub(".mid", "", data$Stimuli))
-data$Category <- "NA"
-data[Stimuli < 5]$Category <- "art_only"
-data[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
-data[Stimuli > 8 & Stimuli < 13]$Category <- "high"
-data[Stimuli > 12]$Category <- "low"
+dt_ioi$Stimuli <- as.numeric(gsub(".mid", "", dt_ioi$Stimuli))
+dt_ioi$Articulation2 <- "NA"
+dt_ioi$Dynamics2 <- "NA"
+dt_ioi[Stimuli < 5]$Articulation2 <- "Present"
+dt_ioi[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
+dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
+dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_ioi[Stimuli > 12]$Articulation2 <- "Absent"
+dt_ioi[Stimuli > 12]$Dynamics2 <- "Absent"
 
-# assign articulation (yes/no) & dynamics (yes/no)
-data$Articulation <- "NA"
-data$Dynamics <- "NA"
-data[Category == "art_only"]$Articulation <- "Present"
-data[Category == "art_only"]$Dynamics <- "Absent"
-data[Category == "dyn_only"]$Articulation <- "Absent"
-data[Category == "dyn_only"]$Dynamics <- "Present"
-data[Category == "high"]$Articulation <- "Present"
-data[Category == "high"]$Dynamics <- "Present"
-data[Category == "low"]$Articulation <- "Absent"
-data[Category == "low"]$Dynamics <- "Absent"
+dt_kot$Stimuli <- as.numeric(gsub(".mid", "", dt_kot$Stimuli))
+dt_kot$Articulation2 <- "NA"
+dt_kot$Dynamics2 <- "NA"
+dt_kot[Stimuli < 5]$Articulation2 <- "Present"
+dt_kot[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
+dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
+dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_kot[Stimuli > 12]$Articulation2 <- "Absent"
+dt_kot[Stimuli > 12]$Dynamics2 <- "Absent"
 
-# IOIs
-dt_ioi <- data
-dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))
+dt_vel$Stimuli <- as.numeric(gsub(".mid", "", dt_vel$Stimuli))
+dt_vel$Articulation2 <- "NA"
+dt_vel$Dynamics2 <- "NA"
+dt_vel[Stimuli < 5]$Articulation2 <- "Present"
+dt_vel[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
+dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
+dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_vel[Stimuli > 12]$Articulation2 <- "Absent"
+dt_vel[Stimuli > 12]$Dynamics2 <- "Absent"
 
-# remove the first note
-dt_ioi <- dt_ioi[RowNr != 1]
-
-# assign Interval
-dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)
-
-# define intervals
-ls_l_f <- list(c(1:4), c(9:16), c(21:24), c(40:46))
-ls_s_p <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
-change_1 <- c(5, 17, 47) # legato to staccato / forte to piano
-change_2 <- c(8, 20, 39) # staccato to legato / piano to forte
-
-# assign Subcomponents
-dt_ioi$Subcomponent <- "NA"
-# legato & forte
-for (phrase in 1:length(ls_l_f)){
-  for (note in 1:length(ls_l_f[[phrase]])){
-    dt_ioi[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
-  }
-}
-# staccato & piano
-for (phrase in 1:length(ls_s_p)){
-  for (note in 1:length(ls_s_p[[phrase]])){
-    dt_ioi[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
-  }
-}
-for (number in change_1){
-  dt_ioi[Interval == number]$Subcomponent <- "LFtoPS"
-}
-for (number in change_2){
-  dt_ioi[Interval == number]$Subcomponent <- "SPtoLF"
-}
-
-# KOT
-
-# KV
-dt_vel <- data
-dt_vel$Diff <- diff(c(0, dt_vel$Velocity))
-
-# remove the first note
-dt_vel_diff <- dt_vel[RowNr != 1]
-dt_vel$Diff <- NULL # remove Diff from dt_vel
-
-# assign Interval
-dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)
-
-# for each note (velocity)
-ls_l_f_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
-ls_s_p_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
-
-# assign Subcomponents
-# for each note
-dt_vel$Subcomponent <- "NA"
-# legato & forte
-for (phrase in 1:length(ls_l_f_2)){
-  for (note in 1:length(ls_l_f_2[[phrase]])){
-    dt_vel[RowNr == ls_l_f_2[[phrase]][note]]$Subcomponent <- "L-F"
-  }
-}
-# staccato & piano
-for (phrase in 1:length(ls_s_p_2)){
-  for (note in 1:length(ls_s_p_2[[phrase]])){
-    dt_vel[RowNr == ls_s_p_2[[phrase]][note]]$Subcomponent <- "S-P"
-  }
-}
-
-dt_vel_diff$Subcomponent <- "NA"
-# legato & forte
-for (phrase in 1:length(ls_l_f)){
-  for (note in 1:length(ls_l_f[[phrase]])){
-    dt_vel_diff[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
-  }
-}
-# staccato & piano
-for (phrase in 1:length(ls_s_p)){
-  for (note in 1:length(ls_s_p[[phrase]])){
-    dt_vel_diff[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
-  }
-}
-for (number in change_1){
-  dt_vel_diff[Interval == number]$Subcomponent <- "LFtoSP"
-}
-for (number in change_2){
-  dt_vel_diff[Interval == number]$Subcomponent <- "SPtoLF"
-}
+dt_vel_diff$Stimuli <- as.numeric(gsub(".mid", "", dt_vel_diff$Stimuli))
+dt_vel_diff$Articulation2 <- "NA"
+dt_vel_diff$Dynamics2 <- "NA"
+dt_vel_diff[Stimuli < 5]$Articulation2 <- "Present"
+dt_vel_diff[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
+dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
+dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_vel_diff[Stimuli > 12]$Articulation2 <- "Absent"
+dt_vel_diff[Stimuli > 12]$Dynamics2 <- "Absent"
 ```
@@
 # for each individual
-ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Stimuli, Articulation, Dynamics)]
-ioi <- ioi[order(SubNr, Stimuli, Articulation, Dynamics)]
+ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2)]
+
+ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2)]
 ioi
@@
 ```{r ioi-plot, echo = FALSE, fig.width = 4.5}
-ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+ggboxplot(ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation2, Dynamics2)], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
 ```
@@
 # group mean
-ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
-ioi_all <- ioi_all[order(Articulation, Dynamics)]
+ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2)]
 ioi_all
@@
 ```{r ioi-all-plot, echo = FALSE}
-ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+ggboxplot(ioi, "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+```
+
+# Articulation
+## 1. Individual
+```{r kot, echo = FALSE}
+# for each individual
+kot_trial <- dt_kot[Articulation == "Legato" | Articulation == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Articulation)]
+
+kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Articulation)]
+kot
+```
+
+### Plot
+#### Boxplot
+
+```{r kot-plot,  echo = FALSE, fig.width = 4.5}
+ggboxplot(kot_trial[Articulation == "Legato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+
+ggboxplot(kot_trial[Articulation == "Staccato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
+```
+
+## 2. All - group
+```{r kot-all, echo = FALSE}
+# group mean
+kot_all <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Articulation)]
+kot_all
+```
+
+### Plot
+#### Boxplot
+
+```{r kot-all-plot, echo = FALSE}
+ggboxplot(kot[Articulation == "Legato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+
+ggboxplot(kot[Articulation == "Staccato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
 ```
@@
 # for each individual
-vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
-vel <- vel[order(SubNr, Stimuli, Articulation, Dynamics)]
+vel_trial <- dt_vel[Dynamics != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
+
+vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Dynamics)]
 vel
@@
 ```{r vel-plot, echo = FALSE, fig.width = 4.5}
-ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
-ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
+ggboxplot(vel_trial[Dynamics == "Forte"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
+
+ggboxplot(vel_trial[Dynamics == "Piano"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
 ```
@@
 # group mean
-vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
-vel_all <- vel_all[order(Articulation, Dynamics, Subcomponent)]
+vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Dynamics)]
 vel_all
@@
 ```{r vel-all-plot,  echo = FALSE}
-ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
-ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
+ggboxplot(vel[Dynamics == "Forte"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
+
+ggboxplot(vel[Dynamics == "Piano"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
 ```
 
-# KV-Diff
+# Dynamics Difference
 ## 1. Individual
-
 ```{r vel-diff, echo = FALSE}
 # for each individual
-vel_diff <- dt_vel_diff[Subcomponent == "LFtoSP" | Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
-vel_diff <- vel_diff[order(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
+vel_diff_trial <- dt_vel_diff[Dynamics == "FtoP" | Dynamics == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
+
+vel_diff <- vel_diff_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Dynamics)]
 vel_diff
@@
 ```{r vel-diff-plot, echo = FALSE, fig.width = 4.5}
-ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (LFtoSP)")
-ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (SPtoLF)")
+ggboxplot(vel_diff_trial[Dynamics == "FtoP"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (FtoP)")
+
+ggboxplot(vel_diff_trial[Dynamics == "PtoF"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (PtoF)")
 ```
@@
 # group mean
-vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
-vel_diff_all <- vel_diff_all[order(Articulation, Dynamics, Subcomponent)]
+vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Dynamics)]
 vel_diff_all
@@
 ```{r vel-diff-all-plot,  echo = FALSE}
-ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (LF to SP)")
-ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (SP to LF)")
+ggboxplot(vel_diff[Dynamics == "FtoP"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")
+
+ggboxplot(vel_diff[Dynamics == "PtoF"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
 ```

### 2022-09-15 analysis/primary.Rmd
add sequence plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r setup, include = FALSE}
+# set chunk option
+knitr::opts_chunk$set(echo = FALSE)
+
 # packages
@@
 dt_ioi$Dynamics2 <- "NA"
+dt_ioi$Category <- "NA"
 dt_ioi[Stimuli < 5]$Articulation2 <- "Present"
 dt_ioi[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_ioi[Stimuli < 5]$Category <- "art_only"
 dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
 dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_ioi[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
 dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
 dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_ioi[Stimuli > 8 & Stimuli < 13]$Category <- "both"
 dt_ioi[Stimuli > 12]$Articulation2 <- "Absent"
 dt_ioi[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_ioi[Stimuli > 12]$Category <- "none"
 
@@
 dt_kot$Dynamics2 <- "NA"
+dt_kot$Category <- "NA"
 dt_kot[Stimuli < 5]$Articulation2 <- "Present"
 dt_kot[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_kot[Stimuli < 5]$Category <- "art_only"
 dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
 dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_kot[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
 dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
 dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_kot[Stimuli > 8 & Stimuli < 13]$Category <- "both"
 dt_kot[Stimuli > 12]$Articulation2 <- "Absent"
 dt_kot[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_kot[Stimuli > 12]$Category <- "none"
 
@@
 dt_vel$Dynamics2 <- "NA"
+dt_vel$Category <- "NA"
 dt_vel[Stimuli < 5]$Articulation2 <- "Present"
 dt_vel[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_vel[Stimuli < 5]$Category <- "art_only"
 dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
 dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_vel[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
 dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
 dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_vel[Stimuli > 8 & Stimuli < 13]$Category <- "both"
 dt_vel[Stimuli > 12]$Articulation2 <- "Absent"
 dt_vel[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_vel[Stimuli > 12]$Category <- "none"
 
@@
 dt_vel_diff$Dynamics2 <- "NA"
+dt_vel_diff$Category <- "NA"
 dt_vel_diff[Stimuli < 5]$Articulation2 <- "Present"
 dt_vel_diff[Stimuli < 5]$Dynamics2 <- "Absent"
+dt_vel_diff[Stimuli < 5]$Category <- "art_only"
 dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
 dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
+dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
 dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
 dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
+dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Category <- "both"
 dt_vel_diff[Stimuli > 12]$Articulation2 <- "Absent"
 dt_vel_diff[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_vel_diff[Stimuli > 12]$Category <- "none"
 ```
@@
 
-```{r ioi, echo = FALSE}
+```{r ioi}
 # for each individual
@@
 
-```{r ioi-plot, echo = FALSE, fig.width = 4.5}
+```{r ioi-plot, fig.width = 4.5}
 ggboxplot(ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation2, Dynamics2)], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
@@
 
-```{r ioi-all, echo = FALSE}
+```{r ioi-all}
 # group mean
@@
 
-```{r ioi-all-plot, echo = FALSE}
+```{r ioi-all-plot}
 ggboxplot(ioi, "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+```
+
+## 3. Sequence plots
+
+```{r seq-ioi, fig.width = 12, fig.height = 2.4}
+ioi_seq <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
+
+ggline(ioi_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI") + scale_x_continuous(breaks = seq(1,71,1))
 ```
@@
 ## 1. Individual
+
 ```{r kot, echo = FALSE}
@@
 ## 2. All - group
+
 ```{r kot-all, echo = FALSE}
@@
 
+## 3. Sequence plots
+
+```{r seq-kot, fig.width = 12, fig.height = 2.4}
+kot_seq <- dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
+
+ggline(kot_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "KOT (ms)", title = "KOT") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
 # Dynamics
 ## 1. Individual
+
 ```{r vel, echo = FALSE}
@@
 ## 2. All - group
+
 ```{r vel-all, echo = FALSE}
@@
 
+## 3. Sequence plots
+
+```{r seq-vel, fig.width = 12, fig.height = 2.4}
+vel_seq <- dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, Articulation, Dynamics, RowNr)]
+
+ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Velocity (0-127)", title = "Velocity") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
 # Dynamics Difference
 ## 1. Individual
+
 ```{r vel-diff, echo = FALSE}
@@
 ```
+
+## 3. Sequence plots
+
+```{r seq-vel-diff, fig.width = 12, fig.height = 2.4}
+vel_diff_seq <- dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
+
+ggline(vel_diff_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "Difference", title = "Velocity Difference") + scale_x_continuous(breaks = seq(1,71,1))
+```

### 2022-09-16 analysis/primary.Rmd
add stats / change the fig size
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# stats
+if (!require("afex")) {install.packages("afex"); require("afex")}
 # plot
@@
 
-```{r seq-ioi, fig.width = 12, fig.height = 2.4}
+```{r seq-ioi, fig.width = 7, fig.height = 2}
 ioi_seq <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
@@
 ggline(ioi_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+```{r}
+aov_ez(
+  data = ioi,
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
 ```
@@
 
-```{r seq-kot, fig.width = 12, fig.height = 2.4}
+```{r seq-kot, fig.width = 7, fig.height = 2}
 kot_seq <- dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
@@
 ggline(kot_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "KOT (ms)", title = "KOT") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+1) Legato
+
+```{r}
+aov_ez(
+  data = kot[Articulation == "Legato"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
+```
+
+2) Staccato
+
+```{r}
+aov_ez(
+  data = kot[Articulation == "Staccato"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
 ```
@@
 
-```{r seq-vel, fig.width = 12, fig.height = 2.4}
+```{r seq-vel, fig.width = 7, fig.height = 2}
 vel_seq <- dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, Articulation, Dynamics, RowNr)]
@@
 ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Velocity (0-127)", title = "Velocity") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+1) Forte
+
+```{r}
+aov_ez(
+  data = vel[Dynamics == "Forte"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
+```
+
+2) Piano
+
+```{r}
+aov_ez(
+  data = vel[Dynamics == "Piano"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
 ```
@@
 
-```{r seq-vel-diff, fig.width = 12, fig.height = 2.4}
+```{r seq-vel-diff, fig.width = 7, fig.height = 2}
 vel_diff_seq <- dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
@@
 ```
+
+## 4. Stats
+
+1) FtoP
+
+```{r}
+aov_ez(
+  data = vel_diff[Dynamics == "FtoP"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
+```
+
+2) PtoF
+
+```{r}
+aov_ez(
+  data = vel_diff[Dynamics == "PtoF"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation2", "Dynamics2")
+)
+```

### 2022-09-16 analysis/firstInstance.Rmd
add analysis only with the first instance
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-09-16 analysis/primary.Rmd
@@
 ---
-title: 'Primary analysis'
+title: 'Analysis only with the first instance'
 output:
@@
 
-Description: This is a summary of primary analysis.
+Description: This is a summary of analysis which only include the performance with the first instance of each student (category).
 
@@
 
+```{r, include = FALSE}
+info <- dt_ioi[, .(N = .N), by = .(SubNr, TrialNr, Stimuli, Category)]
+info
+
+dt_ioi_first <- data.table()
+for (subnr in unique(info$SubNr)){
+  for (category in unique(info[SubNr == subnr]$Category)){
+    trials <- unique(info[SubNr == subnr & Category == category]$TrialNr)
+    current <- dt_ioi[SubNr == subnr & Category == category & TrialNr == min(trials)]
+    dt_ioi_first <- rbind(dt_ioi_first, current)
+  }
+}
+
+dt_kot_first <- data.table()
+for (subnr in unique(info$SubNr)){
+  for (category in unique(info[SubNr == subnr]$Category)){
+    trials <- unique(info[SubNr == subnr & Category == category]$TrialNr)
+    current <- dt_kot[SubNr == subnr & Category == category & TrialNr == min(trials)]
+    dt_kot_first <- rbind(dt_kot_first, current)
+  }
+}
+
+dt_vel_first <- data.table()
+for (subnr in unique(info$SubNr)){
+  for (category in unique(info[SubNr == subnr]$Category)){
+    trials <- unique(info[SubNr == subnr & Category == category]$TrialNr)
+    current <- dt_vel[SubNr == subnr & Category == category & TrialNr == min(trials)]
+    dt_vel_first <- rbind(dt_vel_first, current)
+  }
+}
+
+dt_vel_diff_first <- data.table()
+for (subnr in unique(info$SubNr)){
+  for (category in unique(info[SubNr == subnr]$Category)){
+    trials <- unique(info[SubNr == subnr & Category == category]$TrialNr)
+    current <- dt_vel_diff[SubNr == subnr & Category == category & TrialNr == min(trials)]
+    dt_vel_diff_first <- rbind(dt_vel_diff_first, current)
+  }
+}
+```
+
 # IOI 
@@
 # for each individual
-ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2)]
+ioi_trial <- dt_ioi_first[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2)]
 
@@
 # for each individual
-kot_trial <- dt_kot[Articulation == "Legato" | Articulation == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Articulation)]
+kot_trial <- dt_kot_first[Articulation == "Legato" | Articulation == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Articulation)]
 
@@
 # for each individual
-vel_trial <- dt_vel[Dynamics != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
+vel_trial <- dt_vel_first[Dynamics != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
 
@@
 # for each individual
-vel_diff_trial <- dt_vel_diff[Dynamics == "FtoP" | Dynamics == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
+vel_diff_trial <- dt_vel_diff_first[Dynamics == "FtoP" | Dynamics == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
 

### 2022-09-21 analysis/preprosessor/predata/filtering.Rmd
update predata
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-08-19 analysis/filtering.Rmd
@@
 # read a text file for ideal performance
-dt_ideal <- read.table(here("ideal.txt"))
+dt_ideal <- read.table(here("../ideal.txt"))
 colnames(dt_ideal) <- "Pitch"
@@
 # create a list of data file names
-lf <- list.files(here("raw_data"), pattern = "txt")
+lf <- list.files(here("../raw_data"), pattern = "txt")
 
@@
 for (i in 1:length(lf)){
-  data_i <- fread(file.path(here("raw_data"), lf[i]), header = F, sep = " ", dec = ".")
+  data_i <- fread(file.path(here("../raw_data"), lf[i]), header = F, sep = " ", dec = ".")
   raw_data <- rbind(raw_data, data_i)
@@
 # onset and offset for experimental session
-dt_onset <- dt_all[Key_OnOff == 1 & Session == "experiment"]
-dt_offset <- dt_all[Key_OnOff == 0 & Session == "experiment"]
+dt_onset <- dt_all[Key_OnOff == 1 & Session == "pre"]
+dt_offset <- dt_all[Key_OnOff == 0 & Session == "pre"]
 ```
@@
 error_sub_onset <- dt_error_onset[startsWith(Reason, "Substituted")]
-
 dt_correct_onset_4 <- data.table()
-for (row in 1:nrow(error_sub_onset)){
-  current <- dt_onset[SubNr == error_sub_onset$SubNr[row] & TrialNr == error_sub_onset$TrialNr[row]]
-  decision = 2 
-  correction = 0 # # of correction for reporting stats
-  while (decision == 2){
-    print(sprintf("SubNr: %s, TrialNr: %s", unique(current$SubNr), unique(current$TrialNr)))
-    print("----- First check -----")
-    current <- edit(current, dt_ideal)
-    print("----- Correction check -----")
-    edit(current, dt_ideal)
-    decision <- menu(c("y", "n", "other"), title = "Save the current data? (to continue, enter 'n')")
-    if (decision == 1){
-      correction = correction + 1
-      error_sub_onset$CorrectionNr[row] <- correction
-      dt_correct_onset_4 <- rbind(dt_correct_onset_4, current[, -c(5:6)])
-    } else if (decision == 3) {
-      error_sub_onset$CorrectionNr[row] <- readline(prompt = "Reason?: ")
-    } else if (decision == 2){
-      correction = correction + 1
-      print("----- Continue correction -----")
-    }
-  }
-}
 
@@
 ```{r}
-error_sub_offset <- fread(here("filtered/error_sub_offset.txt"))
+error_sub_offset <- fread(here("filtered", "error_sub_offset.txt"))
 error_sub_offset
@@
 
-```{r}
+```{r, include = FALSE}
 # combine error
@@
 error_ind_onset <- error_onset[startsWith(CorrectionNr, "Check")]
-error_ind_onset
 ```
 
-- SubNr: 12, TrialNr: 13
+- SubNr: 20, TrialNr: 2
 
-```{r}
+```{r, include = FALSE, eval = FALSE}
 current_1 <- dt_onset[SubNr == error_ind_onset$SubNr[1] & TrialNr == error_ind_onset$TrialNr[1]]
@@
 current_1 <- edit(current_1, dt_ideal)
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
 graph(current_1)
 
-# corrected the second error (substituted note)
+# remove the second error
 current_1 <- edit(current_1, dt_ideal)
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- dt_ideal$Pitch
 graph(current_1)
 
-error_ind_onset$CorrectionNr[1] <- 2
+# correct two notes
+current_1 <- edit(current_1, dt_ideal)
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- dt_ideal$Pitch
+graph(current_1)
+
+error_ind_onset$CorrectionNr[1] <- 4
 ```
 
-- SubNr: 10, TrialNr: 10
+- SubNr: 20, TrialNr: 1
+Excluded because the participant stopped playing in the middle.
+
+```{r, include = FALSE, eval = FALSE}
+error_ind_onset$CorrectionNr[2] <- "Excluded"
+```
+
+```{r, include = FALSE, eval = FALSE}
+dt_correct_onset_5 <- current_1
+dt_correct_onset_5 <- dt_correct_onset_5[, -c(5:6)]
+
+# export csv
+fwrite(dt_correct_onset_5, file = here("filtered", "dt_correct_onset_5.txt"))
+fwrite(error_ind_onset, file = here("filtered", "error_ind_onset.txt"))
+```
 
 ```{r}
-current_2 <- dt_onset[SubNr == error_ind_onset$SubNr[2] & TrialNr == error_ind_onset$TrialNr[2]]
-current_2 <- insert_na(current_2, dt_ideal)
-current_2$RowNr <- c(1:nrow(current_2))
-current_2$Ideal <- dt_ideal$Pitch
-graph(current_2)
-
-# insert one skipped note
-current_2 <- edit(current_2, dt_ideal)
-graph(current_2)
-
-# remove an extra note
-current_2 <- edit(current_2, dt_ideal)
-graph(current_2)
-
-error_ind_onset$CorrectionNr[2] <- 2
-```
-
-- SubNr
-
-```{r}
-
+error_ind_onset <- fread(here("filtered", "error_ind_onset.txt"))
+error_ind_onset
 ```
@@
 error_ind_offset <- error_offset[startsWith(CorrectionNr, "Check")]
+```
+
+- SubNr: 11, TrialNr: 1
+
+```{r, include = FALSE, eval = FALSE}
+current_1 <- dt_offset[SubNr == error_ind_offset$SubNr[1] & TrialNr == error_ind_offset$TrialNr[1]]
+current_1$RowNr <- c(1:nrow(current_1))
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# correct two notes
+current_1 <- edit(current_1, dt_ideal)
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# remove one note
+current_1 <- edit(current_1, dt_ideal)
+current_1$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_1) - nrow(dt_ideal))))
+graph(current_1)
+
+# remove one note
+current_1 <- edit(current_1, dt_ideal)
+current_1$Ideal <- dt_ideal$Pitch
+graph(current_1)
+
+error_ind_offset$CorrectionNr[1] <- 4
+```
+
+- SubNr: 20, TrialNr: 2
+
+```{r, include = FALSE, eval = FALSE}
+current_2 <- dt_offset[SubNr == error_ind_offset$SubNr[2] & TrialNr == error_ind_offset$TrialNr[2]]
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_2) - nrow(dt_ideal))))
+graph(current_2)
+
+# remove one note
+current_2 <- edit(current_2, dt_ideal)
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- c(dt_ideal$Pitch, rep(NA, abs(nrow(current_2) - nrow(dt_ideal))))
+graph(current_2)
+
+# remove one note
+current_2 <- edit(current_2, dt_ideal)
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+# correct two notes
+current_2 <- edit(current_2, dt_ideal)
+current_2$RowNr <- c(1:nrow(current_2))
+current_2$Ideal <- dt_ideal$Pitch
+graph(current_2)
+
+error_ind_offset$CorrectionNr[2] <- 4
+```
+
+- SubNr: 20, TrialNr: 1
+Excluded because the participant stopped playing in the middle.
+
+```{r, include = FALSE, eval = FALSE}
+error_ind_offset$CorrectionNr[3] <- "Excluded"
+```
+
+```{r, include = FALSE, eval = FALSE}
+dt_correct_offset_5 <- rbind(current_1, current_2)
+dt_correct_offset_5 <- dt_correct_offset_5[, -c(5:6)]
+
+# export csv
+fwrite(dt_correct_offset_5, file = here("filtered", "dt_correct_offset_5.txt"))
+fwrite(error_ind_offset, file = here("filtered", "error_ind_offset.txt"))
+```
+
+```{r}
+error_ind_offset <- fread(here("filtered", "error_ind_offset.txt"))
 error_ind_offset
 ```
+
+```{r, include = FALSE, eval = FALSE}
+dt_correct_onset_1 <- fread(here("filtered", "dt_correct_onset_1.txt"))
+dt_correct_onset_2 <- fread(here("filtered", "dt_correct_onset_2.txt"))
+dt_correct_onset_3 <- fread(here("filtered", "dt_correct_onset_3.txt"))
+dt_correct_onset_4 <- fread(here("filtered", "dt_correct_onset_4.txt"))
+dt_correct_onset_5 <- fread(here("filtered", "dt_correct_onset_5.txt"))
+
+# combine all
+dt_correct_onset <- rbind(dt_correct_onset_1, dt_correct_onset_2, dt_correct_onset_3, dt_correct_onset_4, dt_correct_onset_5)
+
+# export csv
+fwrite(dt_correct_onset, file = here("filtered", "dt_correct_onset.txt"))
+
+dt_correct_offset_1 <- fread(here("filtered", "dt_correct_offset_1.txt"))
+dt_correct_offset_2 <- fread(here("filtered", "dt_correct_offset_2.txt"))
+dt_correct_offset_3 <- fread(here("filtered", "dt_correct_offset_3.txt"))
+dt_correct_offset_4 <- fread(here("filtered", "dt_correct_offset_4.txt"))
+dt_correct_offset_5 <- fread(here("filtered", "dt_correct_offset_5.txt"))
+
+# combine all
+dt_correct_offset <- rbind(dt_correct_offset_1, dt_correct_offset_2, dt_correct_offset_3, dt_correct_offset_4, dt_correct_offset_5)
+
+# export csv
+fwrite(dt_correct_offset, file = here("filtered", "dt_correct_offset.txt"))
+```
+
+# Checking
+Check whether data look okay.
+
+```{r, include = FALSE}
+rm(list=ls(all=T)) # clear all
+
+# set working directory
+if (!require("here")) {install.packages("here"); require("here")}
+here::i_am("filtering.Rmd")
+
+# read functions
+source(here("function.R"))
+
+# read a text file for ideal performance
+dt_ideal <- read.table(here("../ideal.txt"))
+colnames(dt_ideal) <- "Pitch"
+dt_ideal$RowNr <- 1:nrow(dt_ideal)
+setcolorder(dt_ideal, c(2, 1))
+
+# read txt files
+dt_correct_onset <- fread(here("filtered", "dt_correct_onset.txt"))
+dt_correct_offset <- fread(here("filtered", "dt_correct_offset.txt"))
+```
+
+1) Onset
+```{r}
+print(checker(dt_correct_onset, dt_ideal))
+```
+
+2) Offset
+```{r}
+print(checker(dt_correct_offset, dt_ideal))
+```

### 2022-09-21 analysis/preprosessor/trimming.Rmd
change variable labels to Subcomponent1/2
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-09-15 analysis/trimming.Rmd
@@
 # assign Articulation/Dynamics
-dt_ioi$Articulation <- "NA"
-dt_ioi$Dynamics <- "NA"
+dt_ioi$Subcomponent1 <- "NA"
+dt_ioi$Subcomponent2 <- "NA"
 # Legato/Forte
@@
   for (note in 1:length(ls_legato[[phrase]])){
-    dt_ioi[Interval == ls_legato[[phrase]][note]]$Articulation <- "Legato"
-    dt_ioi[Interval == ls_legato[[phrase]][note]]$Dynamics <- "Forte"
+    dt_ioi[Interval == ls_legato[[phrase]][note]]$Subcomponent1 <- "Legato"
+    dt_ioi[Interval == ls_legato[[phrase]][note]]$Subcomponent2 <- "Forte"
   }
@@
   for (note in 1:length(ls_staccato[[phrase]])){
-    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Articulation <- "Staccato"
-    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Dynamics <- "Piano"
+    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Subcomponent1 <- "Staccato"
+    dt_ioi[Interval == ls_staccato[[phrase]][note]]$Subcomponent2 <- "Piano"
   }
@@
 for (number in change_1){
-  dt_ioi[Interval == number]$Articulation <- "LtoS"
-  dt_ioi[ Interval == number]$Dynamics <- "FtoP"
+  dt_ioi[Interval == number]$Subcomponent1 <- "LtoS"
+  dt_ioi[ Interval == number]$Subcomponent2 <- "FtoP"
 }
 for (number in change_2){
-  dt_ioi[Interval == number]$Articulation <- "StoL"
-  dt_ioi[Interval == number]$Dynamics <- "PtoF"
+  dt_ioi[Interval == number]$Subcomponent1 <- "StoL"
+  dt_ioi[Interval == number]$Subcomponent2 <- "PtoF"
 }
@@
 ```{r}
-dt_ioi_subset <- dt_ioi[Articulation != "NA" & !is.na(IOI)]
+dt_ioi_subset <- dt_ioi[Subcomponent1 != "NA" & !is.na(IOI)]
 
-gghistogram(dt_ioi_subset, x = "IOI", color = "Articulation", bins = 150)
+gghistogram(dt_ioi_subset, x = "IOI", color = "Subcomponent1", bins = 150)
 ```
@@
 dt_ioi_subset$Boundary <- "No"
-dt_ioi_subset[Articulation == "LtoS" | Articulation == "StoL"]$Boundary <- "Yes"
+dt_ioi_subset[Subcomponent1 == "LtoS" | Subcomponent1 == "StoL"]$Boundary <- "Yes"
 
@@
 
-gghistogram(dt_ioi_trimmed, color = "Articulation", x = "IOI", bins = 150)
+gghistogram(dt_ioi_trimmed, color = "Subcomponent1", x = "IOI", bins = 150)
 ```
@@
 # KOT
-```{r}
+```{r, include = FALSE}
 dt_kot <- dt_onset
 
-# Offset 1 - Onset 2
+# calculate KOT (Offset 1 - Onset 2)
 dt_kot$KOT <- NA
@@
 # assign Articulation/Dynamics
-dt_kot$Articulation <- "NA"
-dt_kot$Dynamics <- "NA"
+dt_kot$Subcomponent1 <- "NA"
+dt_kot$Subcomponent2 <- "NA"
 # Legato/Forte
@@
   for (note in 1:length(ls_legato[[phrase]])){
-    dt_kot[Interval == ls_legato[[phrase]][note]]$Articulation <- "Legato"
-    dt_kot[Interval == ls_legato[[phrase]][note]]$Dynamics <- "Forte"
+    dt_kot[Interval == ls_legato[[phrase]][note]]$Subcomponent1 <- "Legato"
+    dt_kot[Interval == ls_legato[[phrase]][note]]$Subcomponent2 <- "Forte"
   }
@@
   for (note in 1:length(ls_staccato[[phrase]])){
-    dt_kot[Interval == ls_staccato[[phrase]][note]]$Articulation <- "Staccato"
-    dt_kot[Interval == ls_staccato[[phrase]][note]]$Dynamics <- "Piano"
+    dt_kot[Interval == ls_staccato[[phrase]][note]]$Subcomponent1 <- "Staccato"
+    dt_kot[Interval == ls_staccato[[phrase]][note]]$Subcomponent2 <- "Piano"
   }
@@
 for (number in change_1){
-  dt_kot[Interval == number]$Articulation <- "LtoS"
-  dt_kot[ Interval == number]$Dynamics <- "FtoP"
+  dt_kot[Interval == number]$Subcomponent1 <- "LtoS"
+  dt_kot[ Interval == number]$Subcomponent2 <- "FtoP"
 }
 for (number in change_2){
-  dt_kot[Interval == number]$Articulation <- "StoL"
-  dt_kot[Interval == number]$Dynamics <- "PtoF"
+  dt_kot[Interval == number]$Subcomponent1 <- "StoL"
+  dt_kot[Interval == number]$Subcomponent2 <- "PtoF"
 }
@@
 ```{r}
-dt_kot_subset <- dt_kot[Articulation != "NA" & !is.na(KOT)]
+dt_kot_subset <- dt_kot[Subcomponent1 != "NA" & !is.na(KOT)]
 
-gghistogram(dt_kot_subset, x = "KOT", color = "Articulation", bins = 150)
+gghistogram(dt_kot_subset, x = "KOT", color = "Subcomponent1", bins = 150)
 ```
@@
 ```{r}
-kot_subcomponent <- dt_kot_subset[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = Articulation]
+kot_subcomponent <- dt_kot_subset[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = Subcomponent1]
 dt_kot_trimmed <- data.table()
-for (subcomponent in unique(dt_kot_subset$Articulation)){
-  upper <- kot_subcomponent[Articulation == subcomponent]$Mean+3*kot_subcomponent[Articulation == subcomponent]$SD
-  lower <- kot_subcomponent[Articulation == subcomponent]$Mean-3*kot_subcomponent[Articulation == subcomponent]$SD
-  dt_current <- dt_kot_subset[Articulation == subcomponent & KOT < upper & KOT > lower]
+for (subcomponent in unique(dt_kot_subset$Subcomponent1)){
+  upper <- kot_subcomponent[Subcomponent1 == subcomponent]$Mean+3*kot_subcomponent[Subcomponent1 == subcomponent]$SD
+  lower <- kot_subcomponent[Subcomponent1 == subcomponent]$Mean-3*kot_subcomponent[Subcomponent1 == subcomponent]$SD
+  dt_current <- dt_kot_subset[Subcomponent1 == subcomponent & KOT < upper & KOT > lower]
   dt_kot_trimmed <- rbind(dt_kot_trimmed, dt_current)
@@
 
-gghistogram(dt_kot_trimmed, x = "KOT", color = "Articulation", bins = 150)
+gghistogram(dt_kot_trimmed, x = "KOT", color = "Subcomponent1", bins = 150)
 ```
+
+Remove `r nrow(dt_kot_subset)-nrow(dt_kot_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_kot_subset)-nrow(dt_kot_trimmed))/nrow(dt_kot_subset))` % of total responses
 
@@
 # KV
+```{r, include = FALSE}
+# calculate Diff (acceleration - velocity difference between notes)
+dt_vel <- dt_onset
+dt_vel$Diff <- diff(c(0, dt_vel$Velocity))
+
+# remove the first note
+dt_vel_diff <- dt_vel[RowNr != 1]
+dt_vel$Diff <- NULL # remove Diff from dt_vel
+
+# assign Subcomponents
+# for each note
+dt_vel$Subcomponent1 <- "NA"
+dt_vel$Subcomponent2 <- "NA"
+# Legato/Forte
+for (phrase in 1:length(ls_legato2)){
+  for (note in 1:length(ls_legato2[[phrase]])){
+    dt_vel[RowNr == ls_legato2[[phrase]][note]]$Subcomponent1 <- "Legato"
+    dt_vel[RowNr  == ls_legato2[[phrase]][note]]$Subcomponent2 <- "Forte"
+  }
+}
+# Staccato/Piano
+for (phrase in 1:length(ls_staccato2)){
+  for (note in 1:length(ls_staccato2[[phrase]])){
+    dt_vel[RowNr  == ls_staccato2[[phrase]][note]]$Subcomponent1 <- "Staccato"
+    dt_vel[RowNr  == ls_staccato2[[phrase]][note]]$Subcomponent2 <- "Piano"
+  }
+}
+```
+
+## Histogram
+### Before removing outliers
+```{r}
+dt_vel_subset <- dt_vel[Subcomponent2 != "NA" & !is.na(Velocity)]
+
+gghistogram(dt_vel_subset, x = "Velocity", color = "Subcomponent2", bins = 150)
+```
+
+### After removing outliers
+```{r}
+vel_subcomponent <- dt_vel_subset[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = Subcomponent2]
+dt_vel_trimmed <- data.table()
+for (subcomponent in unique(dt_vel_subset$Subcomponent2)){
+  upper <- vel_subcomponent[Subcomponent2 == subcomponent]$Mean+3*vel_subcomponent[Subcomponent2 == subcomponent]$SD
+  lower <- vel_subcomponent[Subcomponent2 == subcomponent]$Mean-3*vel_subcomponent[Subcomponent2 == subcomponent]$SD
+  dt_current <- dt_vel_subset[Subcomponent2 == subcomponent & Velocity < upper & Velocity > lower]
+  dt_vel_trimmed <- rbind(dt_vel_trimmed, dt_current)
+}
+
+gghistogram(dt_vel_trimmed, x = "Velocity", color = "Subcomponent2", bins = 150)
+```
+
+Remove `r nrow(dt_vel_subset)-nrow(dt_vel_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_vel_subset)-nrow(dt_vel_trimmed))/nrow(dt_vel_subset))` % of total responses
+
+```{r, include = FALSE}
+fwrite(dt_vel_trimmed, file = here("trimmed", "dt_vel_trimmed.txt"), row.names = F)
+```
 
 # KV Difference
+```{r, include = FALSE}
+# assign Interval
+dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)
+
+# assign Subcomponents
+# for intervals
+dt_vel_diff$Subcomponent1 <- "NA"
+dt_vel_diff$Subcomponent2 <- "NA"
+# Legato/Forte
+for (phrase in 1:length(ls_legato)){
+  for (note in 1:length(ls_legato[[phrase]])){
+    dt_vel_diff[Interval == ls_legato[[phrase]][note]]$Subcomponent1 <- "Legato"
+    dt_vel_diff[Interval == ls_legato[[phrase]][note]]$Subcomponent2 <- "Forte"
+  }
+}
+# Staccato/Piano
+for (phrase in 1:length(ls_staccato)){
+  for (note in 1:length(ls_staccato[[phrase]])){
+    dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$Subcomponent1 <- "Staccato"
+    dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$Subcomponent2 <- "Piano"
+  }
+}
+
+# assign Subcomponent Change
+for (number in change_1){
+  dt_vel_diff[Interval == number]$Subcomponent1 <- "LtoS"
+  dt_vel_diff[ Interval == number]$Subcomponent2 <- "FtoP"
+}
+for (number in change_2){
+  dt_vel_diff[Interval == number]$Subcomponent1 <- "StoL"
+  dt_vel_diff[Interval == number]$Subcomponent2 <- "PtoF"
+}
+
+```
+
+## Histogram
+### Before removing outliers
+```{r}
+dt_vel_diff_subset <- dt_vel_diff[Subcomponent2 != "NA" & !is.na(Diff)]
+
+gghistogram(dt_vel_diff_subset, x = "Diff", color = "Subcomponent2", bins = 150)
+```
+
+### After removing outliers
+```{r}
+vel_diff_subcomponent <- dt_vel_diff_subset[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = Subcomponent2]
+dt_vel_diff_trimmed <- data.table()
+for (subcomponent in unique(dt_vel_diff_subset$Subcomponent2)){
+  upper <- vel_diff_subcomponent[Subcomponent2 == subcomponent]$Mean+3*vel_diff_subcomponent[Subcomponent2 == subcomponent]$SD
+  lower <- vel_diff_subcomponent[Subcomponent2 == subcomponent]$Mean-3*vel_diff_subcomponent[Subcomponent2 == subcomponent]$SD
+  dt_current <- dt_vel_diff_subset[Subcomponent2 == subcomponent & Diff < upper & Diff > lower]
+  dt_vel_diff_trimmed <- rbind(dt_vel_diff_trimmed, dt_current)
+}
+
+gghistogram(dt_vel_diff_trimmed, x = "Diff", color = "Subcomponent2", bins = 150)
+```
+
+Remove `r nrow(dt_vel_diff_subset)-nrow(dt_vel_diff_trimmed)` responses beyond +- 3SD / `r 100*((nrow(dt_vel_diff_subset)-nrow(dt_vel_diff_trimmed))/nrow(dt_vel_diff_subset))` % of total responses
+
+```{r, include = FALSE}
+fwrite(dt_vel_diff_trimmed, file = here("trimmed", "dt_vel_diff_trimmed.txt"), row.names = F)
+```

### 2022-09-21 analysis/baseline.Rmd
upload comparison between baseline and test performance
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Comparison between baseline and test performance'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of primary analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

```{r setup, include = FALSE}
# set chunk option
knitr::opts_chunk$set(echo = FALSE)

# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# stats
if (!require("afex")) {install.packages("afex"); require("afex")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r file, include = FALSE}
data_ioi = "preprosessor/trimmed/dt_ioi_trimmed.txt"
data_kot = "preprosessor/trimmed/dt_kot_trimmed.txt"
data_vel = "preprosessor/trimmed/dt_vel_trimmed.txt"
data_vel_diff = "preprosessor/trimmed/dt_vel_diff_trimmed.txt"
data_baseline_ioi = "preprosessor/predata/trimmed/dt_ioi_trimmed.txt"
data_baseline_kot = "preprosessor/predata/trimmed/dt_kot_trimmed.txt"
data_baseline_vel = "preprosessor/predata/trimmed/dt_vel_trimmed.txt"
data_baseline_vel_diff = "preprosessor/predata/trimmed/dt_vel_diff_trimmed.txt"
```

```{r prep, include = FALSE}
dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
dt_vel <- fread(data_vel, header = T, sep = ",", dec = ".")
dt_vel_diff <- fread(data_vel_diff, header = T, sep = ",", dec = ".")

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# assign stimuli categories
dt_ioi$Stimuli <- as.numeric(gsub(".mid", "", dt_ioi$Stimuli))
dt_ioi$Articulation <- "NA"
dt_ioi$Dynamics <- "NA"
dt_ioi$Category <- "NA"
dt_ioi[Stimuli < 5]$Articulation <- "Present"
dt_ioi[Stimuli < 5]$Dynamics <- "Absent"
dt_ioi[Stimuli < 5]$Category <- "art_only"
dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
dt_ioi[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
dt_ioi[Stimuli > 8 & Stimuli < 13]$Category <- "both"
dt_ioi[Stimuli > 12]$Articulation <- "Absent"
dt_ioi[Stimuli > 12]$Dynamics <- "Absent"
dt_ioi[Stimuli > 12]$Category <- "none"

dt_kot$Stimuli <- as.numeric(gsub(".mid", "", dt_kot$Stimuli))
dt_kot$Articulation <- "NA"
dt_kot$Dynamics <- "NA"
dt_kot$Category <- "NA"
dt_kot[Stimuli < 5]$Articulation <- "Present"
dt_kot[Stimuli < 5]$Dynamics <- "Absent"
dt_kot[Stimuli < 5]$Category <- "art_only"
dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
dt_kot[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
dt_kot[Stimuli > 8 & Stimuli < 13]$Category <- "both"
dt_kot[Stimuli > 12]$Articulation <- "Absent"
dt_kot[Stimuli > 12]$Dynamics <- "Absent"
dt_kot[Stimuli > 12]$Category <- "none"

dt_vel$Stimuli <- as.numeric(gsub(".mid", "", dt_vel$Stimuli))
dt_vel$Articulation <- "NA"
dt_vel$Dynamics <- "NA"
dt_vel$Category <- "NA"
dt_vel[Stimuli < 5]$Articulation <- "Present"
dt_vel[Stimuli < 5]$Dynamics <- "Absent"
dt_vel[Stimuli < 5]$Category <- "art_only"
dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
dt_vel[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
dt_vel[Stimuli > 8 & Stimuli < 13]$Category <- "both"
dt_vel[Stimuli > 12]$Articulation <- "Absent"
dt_vel[Stimuli > 12]$Dynamics <- "Absent"
dt_vel[Stimuli > 12]$Category <- "none"

dt_vel_diff$Stimuli <- as.numeric(gsub(".mid", "", dt_vel_diff$Stimuli))
dt_vel_diff$Articulation <- "NA"
dt_vel_diff$Dynamics <- "NA"
dt_vel_diff$Category <- "NA"
dt_vel_diff[Stimuli < 5]$Articulation <- "Present"
dt_vel_diff[Stimuli < 5]$Dynamics <- "Absent"
dt_vel_diff[Stimuli < 5]$Category <- "art_only"
dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Category <- "both"
dt_vel_diff[Stimuli > 12]$Articulation <- "Absent"
dt_vel_diff[Stimuli > 12]$Dynamics <- "Absent"
dt_vel_diff[Stimuli > 12]$Category <- "none"
```

```{r, include = FALSE}
# combine data with baseline performance
dt_bl_ioi <- fread(data_baseline_ioi, header = T, sep = ",", dec = ".")
dt_bl_kot <- fread(data_baseline_kot, header = T, sep = ",", dec = ".")
dt_bl_vel <- fread(data_baseline_vel, header = T, sep = ",", dec = ".")
dt_bl_vel_diff <- fread(data_baseline_vel_diff, header = T, sep = ",", dec = ".")

dt_bl_ioi$Articulation <- "Baseline"
dt_bl_kot$Articulation <- "Baseline"
dt_bl_vel$Articulation <- "Baseline"
dt_bl_vel_diff$Articulation <- "Baseline"

dt_bl_ioi$Dynamics <- "Baseline"
dt_bl_kot$Dynamics <- "Baseline"
dt_bl_vel$Dynamics <- "Baseline"
dt_bl_vel_diff$Dynamics <- "Baseline"

dt_bl_ioi$Category <- "baseline"
dt_bl_kot$Category <- "baseline"
dt_bl_vel$Category <- "baseline"
dt_bl_vel_diff$Category <- "baseline"

dt_ioi <- rbind(dt_ioi, dt_bl_ioi)
dt_kot <- rbind(dt_kot, dt_bl_kot)
dt_vel <- rbind(dt_vel, dt_bl_vel)
dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
```

# IOI 
## 1. Individual
```{r ioi}
# for each individual
ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]

ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category)]
setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
ioi
```

### Plot
#### Boxplot
```{r ioi-plot, fig.width = 6}
ggboxplot(ioi_trial, "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

## 2. All - group

```{r ioi-all}
# group mean
ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category)]
ioi_all
```

### Plot
#### Box plot

```{r ioi-all-plot}
ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

## 3. Sequence plots

```{r seq-ioi, fig.width = 7, fig.height = 2}
ioi_seq <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Interval)]

ggline(ioi_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI") + scale_x_continuous(breaks = seq(1,71,1))
```

## 4. Stats

```{r}
aov_ez(
  data = ioi,
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

# Articulation
## 1. Individual

```{r kot, echo = FALSE}
# for each individual
kot_trial <- dt_kot[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]

kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent1)]
setorder(kot, "SubNr", "Articulation", "Dynamics", "Category")
kot
```

### Plot
#### Boxplot

```{r kot-plot,  echo = FALSE, fig.width = 6}
ggboxplot(kot_trial[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")

ggboxplot(kot_trial[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

## 2. All - group

```{r kot-all, echo = FALSE}
# group mean
kot_all <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent1)]
kot_all
```

### Plot
#### Boxplot

```{r kot-all-plot, echo = FALSE}
ggboxplot(kot[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")

ggboxplot(kot[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

## 3. Sequence plots

```{r seq-kot, fig.width = 7, fig.height = 2}
kot_seq <- dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Interval)]

ggline(kot_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "KOT (ms)", title = "KOT") + scale_x_continuous(breaks = seq(1,71,1))
```

## 4. Stats

1) Legato

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Legato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

2) Staccato

```{r}
aov_ez(
  data = kot[Subcomponent1 == "Staccato"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

# Dynamics
## 1. Individual

```{r vel, echo = FALSE}
# for each individual
vel_trial <- dt_vel[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]

vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
setorder(vel, "SubNr", "Dynamics", "Articulation", "Category")
vel
```

### Plot
#### Boxplot

```{r vel-plot, echo = FALSE, fig.width = 6}
ggboxplot(vel_trial[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")

ggboxplot(vel_trial[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
```

## 2. All - group

```{r vel-all, echo = FALSE}
# group mean
vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent2)]
vel_all
```

### Plot
#### Boxplot

```{r vel-all-plot,  echo = FALSE}
ggboxplot(vel[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")

ggboxplot(vel[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

## 3. Sequence plots

```{r seq-vel, fig.width = 7, fig.height = 2}
vel_seq <- dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, RowNr)]

ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Velocity (0-127)", title = "Velocity") + scale_x_continuous(breaks = seq(1,72,1))
```

## 4. Stats

1) Forte

```{r}
aov_ez(
  data = vel[Subcomponent2 == "Forte"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

2) Piano

```{r}
aov_ez(
  data = vel[Subcomponent2 == "Piano"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

# Dynamics Difference
## 1. Individual

```{r vel-diff, echo = FALSE}
# for each individual
vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]

vel_diff <- vel_diff_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
setorder(vel_diff, "SubNr", "Dynamics", "Articulation", "Category")
vel_diff
```

### Plot
#### Boxplot

```{r vel-diff-plot, echo = FALSE, fig.width = 6}
ggboxplot(vel_diff_trial[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (FtoP)")

ggboxplot(vel_diff_trial[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (PtoF)")
```

## 2. All - group
```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent2)]
vel_diff_all
```

### Plot
#### Boxplot

```{r vel-diff-all-plot,  echo = FALSE}
ggboxplot(vel_diff[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")

ggboxplot(vel_diff[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
```

## 3. Sequence plots

```{r seq-vel-diff, fig.width = 7, fig.height = 2}
vel_diff_seq <- dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Interval)]

ggline(vel_diff_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "Difference", title = "Velocity Difference") + scale_x_continuous(breaks = seq(1,71,1))
```

## 4. Stats

1) FtoP

```{r}
aov_ez(
  data = vel_diff[Subcomponent2 == "FtoP"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

2) PtoF

```{r}
aov_ez(
  data = vel_diff[Subcomponent2 == "PtoF"],
  id = "SubNr",
  dv = "Mean",
  within = c("Category")
)
```

### 2022-09-21 analysis/primary.Rmd
change variable labels
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-09-15 analysis/primary.Rmd
@@
 dt_ioi$Stimuli <- as.numeric(gsub(".mid", "", dt_ioi$Stimuli))
-dt_ioi$Articulation2 <- "NA"
-dt_ioi$Dynamics2 <- "NA"
-dt_ioi[Stimuli < 5]$Articulation2 <- "Present"
-dt_ioi[Stimuli < 5]$Dynamics2 <- "Absent"
-dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
-dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
-dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
-dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
-dt_ioi[Stimuli > 12]$Articulation2 <- "Absent"
-dt_ioi[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_ioi$Articulation <- "NA"
+dt_ioi$Dynamics <- "NA"
+dt_ioi$Category <- "NA"
+dt_ioi[Stimuli < 5]$Articulation <- "Present"
+dt_ioi[Stimuli < 5]$Dynamics <- "Absent"
+dt_ioi[Stimuli < 5]$Category <- "art_only"
+dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
+dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
+dt_ioi[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
+dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
+dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
+dt_ioi[Stimuli > 8 & Stimuli < 13]$Category <- "both"
+dt_ioi[Stimuli > 12]$Articulation <- "Absent"
+dt_ioi[Stimuli > 12]$Dynamics <- "Absent"
+dt_ioi[Stimuli > 12]$Category <- "none"
 
 dt_kot$Stimuli <- as.numeric(gsub(".mid", "", dt_kot$Stimuli))
-dt_kot$Articulation2 <- "NA"
-dt_kot$Dynamics2 <- "NA"
-dt_kot[Stimuli < 5]$Articulation2 <- "Present"
-dt_kot[Stimuli < 5]$Dynamics2 <- "Absent"
-dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
-dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
-dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
-dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
-dt_kot[Stimuli > 12]$Articulation2 <- "Absent"
-dt_kot[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_kot$Articulation <- "NA"
+dt_kot$Dynamics <- "NA"
+dt_kot$Category <- "NA"
+dt_kot[Stimuli < 5]$Articulation <- "Present"
+dt_kot[Stimuli < 5]$Dynamics <- "Absent"
+dt_kot[Stimuli < 5]$Category <- "art_only"
+dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
+dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
+dt_kot[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
+dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
+dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
+dt_kot[Stimuli > 8 & Stimuli < 13]$Category <- "both"
+dt_kot[Stimuli > 12]$Articulation <- "Absent"
+dt_kot[Stimuli > 12]$Dynamics <- "Absent"
+dt_kot[Stimuli > 12]$Category <- "none"
 
 dt_vel$Stimuli <- as.numeric(gsub(".mid", "", dt_vel$Stimuli))
-dt_vel$Articulation2 <- "NA"
-dt_vel$Dynamics2 <- "NA"
-dt_vel[Stimuli < 5]$Articulation2 <- "Present"
-dt_vel[Stimuli < 5]$Dynamics2 <- "Absent"
-dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
-dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
-dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
-dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
-dt_vel[Stimuli > 12]$Articulation2 <- "Absent"
-dt_vel[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_vel$Articulation <- "NA"
+dt_vel$Dynamics <- "NA"
+dt_vel$Category <- "NA"
+dt_vel[Stimuli < 5]$Articulation <- "Present"
+dt_vel[Stimuli < 5]$Dynamics <- "Absent"
+dt_vel[Stimuli < 5]$Category <- "art_only"
+dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
+dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
+dt_vel[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
+dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
+dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
+dt_vel[Stimuli > 8 & Stimuli < 13]$Category <- "both"
+dt_vel[Stimuli > 12]$Articulation <- "Absent"
+dt_vel[Stimuli > 12]$Dynamics <- "Absent"
+dt_vel[Stimuli > 12]$Category <- "none"
 
 dt_vel_diff$Stimuli <- as.numeric(gsub(".mid", "", dt_vel_diff$Stimuli))
-dt_vel_diff$Articulation2 <- "NA"
-dt_vel_diff$Dynamics2 <- "NA"
-dt_vel_diff[Stimuli < 5]$Articulation2 <- "Present"
-dt_vel_diff[Stimuli < 5]$Dynamics2 <- "Absent"
-dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation2 <- "Absent"
-dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics2 <- "Present"
-dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation2 <- "Present"
-dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics2 <- "Present"
-dt_vel_diff[Stimuli > 12]$Articulation2 <- "Absent"
-dt_vel_diff[Stimuli > 12]$Dynamics2 <- "Absent"
+dt_vel_diff$Articulation <- "NA"
+dt_vel_diff$Dynamics <- "NA"
+dt_vel_diff$Category <- "NA"
+dt_vel_diff[Stimuli < 5]$Articulation <- "Present"
+dt_vel_diff[Stimuli < 5]$Dynamics <- "Absent"
+dt_vel_diff[Stimuli < 5]$Category <- "art_only"
+dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
+dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
+dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
+dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
+dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
+dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Category <- "both"
+dt_vel_diff[Stimuli > 12]$Articulation <- "Absent"
+dt_vel_diff[Stimuli > 12]$Dynamics <- "Absent"
+dt_vel_diff[Stimuli > 12]$Category <- "none"
 ```
@@
 
-```{r ioi, echo = FALSE}
+```{r ioi}
 # for each individual
-ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2)]
+ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
 
-ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2)]
+ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category)]
+setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
 ioi
@@
 
-```{r ioi-plot, echo = FALSE, fig.width = 4.5}
-ggboxplot(ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation2, Dynamics2)], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+```{r ioi-plot, fig.width = 6}
+ggboxplot(ioi_trial, "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
 ```
@@
 
-```{r ioi-all, echo = FALSE}
+```{r ioi-all}
 # group mean
-ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2)]
+ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
 ioi_all
@@
 
-```{r ioi-all-plot, echo = FALSE}
-ggboxplot(ioi, "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+```{r ioi-all-plot}
+ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+```
+
+## 3. Sequence plots
+
+```{r seq-ioi, fig.width = 7, fig.height = 2}
+ioi_seq <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Interval)]
+
+ggline(ioi_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+```{r}
+aov_ez(
+  data = ioi,
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
 ```
@@
 ## 1. Individual
+
 ```{r kot, echo = FALSE}
 # for each individual
-kot_trial <- dt_kot[Articulation == "Legato" | Articulation == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Articulation)]
+kot_trial <- dt_kot[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]
 
-kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Articulation)]
+kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent1)]
+setorder(kot, "SubNr", "Subcomponent1", "Articulation", "Dynamics", "Category")
 kot
@@
 
-```{r kot-plot,  echo = FALSE, fig.width = 4.5}
-ggboxplot(kot_trial[Articulation == "Legato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+```{r kot-plot,  echo = FALSE, fig.width = 6}
+ggboxplot(kot_trial[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
 
-ggboxplot(kot_trial[Articulation == "Staccato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
+ggboxplot(kot_trial[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
 ```
@@
 ## 2. All - group
+
 ```{r kot-all, echo = FALSE}
 # group mean
-kot_all <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Articulation)]
+kot_all <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent1)]
 kot_all
@@
 ```{r kot-all-plot, echo = FALSE}
-ggboxplot(kot[Articulation == "Legato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+ggboxplot(kot[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
 
-ggboxplot(kot[Articulation == "Staccato"], "Articulation2", "Mean", color = "Dynamics2", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
+ggboxplot(kot[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
+```
+
+## 3. Sequence plots
+
+```{r seq-kot, fig.width = 7, fig.height = 2}
+kot_seq <- dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Articulation, Dynamics, Interval)]
+
+ggline(kot_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "KOT (ms)", title = "KOT") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+1) Legato
+
+```{r}
+aov_ez(
+  data = kot[Subcomponent1 == "Legato"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+
+2) Staccato
+
+```{r}
+aov_ez(
+  data = kot[Subcomponent1 == "Staccato"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
 ```
@@
 ## 1. Individual
+
 ```{r vel, echo = FALSE}
 # for each individual
-vel_trial <- dt_vel[Dynamics != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
+vel_trial <- dt_vel[Dynamics != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
 
-vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Dynamics)]
+vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
+setorder(vel, "SubNr", "Subcomponent2", "Dynamics", "Articulation", "Category")
 vel
@@
 
-```{r vel-plot, echo = FALSE, fig.width = 4.5}
-ggboxplot(vel_trial[Dynamics == "Forte"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
+```{r vel-plot, echo = FALSE, fig.width = 6}
+ggboxplot(vel_trial[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
 
-ggboxplot(vel_trial[Dynamics == "Piano"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
+ggboxplot(vel_trial[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
 ```
@@
 ## 2. All - group
+
 ```{r vel-all, echo = FALSE}
 # group mean
-vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Dynamics)]
+vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent2)]
 vel_all
@@
 ```{r vel-all-plot,  echo = FALSE}
-ggboxplot(vel[Dynamics == "Forte"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
+ggboxplot(vel[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
 
-ggboxplot(vel[Dynamics == "Piano"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
+ggboxplot(vel[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
+```
+
+## 3. Sequence plots
+
+```{r seq-vel, fig.width = 7, fig.height = 2}
+vel_seq <- dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, RowNr)]
+
+ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Velocity (0-127)", title = "Velocity") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+1) Forte
+
+```{r}
+aov_ez(
+  data = vel[Subcomponent2 == "Forte"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+
+2) Piano
+
+```{r}
+aov_ez(
+  data = vel[Subcomponent2 == "Piano"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
 ```
@@
 ## 1. Individual
+
 ```{r vel-diff, echo = FALSE}
 # for each individual
-vel_diff_trial <- dt_vel_diff[Dynamics == "FtoP" | Dynamics == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation2, Dynamics2, Dynamics)]
+vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
 
-vel_diff <- vel_diff_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation2, Dynamics2, Dynamics)]
+vel_diff <- vel_diff_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
+setorder(vel_diff, "SubNr", "Subcomponent2", "Dynamics", "Articulation", "Category")
 vel_diff
@@
 
-```{r vel-diff-plot, echo = FALSE, fig.width = 4.5}
-ggboxplot(vel_diff_trial[Dynamics == "FtoP"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (FtoP)")
+```{r vel-diff-plot, echo = FALSE, fig.width = 6}
+ggboxplot(vel_diff_trial[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (FtoP)")
 
-ggboxplot(vel_diff_trial[Dynamics == "PtoF"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (PtoF)")
+ggboxplot(vel_diff_trial[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (PtoF)")
 ```
@@
 # group mean
-vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation2, Dynamics2, Dynamics)]
+vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category, Subcomponent2)]
 vel_diff_all
@@
 ```{r vel-diff-all-plot,  echo = FALSE}
-ggboxplot(vel_diff[Dynamics == "FtoP"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")
+ggboxplot(vel_diff[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")
 
-ggboxplot(vel_diff[Dynamics == "PtoF"], "Dynamics2", "Mean", color = "Articulation2", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
+ggboxplot(vel_diff[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
 ```
+
+## 3. Sequence plots
+
+```{r seq-vel-diff, fig.width = 7, fig.height = 2}
+vel_diff_seq <- dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Interval)]
+
+ggline(vel_diff_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Interval", ylab = "Difference", title = "Velocity Difference") + scale_x_continuous(breaks = seq(1,71,1))
+```
+
+## 4. Stats
+
+1) FtoP
+
+```{r}
+aov_ez(
+  data = vel_diff[Subcomponent2 == "FtoP"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+
+2) PtoF
+
+```{r}
+aov_ez(
+  data = vel_diff[Subcomponent2 == "PtoF"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```

### 2022-09-21 analysis/both_none.Rmd
upload analysis to compare both and none categories
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-09-21 analysis/baseline.Rmd
@@
 ---
-title: 'Comparison between baseline and test performance'
+title: 'Comparison between both and none categories'
 output:
@@
 
-Description: This is a summary of primary analysis.
+Description: This is a summary of analysis comparing between both and none categories.
 
@@
 data_vel_diff = "preprosessor/trimmed/dt_vel_diff_trimmed.txt"
-data_baseline_ioi = "preprosessor/predata/trimmed/dt_ioi_trimmed.txt"
-data_baseline_kot = "preprosessor/predata/trimmed/dt_kot_trimmed.txt"
-data_baseline_vel = "preprosessor/predata/trimmed/dt_vel_trimmed.txt"
-data_baseline_vel_diff = "preprosessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 ```
@@
 
-```{r, include = FALSE}
-# combine data with baseline performance
-dt_bl_ioi <- fread(data_baseline_ioi, header = T, sep = ",", dec = ".")
-dt_bl_kot <- fread(data_baseline_kot, header = T, sep = ",", dec = ".")
-dt_bl_vel <- fread(data_baseline_vel, header = T, sep = ",", dec = ".")
-dt_bl_vel_diff <- fread(data_baseline_vel_diff, header = T, sep = ",", dec = ".")
-
-dt_bl_ioi$Articulation <- "Baseline"
-dt_bl_kot$Articulation <- "Baseline"
-dt_bl_vel$Articulation <- "Baseline"
-dt_bl_vel_diff$Articulation <- "Baseline"
-
-dt_bl_ioi$Dynamics <- "Baseline"
-dt_bl_kot$Dynamics <- "Baseline"
-dt_bl_vel$Dynamics <- "Baseline"
-dt_bl_vel_diff$Dynamics <- "Baseline"
-
-dt_bl_ioi$Category <- "baseline"
-dt_bl_kot$Category <- "baseline"
-dt_bl_vel$Category <- "baseline"
-dt_bl_vel_diff$Category <- "baseline"
-
-dt_ioi <- rbind(dt_ioi, dt_bl_ioi)
-dt_kot <- rbind(dt_kot, dt_bl_kot)
-dt_vel <- rbind(dt_vel, dt_bl_vel)
-dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
-```
-
 # IOI 
@@
 # for each individual
-ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
+ioi_trial <- dt_ioi[Category == "none" | Category == "both", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
 
@@
 ```{r seq-ioi, fig.width = 7, fig.height = 2}
-ioi_seq <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Interval)]
+ioi_seq <- dt_ioi[Category == "none" | Category == "both", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Interval)]
 
@@
 )
+
 ```
@@
 # for each individual
-kot_trial <- dt_kot[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]
+dt_subset_kot <- dt_kot[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato"]
+kot_trial <- dt_subset_kot[Category == "none" | Category == "both", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]
 
@@
 ```{r seq-kot, fig.width = 7, fig.height = 2}
-kot_seq <- dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Interval)]
+kot_seq <- dt_kot[Category == "none" | Category == "both", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Interval)]
 
@@
 # for each individual
-vel_trial <- dt_vel[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
+dt_subset_vel <- dt_vel[Subcomponent2 != "NA"]
+vel_trial <- dt_subset_vel[Category == "none" | Category == "both", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
 
@@
 ```{r seq-vel, fig.width = 7, fig.height = 2}
-vel_seq <- dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, RowNr)]
+vel_seq <- dt_vel[Category == "none" | Category == "both", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, RowNr)]
 
@@
 # for each individual
-vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
+dt_subset_vel_diff <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF"]
+vel_diff_trial <- dt_subset_vel_diff[Category == "none" | Category == "both", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
 
@@
 ```{r seq-vel-diff, fig.width = 7, fig.height = 2}
-vel_diff_seq <- dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Interval)]
+vel_diff_seq <- dt_vel_diff[Category == "none" | Category == "both", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Interval)]
 

### 2022-09-25 paper/manuscript.Rmd
update a template for paper
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Atsuko Tominaga"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

# Methods

## Participants

## Material

## Procedure

## Data analysis

# Results

# Discussion

### 2022-09-25 paper/manuscript.tex
update a template for paper
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={The title},
  pdfauthor={Atsuko Tominaga1 \& Ernst-August Doelle1,2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The title}
\author{Atsuko Tominaga\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
\date{}

\shorttitle{Title}

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. Atsuko Tominaga: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing, Supervision.

Correspondence concerning this article should be addressed to Atsuko Tominaga, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}

\abstract{%
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{material}{%
\subsection{Material}\label{material}}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\end{document}

### 2022-09-25 material/expstimuli/audio/original/averaging/low/old/20_21/average.R
.gitignore fix
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,163 +1,0 @@
-# /usr/local/bin/r
-# Average onset, duration, key velocity for normative performance
-# Created: 06/05/2020
-
-# packages
-# data manipulation
-if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-
-# function
-source("./function.R")
-
-# create a folder to store stimuli
-foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
-dir.create(foldername)
-
-# read csv/txt
-dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
-dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
-#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
-#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
-duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
-kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
-dt_ideal <- fread("./ideal.txt", header = F)
-
-# duration and kv
-valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
-valid_du_kv$Duplicated <- duplicated(valid_du_kv)
-
-# valid performances
-valid <- valid_du_kv[Duplicated == TRUE]
-
-### create 8 instances! ###
-# random sampling
-valid$Sample <- sample(c(1:31), replace = FALSE)
-print(valid)
-fwrite(valid, paste(foldername, "valid.txt", sep = ""))
-
-# 1. average IOIs
-dt_ioi_instance  <- data.table()
-counter = 0
-for (i in 1:8){
-  stim <- combining_onset(dt_onset, valid, i)
-  
-  # calculate normIOI
-  stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
-  stim$normIOI <- stim$IOI/stim$Tempo
-  # IOI = 0 if RowNr == 1
-  stim[RowNr == 1]$normIOI <- 0
-  
[…]

### 2022-09-25 material/expstimuli/audio/original/averaging/low/old/20_21/function.R

@@ -1,47 +1,0 @@
-# averaging function (create 8 instances - averaging 3 performances)
-# packages
-if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-
-# valid: information(data.table) about valid performances
-# number: loop number (if outside loop, enter 0)
-
-combining_onset <- function(dt_onset, valid, number){
-  SubNr1 <- valid[Sample == number+counter]$SubNr
-  SubNr2 <- valid[Sample == number+1+counter]$SubNr
-  SubNr3 <- valid[Sample == number+2+counter]$SubNr
-  TrialNr1 <- valid[Sample == number+counter]$TrialNr
-  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
-  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
-  stim_1 <- dt_onset[SubNr == SubNr1 & TrialNr == TrialNr1]
-  stim_2 <- dt_onset[SubNr == SubNr2 & TrialNr == TrialNr2]
-  stim_3 <- dt_onset[SubNr == SubNr3 & TrialNr == TrialNr3]
-  outcome <- rbind(stim_1, stim_2, stim_3)
-  outcome$RowNr <- rep(c(1:72), 3)
-  
-  # tempo labels
-  outcome[Tempo == 120]$Tempo <- 250
-  outcome[Tempo == 110]$Tempo <- 273
-  outcome[Tempo == 100]$Tempo <- 300
-  return(outcome)
-}
-
-combining_offset <- function(dt_offset, valid, number){
-  SubNr1 <- valid[Sample == number+counter]$SubNr
-  SubNr2 <- valid[Sample == number+1+counter]$SubNr
-  SubNr3 <- valid[Sample == number+2+counter]$SubNr
-  TrialNr1 <- valid[Sample == number+counter]$TrialNr
-  TrialNr2 <- valid[Sample == number+1+counter]$TrialNr
-  TrialNr3 <- valid[Sample == number+2+counter]$TrialNr
-  stim_1 <- dt_offset[SubNr == SubNr1 & TrialNr == TrialNr1]
-  stim_2 <- dt_offset[SubNr == SubNr2 & TrialNr == TrialNr2]
-  stim_3 <- dt_offset[SubNr == SubNr3 & TrialNr == TrialNr3]
-  outcome <- rbind(stim_1, stim_2, stim_3)
[…]

### 2022-09-25 material/expstimuli/audio/original/averaging/low/old/original/average.R

@@ -1,148 +1,0 @@
-# /usr/local/bin/r
-# Average onset, duration, key velocity for normative performance
-# Created: 06/05/2020
-
-# packages
-# data manipulation
-if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-
-# function
-source("./function.R")
-
-# create a folder to store stimuli
-foldername = paste(format(Sys.time(), "%s-%d%m%y"), "/", sep = "") # current time
-dir.create(foldername)
-
-# read csv/txt
-dt_onset <- fread("../analysis/stim_n/preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".")
-dt_offset <- fread("../analysis/stim_n/preprocessor/filtered/data_offset.csv", header = T, sep = ",", dec = ".")
-#onset_valid <- fread("../analysis/stim_n/onset_valid.csv", header = T, sep = ",", dec = ".")
-#offset_valid <- fread("../analysis/stim_n/offset_valid.csv", header = T, sep = ",", dec = ".")
-duration_valid <- fread("../analysis/stim_n/duration_valid.csv", header = T, sep = ",", dec = ".")
-kv_valid <- fread("../analysis/stim_n/kv_valid.csv", header = T, sep = ",", dec = ".")
-dt_ideal <- fread("./ideal.txt", header = F)
-
-# duration and kv
-valid_du_kv <- rbind(duration_valid[, c("SubNr", "TrialNr")], kv_valid[, c("SubNr", "TrialNr")])
-valid_du_kv$Duplicated <- duplicated(valid_du_kv)
-
-# valid performances
-valid <- valid_du_kv[Duplicated == TRUE]
-
-### create 8 instances! ###
-# random sampling
-valid$Sample <- sample(c(1:31), replace = FALSE)
-print(valid)
-fwrite(valid, paste(foldername, "valid.txt", sep = ""))
-
-# 1. average IOIs
-dt_ioi_instance  <- data.table()
-counter = 0
-for (i in 1:8){
-  stim <- combining_onset(dt_onset, valid, i)
-  
-  # calculate normIOI
-  stim$IOI <- diff(c(0, stim$TimeStamp))# convert bpm to ms
-  stim$normIOI <- stim$IOI/stim$Tempo
-  # IOI = 0 if RowNr == 1
-  stim[RowNr == 1]$normIOI <- 0
-  
[…]

### 2022-09-26 paper/manuscript.Rmd
updated methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
-    email         : "my@email.com"
+    address       : "Quellenstraße 51, 1100 Vienna, Austria"
+    email         : "tominaga_atsuko@phd.ceu.edu"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
       - "Writing - Review & Editing"
-  - name          : "Ernst-August Doelle"
-    affiliation   : "1,2"
+  - name          : "Günther Knoblich"
+    affiliation   : "1"
+    role:
+      - "Writing - Review & Editing"
+      - "Supervision"
+  - name          : "Natalie Sebanz"
+    affiliation   : "1"
     role:
@@
   - id            : "1"
-    institution   : "Wilhelm-Wundt-University"
-  - id            : "2"
-    institution   : "Konstanz Business School"
+    institution   : "Department of Cognitive Science, Central European University"
 
@@
 
-# Methods
+## Methods
+
+<!-- ```{r methods, include = FALSE} -->
+<!-- source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE) -->
+<!-- ``` -->
 
@@
 
-## Material
+20 participants were recruited who ---.
+Most participants were right-handed (left: ---).
+The mean age of the participants was --- (*SD* = ---).
+They had --- years of practice on average.
+All participants were recruited through an online participant platform (SONA system, https://www.sona-systems.com).
+The study (No.
+---) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
+
+## Apparatus and stimuli
+
+The experiment was programmed in Max/MSP (---; https://cycling74.com/products/max) on a Mac Book Pro with Mac OS X Catalina 10.15.6.
+A weighted Yamaha MIDI digital piano was used to record participants' performance.
+All auditory feedback was given to participants through headphones (Audio-Technica ATH-M50X).
+Sheet music was displayed on a computer monitor in front of the participants.
+
+Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) because it is relatively simple in terms of motor skills.
+The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
+The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
+Original sheet music was used for the purpose of practice (*Fig A*).
+Two expressive notations were added to the original sheet music for the experiment (*Fig B*).
+These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
+The fingering was also assigned and confirmed by the same doctoral student.
 
@@
 
+Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and performed it without pitch errors while implementing notated expressions in the experiment.
+
 ## Data analysis
+
+In each performance (trial), the pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers. We quantified performances in terms of tempo,
 

### 2022-09-26 paper/manuscript.tex
updated methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   pdftitle={The title},
-  pdfauthor={Atsuko Tominaga1 \& Ernst-August Doelle1,2},
+  pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
   pdflang={en-EN},
@@
 \title{The title}
-\author{Atsuko Tominaga\textsuperscript{1} \& Ernst-August Doelle\textsuperscript{1,2}}
+\author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
 \date{}
@@
 The authors made the following contributions.
-Atsuko Tominaga: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing, Supervision.
+Atsuko Tominaga: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Günther Knoblich: Writing - Review \& Editing, Supervision; Natalie Sebanz: Writing - Review \& Editing, Supervision.
 
-Correspondence concerning this article should be addressed to Atsuko Tominaga, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
+Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria.
+E-mail: \href{mailto:tominaga_atsuko@phd.ceu.edu}{\nolinkurl{tominaga\_atsuko@phd.ceu.edu}}
 
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Wilhelm-Wundt-University\\\textsuperscript{2} Konstanz Business School}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Cognitive Science, Central European University}
 
@@
 \hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\subsection{Methods}\label{methods}}
 
@@
 
-\hypertarget{material}{%
-\subsection{Material}\label{material}}
+20 participants were recruited who ---.
+Most participants were right-handed (left: ---).
+The mean age of the participants was --- (\emph{SD} = ---).
+They had --- years of practice on average.
+All participants were recruited through an online participant platform (SONA system, \url{https://www.sona-systems.com}).
+The study (No.~---) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
+
+\hypertarget{apparatus-and-stimuli}{%
+\subsection{Apparatus and stimuli}\label{apparatus-and-stimuli}}
+
+The experiment was programmed in Max/MSP (---; \url{https://cycling74.com/products/max}) on a Mac Book Pro with Mac OS X Catalina 10.15.6.
+A weighted Yamaha MIDI digital piano was used to record participants' performance.
+All auditory feedback was given to participants through headphones (Audio-Technica ATH-M50X).
+Sheet music was displayed on a computer monitor in front of the participants.
+
+Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) because it is relatively simple in terms of motor skills.
+The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
+The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
+Original sheet music was used for the purpose of practice (\emph{Fig A}).
+Two expressive notations were added to the original sheet music for the experiment (\emph{Fig B}).
+These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
+The fingering was also assigned and confirmed by the same doctoral student.
 
@@
 
+Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and performed it without pitch errors while implementing notated expressions in the experiment.
+
 \hypertarget{data-analysis}{%
 \subsection{Data analysis}\label{data-analysis}}
+
+In each performance (trial), the pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers. We quantified performances in terms of tempo,
 

### 2022-09-26 analysis/demographics/questionnaire.Rmd
update demographic information
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Demographics'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
dt <- dt[-c(1,2),] # exclude test/pilot data

# change some colnames
colnames(dt)[2] <- "SubNr"
colnames(dt)[3] <- "Age"
colnames(dt)[4] <- "Gender"
colnames(dt)[7] <- "Handedness"
colnames(dt)[15] <- "PianoTotalPractice"
colnames(dt)[25] <- "TeachingPiano" # there are only people who have taught the piano
colnames(dt)[27] <- "TeachingPianoYears"

dt$TeachingPianoYears <- c(7, 1, 12, 5, 30, NA, 11, 2, 5, 3, 6, NA, 3, 6, 8, 9, 3, NA, 5, 3)

# change some characteristics
dt$Age <- as.numeric(dt$Age)
dt$PianoTotalPractice <- as.numeric(dt$PianoTotalPractice)

# exclude participants
dt_included <- dt

print(dt_included)
```

## Descriptive stats for included participants
Note: A different questionnaire was used for only SubNr 21 and adjusted for necessary columns (but not others; see original csv for the precise information)

- Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
- Male: `r nrow(dt_included[Gender == "Male"])` (mean age: `r round(mean(dt_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_included[Gender == "Female"])` (mean age: `r round(mean(dt_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_included$PianoTotalPractice), 2)` years (sd: `r round(sd(dt_included$PianoTotalPractice), 2)`)
- Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)
- Teaching experience (piano): Yes (`r nrow(dt_included[TeachingPiano == "Yes"])`, mean years of teaching: `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`, sd: `r round(sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`), No (`r nrow(dt_included[TeachingPiano == "No"])`)

## Some descriptive questions

1. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

```{r export, include = FALSE}
knitr::purl("questionnaire.Rmd")
```

### 2022-09-26 analysis/demographics/questionnaire.R

@@ -1,0 +1,49 @@
+## ----setup, include = FALSE---------------------------------
+# packages
+# data manipulation
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+# plot
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+
+
+## ----file, include = FALSE----------------------------------
+filename_q = "./questionnaire.csv"
+
+
+## ----extract, echo = FALSE----------------------------------
+# read csv files
+dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
+dt <- dt[-c(1,2),] # exclude test/pilot data
+
+# change some colnames
+colnames(dt)[2] <- "SubNr"
+colnames(dt)[3] <- "Age"
+colnames(dt)[4] <- "Gender"
+colnames(dt)[7] <- "Handedness"
+colnames(dt)[15] <- "PianoTotalPractice"
+colnames(dt)[25] <- "TeachingPiano" # there are only people who have taught the piano
+colnames(dt)[27] <- "TeachingPianoYears"
+
+dt$PianoTotalYears <- c()
+
+# change some characteristics
+dt$Age <- as.numeric(dt$Age)
+dt$PianoTotalPractice <- as.numeric(dt$PianoTotalPractice)
+
+# exclude participants
+dt_included <- dt
+
+print(dt_included)
+
+
+## ----1, echo = FALSE----------------------------------------
+data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
+
+
+## ----2, echo = FALSE----------------------------------------
+data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
+
+
+## ----export, include = FALSE--------------------------------
+knitr::purl("questionnaire.Rmd")
+

### 2022-09-27 paper/manuscript.Rmd
update participants
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r setup, include = FALSE}
-library("papaja")
-r_refs("r-references.bib")
-```
-
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+```{r set-up, include = FALSE}
+if (!require("papaja")) {install.packages("papaja"): require("papaja")}
+if (!require("here")) {install.packages("here"): require("here")}
+if (!require("magick")) {install.packages("magick"): require("magick")}
 ```
@@
 
-<!-- ```{r methods, include = FALSE} -->
-<!-- source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE) -->
-<!-- ``` -->
+```{r methods, include = FALSE}
+source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE)
+source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE)
+```
 
@@
 
-20 participants were recruited who ---.
-Most participants were right-handed (left: ---).
-The mean age of the participants was --- (*SD* = ---).
-They had --- years of practice on average.
+`r nrow(dt_included)` participants were recruited who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
+Most participants were right-handed (left: `r nrow(dt_included[Handedness == "Left"])`, ambidextrous: `r nrow(dt_included[Handedness == "Both"])`).
+The mean age of the participants was `r round(mean(dt_included$Age), 2)` (*SD* = `r round(sd(dt_included$Age), 2)`).
+They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
+`r nrow(dt_included[TeachingPiano == "Yes"])` participants had teaching experience in piano (*M* = `r round( mean(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)` years, *SD* = `r round(sd(dt_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`).
 All participants were recruited through an online participant platform (SONA system, https://www.sona-systems.com).
 The study (No.
----) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
+2020_05) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
 
@@
 
-The experiment was programmed in Max/MSP (---; https://cycling74.com/products/max) on a Mac Book Pro with Mac OS X Catalina 10.15.6.
+The experiment was programmed in Max/MSP (8.1.11; https://cycling74.com/products/max) on a Mac Book Pro with Mac OS X Catalina 10.15.7.
 A weighted Yamaha MIDI digital piano was used to record participants' performance.

### 2022-09-27 paper/manuscript.tex
update participants
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-20 participants were recruited who ---.
-Most participants were right-handed (left: ---).
-The mean age of the participants was --- (\emph{SD} = ---).
-They had --- years of practice on average.
+20 participants were recruited who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
+Most participants were right-handed (left: 2, ambidextrous: 2).
+The mean age of the participants was 28.25 (\emph{SD} = 10.95).
+They had 21.55 years of practice on average (\emph{SD} = 11.59).
+17 participants had teaching experience in piano (\emph{M} = 7 years, \emph{SD} = 6.68).
 All participants were recruited through an online participant platform (SONA system, \url{https://www.sona-systems.com}).
-The study (No.~---) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
+The study (No.~2020\_05) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
 
@@
 
-The experiment was programmed in Max/MSP (---; \url{https://cycling74.com/products/max}) on a Mac Book Pro with Mac OS X Catalina 10.15.6.
+The experiment was programmed in Max/MSP (8.1.11; \url{https://cycling74.com/products/max}) on a Mac Book Pro with Mac OS X Catalina 10.15.7.
 A weighted Yamaha MIDI digital piano was used to record participants' performance.

### 2022-09-27 analysis/baseline.Rmd
correct some comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
correct typo on filename...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Description: This is a summary of primary analysis.
+Description: This is a summary of analysis comparing between baseline and test performance.
 
@@
 ```{r file, include = FALSE}
-data_ioi = "preprosessor/trimmed/dt_ioi_trimmed.txt"
-data_kot = "preprosessor/trimmed/dt_kot_trimmed.txt"
-data_vel = "preprosessor/trimmed/dt_vel_trimmed.txt"
-data_vel_diff = "preprosessor/trimmed/dt_vel_diff_trimmed.txt"
-data_baseline_ioi = "preprosessor/predata/trimmed/dt_ioi_trimmed.txt"
-data_baseline_kot = "preprosessor/predata/trimmed/dt_kot_trimmed.txt"
-data_baseline_vel = "preprosessor/predata/trimmed/dt_vel_trimmed.txt"
-data_baseline_vel_diff = "preprosessor/predata/trimmed/dt_vel_diff_trimmed.txt"
+data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
+data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
+data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
+data_vel_diff = "preprocessor/trimmed/dt_vel_diff_trimmed.txt"
+data_baseline_ioi = "preprocessor/predata/trimmed/dt_ioi_trimmed.txt"
+data_baseline_kot = "preprocessor/predata/trimmed/dt_kot_trimmed.txt"
+data_baseline_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
+data_baseline_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 ```

### 2022-09-27 analysis/both_none.Rmd
correct typo on filename...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ```{r file, include = FALSE}
-data_ioi = "preprosessor/trimmed/dt_ioi_trimmed.txt"
-data_kot = "preprosessor/trimmed/dt_kot_trimmed.txt"
-data_vel = "preprosessor/trimmed/dt_vel_trimmed.txt"
-data_vel_diff = "preprosessor/trimmed/dt_vel_diff_trimmed.txt"
+data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
+data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
+data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
+data_vel_diff = "preprocessor/trimmed/dt_vel_diff_trimmed.txt"
 ```

### 2022-09-27 analysis/preprocessor/filtering.Rmd
correct typo on filename...
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-09-14 analysis/filtering.Rmd
@@
 
-```{r onset}
+```{r onset, echo = FALSE}
 dt_error_onset <- checker(dt_onset, dt_ideal)
@@
 ```
+
+`r nrow(dt_error_onset)` trials had at least one pitch error (`r round(nrow(dt_error_onset)/(20*16)*100, 2)` %).
 
@@
 ```
+
+`r nrow(dt_error_offset)` trials had at least one pitch error (`r round(nrow(dt_error_offset)/(20*16)*100, 2)` %).
 
@@
 
-```{r}
+```{r, echo = FALSE}
 error_extra_onset <- fread(here("filtered/error_extra_onset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_extra_offset <- fread(here("filtered", "error_extra_offset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_missing_onset <- fread(here("filtered", "error_missing_onset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_missing_offset <- fread(here("filtered", "error_missing_offset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_sub_onset <- fread(here("filtered", "error_sub_onset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_sub_offset <- fread(here("filtered", "error_sub_offset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_ind_onset <- fread(here("filtered", "error_ind_onset.txt"))
@@
 
-```{r}
+```{r, echo = FALSE}
 # combine error
@@
 ```{r, include = FALSE, eval = FALSE}
-current_1 <- dt_onset[SubNr == error_ind_offset$SubNr[1] & TrialNr == error_ind_offset$TrialNr[1]]
+current_1 <- dt_offset[SubNr == error_ind_offset$SubNr[1] & TrialNr == error_ind_offset$TrialNr[1]]
 current_1$RowNr <- c(1:nrow(current_1))
@@
 
-```{r}
+```{r, echo = FALSE}
 error_ind_offset <- fread(here("filtered", "error_ind_offset.txt"))
@@
 
-# Checking
-Check whether data look okay.
+<!-- # Checking -->
+<!-- Check whether data look okay. -->
 
-```{r, include = FALSE}
-rm(list=ls(all=T)) # clear all
+<!-- ```{r, include = FALSE} -->
+<!-- rm(list=ls(all=T)) # clear all -->
 
-# set working directory
-if (!require("here")) {install.packages("here"); require("here")}
-here::i_am("filtering.Rmd")
+<!-- # set working directory -->
+<!-- if (!require("here")) {install.packages("here"); require("here")} -->
+<!-- here::i_am("filtering.Rmd") -->
 
-# read functions
-source(here("function.R"))
+<!-- # read functions -->
+<!-- source(here("function.R")) -->
 
-# read a text file for ideal performance
-dt_ideal <- read.table(here("ideal.txt"))
-colnames(dt_ideal) <- "Pitch"
-dt_ideal$RowNr <- 1:nrow(dt_ideal)
-setcolorder(dt_ideal, c(2, 1))
+<!-- # read a text file for ideal performance -->
+<!-- dt_ideal <- read.table(here("ideal.txt")) -->
+<!-- colnames(dt_ideal) <- "Pitch" -->
+<!-- dt_ideal$RowNr <- 1:nrow(dt_ideal) -->
+<!-- setcolorder(dt_ideal, c(2, 1)) -->
 
-# read txt files
-dt_correct_onset <- fread(here("filtered", "dt_correct_onset.txt"))
-dt_correct_offset <- fread(here("filtered", "dt_correct_offset.txt"))
+<!-- # read txt files -->
+<!-- dt_correct_onset <- fread(here("filtered", "dt_correct_onset.txt")) -->
+<!-- dt_correct_offset <- fread(here("filtered", "dt_correct_offset.txt")) -->
+<!-- ``` -->
+
+<!-- 1) Onset -->
+<!-- ```{r, echo = FALSE} -->
+<!-- print(checker(dt_correct_onset, dt_ideal)) -->
+<!-- ``` -->
+
+<!-- 2) Offset -->
+<!-- ```{r, echo = FALSE} -->
+<!-- print(checker(dt_correct_offset, dt_ideal)) -->
+<!-- ``` -->
+
+```{r export, include = FALSE}
+knitr::purl("filtering.Rmd")
 ```
-
-1) Onset
-```{r}
-print(checker(dt_correct_onset, dt_ideal))
-```
-
-2) Offset
-```{r}
-print(checker(dt_correct_offset, dt_ideal))
-```

### 2022-09-27 analysis/demographics/questionnaire.R

@@ -1,16 +1,14 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-# plot
-if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 filename_q = "./questionnaire.csv"
 
 
-## ----extract, echo = FALSE----------------------------------
+## ----extract, echo = FALSE----------------------------------------------------
 # read csv files
 dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
 dt <- dt[-c(1,2),] # exclude test/pilot data
@@ -24,7 +22,7 @@ colnames(dt)[15] <- "PianoTotalPractice"
 colnames(dt)[25] <- "TeachingPiano" # there are only people who have taught the piano
 colnames(dt)[27] <- "TeachingPianoYears"
 
-dt$PianoTotalYears <- c()
+dt$TeachingPianoYears <- c(7, 1, 12, 5, 30, NA, 11, 2, 5, 3, 6, NA, 3, 6, 8, 9, 3, NA, 5, 3)
 
 # change some characteristics
 dt$Age <- as.numeric(dt$Age)
@@ -36,14 +34,10 @@ dt_included <- dt
 print(dt_included)
 
 
-## ----1, echo = FALSE----------------------------------------
-data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
-
-
-## ----2, echo = FALSE----------------------------------------
+## ----2, echo = FALSE----------------------------------------------------------
 data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
 
 
-## ----export, include = FALSE--------------------------------
+## ----export, include = FALSE--------------------------------------------------
 knitr::purl("questionnaire.Rmd")
 

### 2022-09-27 analysis/preprocessor/filtering.R

@@ -1,0 +1,850 @@
+## ----setup, include = FALSE---------------------------------------------------
+# set chunk option
+knitr::opts_chunk$set(echo = FALSE)
+
+# set working directory
+if (!require("here")) {install.packages("here"); require("here")}
+here::i_am("filtering.Rmd")
+
+# create a folder if not exists
+if (!file.exists(here("filtered"))){
+  dir.create(here("filtered"))
+}
+
+# read functions
+source(here("function.R"))
+
+# read a text file for ideal performance
+dt_ideal <- read.table(here("ideal.txt"))
+colnames(dt_ideal) <- "Pitch"
+dt_ideal$RowNr <- 1:nrow(dt_ideal)
+setcolorder(dt_ideal, c(2, 1))
+
+# create a list of data file names
+lf <- list.files(here("raw_data"), pattern = "txt")
+
+# create raw_data - merge all data files into one
+raw_data <- data.table()
+for (i in 1:length(lf)){
+  data_i <- fread(file.path(here("raw_data"), lf[i]), header = F, sep = " ", dec = ".")
+  raw_data <- rbind(raw_data, data_i)
+}
+
+# add column namesls
+colnames(raw_data) <- c("NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "SubNr", "TrialNr", "Stimuli", "Session")
+
+# clean raw_data
+raw_data$NoteNr <- as.numeric(gsub(",", "", raw_data$NoteNr))
+raw_data$Session <- gsub(";", "", raw_data$Session)
+
+# sort by SubNr, TrialNr
+raw_data <- raw_data[order(raw_data$SubNr, raw_data$TrialNr),]
+
+# raw_data without metronome
+dt_all <- raw_data[Pitch != 31 & Pitch != 34]
+
+# onset and offset for experimental session
+dt_onset <- dt_all[Key_OnOff == 1 & Session == "experiment"]
+dt_offset <- dt_all[Key_OnOff == 0 & Session == "experiment"]
+
+
+## ----onset, echo = FALSE------------------------------------------------------
+dt_error_onset <- checker(dt_onset, dt_ideal)
+
+# mark imperfect performances
+dt_onset$Error <- 0
+for (error in 1:nrow(dt_error_onset)){
[…]

### 2022-09-27 analysis/preprocessor/function.R

@@ -1,0 +1,116 @@
+# install and load required packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+if (!require("tibble")) {install.packages("tibble"); require("tibble")}
+if (!require("editData")) {install.packages("editData"); require("editData")}
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+
+# ggplot settings
+theme_set(theme_classic())
+
+### check whether a trial contains pitch errors
+# data: data of the current trial
+# ideal: dt_ideal
+checker <- function(data, ideal){
+  dt_errors <- data.table() # return data of erroneous trials - SubNr/TrialNr/Reason
+  for (subject in unique(data$SubNr)){
+    for (trial in c(1:16)){
+      current <- data[SubNr == subject & TrialNr == trial]
+      if (nrow(current) != 0){ # current data not empty
+        if (nrow(current) != nrow(ideal)){ # extra/missing note
+          if (nrow(current) > nrow(ideal)){
+            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Extra Notes")) 
+          } else if (nrow(current) < nrow(ideal)){
+            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Missing Notes")) 
+          }
+        } else if (nrow(current) == nrow(ideal)){ # substituted note
+          for (note in 1:nrow(ideal)){
+            if (current$Pitch[note] != ideal$Pitch[note]){
+              dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Substituted Notes - RowNr ", as.character(note), sep = "")))
+              break
+            }
+          }
+        }
+      } else if (nrow(current) == 0){ # current data empty
+        dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Missing Trial")))
+      }
+    }
+  }
+  if (nrow(dt_errors) != 0){
+    colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
+    return(dt_errors)
+  } else if (nrow(dt_errors) == 0){
+    return("No errors!")
+  }
+}
+
+### edit data
+# data: data of the current trial
+# ideal: dt_ideal
+edit <- function(data, ideal){
+  data$RowNr <- c(1:nrow(data))
[…]

### 2022-09-27 analysis/preprocessor/predata/function.R

@@ -1,0 +1,116 @@
+# install and load required packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+if (!require("tibble")) {install.packages("tibble"); require("tibble")}
+if (!require("editData")) {install.packages("editData"); require("editData")}
+if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
+
+# ggplot settings
+theme_set(theme_classic())
+
+### check whether a trial contains pitch errors
+# data: data of the current trial
+# ideal: dt_ideal
+checker <- function(data, ideal){
+  dt_errors <- data.table() # return data of erroneous trials - SubNr/TrialNr/Reason
+  for (subject in unique(data$SubNr)){
+    for (trial in unique(data[SubNr == subject]$TrialNr)){
+      current <- data[SubNr == subject & TrialNr == trial]
+      if (nrow(current) != 0){ # current data not empty
+        if (nrow(current) != nrow(ideal)){ # extra/missing note
+          if (nrow(current) > nrow(ideal)){
+            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Extra Notes")) 
+          } else if (nrow(current) < nrow(ideal)){
+            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Missing Notes")) 
+          }
+        } else if (nrow(current) == nrow(ideal)){ # substituted note
+          for (note in 1:nrow(ideal)){
+            if (current$Pitch[note] != ideal$Pitch[note]){
+              dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Substituted Notes - RowNr ", as.character(note), sep = "")))
+              break
+            }
+          }
+        }
+      } else if (nrow(current) == 0){ # current data empty
+        dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Missing Trial")))
+      }
+    }
+  }
+  if (nrow(dt_errors) != 0){
+    colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
+    return(dt_errors)
+  } else if (nrow(dt_errors) == 0){
+    return("No errors!")
+  }
+}
+
+### edit data
+# data: data of the current trial
+# ideal: dt_ideal
+edit <- function(data, ideal){
+  data$RowNr <- c(1:nrow(data))
[…]

### 2022-09-27 analysis/preprocessor/trimming.R

@@ -1,0 +1,299 @@
+## ----setup, include = FALSE--------------------------------
+# set chunk option
+knitr::opts_chunk$set(echo = FALSE)
+
+# set working directory
+if (!require("here")) {install.packages("here"); require("here")}
+here::i_am("trimming.Rmd")
+
+# packages
+if (!require("data.table")) {install.packages("data.table"); require("data.table")}
+if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
+
+# create a folder if not exists
+if (!file.exists(here("trimmed"))){
+  dir.create(here("trimmed"))
+}
+
+# read txt files
+dt_onset <- fread(here("filtered", "dt_correct_onset.txt"))
+dt_offset <- fread(here("filtered", "dt_correct_offset.txt"))
+
+# sort data
+setorder(dt_onset, cols = "SubNr", "TrialNr", "RowNr")
+setorder(dt_offset, cols = "SubNr", "TrialNr", "RowNr")
+
+
+## ---- include = FALSE--------------------------------------
+# define subcomponents
+# for intervals (Legato/Forte, Staccato/Piano)
+ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
+ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
+
+# for each note (velocity)
+ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
+ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
+
+# define subcomponent change (LtoS, FtoP)
+change_1 <- c(5, 17, 47)
+# define subcomponent change (StoL, PtoF)
+change_2 <- c(8, 20, 39)
+
+
+## ---- include = FALSE--------------------------------------
+# calculate IOIs
+dt_ioi <- dt_onset
+dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))
+
+# remove the first note
+dt_ioi <- dt_ioi[RowNr != 1]
+
+# assign Interval
+dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)
+
+# assign Articulation/Dynamics
+dt_ioi$Subcomponent1 <- "NA"
+dt_ioi$Subcomponent2 <- "NA"
[…]

### 2022-09-27 analysis/preprosessor/function.R

@@ -1,116 +1,0 @@
-# install and load required packages
-if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-if (!require("tibble")) {install.packages("tibble"); require("tibble")}
-if (!require("editData")) {install.packages("editData"); require("editData")}
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-
-# ggplot settings
-theme_set(theme_classic())
-
-### check whether a trial contains pitch errors
-# data: data of the current trial
-# ideal: dt_ideal
-checker <- function(data, ideal){
-  dt_errors <- data.table() # return data of erroneous trials - SubNr/TrialNr/Reason
-  for (subject in unique(data$SubNr)){
-    for (trial in c(1:16)){
-      current <- data[SubNr == subject & TrialNr == trial]
-      if (nrow(current) != 0){ # current data not empty
-        if (nrow(current) != nrow(ideal)){ # extra/missing note
-          if (nrow(current) > nrow(ideal)){
-            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Extra Notes")) 
-          } else if (nrow(current) < nrow(ideal)){
-            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Missing Notes")) 
-          }
-        } else if (nrow(current) == nrow(ideal)){ # substituted note
-          for (note in 1:nrow(ideal)){
-            if (current$Pitch[note] != ideal$Pitch[note]){
-              dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Substituted Notes - RowNr ", as.character(note), sep = "")))
-              break
-            }
-          }
-        }
-      } else if (nrow(current) == 0){ # current data empty
-        dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Missing Trial")))
-      }
-    }
-  }
-  if (nrow(dt_errors) != 0){
-    colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
-    return(dt_errors)
-  } else if (nrow(dt_errors) == 0){
-    return("No errors!")
-  }
-}
-
-### edit data
-# data: data of the current trial
-# ideal: dt_ideal
-edit <- function(data, ideal){
-  data$RowNr <- c(1:nrow(data))
[…]

### 2022-09-27 analysis/preprosessor/predata/function.R

@@ -1,116 +1,0 @@
-# install and load required packages
-if (!require("data.table")) {install.packages("data.table"); require("data.table")}
-if (!require("tibble")) {install.packages("tibble"); require("tibble")}
-if (!require("editData")) {install.packages("editData"); require("editData")}
-if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
-
-# ggplot settings
-theme_set(theme_classic())
-
-### check whether a trial contains pitch errors
-# data: data of the current trial
-# ideal: dt_ideal
-checker <- function(data, ideal){
-  dt_errors <- data.table() # return data of erroneous trials - SubNr/TrialNr/Reason
-  for (subject in unique(data$SubNr)){
-    for (trial in unique(data[SubNr == subject]$TrialNr)){
-      current <- data[SubNr == subject & TrialNr == trial]
-      if (nrow(current) != 0){ # current data not empty
-        if (nrow(current) != nrow(ideal)){ # extra/missing note
-          if (nrow(current) > nrow(ideal)){
-            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Extra Notes")) 
-          } else if (nrow(current) < nrow(ideal)){
-            dt_errors <- rbind(dt_errors, data.table(subject, trial, "Missing Notes")) 
-          }
-        } else if (nrow(current) == nrow(ideal)){ # substituted note
-          for (note in 1:nrow(ideal)){
-            if (current$Pitch[note] != ideal$Pitch[note]){
-              dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Substituted Notes - RowNr ", as.character(note), sep = "")))
-              break
-            }
-          }
-        }
-      } else if (nrow(current) == 0){ # current data empty
-        dt_errors <- rbind(dt_errors, data.table(subject, trial, paste("Missing Trial")))
-      }
-    }
-  }
-  if (nrow(dt_errors) != 0){
-    colnames(dt_errors) <- c("SubNr", "TrialNr", "Reason")
-    return(dt_errors)
-  } else if (nrow(dt_errors) == 0){
-    return("No errors!")
-  }
-}
-
-### edit data
-# data: data of the current trial
-# ideal: dt_ideal
-edit <- function(data, ideal){
-  data$RowNr <- c(1:nrow(data))
[…]

### 2022-09-27 analysis/preprocessor/filtering.Rmd
add information about corrected responses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-- Extra notes
-- Missing notes
-- Substituted notes
+- Extra notes (Onset: `r round(nrow(error_extra_onset)/(20*16)*100, 2)` %, Offset: `r round(nrow(error_extra_offset)/(20*16)*100, 2)` %)
+- Missing notes (Onset:  `r round(nrow(error_missing_onset)/(20*16)*100, 2)` %, Offset: `r round(nrow(error_missing_offset)/(20*16)*100, 2)` %)
+- Substituted notes (Onset: `r round(nrow(error_sub_offset)/(20*16)*100, 2)` %, Offset: `r round(nrow(error_sub_offset)/(20*16)*100, 2)` %)
 
@@
 
-```{r, include = FALSE, eval = FALSE}
+```{r, include = FALSE}
 dt_correct_onset_1 <- fread(here("filtered", "dt_correct_onset_1.txt"))
@@
 
+# Corrected responses
+```{r, include = FALSE}
+extra_onset <- sum(as.numeric(error_extra_onset[CorrectionNr != "Check individually"]$CorrectionNr))
+extra_offset <- sum(as.numeric(error_extra_offset[CorrectionNr != "Check individually"]$CorrectionNr))
+missing_onset <- sum(as.numeric(error_missing_onset[CorrectionNr != "Check individually"]$CorrectionNr))
+missing_offset <- sum(as.numeric(error_missing_offset[CorrectionNr != "Check individually"]$CorrectionNr))
+sub_onset <- sum(as.numeric(error_sub_onset[CorrectionNr != "Check individually"]$CorrectionNr))
+sub_offset <- sum(as.numeric(error_sub_offset[CorrectionNr != "Check individually"]$CorrectionNr))
+mixed_onset <- sum(as.numeric(error_ind_onset$CorrectionNr))
+mixed_offset <- sum(as.numeric(error_ind_offset$CorrectionNr))
+```
+
+- Extra notes (Onset: `r extra_onset`, Offset: `r extra_offset`)
+- Missing notes (Onset: `r missing_onset`, Offset: `r missing_offset`)
+- Substituted notes (Onset: `r sub_onset`, Offset: `r sub_offset`)
+- Mixed (Onset: `r mixed_onset`, Offset: `r mixed_offset`)
+- Total (Onset: `r sum(extra_onset, missing_onset, sub_onset, mixed_onset)`, Offset: `r sum(extra_offset, missing_offset, sub_offset, mixed_offset)`)
+- In total, `r round(sum(extra_onset, missing_onset, sub_onset, mixed_onset)/nrow(dt_onset)*100, 2)` % for onsets and `r round(sum(extra_offset, missing_offset, sub_offset, mixed_offset)/nrow(dt_offset)*100, 2)` % for offsets are corrected.
+
 ```{r export, include = FALSE}

### 2022-09-27 analysis/preprocessor/filtering.R

@@ -819,30 +819,41 @@ error_ind_offset <- fread(here("filtered", "error_ind_offset.txt"))
 error_ind_offset
 
 
-## ---- include = FALSE, eval = FALSE-------------------------------------------
-## dt_correct_onset_1 <- fread(here("filtered", "dt_correct_onset_1.txt"))
-## dt_correct_onset_2 <- fread(here("filtered", "dt_correct_onset_2.txt"))
-## dt_correct_onset_3 <- fread(here("filtered", "dt_correct_onset_3.txt"))
-## dt_correct_onset_4 <- fread(here("filtered", "dt_correct_onset_4.txt"))
-## dt_correct_onset_5 <- fread(here("filtered", "dt_correct_onset_5.txt"))
-## 
-## # combine all
-## dt_correct_onset <- rbind(dt_correct_onset_1, dt_correct_onset_2, dt_correct_onset_3, dt_correct_onset_4, dt_correct_onset_5)
-## 
-## # export csv
-## fwrite(dt_correct_onset, file = here("filtered", "dt_correct_onset.txt"))
-## 
-## dt_correct_offset_1 <- fread(here("filtered", "dt_correct_offset_1.txt"))
-## dt_correct_offset_2 <- fread(here("filtered", "dt_correct_offset_2.txt"))
-## dt_correct_offset_3 <- fread(here("filtered", "dt_correct_offset_3.txt"))
-## dt_correct_offset_4 <- fread(here("filtered", "dt_correct_offset_4.txt"))
-## dt_correct_offset_5 <- fread(here("filtered", "dt_correct_offset_5.txt"))
-## 
-## # combine all
-## dt_correct_offset <- rbind(dt_correct_offset_1, dt_correct_offset_2, dt_correct_offset_3, dt_correct_offset_4, dt_correct_offset_5)
-## 
-## # export csv
-## fwrite(dt_correct_offset, file = here("filtered", "dt_correct_offset.txt"))
+## ---- include = FALSE---------------------------------------------------------
+dt_correct_onset_1 <- fread(here("filtered", "dt_correct_onset_1.txt"))
+dt_correct_onset_2 <- fread(here("filtered", "dt_correct_onset_2.txt"))
+dt_correct_onset_3 <- fread(here("filtered", "dt_correct_onset_3.txt"))
+dt_correct_onset_4 <- fread(here("filtered", "dt_correct_onset_4.txt"))
[…]

### 2022-09-27 paper/manuscript.Rmd
updated data analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-bibliography      : "r-references.bib"
+#bibliography      : ""
 
 floatsintext      : no
-linenumbers       : yes
+linenumbers       : no
 draft             : no
@@
 source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE)
-source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE)
+source(here("analysis/preprocessor", "filtering.R"), chdir = TRUE)
 ```
@@
 
-`r nrow(dt_included)` participants were recruited who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
+We recruited `r nrow(dt_included)` participants who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
 Most participants were right-handed (left: `r nrow(dt_included[Handedness == "Left"])`, ambidextrous: `r nrow(dt_included[Handedness == "Both"])`).
@@
 A weighted Yamaha MIDI digital piano was used to record participants' performance.
+The pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers.
 All auditory feedback was given to participants through headphones (Audio-Technica ATH-M50X).
@@
 
-Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) because it is relatively simple in terms of motor skills.
+We made artificial recordings to manipulate the skill levels of imaginary students.
+We made four different types of stimuli.
+In Tominaga et al.
+(2022), participants were asked to perform the piece with either articulation or dynamics in two different conditions.
+In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson).
+In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert).
+
+Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) and is relatively simple in terms of motor skills.
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
@@
 
-In each performance (trial), the pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers. We quantified performances in terms of tempo,
+The dependent variables were computed for data analysis.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
+A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
+Tone intensity is assessed  by key velocity (KV) and measures the loudness of a musical note
+
+Data processing and statistical analysis were performed in R version 4.0.5.
+For statistical analysis, we included 8th notes with expressive notations only.
+Pitch errors were identified by comparing the sequence of  musical notes produced by a participant with the sequence of musical notes according to the sheet music.
+Pitch errors included either, extra, missing or substituted tones and were manually removed by using the *editData* R package.
+For note onsets,  `r round(nrow(dt_error_onset)/(20*16)*100, 2)` % of the trials contained at least one pitch error (extra notes: `r round(nrow(error_extra_onset)/(20*16)*100, 2)` %, missing notes:`r round(nrow(error_missing_onset)/(20*16)*100, 2)` %, substituted notes: `r round(nrow(error_sub_onset)/(20*16)*100, 2)` %).
+For note offsets,  `r round(nrow(dt_error_offset)/(20*16)*100, 2)` % of the trials contained at least one pitch error (extra notes: `r round(nrow(error_extra_offset)/(20*16)*100, 2)` %, missing notes: `r round(nrow(error_missing_offset)/(20*16)*100, 2)` %, substituted notes: `r round(nrow(error_sub_offset)/(20*16)*100, 2)` %).
+We found that some participants did not precisely follow the sheet music (e.g., they held some notes longer than notated), therefore the order of offsets did not correspond to that of onsets.
+We considered these as errors and removed the erroneous notes even if the order of onsets was correct.
+As a result, less than 1 % of total responses were corrected.
+After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference , defined as values more than 3 standard deviations from the mean of each dependent variable.
+For each dependent variable, this resulted in less than 2 % of overall responses being removed as outliers.
 

### 2022-09-27 paper/manuscript.tex
updated data analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \makeatother
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
@@
 
-20 participants were recruited who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
+We recruited 20 participants who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
 Most participants were right-handed (left: 2, ambidextrous: 2).
@@
 
-Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) because it is relatively simple in terms of motor skills.
+We made artificial recordings to manipulate the skill levels of imaginary students.
+We made four different types of stimuli.
+In Tominaga et al.~(2022), participants were asked to perform the piece with either articulation or dynamics in two different conditions.
+In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson).
+In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert).
+
+Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) and is relatively simple in terms of motor skills.
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
@@
 
-In each performance (trial), the pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers. We quantified performances in terms of tempo,
+The dependent variables were computed for data analysis.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
+Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
+A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
+Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note
+
+Data processing and statistical analysis were performed in R version 4.0.5.
+For statistical analysis, we included 8th notes with expressive notations only.
+Pitch errors were identified by comparing the sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
+Pitch errors included either, extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
+For note onsets, 31.87 \% of the trials contained at least one pitch error (extra notes: 5.94 \%, missing \url{notes:24.38} \%, substituted notes: 1.56 \%).
+For note offsets, 35.31 \% of the trials contained at least one pitch error (extra notes: 5.94 \%, missing notes: 24.38 \%, substituted notes: 5 \%).
+We found that some participants did not precisely follow the sheet music (e.g., they held some notes longer than notated), therefore the order of offsets did not correspond to that of onsets.
+We considered these as errors and removed the erroneous notes even if the order of onsets was correct.
+As a result, less than 1 \% of total responses were corrected.
+After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference , defined as values more than 3 standard deviations from the mean of each dependent variable.
+For each dependent variable, this resulted in less than 2 \% of overall responses being removed as outliers.
 

### 2022-09-28 paper/manuscript.Rmd
added apparatus and stimuli / first draft  of methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-We recruited `r nrow(dt_included)` participants who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
+We recruited `r nrow(dt_included)` participants who already had a degree (above bachelor's or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
 Most participants were right-handed (left: `r nrow(dt_included[Handedness == "Left"])`, ambidextrous: `r nrow(dt_included[Handedness == "Both"])`).
-The mean age of the participants was `r round(mean(dt_included$Age), 2)` (*SD* = `r round(sd(dt_included$Age), 2)`).
+The mean age of the participants was `r round(mean(dt_included$Age), 2)` years (*SD* = `r round(sd(dt_included$Age), 2)`).
 They had `r round(mean(dt_included$PianoTotalPractice), 2)` years of practice on average (*SD* = `r round(sd(dt_included$PianoTotalPractice), 2)`).
@@
 
-We made artificial recordings to manipulate the skill levels of imaginary students.
-We made four different types of stimuli.
-In Tominaga et al.
-(2022), participants were asked to perform the piece with either articulation or dynamics in two different conditions.
-In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson).
-In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert).
-
-Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) and is relatively simple in terms of motor skills.
+As stimuli, one piece of music, which was used in our previous experiment, was selected (for details, Tominaga et al., 2022).
+The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major.
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Original sheet music was used for the purpose of practice (*Fig A*).
-Two expressive notations were added to the original sheet music for the experiment (*Fig B*).
-These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
-The fingering was also assigned and confirmed by the same doctoral student.
+Two expressive notations (i.e., articulation and dynamics) were added as *Fig 1*.
+
+We created artificial recordings of students to manipulate their skill levels.
+In this experiment, we made the recordings by changing two factors of articulation (present or absent) and dynamics (present or absent).
+Therefore, there were four types of recordings.
+We aimed to create a quarter of the recordings where both articulation and dynamics were implemented.
+The second quarter included the recordings where only articulation was implemented whereas dynamics was missing.
+The third quarter included the recordings where only dynamics was implemented whereas articulation was missing.
+The last quarter included the recordings where neither articulation nor dynamics was implemented.
+We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
+How we generated the stimuli is written in *Supplementary Material*.
+
+<!-- ```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stimuli}Sheet music.
+Articulation.
+For articulation notation, the curved line (slur) indicates legato and the dots indicate staccato.
+For dynamics notation, the symbol `f' denotes forte and the symbol `p' denotes piano.
+For data analysis, only the 8th notes were included.", echo = FALSE} -->
+<!-- stim <- image_border(image_read(here("paper/image", "stim_m.png")), "#FFFFFF", "140x140") -->
+<!-- stim -->
+<!-- ``` -->
 
@@
 
-Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and performed it without pitch errors while implementing notated expressions in the experiment.
+Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and perform it without pitch errors while implementing notated expressions in the experiment.
+
+First, we recorded participants' baseline performance by asking them to perform the piece without listening to any students' performance.
+A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
+Sheet music (*Fig 1*) was displayed in front of the participants.
+Also, they were told to perform the piece expressively with their interpretation.
+This instruction was given to make sure that they paid attention to expressive aspects of the performance.
+Each participant performed the piece twice.
+
+After we recorded the  participants' baseline performance, participants were informed that they were going to listen to a number of recordings from 16 different students, who were learning musical expressive techniques.
+Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to the student (i.e., recording).
+In total, there were 16 trials and participants played the piece for each recording only once.
+The order of the recordings was randomised for each participant.
+A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
+
+After participants completed the 16 trials, they were asked to perform the piece in the same situation as we recorded their baseline performance.
+At the end of the experiment, participants filled in a questionnaire asking about their demographic information and experience in piano performance/teaching.
 
@@
 
-The dependent variables were computed for data analysis.
+The dependent variables were computed from MIDI data for data analysis.
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
@@
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
-Tone intensity is assessed  by key velocity (KV) and measures the loudness of a musical note
+Tone intensity is assessed  by key velocity (KV) and measures the loudness of a musical note.
+A higher value indicates forte styles whereas a lower value indicates piano styles.
+The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
+Also, KV difference was calculated by subtracting the KV value of the current note from that of the following note.
+We particularly focused on specific points where each subcomponent changed from one to the other (i.e., forte to piano or piano to forte) to measure dynamics contrast between forte and piano.
 
@@
 Pitch errors included either, extra, missing or substituted tones and were manually removed by using the *editData* R package.
-For note onsets,  `r round(nrow(dt_error_onset)/(20*16)*100, 2)` % of the trials contained at least one pitch error (extra notes: `r round(nrow(error_extra_onset)/(20*16)*100, 2)` %, missing notes:`r round(nrow(error_missing_onset)/(20*16)*100, 2)` %, substituted notes: `r round(nrow(error_sub_onset)/(20*16)*100, 2)` %).
+For note onsets,  `r round(nrow(dt_error_onset)/(20*16)*100, 2)` % of the trials contained at least one pitch error (extra notes: `r round(nrow(error_extra_onset)/(20*16)*100, 2)` %, missing notes: `r round(nrow(error_missing_onset)/(20*16)*100, 2)` %, substituted notes: `r round(nrow(error_sub_onset)/(20*16)*100, 2)` %).
 For note offsets,  `r round(nrow(dt_error_offset)/(20*16)*100, 2)` % of the trials contained at least one pitch error (extra notes: `r round(nrow(error_extra_offset)/(20*16)*100, 2)` %, missing notes: `r round(nrow(error_missing_offset)/(20*16)*100, 2)` %, substituted notes: `r round(nrow(error_sub_offset)/(20*16)*100, 2)` %).
@@
 As a result, less than 1 % of total responses were corrected.
-After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference , defined as values more than 3 standard deviations from the mean of each dependent variable.
+After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference, defined as values more than 3 standard deviations from the mean of each dependent variable.
 For each dependent variable, this resulted in less than 2 % of overall responses being removed as outliers.
+
+We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.
+absent) and Dynamics (present or absent).
+The *aov_ez* function in the *afex* R package was used for a repeated-measures ANOVA.
+For post-doc comparisons on the estimated marginal means, we used the  *emmeans* R package.
 

### 2022-09-28 paper/manuscript.tex
added apparatus and stimuli / first draft  of methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-We recruited 20 participants who already had a degree (above bachelor or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
+We recruited 20 participants who already had a degree (above bachelor's or equivalent) in piano performance/teaching or were studying advanced piano performance at a music school.
 Most participants were right-handed (left: 2, ambidextrous: 2).
-The mean age of the participants was 28.25 (\emph{SD} = 10.95).
+The mean age of the participants was 28.25 years (\emph{SD} = 10.95).
 They had 21.55 years of practice on average (\emph{SD} = 11.59).
@@
 
-We made artificial recordings to manipulate the skill levels of imaginary students.
-We made four different types of stimuli.
-In Tominaga et al.~(2022), participants were asked to perform the piece with either articulation or dynamics in two different conditions.
-In the teaching condition, participants were instructed to perform the piece with the designated expressive technique as if they were teaching it to students (e.g., in a lesson).
-In the performing condition, participants were instructed to perform the piece with the designated expressive technique as if they were performing it to an audience (e.g., in a concert).
-
-Clementi's Sonatina Op.36 (No.3) in C major was selected as a stimulus because it contains our targeted expressions (i.e., articulation and dynamics) and is relatively simple in terms of motor skills.
+As stimuli, one piece of music, which was used in our previous experiment, was selected (for details, Tominaga et al., 2022).
+The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major.
 The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Original sheet music was used for the purpose of practice (\emph{Fig A}).
-Two expressive notations were added to the original sheet music for the experiment (\emph{Fig B}).
-These excerpts were confirmed to be musically natural by a doctoral student in piano performance at Liszt Ferenc Academy of Music in Hungary.
-The fingering was also assigned and confirmed by the same doctoral student.
+Two expressive notations (i.e., articulation and dynamics) were added as \emph{Fig 1}.
+
+We created artificial recordings of students to manipulate their skill levels.
+In this experiment, we made the recordings by changing two factors of articulation (present or absent) and dynamics (present or absent).
+Therefore, there were four types of recordings.
+We aimed to create a quarter of the recordings where both articulation and dynamics were implemented.
+The second quarter included the recordings where only articulation was implemented whereas dynamics was missing.
+The third quarter included the recordings where only dynamics was implemented whereas articulation was missing.
+The last quarter included the recordings where neither articulation nor dynamics was implemented.
+We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
+How we generated the stimuli is written in \emph{Supplementary Material}.
 
@@
 
-Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and performed it without pitch errors while implementing notated expressions in the experiment.
+Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and perform it without pitch errors while implementing notated expressions in the experiment.
+
+First, we recorded participants' baseline performance by asking them to perform the piece without listening to any students' performance.
+A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
+Sheet music (\emph{Fig 1}) was displayed in front of the participants.
+Also, they were told to perform the piece expressively with their interpretation.
+This instruction was given to make sure that they paid attention to expressive aspects of the performance.
+Each participant performed the piece twice.
+
+After we recorded the participants' baseline performance, participants were informed that they were going to listen to a number of recordings from 16 different students, who were learning musical expressive techniques.
+Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to the student (i.e., recording).
+In total, there were 16 trials and participants played the piece for each recording only once.
+The order of the recordings was randomised for each participant.
+A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
+
+After participants completed the 16 trials, they were asked to perform the piece in the same situation as we recorded their baseline performance.
+At the end of the experiment, participants filled in a questionnaire asking about their demographic information and experience in piano performance/teaching.
 
@@
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
-Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note
+Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note.
+A higher value indicates forte styles whereas a lower value indicates piano styles.
+The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
+Also, KV difference was calculated by subtracting the KV value of the current note from that of the following note.
+We particularly focused on specific points where each subcomponent changed from one to the other (i.e., forte to piano or piano to forte) to measure dynamics contrast between forte and piano.
 
@@
 Pitch errors included either, extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
-For note onsets, 31.87 \% of the trials contained at least one pitch error (extra notes: 5.94 \%, missing \url{notes:24.38} \%, substituted notes: 1.56 \%).
+For note onsets, 31.87 \% of the trials contained at least one pitch error (extra notes: 5.94 \%, missing notes: 24.38 \%, substituted notes: 1.56 \%).
 For note offsets, 35.31 \% of the trials contained at least one pitch error (extra notes: 5.94 \%, missing notes: 24.38 \%, substituted notes: 5 \%).
@@
 As a result, less than 1 \% of total responses were corrected.
-After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference , defined as values more than 3 standard deviations from the mean of each dependent variable.
+After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference, defined as values more than 3 standard deviations from the mean of each dependent variable.
 For each dependent variable, this resulted in less than 2 \% of overall responses being removed as outliers.
+
+We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present or absent).
+The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
+For post-doc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.
 

### 2022-10-06 analysis/baseline.Rmd
shortened labelling process
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 data_vel_diff = "preprocessor/trimmed/dt_vel_diff_trimmed.txt"
-data_baseline_ioi = "preprocessor/predata/trimmed/dt_ioi_trimmed.txt"
-data_baseline_kot = "preprocessor/predata/trimmed/dt_kot_trimmed.txt"
-data_baseline_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
-data_baseline_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
+data_bl_ioi = "preprocessor/predata/trimmed/dt_ioi_trimmed.txt"
+data_bl_kot = "preprocessor/predata/trimmed/dt_kot_trimmed.txt"
+data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
+data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 ```
@@
 ```{r prep, include = FALSE}
+# read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
@@
 
-# SubNr as a factor
-dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
+dt1.list <- list(dt_ioi, dt_kot, dt_vel, dt_vel_diff) # make a list of data tables
 
-# assign stimuli categories
-dt_ioi$Stimuli <- as.numeric(gsub(".mid", "", dt_ioi$Stimuli))
-dt_ioi$Articulation <- "NA"
-dt_ioi$Dynamics <- "NA"
-dt_ioi$Category <- "NA"
-dt_ioi[Stimuli < 5]$Articulation <- "Present"
-dt_ioi[Stimuli < 5]$Dynamics <- "Absent"
-dt_ioi[Stimuli < 5]$Category <- "art_only"
-dt_ioi[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
-dt_ioi[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
-dt_ioi[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
-dt_ioi[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
-dt_ioi[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
-dt_ioi[Stimuli > 8 & Stimuli < 13]$Category <- "both"
-dt_ioi[Stimuli > 12]$Articulation <- "Absent"
-dt_ioi[Stimuli > 12]$Dynamics <- "Absent"
-dt_ioi[Stimuli > 12]$Category <- "none"
+# add labels to data tables
+label <- function(data){
+  data[, "SubNr" := as.factor(data$SubNr)]
+  data[, "Stimuli" := as.numeric(gsub(".mid", "", data$Stimuli))]
+  data[, c("Articulation", "Dynamics", "Category") := .("NA", "NA", "NA")]
+  data[Stimuli < 5, c("Articulation", "Dynamics", "Category") := .("Present", "Absent", "art_only")]
+  data[Stimuli > 4 & Stimuli < 9, c("Articulation", "Dynamics", "Category") := .("Absent", "Present", "dyn_only")]
+  data[Stimuli > 8 & Stimuli < 13, c("Articulation", "Dynamics", "Category") := .("Present", "Present", "both")]
+  data[Stimuli > 12, c("Articulation", "Dynamics", "Category") := .("Absent", "Absent", "none")]
+}
 
-dt_kot$Stimuli <- as.numeric(gsub(".mid", "", dt_kot$Stimuli))
-dt_kot$Articulation <- "NA"
-dt_kot$Dynamics <- "NA"
-dt_kot$Category <- "NA"
-dt_kot[Stimuli < 5]$Articulation <- "Present"
-dt_kot[Stimuli < 5]$Dynamics <- "Absent"
-dt_kot[Stimuli < 5]$Category <- "art_only"
-dt_kot[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
-dt_kot[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
-dt_kot[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
-dt_kot[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
-dt_kot[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
-dt_kot[Stimuli > 8 & Stimuli < 13]$Category <- "both"
-dt_kot[Stimuli > 12]$Articulation <- "Absent"
-dt_kot[Stimuli > 12]$Dynamics <- "Absent"
-dt_kot[Stimuli > 12]$Category <- "none"
+# perform the function above
+res <- lapply(dt1.list, label)
 
-dt_vel$Stimuli <- as.numeric(gsub(".mid", "", dt_vel$Stimuli))
-dt_vel$Articulation <- "NA"
-dt_vel$Dynamics <- "NA"
-dt_vel$Category <- "NA"
-dt_vel[Stimuli < 5]$Articulation <- "Present"
-dt_vel[Stimuli < 5]$Dynamics <- "Absent"
-dt_vel[Stimuli < 5]$Category <- "art_only"
-dt_vel[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
-dt_vel[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
-dt_vel[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
-dt_vel[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
-dt_vel[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
-dt_vel[Stimuli > 8 & Stimuli < 13]$Category <- "both"
-dt_vel[Stimuli > 12]$Articulation <- "Absent"
-dt_vel[Stimuli > 12]$Dynamics <- "Absent"
-dt_vel[Stimuli > 12]$Category <- "none"
+# read data (baseline performances)
+dt_bl_ioi <- fread(data_bl_ioi, header = T, sep = ",", dec = ".")
+dt_bl_kot <- fread(data_bl_kot, header = T, sep = ",", dec = ".")
+dt_bl_vel <- fread(data_bl_vel, header = T, sep = ",", dec = ".")
+dt_bl_vel_diff <- fread(data_bl_vel_diff, header = T, sep = ",", dec = ".")
 
-dt_vel_diff$Stimuli <- as.numeric(gsub(".mid", "", dt_vel_diff$Stimuli))
-dt_vel_diff$Articulation <- "NA"
-dt_vel_diff$Dynamics <- "NA"
-dt_vel_diff$Category <- "NA"
-dt_vel_diff[Stimuli < 5]$Articulation <- "Present"
-dt_vel_diff[Stimuli < 5]$Dynamics <- "Absent"
-dt_vel_diff[Stimuli < 5]$Category <- "art_only"
-dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Articulation <- "Absent"
-dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Dynamics <- "Present"
-dt_vel_diff[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
-dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Articulation <- "Present"
-dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Dynamics <- "Present"
-dt_vel_diff[Stimuli > 8 & Stimuli < 13]$Category <- "both"
-dt_vel_diff[Stimuli > 12]$Articulation <- "Absent"
-dt_vel_diff[Stimuli > 12]$Dynamics <- "Absent"
-dt_vel_diff[Stimuli > 12]$Category <- "none"
-```
+dt2.list <- list(dt_bl_ioi, dt_bl_kot, dt_bl_vel, dt_bl_vel_diff) # make a list of data tables (baseline performances)
 
-```{r, include = FALSE}
+# add labels to data tables
+label_bl <- function(data){
+  data[, "SubNr" := as.factor(data$SubNr)]
+  data[, c("Articulation", "Dynamics", "Category") := .("Baseline", "Baseline", "baseline")]
+}
+
+# perform the function above
+res_bl <- lapply(dt2.list, label)
+
 # combine data with baseline performance
-dt_bl_ioi <- fread(data_baseline_ioi, header = T, sep = ",", dec = ".")
-dt_bl_kot <- fread(data_baseline_kot, header = T, sep = ",", dec = ".")
-dt_bl_vel <- fread(data_baseline_vel, header = T, sep = ",", dec = ".")
-dt_bl_vel_diff <- fread(data_baseline_vel_diff, header = T, sep = ",", dec = ".")
-
-dt_bl_ioi$Articulation <- "Baseline"
-dt_bl_kot$Articulation <- "Baseline"
-dt_bl_vel$Articulation <- "Baseline"
-dt_bl_vel_diff$Articulation <- "Baseline"
-
-dt_bl_ioi$Dynamics <- "Baseline"
-dt_bl_kot$Dynamics <- "Baseline"
-dt_bl_vel$Dynamics <- "Baseline"
-dt_bl_vel_diff$Dynamics <- "Baseline"
-
-dt_bl_ioi$Category <- "baseline"
-dt_bl_kot$Category <- "baseline"
-dt_bl_vel$Category <- "baseline"
-dt_bl_vel_diff$Category <- "baseline"
-
 dt_ioi <- rbind(dt_ioi, dt_bl_ioi)

### 2022-10-06 analysis/baseline.Rmd
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # perform the function above
-res_bl <- lapply(dt2.list, label)
+res_bl <- lapply(dt2.list, label_bl)
 
@@
 # for each individual
+dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
+dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
+
 ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
@@
 ```{r ioi-all-plot}
-ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
+ggplot(ioi, aes(x = Articulation, y = Mean, color = Dynamics))+ 
+  geom_boxplot(position = position_dodge2(preserve = "single")) +
+  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
+  theme_pubr()
+#ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
 ```
@@
 aov_ez(
-  data = ioi,
+  data = ioi[Category != "baseline"],
   id = "SubNr",
   dv = "Mean",
-  within = c("Category")
+  within = c("Articulation", "Dynamics")
 )
@@
 # for each individual
+dt_kot$Articulation <- factor(dt_kot$Articulation, c("Present", "Absent", "Baseline"))
+dt_kot$Dynamics <- factor(dt_kot$Dynamics, c("Present", "Absent", "Baseline"))
+
 kot_trial <- dt_kot[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]

### 2022-10-07 analysis/baseline.Rmd
update stats for ioi, kot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title: 'Comparison between baseline and test performance'
+title: 'Music adaptation: Analysis'
 output:
@@
 if (!require("afex")) {install.packages("afex"); require("afex")}
+if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 # plot
@@
 
-ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
+ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
 
-ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category)]
+ioi <- ioi_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), CV = mean(CV)), by = .(SubNr, Articulation, Dynamics, Category)]
 setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
@@
 # group mean
-ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Category)]
+ioi_all <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean), CV = mean(CV)), by = .(Articulation, Dynamics, Category)]
 ioi_all
@@
   geom_boxplot(position = position_dodge2(preserve = "single")) +
-  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
+  geom_point(position = position_dodge(0.75)) +
   theme_pubr()
@@
 
+1) Articulation x Dynamics
+
 ```{r}
@@
   within = c("Articulation", "Dynamics")
+)
+```
+
+2) Category: All
+
+```{r}
+aov_ez(
+  data = ioi,
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+```
+
+3) Category: both vs. none vs. baseline
+
+```{r}
+aov_ez(
+  data = ioi[Category == "both" | Category == "none" | Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+```
+
+4) Variability
+
+```{r}
+aov_ez(
+  data = ioi,
+  id = "SubNr",
+  dv = "SD",
+  within = c("Category")
 )
@@
 kot <- kot_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent1)]
-setorder(kot, "SubNr", "Articulation", "Dynamics", "Category")
+setorder(kot, "SubNr", "Subcomponent1", "Articulation", "Dynamics", "Category")
 kot
@@
 ```{r kot-all-plot, echo = FALSE}
-ggboxplot(kot[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
+  geom_boxplot(position = position_dodge2(preserve = "single")) +
+  #geom_point(position = position_dodge(0.75)) +
+  facet_grid( ~ Subcomponent1) +
+  theme_pubr()
 
-ggboxplot(kot[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
+ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
+  geom_boxplot(position = position_dodge2(preserve = "single")) +
+  #geom_point(position = position_dodge(0.75)) +
+  facet_grid( ~ Subcomponent1) +
+  theme_pubr()
+
+#ggboxplot(kot[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
+
+#ggboxplot(kot[Subcomponent1 == "Staccato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
 ```
@@
 ## 4. Stats
+### Legato
 
-1) Legato
+1) Articulation x Dynamics
 
@@
 aov_ez(
+  data = kot[Subcomponent1 == "Legato" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+2) Category: All
+
+```{r}
+leg_ca_all_aov <- aov_ez(
   data = kot[Subcomponent1 == "Legato"],
@@
 )
+leg_ca_all_aov
+
+pairs(emmeans(leg_ca_all_aov, "Category"), adjust = "tukey")
 ```
 
-2) Staccato
+3) Category: both vs. dyn_only vs. baseline
+
+```{r}
+leg_ca_par_aov <- aov_ez(
+  data = kot[Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only" | Subcomponent1 == "Legato" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+leg_ca_par_aov
+
+pairs(emmeans(leg_ca_par_aov, "Category"), adjust = "tukey")
+```
+
+### Staccato
+
+1) Articulation x Dynamics
 
@@
 aov_ez(
+  data = kot[Subcomponent1 == "Staccato" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+2) Category: All
+
+```{r}
+sta_ca_all_aov <- aov_ez(
   data = kot[Subcomponent1 == "Staccato"],
@@
 )
+sta_ca_all_aov
+
+pairs(emmeans(sta_ca_all_aov, "Category"), adjust = "tukey")
 ```
 
+3) Category: both vs. dyn_only vs. baseline
+
+```{r}
+sta_ca_par_aov <- aov_ez(
+  data = kot[Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only" | Subcomponent1 == "Staccato" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+sta_ca_par_aov
+
+pairs(emmeans(sta_ca_par_aov, "Category"), adjust = "tukey")
+```
 # Dynamics
@@
 # for each individual
+dt_vel$Articulation <- factor(dt_vel$Articulation, c("Present", "Absent", "Baseline"))
+dt_vel$Dynamics <- factor(dt_vel$Dynamics, c("Present", "Absent", "Baseline"))
+
 vel_trial <- dt_vel[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
@@
 ```{r vel-all-plot,  echo = FALSE}
-ggboxplot(vel[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
+ggplot(vel, aes(x = Dynamics, y = Mean, color = Articulation))+ 
+  geom_boxplot(position = position_dodge2(preserve = "single")) +
+  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
+  facet_grid( ~ Subcomponent2) +
+  theme_pubr()
 
-ggboxplot(vel[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
+#ggboxplot(vel[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
+
+#ggboxplot(vel[Subcomponent2 == "Piano"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
 ```
@@
 # for each individual
+dt_vel_diff$Articulation <- factor(dt_vel_diff$Articulation, c("Present", "Absent", "Baseline"))
+dt_vel_diff$Dynamics <- factor(dt_vel_diff$Dynamics, c("Present", "Absent", "Baseline"))
+
 vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
@@
 ```{r vel-diff-all-plot,  echo = FALSE}
-ggboxplot(vel_diff[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")
+ggplot(vel_diff, aes(x = Dynamics, y = Mean, color = Articulation))+ 
+  geom_boxplot(position = position_dodge2(preserve = "single")) +
+  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
+  facet_grid( ~ Subcomponent2) +
+  theme_pubr()
 
-ggboxplot(vel_diff[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
+#ggboxplot(vel_diff[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (FtoP)")
+
+#ggboxplot(vel_diff[Subcomponent2 == "PtoF"], "Dynamics", "Mean", color = "Articulation", add = "jitter", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (PtoF)")
 ```

### 2022-10-11 analysis/baseline.Rmd
update plots and stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ## 1. Individual
+
 ```{r ioi}
@@
 ```{r}
-aov_ez(
+ioi_ca_all_aov <- aov_ez(
   data = ioi,
@@
 )
+
+pairs(emmeans(ioi_ca_all_aov, "Category"), adjust = "tukey")
 ```
@@
 ```{r}
-aov_ez(
+ioi_ca_par_aov <- aov_ez(
   data = ioi[Category == "both" | Category == "none" | Category == "baseline"],
@@
 )
+
+pairs(emmeans(ioi_ca_par_aov, "Category"), adjust = "tukey")
 ```
@@
 ```{r kot-all-plot, echo = FALSE}
-ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
+ggplot(kot, aes(x = Articulation, y = Mean, color = Dynamics))+
   geom_boxplot(position = position_dodge2(preserve = "single")) +
-  #geom_point(position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent1) +
-  theme_pubr()
-
-ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
-  geom_boxplot(position = position_dodge2(preserve = "single")) +
-  #geom_point(position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent1) +
+  geom_point(position = position_dodge(0.75)) +
+  facet_grid( ~ Subcomponent1, labeller = as_labeller(c("Legato" = "Legato/Forte", "Staccato" = "Staccato/Piano"))) +
   theme_pubr()
@@
 ```
+
 2) Category: All
@@
 vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
-setorder(vel, "SubNr", "Dynamics", "Articulation", "Category")
+setorder(vel, "Subcomponent2", "SubNr", "Dynamics", "Articulation", "Category")
 vel
@@
   geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent2) +
+  facet_grid( ~ Subcomponent2, labeller = as_labeller(c("Forte" = "Forte/Legato", "Piano" = "Piano/Staccato"))) +
   theme_pubr()
@@
 ## 4. Stats
+### Forte
 
-1) Forte
+1) Articulation x Dynamics
 
@@
 aov_ez(
+  data = vel[Subcomponent2 == "Forte" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+
+2) Category: All
+
+```{r}
+for_ca_all_aov <- aov_ez(
   data = vel[Subcomponent2 == "Forte"],
@@
 )
+for_ca_all_aov
+
+pairs(emmeans(for_ca_all_aov, "Category"), adjust = "tukey")
 ```
 
-2) Piano
+3) Category: both vs. art_only vs. baseline
+
+```{r}
+for_ca_par_aov <- aov_ez(
+  data = vel[Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only" | Subcomponent2 == "Forte" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+for_ca_par_aov
+
+pairs(emmeans(for_ca_par_aov, "Category"), adjust = "tukey")
+```
+
+### Piano
+
+1) Articulation x Dynamics
 
@@
 aov_ez(
+  data = vel[Subcomponent2 == "Piano" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+2) Category: All
+
+```{r}
+pia_ca_all_aov <- aov_ez(
   data = vel[Subcomponent2 == "Piano"],
@@
 )
+pia_ca_all_aov
+
+pairs(emmeans(pia_ca_all_aov, "Category"), adjust = "tukey")
+```
+
+3) Category: both vs. art_only vs. baseline
+
+```{r}
+pia_ca_par_aov <- aov_ez(
+  data = vel[Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only" | Subcomponent2 == "Piano" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+pia_ca_par_aov
+
+pairs(emmeans(pia_ca_par_aov, "Category"), adjust = "tukey")
 ```
@@
 ## 2. All - group
+
 ```{r vel-diff-all, echo = FALSE}
@@
   geom_boxplot(position = position_dodge2(preserve = "single")) +
-  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent2) +
+  geom_point(position = position_dodge(0.75)) +
+  facet_grid( ~ Subcomponent2, labeller = as_labeller(c("FtoP" = "FtoP/PtoF", "PtoF" = "PtoF/StoL"))) +
   theme_pubr()
@@
 ## 4. Stats
+### FtoP
 
-1) FtoP
+1) Articulation x Dynamics
 
@@
 aov_ez(
+  data = vel_diff[Subcomponent2 == "FtoP" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+
+2) Category: All
+
+```{r}
+ftop_ca_all_aov <- aov_ez(
   data = vel_diff[Subcomponent2 == "FtoP"],
@@
 )
+ftop_ca_all_aov
+
+pairs(emmeans(ftop_ca_all_aov, "Category"), adjust = "tukey")
 ```
 
-2) PtoF
+3) Category: both vs. art_only vs. baseline
+
+```{r}
+ftop_ca_par_aov <- aov_ez(
+  data = vel_diff[Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+ftop_ca_par_aov
+
+pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
+```
+
+### PtoF
+
+1) Articulation x Dynamics
 
@@
 aov_ez(
+  data = vel_diff[Subcomponent2 == "PtoF" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+```
+
+2) Category: All
+
+```{r}
+ptof_ca_all_aov <- aov_ez(
   data = vel_diff[Subcomponent2 == "PtoF"],
@@
 )
+ptof_ca_all_aov
+
+pairs(emmeans(ptof_ca_all_aov, "Category"), adjust = "tukey")
 ```
+
+3) Category: both vs. art_only vs. baseline
+
+```{r}
+ftop_ca_par_aov <- aov_ez(
+  data = vel_diff[Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+ftop_ca_par_aov
+
+pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
+```

### 2022-10-14 analysis/onlyfirst.Rmd
update analysis for only first instances
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-10-07 analysis/baseline.Rmd
@@
 ---
-title: 'Music adaptation: Analysis'
+title: 'Music adaptation: Analysis (only first instances)'
 output:
@@
 
-Description: This is a summary of analysis comparing between baseline and test performance.
+Description: This is a summary of analysis which only include the performance with the first instance of each student (category).
 
 - Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
+
+## Experimental Design
+Note: Here, we only look at the first instance of each student (category).
+
+- Independent variables: Articulation (2 levels: Present or Absent) x Dynamics (2 levels: Present or Absent)
+- Dependent variables: Interonset intervals (IOIs), Key-overlap time (KOT), Key velocity (KV) and KV Difference
+
+1. Main predictions ([preregistration](https://osf.io/qbzwt))
+
+- If students performed the piece without articulation (i.e., only dynamics is implemented in the recordings), participants should exaggerate articulation more (i.e., producing longer legato and shorter staccato).
+- If students performed the piece without dynamics (i.e., only articulation is implemented in the recordings), participants should exaggerate dynamics more (i.e., producing louder forte and softer piano).
+- If students performed the piece without articulation and dynamics (i.e., neither articulation nor dynamics is implemented in the recordings), participants should exaggerate both articulation and dynamics more.
+
+2. Excluded participants
+
+- NA
+
+3. Stats packages
+
+- [afex](https://github.com/singmann/afex) for ANOVAs
 
@@
 
+### TrialNr and Stimuli
+
+```{r}
+# select only performances for the first instance of each student (category)
+info <- dt_ioi[, .(N = .N), by = .(SubNr, TrialNr, Stimuli, Category)]
+info
+
+first <- function(data, new_data){
+  for (subnr in unique(info$SubNr)){
+    for (category in unique(info[SubNr == subnr]$Category)){
+      trials <- unique(info[SubNr == subnr & Category == category]$TrialNr)
+      current <- data[SubNr == subnr & Category == category & TrialNr == min(trials)]
+      new_data <- rbind(new_data, current)
+    }
+  }
+  return(new_data)
+}
+
+dt_ioi_first <- data.table() 
+dt_ioi_first <- first(dt_ioi, dt_ioi_first)
+
+dt_kot_first <- data.table()
+dt_kot_first <- first(dt_kot, dt_kot_first)
+
+dt_vel_first <- data.table()
+dt_vel_first <- first(dt_vel, dt_vel_first)
+
+dt_vel_diff_first <- data.table()
+dt_vel_diff_first <- first(dt_vel_diff, dt_vel_diff_first)
+```
+
 # IOI 
 ## 1. Individual
+
 ```{r ioi}
 # for each individual
-dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
-dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
+dt_ioi_first$Articulation <- factor(dt_ioi_first$Articulation, c("Present", "Absent", "Baseline"))
+dt_ioi_first$Dynamics <- factor(dt_ioi_first$Dynamics, c("Present", "Absent", "Baseline"))
 
-ioi_trial <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
+ioi_trial <- dt_ioi_first[, .(N = .N, Mean = mean(IOI), SD = sd(IOI), CV = sd(IOI)/mean(IOI)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category)]
 
@@
 #### Boxplot
+
 ```{r ioi-plot, fig.width = 6}
@@
 
-## 2. All - group
+## 2. Group
 
@@
 ```{r ioi-all-plot}
-ggplot(ioi, aes(x = Articulation, y = Mean, color = Dynamics))+ 
-  geom_boxplot(position = position_dodge2(preserve = "single")) +
-  geom_point(position = position_dodge(0.75)) +
+ggplot(ioi, aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") +
   theme_pubr()
+
 #ggboxplot(ioi, "Articulation", "Mean", color = "Dynamics", add = "jitter", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
@@
 ```{r}
-aov_ez(
+ioi_all_aov <- aov_ez(
   data = ioi[Category != "baseline"],
@@
   dv = "Mean",
-  within = c("Articulation", "Dynamics")
+  within = c("Articulation", "Dynamics"),
+  detailed = TRUE
 )
+
+ioi_all_aov$anova_table
+pairs(emmeans(ioi_all_aov, ~Dynamics|Articulation), adjust = "tukey")
 ```
@@
 )
+
+ioi_ca_all_aov$anova_table
+pairs(emmeans(ioi_ca_all_aov, "Category"), adjust = "tukey")
 ```
@@
 )
+
+ioi_ca_par_aov$anova_table
+pairs(emmeans(ioi_ca_par_aov, "Category"), adjust = "tukey")
 ```
 
-4) Variability
+<!-- 4) Variability -->
 
-```{r}
-aov_ez(
-  data = ioi,
-  id = "SubNr",
-  dv = "SD",
-  within = c("Category")
-)
-```
+<!-- ```{r} -->
+<!-- var_stats <- aov_ez( -->
+<!--   data = ioi, -->
+<!--   id = "SubNr", -->
+<!--   dv = "SD", -->
+<!--   within = c("Category") -->
+<!-- ) -->
+
+<!-- var_stats$anova_table -->
+<!-- ``` -->
 
@@
 # for each individual
-dt_kot$Articulation <- factor(dt_kot$Articulation, c("Present", "Absent", "Baseline"))
-dt_kot$Dynamics <- factor(dt_kot$Dynamics, c("Present", "Absent", "Baseline"))
+dt_kot_first$Articulation <- factor(dt_kot_first$Articulation, c("Present", "Absent", "Baseline"))
+dt_kot_first$Dynamics <- factor(dt_kot_first$Dynamics, c("Present", "Absent", "Baseline"))
 
-kot_trial <- dt_kot[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]
+kot_trial <- dt_kot_first[Subcomponent1 == "Legato" | Subcomponent1 == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent1)]
 
@@
 ```{r kot-all-plot, echo = FALSE}
-ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
-  geom_boxplot(position = position_dodge2(preserve = "single")) +
-  #geom_point(position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent1) +
+ggplot(kot[Subcomponent1 == "Legato"], aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KOT (ms)", title = "KOT - Legato") +
   theme_pubr()
 
-ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics))+ 
-  geom_boxplot(position = position_dodge2(preserve = "single")) +
-  #geom_point(position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent1) +
+ggplot(kot[Subcomponent1 == "Staccato"], aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KOT (ms)", title = "KOT - Staccato") +
   theme_pubr()
@@
 ```{r}
-aov_ez(
+leg_all_aov <- aov_ez(
   data = kot[Subcomponent1 == "Legato" & Category != "baseline"],
@@
 )
+
+leg_all_aov$anova_table
+pairs(emmeans(leg_all_aov, ~Dynamics|Articulation), adjust = "tukey")
 ```
+
 2) Category: All
@@
 
+leg_ca_all_aov$anova_table
 pairs(emmeans(leg_ca_all_aov, "Category"), adjust = "tukey")
@@
 
+leg_ca_par_aov$anova_table
 pairs(emmeans(leg_ca_par_aov, "Category"), adjust = "tukey")
@@
 ```{r}
-aov_ez(
+sta_all_aov <- aov_ez(
   data = kot[Subcomponent1 == "Staccato" & Category != "baseline"],
@@
 )
+
+sta_all_aov$anova_table
+pairs(emmeans(sta_all_aov, ~Dynamics|Articulation), adjust = "tukey")
 ```
+
 2) Category: All
@@
 )
-sta_ca_all_aov
 
+sta_ca_all_aov$anova_table
 pairs(emmeans(sta_ca_all_aov, "Category"), adjust = "tukey")
@@
 )
-sta_ca_par_aov
 
+sta_ca_par_aov$anova_table
 pairs(emmeans(sta_ca_par_aov, "Category"), adjust = "tukey")
 ```
+
 # Dynamics
@@
 # for each individual
-dt_vel$Articulation <- factor(dt_vel$Articulation, c("Present", "Absent", "Baseline"))
-dt_vel$Dynamics <- factor(dt_vel$Dynamics, c("Present", "Absent", "Baseline"))
+dt_vel_first$Articulation <- factor(dt_vel_first$Articulation, c("Present", "Absent", "Baseline"))
+dt_vel_first$Dynamics <- factor(dt_vel_first$Dynamics, c("Present", "Absent", "Baseline"))
 
-vel_trial <- dt_vel[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
+vel_trial <- dt_vel_first[Subcomponent2 != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
 
 vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
-setorder(vel, "SubNr", "Dynamics", "Articulation", "Category")
+setorder(vel, "Subcomponent2", "SubNr", "Dynamics", "Articulation", "Category")
 vel
@@
 ```{r vel-all-plot,  echo = FALSE}
-ggplot(vel, aes(x = Dynamics, y = Mean, color = Articulation))+ 
-  geom_boxplot(position = position_dodge2(preserve = "single")) +
-  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent2) +
+ggplot(vel[Subcomponent2 == "Forte"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
+    geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "Key Velocity (0-127)", title = "KV - Forte") +
+  theme_pubr()
+
+ggplot(vel[Subcomponent2 == "Piano"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
+    geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "Key Velocity (0-127)", title = "KV - Piano") +
   theme_pubr()
@@
 
-ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Velocity (0-127)", title = "Velocity") + scale_x_continuous(breaks = seq(1,72,1))
+ggline(vel_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Category", color = "Category", xlab = "Note Nr", ylab = "Key Velocity (0-127)", title = "KV") + scale_x_continuous(breaks = seq(1,72,1))
 ```
@@
 ## 4. Stats
+### Forte
 
-1) Forte
+1) Articulation x Dynamics
 
 ```{r}
-aov_ez(
+for_all_aov <- aov_ez(
+  data = vel[Subcomponent2 == "Forte" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+
+for_all_aov$anova_table
+pairs(emmeans(for_all_aov, ~Articulation|Dynamics), adjust = "tukey")
+```
+
+2) Category: All
+
+```{r}
+for_ca_all_aov <- aov_ez(
   data = vel[Subcomponent2 == "Forte"],
@@
 )
+
+for_ca_all_aov$anova_table
+pairs(emmeans(for_ca_all_aov, "Category"), adjust = "tukey")
 ```
 
-2) Piano
+3) Category: both vs. art_only vs. baseline
 
 ```{r}
-aov_ez(
+for_ca_par_aov <- aov_ez(
+  data = vel[Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only" | Subcomponent2 == "Forte" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+
+for_ca_par_aov$anova_table
+pairs(emmeans(for_ca_par_aov, "Category"), adjust = "tukey")
+```
+
+### Piano
+
+1) Articulation x Dynamics
+
+```{r}
+pia_all_aov <- aov_ez(
+  data = vel[Subcomponent2 == "Piano" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+
+pia_all_aov$anova_table
+pairs(emmeans(pia_all_aov, ~Articulation|Dynamics), adjust = "tukey")
+```
+
+2) Category: All
+
+```{r}
+pia_ca_all_aov <- aov_ez(
   data = vel[Subcomponent2 == "Piano"],
@@
 )
+
+pia_ca_all_aov$anova_table
+pairs(emmeans(pia_ca_all_aov, "Category"), adjust = "tukey")
+```
+
+3) Category: both vs. art_only vs. baseline
+
+```{r}
+pia_ca_par_aov <- aov_ez(
+  data = vel[Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only" | Subcomponent2 == "Piano" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+
+pia_ca_par_aov$anova_table
+pairs(emmeans(pia_ca_par_aov, "Category"), adjust = "tukey")
 ```
@@
 # for each individual
-dt_vel_diff$Articulation <- factor(dt_vel_diff$Articulation, c("Present", "Absent", "Baseline"))
-dt_vel_diff$Dynamics <- factor(dt_vel_diff$Dynamics, c("Present", "Absent", "Baseline"))
+dt_vel_diff_first$Articulation <- factor(dt_vel_diff_first$Articulation, c("Present", "Absent", "Baseline"))
+dt_vel_diff_first$Dynamics <- factor(dt_vel_diff_first$Dynamics, c("Present", "Absent", "Baseline"))
 
-vel_diff_trial <- dt_vel_diff[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
+vel_diff_trial <- dt_vel_diff_first[Subcomponent2 == "FtoP" | Subcomponent2 == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, TrialNr, Stimuli, Articulation, Dynamics, Category, Subcomponent2)]
 
@@
 
-## 2. All - group
+## 2. Group
+
 ```{r vel-diff-all, echo = FALSE}
@@
 ```{r vel-diff-all-plot,  echo = FALSE}
-ggplot(vel_diff, aes(x = Dynamics, y = Mean, color = Articulation))+ 
-  geom_boxplot(position = position_dodge2(preserve = "single")) +
-  geom_point(binaxis = "y", stackdir = "center", position = position_dodge(0.75)) +
-  facet_grid( ~ Subcomponent2) +
+ggplot(vel_diff[Subcomponent2 == "FtoP"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - Forte to Piano") +
+  theme_pubr()
+
+ggplot(vel_diff[Subcomponent2 == "PtoF"], aes(x = Dynamics, y = Mean, color = Articulation))+ 
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - Piano to Forte") +
   theme_pubr()
@@
 ## 4. Stats
+### FtoP
 
-1) FtoP
+1) Articulation x Dynamics
 
 ```{r}
-aov_ez(
+ftop_all_aov <- aov_ez(
+  data = vel_diff[Subcomponent2 == "FtoP" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+
+ftop_all_aov$anova_table
+pairs(emmeans(ftop_all_aov, ~Articulation|Dynamics), adjust = "tukey")
+```
+
+2) Category: All
+
+```{r}
+ftop_ca_all_aov <- aov_ez(
   data = vel_diff[Subcomponent2 == "FtoP"],
@@
 )
+
+ftop_ca_all_aov$anova_table
+pairs(emmeans(ftop_ca_all_aov, "Category"), adjust = "tukey")
 ```
 
-2) PtoF
+3) Category: both vs. art_only vs. baseline
 
 ```{r}
-aov_ez(
+ftop_ca_par_aov <- aov_ez(
+  data = vel_diff[Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+
+ftop_ca_par_aov$anova_table
+pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
+```
+
+### PtoF
+
+1) Articulation x Dynamics
+
+```{r}
+ptof_all_aov <- aov_ez(
+  data = vel_diff[Subcomponent2 == "PtoF" & Category != "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Articulation", "Dynamics")
+)
+
+ptof_all_aov$anova_table
+pairs(emmeans(ptof_all_aov, ~Articulation|Dynamics), adjust = "tukey")
+```
+
+2) Category: All
+
+```{r}
+ptof_ca_all_aov <- aov_ez(
   data = vel_diff[Subcomponent2 == "PtoF"],
@@
 )
+
+ptof_ca_all_aov$anova_table
+pairs(emmeans(ptof_ca_all_aov, "Category"), adjust = "tukey")
 ```
+
+3) Category: both vs. art_only vs. baseline
+
+```{r}
+ptof_ca_par_aov <- aov_ez(
+  data = vel_diff[Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
+  id = "SubNr",
+  dv = "Mean",
+  within = c("Category")
+)
+
+ptof_ca_par_aov
+pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
+```

### 2022-10-14 analysis/baseline.Rmd
update analysis with baseline
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-10-07 analysis/baseline.Rmd
@@
 - Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
+
+## Experimental Design
+
+- Independent variables: Articulation (2 levels: Present or Absent) x Dynamics (2 levels: Present or Absent)
+- Dependent variables: Interonset intervals (IOIs), Key-overlap time (KOT), Key velocity (KV) and KV Difference
+
+1. Main predictions ([preregistration](https://osf.io/qbzwt))
+
+- If students performed the piece without articulation (i.e., only dynamics is implemented in the recordings), participants should exaggerate articulation more (i.e., producing longer legato and shorter staccato).
+- If students performed the piece without dynamics (i.e., only articulation is implemented in the recordings), participants should exaggerate dynamics more (i.e., producing louder forte and softer piano).
+- If students performed the piece without articulation and dynamics (i.e., neither articulation nor dynamics is implemented in the recordings), participants should exaggerate both articulation and dynamics more.
+
+2. Excluded participants
+
+- NA
+
+3. Stats packages
+
+- [afex](https://github.com/singmann/afex) for ANOVAs
 
@@
 )
+
+ioi_ca_par_aov$anova_table
+pairs(emmeans(ioi_ca_par_aov, "Category"), adjust = "tukey")
 ```
@@
 ```{r}
-aov_ez(
+var_stats <- aov_ez(
   data = ioi,
@@
 )
+
+var_stats$anova_table
 ```

### 2022-10-14 analysis/baseline.R
export R script
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,36 +1,4 @@
----
-title: 'Music adaptation: Analysis (only first instances)'
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of analysis which only include the performance with the first instance of each student (category).
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-## Experimental Design
-Note: Here, we only look at the first instance of each student (category).
-
-- Independent variables: Articulation (2 levels: Present or Absent) x Dynamics (2 levels: Present or Absent)
-- Dependent variables: Interonset intervals (IOIs), Key-overlap time (KOT), Key velocity (KV) and KV Difference
-
-1. Main predictions ([preregistration](https://osf.io/qbzwt))
- 
-- If students performed the piece without articulation (i.e., only dynamics is implemented in the recordings), participants should exaggerate articulation more (i.e., producing longer legato and shorter staccato).
-- If students performed the piece without dynamics (i.e., only articulation is implemented in the recordings), participants should exaggerate dynamics more (i.e., producing louder forte and softer piano).
-- If students performed the piece without articulation and dynamics (i.e., neither articulation nor dynamics is implemented in the recordings), participants should exaggerate both articulation and dynamics more.
-
-2. Excluded participants
-
-- NA
-
-3. Stats packages
-
-- [afex](https://github.com/singmann/afex) for ANOVAs
-
-```{r setup, include = FALSE}
+## ----setup, include = FALSE-----------------------------------------------------------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -42,9 +10,9 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 # plot
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
-```
 
[…]

### 2022-10-14 analysis/onlyfirst.R

@@ -1,36 +1,4 @@
----
-title: 'Music adaptation: Analysis (only first instances)'
-output:
-  html_notebook: default
-editor_options: 
-  chunk_output_type: inline
----
-
-Description: This is a summary of analysis which only include the performance with the first instance of each student (category).
-
-- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`
-
-## Experimental Design
-Note: Here, we only look at the first instance of each student (category).
-
-- Independent variables: Articulation (2 levels: Present or Absent) x Dynamics (2 levels: Present or Absent)
-- Dependent variables: Interonset intervals (IOIs), Key-overlap time (KOT), Key velocity (KV) and KV Difference
-
-1. Main predictions ([preregistration](https://osf.io/qbzwt))
- 
-- If students performed the piece without articulation (i.e., only dynamics is implemented in the recordings), participants should exaggerate articulation more (i.e., producing longer legato and shorter staccato).
-- If students performed the piece without dynamics (i.e., only articulation is implemented in the recordings), participants should exaggerate dynamics more (i.e., producing louder forte and softer piano).
-- If students performed the piece without articulation and dynamics (i.e., neither articulation nor dynamics is implemented in the recordings), participants should exaggerate both articulation and dynamics more.
-
-2. Excluded participants
-
-- NA
-
-3. Stats packages
-
-- [afex](https://github.com/singmann/afex) for ANOVAs
-
-```{r setup, include = FALSE}
+## ----setup, include = FALSE--------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -42,9 +10,9 @@ if (!require("afex")) {install.packages("afex"); require("afex")}
 if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 # plot
[…]

### 2022-10-17 analysis/baseline.Rmd
export R script
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>
correct errors for ptof category analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ptof_ca_par_aov <- aov_ez(
-  data = vel_diff[Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only" | Subcomponent2 == "FtoP" & Category == "baseline"],
+  data = vel_diff[Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only" | Subcomponent2 == "PtoF" & Category == "baseline"],
   id = "SubNr",
@@
 ptof_ca_par_aov
-pairs(emmeans(ftop_ca_par_aov, "Category"), adjust = "tukey")
+pairs(emmeans(ptof_ca_par_aov, "Category"), adjust = "tukey")
 ```
+
+```{r export, include = FALSE}
+knitr::purl("baseline.Rmd")
+```

### 2022-10-17 analysis/baseline.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE-----------------------------------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -12,7 +12,7 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE------------------------------------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
 data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
 data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
@@ -23,7 +23,7 @@ data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
 data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 
 
-## ----prep, include = FALSE------------------------------------------------------------------------------------
+## ----prep, include = FALSE----------------------------------------------------
 # read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
 dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
@@ -70,7 +70,7 @@ dt_vel <- rbind(dt_vel, dt_bl_vel)
 dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
 
 
-## ----ioi------------------------------------------------------------------------------------------------------
+## ----ioi----------------------------------------------------------------------
 # for each individual
 dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
 dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
@@ -82,17 +82,17 @@ setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
 ioi
 
 
-## ----ioi-plot, fig.width = 6----------------------------------------------------------------------------------
+## ----ioi-plot, fig.width = 6--------------------------------------------------
[…]

### 2022-10-17 analysis/onlyfirst.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE--------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -12,7 +12,7 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE---------------------------------
+## ----file, include = FALSE----------------------------------------------------
 data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
 data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
 data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
@@ -23,7 +23,7 @@ data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
 data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 
 
-## ----prep, include = FALSE---------------------------------
+## ----prep, include = FALSE----------------------------------------------------
 # read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
 dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
@@ -70,7 +70,7 @@ dt_vel <- rbind(dt_vel, dt_bl_vel)
 dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
 
 
-## ----------------------------------------------------------
+## -----------------------------------------------------------------------------
 # select only performances for the first instance of each student (category)
 info <- dt_ioi[, .(N = .N), by = .(SubNr, TrialNr, Stimuli, Category)]
 info
@@ -99,7 +99,7 @@ dt_vel_diff_first <- data.table()
 dt_vel_diff_first <- first(dt_vel_diff, dt_vel_diff_first)
 
 
-## ----ioi---------------------------------------------------
+## ----ioi----------------------------------------------------------------------
 # for each individual
 dt_ioi_first$Articulation <- factor(dt_ioi_first$Articulation, c("Present", "Absent", "Baseline"))
 dt_ioi_first$Dynamics <- factor(dt_ioi_first$Dynamics, c("Present", "Absent", "Baseline"))
[…]

### 2022-10-17 paper/manuscript.Rmd
update results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_word
 ---
@@
 if (!require("magick")) {install.packages("magick"): require("magick")}
+library(ggsci)
 ```
@@
 The experiment was programmed in Max/MSP (8.1.11; https://cycling74.com/products/max) on a Mac Book Pro with Mac OS X Catalina 10.15.7.
-A weighted Yamaha MIDI digital piano was used to record participants' performance.
+A weighted Yamaha MIDI digital piano was used to record participants' performances.
 The pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers.
@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Two expressive notations (i.e., articulation and dynamics) were added as *Fig 1*.
+Two expressive notations (i.e., articulation and dynamics) were added as *Fig \ref{fig:stim}*.
 
@@
 
-<!-- ```{r stim-1, out.width = "100%", fig.cap = "\\label{fig:stimuli}Sheet music.
+```{r stimli, out.width = "100%", fig.cap = "\\label{fig:stim}Sheet music.
 Articulation.
@@
 For dynamics notation, the symbol `f' denotes forte and the symbol `p' denotes piano.
-For data analysis, only the 8th notes were included.", echo = FALSE} -->
-<!-- stim <- image_border(image_read(here("paper/image", "stim_m.png")), "#FFFFFF", "140x140") -->
-<!-- stim -->
-<!-- ``` -->
+For data analysis, only the 8th notes were included.", echo = FALSE}
+here::i_am("README.md")
+stim <- image_border(image_read(here("paper/image", "stim_m.png")), "#FFFFFF", "140x140")
+stim
+```
 
@@
 A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
-Sheet music (*Fig 1*) was displayed in front of the participants.
-Also, they were told to perform the piece expressively with their interpretation.
+Sheet music (*Fig \ref{fig:stim}*) was displayed in front of the participants.
+Also, they were told to perform the piece expressively with their interpretation and do their best as a performer.
 This instruction was given to make sure that they paid attention to expressive aspects of the performance.
@@
 We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.
-absent) and Dynamics (present or absent).
+absent) and Dynamics (present or absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
 The *aov_ez* function in the *afex* R package was used for a repeated-measures ANOVA.
@@
 
+All effects are reported as significant at *p* < .05. For KOT, KV and KV Difference, we performed two-way ANOVAs separately for each subcomponent (i.e., legato, staccato, forte, piano).
+
+```{r results-1, include = FALSE}
+source(here("analysis", "baseline.R"), chdir = TRUE)
+```
+
+## IOIs
+
+Neither main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* = `r ioi_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* = `r ioi_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* = `r ioi_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
+Therefore, participants did not change the tempo depending on the types of the recordings (*Fig \ref{fig:ioi-1}*).
+
+```{r ioi-plot, include = FALSE}
+ioi_all_plot <- ggplot(ioi[Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_all_plot.png"), plot = ioi_all_plot, dpi = 300, height = 5)
+img_ioi_all_plot <- image_read(here("paper/image", "ioi_all_plot.png"), density = 300)
+```
+
+## KOT
+### Legato
+
+There was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* = `r leg_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* = `r leg_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* = `r leg_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
+Participants produced longer legato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, left).
+
+### Staccato
+
+There was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* = `r sta_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r sta_all_aov$anova_table[2,1]`, `r sta_all_aov$anova_table[2,2]`) = `r sta_all_aov$anova_table[2,4]`, *p* = `r sta_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r sta_all_aov$anova_table[3,1]`, `r sta_all_aov$anova_table[3,2]`) = `r sta_all_aov$anova_table[3,4]`, *p* = `r sta_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[3,5])`).
+Participants produced shorter staccato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, right).
+
+```{r kot-plot, include = FALSE}
+leg_all_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KOT (ms)", title = "KOT - Legato") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+sta_all_plot <- ggplot(kot[Subcomponent1 == "Staccato" & Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KOT (ms)", title = "KOT - Staccato") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "leg_all_plot.png"), plot = leg_all_plot, dpi = 300, height = 5)
+img_leg_all_plot <- image_read(here("paper/image", "leg_all_plot.png"), density = 300)
+
+ggsave(here("paper/image", "sta_all_plot.png"), plot = sta_all_plot, dpi = 300, height = 5)
+img_sta_all_plot <- image_read(here("paper/image", "sta_all_plot.png"), density = 300)
+```
+
+## KV
+### Forte
+
+Neither main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* = `r for_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* = `r for_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* = `r for_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
+Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (*Fig \ref{fig:vel-1}*, left).
+
+### Piano
+
+There was a significant main effect of Articulation (*F*(`r pia_all_aov$anova_table[1,1]`, `r pia_all_aov$anova_table[1,2]`) = `r pia_all_aov$anova_table[1,4]`, *p* = `r pia_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[1,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r pia_all_aov$anova_table[2,1]`, `r pia_all_aov$anova_table[2,2]`) = `r pia_all_aov$anova_table[2,4]`, *p* = `r pia_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r pia_all_aov$anova_table[3,1]`, `r pia_all_aov$anova_table[3,2]`) = `r pia_all_aov$anova_table[3,4]`, *p* = `r pia_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[3,5])`).
+Participants produced softer piano when listening to the recordings where articulation was implemented (*Fig \ref{fig:vel-1}*, right).
+
+```{r vel-plot, include = FALSE}
+for_all_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category != "baseline"], aes(x = Dynamics, y = Mean, color = Articulation)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KV (0-127)", title = "KV - Forte") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+pia_all_plot <- ggplot(vel[Subcomponent2 == "Piano" & Category != "baseline"], aes(x = Dynamics, y = Mean, color = Articulation)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KV (0-127)", title = "KV - Piano") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "for_all_plot.png"), plot = for_all_plot, dpi = 300, height = 5)
+img_for_all_plot <- image_read(here("paper/image", "for_all_plot.png"), density = 300)
+
+ggsave(here("paper/image", "pia_all_plot.png"), plot = pia_all_plot, dpi = 300, height = 5)
+img_pia_all_plot <- image_read(here("paper/image", "pia_all_plot.png"), density = 300)
+```
+
+## KV Difference
+### Forte to Piano
+
+Neither main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* = `r ftop_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* = `r ftop_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* = `r ftop_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
+
+### Piano to Forte
+
+Neither main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* = `r ptof_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* = `r ptof_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* = `r ptof_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
+
+These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of the recording (*Fig \ref{fig:vel-diff-1}*).
+
+```{r vel-diff-plot, include = FALSE}
+ftop_all_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category != "baseline"], aes(x = Dynamics, y = Mean, color = Articulation)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - FtoP") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+ptof_all_plot <- ggplot(vel_diff[Subcomponent2 == "PtoF" & Category != "baseline"], aes(x = Dynamics, y = Mean, color = Articulation)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoP") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ftop_all_plot.png"), plot = ftop_all_plot, dpi = 300, height = 5)
+img_ftop_all_plot <- image_read(here("paper/image", "ftop_all_plot.png"), density = 300)
+
+ggsave(here("paper/image", "ptof_all_plot.png"), plot = ptof_all_plot, dpi = 300, height = 5)
+img_ptof_all_plot <- image_read(here("paper/image", "ptof_all_plot.png"), density = 300)
+```
+
+```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}IOIs (ms). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+img_ioi_1 <- image_scale(img_ioi_all_plot, "2200")
+image_write(img_ioi_1, path = here("paper/image", "Fig2.tiff"), format = "tiff")
+img_ioi_1
+```
+
+```{r plot-kot-1, out.width = "100%", fig.cap = "\\label{fig:kot-1}KOT(ms) for each subcomponent; legato (left) and staccato (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_kot_1 <- image_append(c(img_leg_all_plot, img_sta_all_plot))
+img_kot_1 <- image_scale(img_kot_1, "2200")
+image_write(img_kot_1, path = here("paper/image", "Fig3.tiff"), format = "tiff")
+img_kot_1
+```
+
+```{r plot-vel-1, out.width = "100%", fig.cap = "\\label{fig:vel-1}KV (0-127) for each subcomponent; forte (left) and piano (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_vel_1 <- image_append(c(img_for_all_plot, img_pia_all_plot))
+img_vel_1 <- image_scale(img_vel_1, "2200")
+image_write(img_vel_1, path = here("paper/image", "Fig4.tiff"), format = "tiff")
+img_vel_1
+```
+
+```{r plot-vel-diff-1, out.width = "100%", fig.cap = "\\label{fig:vel-diff-1}KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_vel_diff_1 <- image_append(c(img_ftop_all_plot, img_ptof_all_plot))
+img_vel_diff_1 <- image_scale(img_vel_diff_1, "2200")
+image_write(img_vel_diff_1, path = here("paper/image", "Fig5.tiff"), format = "tiff")
+img_vel_diff_1
+```
+
+## Comparison with baseline performance
+In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced before listening to any of the recordings), we performed one-way repeated-measures ANOVAs.
+For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent.
+To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer.
+By comparing with the baseline performance, we can investigate how didactic intentions influence participants' performances.
+
+## IOIs
+
+We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of students' skill levels.
+We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
+
+There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* = `r ioi_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (*Fig \ref{fig:ioi-2}*).
+
+```{r ioi-bl-plot, include = FALSE}
+ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_ca_plot.png"), plot = ioi_ca_plot, dpi = 300, height = 5)
+img_ioi_ca_plot <- image_read(here("paper/image", "ioi_ca_plot.png"), density = 300)
+```
+
+## KOT
+
+We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not on each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
+
+### Legato
+
+There was a significant main effect of Category (*F*(`r leg_ca_par_aov$anova_table[1,1]`, `r leg_ca_par_aov$anova_table[1,2]`) = `r leg_ca_par_aov$anova_table[1,4]`, *p* = `r leg_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001 , baseline and dynamics-only: *p* = .008), suggesting that participants produced longer legato when they had an intention to teach (*Fig \ref{fig:kot-2}*, left).
+
+### Staccato
+
+There was no significant main effect of Category (*F*(`r sta_ca_par_aov$anova_table[1,1]`, `r sta_ca_par_aov$anova_table[1,2]`) = `r sta_ca_par_aov$anova_table[1,4]`, *p* = `r sta_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (*Fig \ref{fig:kot-2}*, right).
+
+```{r kot-bl-plot, include = FALSE}
+leg_ca_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category == "baseline" | Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(x = "Category", y = "KOT (ms)", title = "KOT - Legato") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+sta_ca_plot <- ggplot(kot[Subcomponent1 == "Staccato" & Category == "baseline" | Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(x = "Category", y = "KOT (ms)", title = "KOT - Staccato") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "leg_ca_plot.png"), plot = leg_ca_plot, dpi = 300, height = 5)
+img_leg_ca_plot <- image_read(here("paper/image", "leg_ca_plot.png"), density = 300)
+
+ggsave(here("paper/image", "sta_ca_plot.png"), plot = sta_ca_plot, dpi = 300, height = 5)
+img_sta_ca_plot <- image_read(here("paper/image", "sta_ca_plot.png"), density = 300)
+```
+
+## KV
+
+We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+### Forte
+
+There was no significant main effect of Category (*F*(`r for_ca_par_aov$anova_table[1,1]`, `r for_ca_par_aov$anova_table[1,2]`) = `r for_ca_par_aov$anova_table[1,4]`, *p* = `r for_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (*Fig \ref{fig:vel-2}*, left).
+
+### Piano
+
+There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* = `r pia_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05 , baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had an intention to teach (*Fig \ref{fig:vel-2}*, right).
+
+```{r vel-bl-plot, include = FALSE}
+for_ca_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category == "baseline" | Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(x = "Category", y = "KV (0-127)", title = "KV - Forte") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+pia_ca_plot <- ggplot(vel[Subcomponent2 == "Piano" & Category == "baseline" | Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(x = "Category", y = "KV (0-127)", title = "KV - Piano") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "for_ca_plot.png"), plot = for_ca_plot, dpi = 300, height = 5)
+img_for_ca_plot <- image_read(here("paper/image", "for_ca_plot.png"), density = 300)
+
+ggsave(here("paper/image", "pia_ca_plot.png"), plot = pia_ca_plot, dpi = 300, height = 5)
+img_pia_ca_plot <- image_read(here("paper/image", "pia_ca_plot.png"), density = 300)
+```
+
+## KV Difference
+
+We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+### Forte to Piano
+
+There was no significant main effect of Category (*F*(`r ftop_ca_par_aov$anova_table[1,1]`, `r ftop_ca_par_aov$anova_table[1,2]`) = `r ftop_ca_par_aov$anova_table[1,4]`, *p* = `r ftop_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
+
+### Piano to Forte
+
+There was no significant main effect of Category (*F*(`r ptof_ca_par_aov$anova_table[1,1]`, `r ptof_ca_par_aov$anova_table[1,2]`) = `r ptof_ca_par_aov$anova_table[1,4]`, *p* = `r ptof_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
+
+These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had an intention to teach (*Fig \ref{fig:vel-diff-2}*).
+
+```{r vel-diff-bl-plot, include = FALSE}
+ftop_ca_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category == "baseline" | Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - FtoP") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+ptof_ca_plot <- ggplot(vel_diff[Subcomponent2 == "PtoF" & Category == "baseline" | Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - PtoF") +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ftop_ca_plot.png"), plot = ftop_ca_plot, dpi = 300, height = 5)
+img_ftop_ca_plot <- image_read(here("paper/image", "ftop_ca_plot.png"), density = 300)
+
+ggsave(here("paper/image", "ptof_ca_plot.png"), plot = ptof_ca_plot, dpi = 300, height = 5)
+img_ptof_ca_plot <- image_read(here("paper/image", "ptof_ca_plot.png"), density = 300)
+```
+
+```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+img_ioi_2 <- image_scale(img_ioi_ca_plot, "2200")
+image_write(img_ioi_2, path = here("paper/image", "Fig6.tiff"), format = "tiff")
+img_ioi_2
+```
+
+```{r plot-kot-2, out.width = "100%", fig.cap = "\\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_kot_2 <- image_append(c(img_leg_ca_plot, img_sta_ca_plot))
+img_kot_2 <- image_scale(img_kot_2, "2200")
+image_write(img_kot_2, path = here("paper/image", "Fig7.tiff"), format = "tiff")
+img_kot_2
+```
+
+```{r plot-vel-2, out.width = "100%", fig.cap = "\\label{fig:vel-2}Comparison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_vel_2 <- image_append(c(img_for_ca_plot, img_pia_ca_plot))
+img_vel_2 <- image_scale(img_vel_2, "2200")
+image_write(img_vel_2, path = here("paper/image", "Fig8.tiff"), format = "tiff")
+img_vel_2
+```
+
+```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Comparison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_vel_diff_2 <- image_append(c(img_ftop_ca_plot, img_ptof_ca_plot))
+img_vel_diff_2 <- image_scale(img_vel_diff_2, "2200")
+image_write(img_vel_diff_2, path = here("paper/image", "Fig9.tiff"), format = "tiff")
+img_vel_diff_2
+```
+
+```{r results-2, include = FALSE}
+source(here("analysis", "onlyfirst.R"), chdir = TRUE)
+```
+
 # Discussion
+
+# References
+
+# Acknowledgement

### 2022-10-17 paper/manuscript.tex
update results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Two expressive notations (i.e., articulation and dynamics) were added as \emph{Fig 1}.
+Two expressive notations (i.e., articulation and dynamics) were added as \emph{Fig \ref{fig:stim}}.
 
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimli-1} \caption{\label{fig:stim}Sheet music.
+Articulation.
+For articulation notation, the curved line (slur) indicates legato and the dots indicate staccato.
+For dynamics notation, the symbol `f' denotes forte and the symbol `p' denotes piano.
+For data analysis, only the 8th notes were included.}\label{fig:stimli}
+\end{figure}
+
 \hypertarget{procedure}{%
@@
 A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
-Sheet music (\emph{Fig 1}) was displayed in front of the participants.
-Also, they were told to perform the piece expressively with their interpretation.
+Sheet music (\emph{Fig \ref{fig:stim}}) was displayed in front of the participants.
+Also, they were told to perform the piece expressively with their interpretation and do their best as a performer.
 This instruction was given to make sure that they paid attention to expressive aspects of the performance.
@@
 
-We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present or absent).
+We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present or absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
 The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
@@
 
+All effects are reported as significant at \emph{p} \textless{} .05. For KOT, KV and KV Difference, we performed two-way ANOVAs separately for each subcomponent (i.e., legato, staccato, forte, piano).
+
+\hypertarget{iois}{%
+\subsection{IOIs}\label{iois}}
+
+Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.11, \(\eta_G^2\) = 0.010) and Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.09, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
+Therefore, participants did not change the tempo depending on the types of the recordings (\emph{Fig \ref{fig:ioi-1}}).
+
+\hypertarget{kot}{%
+\subsection{KOT}\label{kot}}
+
+\hypertarget{legato}{%
+\subsubsection{Legato}\label{legato}}
+
+There was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.03, \(\eta_G^2\) = 0.002).
+However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002).
+Participants produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).
+
+\hypertarget{staccato}{%
+\subsubsection{Staccato}\label{staccato}}
+
+There was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.04, \(\eta_G^2\) = 0.009).
+However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.09, \(\eta_G^2\) = 0.001).
+Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).
+
+\hypertarget{kv}{%
+\subsection{KV}\label{kv}}
+
+\hypertarget{forte}{%
+\subsubsection{Forte}\label{forte}}
+
+Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.21, \(\eta_G^2\) = 0.002) and Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
+Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (\emph{Fig \ref{fig:vel-1}}, left).
+
+\hypertarget{piano}{%
+\subsubsection{Piano}\label{piano}}
+
+There was a significant main effect of Articulation (\emph{F}(1, 19) = 7.18, \emph{p} = 0.01, \(\eta_G^2\) = 0.060).
+However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000).
+Participants produced softer piano when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).
+
+\hypertarget{kv-difference}{%
+\subsection{KV Difference}\label{kv-difference}}
+
+\hypertarget{forte-to-piano}{%
+\subsubsection{Forte to Piano}\label{forte-to-piano}}
+
+Neither main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.08, \(\eta_G^2\) = 0.012) and Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.34, \(\eta_G^2\) = 0.001).
+
+\hypertarget{piano-to-forte}{%
+\subsubsection{Piano to Forte}\label{piano-to-forte}}
+
+Neither main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.15, \(\eta_G^2\) = 0.005) and Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
+
+These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of the recording (\emph{Fig \ref{fig:vel-diff-1}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-1-1} \caption{\label{fig:kot-1}KOT(ms) for each subcomponent; legato (left) and staccato (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-1-1} \caption{\label{fig:vel-1}KV (0-127) for each subcomponent; forte (left) and piano (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1-1} \caption{\label{fig:vel-diff-1}KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-1}
+\end{figure}
+
+\hypertarget{comparison-with-baseline-performance}{%
+\subsection{Comparison with baseline performance}\label{comparison-with-baseline-performance}}
+
+In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced before listening to any of the recordings), we performed one-way repeated-measures ANOVAs.
+For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent.
+To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer.
+By comparing with the baseline performance, we can investigate how didactic intentions influence participants' performances.
+
+\hypertarget{iois-1}{%
+\subsection{IOIs}\label{iois-1}}
+
+We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of students' skill levels.
+We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
+
+There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse--Geisser corrected), suggesting that participants kept the tempo regardless of whether they had an intention to teach (\emph{Fig \ref{fig:ioi-2}}).
+
+\hypertarget{kot-1}{%
+\subsection{KOT}\label{kot-1}}
+
+We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not on each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
+
+\hypertarget{legato-1}{%
+\subsubsection{Legato}\label{legato-1}}
+
+There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} = 0.00, \(\eta_G^2\) = 0.060; Greenhouse--Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
+
+\hypertarget{staccato-1}{%
+\subsubsection{Staccato}\label{staccato-1}}
+
+There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse--Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, right).
+
+\hypertarget{kv-1}{%
+\subsection{KV}\label{kv-1}}
+
+We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+\hypertarget{forte-1}{%
+\subsubsection{Forte}\label{forte-1}}
+
+There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse--Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, left).
+
+\hypertarget{piano-1}{%
+\subsubsection{Piano}\label{piano-1}}
+
+There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.01, \(\eta_G^2\) = 0.072; Greenhouse--Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
+
+\hypertarget{kv-difference-1}{%
+\subsection{KV Difference}\label{kv-difference-1}}
+
+We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+\hypertarget{forte-to-piano-1}{%
+\subsubsection{Forte to Piano}\label{forte-to-piano-1}}
+
+There was no significant main effect of Category (\emph{F}(1.53, 29.15) = 1.73, \emph{p} = 0.20, \(\eta_G^2\) = 0.014; Greenhouse--Geisser corrected).
+
+\hypertarget{piano-to-forte-1}{%
+\subsubsection{Piano to Forte}\label{piano-to-forte-1}}
+
+There was no significant main effect of Category (\emph{F}(1.28, 24.39) = 2.27, \emph{p} = 0.14, \(\eta_G^2\) = 0.013; Greenhouse--Geisser corrected).
+
+These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-diff-2}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comaprison with the baseline performance in terms of IOIs (ms).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comaprison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Comaprison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Comaprison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-2}
+\end{figure}
+
 \hypertarget{discussion}{%
@@
 
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
+
 \end{document}

### 2022-10-31 paper/manuscript.Rmd
update results (analysis, floating points)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
     email         : "tominaga_atsuko@phd.ceu.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - "Conceptualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
+    # role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+    #   - "Conceptualization"
+    #   - "Writing - Original Draft Preparation"
+    #   - "Writing - Review & Editing"
   - name          : "Günther Knoblich"
     affiliation   : "1"
-    role:
-      - "Writing - Review & Editing"
-      - "Supervision"
+    # role:
+    #   - "Writing - Review & Editing"
+    #   - "Supervision"
   - name          : "Natalie Sebanz"
     affiliation   : "1"
-    role:
-      - "Writing - Review & Editing"
-      - "Supervision"
+    # role:
+    #   - "Writing - Review & Editing"
+    #   - "Supervision"
 
@@
 
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+# authornote: |
 
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
+  AAA
 
@@
 classoption       : "man"
-output            : papaja::apa6_word
+output            : papaja::apa6_pdf
 ---
@@
 if (!require("magick")) {install.packages("magick"): require("magick")}
-library(ggsci)
+if (!require("ggsci")) {install.packages("ggsci"): require("ggsci")}
 ```
@@
 ```{r methods, include = FALSE} 
-source(here("analysis/demographics", "questionnaire.R"), chdir = TRUE)
+here::i_am("README.md")
+source(here("analysis/demographics" , "questionnaire.R"), chdir = TRUE)
 source(here("analysis/preprocessor", "filtering.R"), chdir = TRUE)
@@
 ```{r results-1, include = FALSE}
+here::i_am("README.md")
 source(here("analysis", "baseline.R"), chdir = TRUE)
@@
 
-Neither main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* = `r ioi_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* = `r ioi_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* = `r ioi_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
+Neither main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
 Therefore, participants did not change the tempo depending on the types of the recordings (*Fig \ref{fig:ioi-1}*).
@@
 
-There was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* = `r leg_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
-However, there was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* = `r leg_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* = `r leg_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
+There was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* `r if(leg_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* `r if(leg_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
 Participants produced longer legato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, left).
@@
 
-There was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* = `r sta_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
-However, there was no significant main effect of Dynamics (*F*(`r sta_all_aov$anova_table[2,1]`, `r sta_all_aov$anova_table[2,2]`) = `r sta_all_aov$anova_table[2,4]`, *p* = `r sta_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r sta_all_aov$anova_table[3,1]`, `r sta_all_aov$anova_table[3,2]`) = `r sta_all_aov$anova_table[3,4]`, *p* = `r sta_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[3,5])`).
+There was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* `r if(sta_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r sta_all_aov$anova_table[2,1]`, `r sta_all_aov$anova_table[2,2]`) = `r sta_all_aov$anova_table[2,4]`, *p* `r if(sta_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r sta_all_aov$anova_table[3,1]`, `r sta_all_aov$anova_table[3,2]`) = `r sta_all_aov$anova_table[3,4]`, *p* `r if(sta_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[3,5])`).
 Participants produced shorter staccato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, right).
@@
 
-Neither main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* = `r for_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* = `r for_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* = `r for_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
+Neither main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
 Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (*Fig \ref{fig:vel-1}*, left).
@@
 
-There was a significant main effect of Articulation (*F*(`r pia_all_aov$anova_table[1,1]`, `r pia_all_aov$anova_table[1,2]`) = `r pia_all_aov$anova_table[1,4]`, *p* = `r pia_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[1,5])`).
-However, there was no significant main effect of Dynamics (*F*(`r pia_all_aov$anova_table[2,1]`, `r pia_all_aov$anova_table[2,2]`) = `r pia_all_aov$anova_table[2,4]`, *p* = `r pia_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r pia_all_aov$anova_table[3,1]`, `r pia_all_aov$anova_table[3,2]`) = `r pia_all_aov$anova_table[3,4]`, *p* = `r pia_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[3,5])`).
+There was a significant main effect of Articulation (*F*(`r pia_all_aov$anova_table[1,1]`, `r pia_all_aov$anova_table[1,2]`) = `r pia_all_aov$anova_table[1,4]`, *p* `r if(pia_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[1,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r pia_all_aov$anova_table[2,1]`, `r pia_all_aov$anova_table[2,2]`) = `r pia_all_aov$anova_table[2,4]`, *p* `r if(pia_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r pia_all_aov$anova_table[3,1]`, `r pia_all_aov$anova_table[3,2]`) = `r pia_all_aov$anova_table[3,4]`, *p* `r if(pia_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[3,5])`).
 Participants produced softer piano when listening to the recordings where articulation was implemented (*Fig \ref{fig:vel-1}*, right).
@@
 
-Neither main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* = `r ftop_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* = `r ftop_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* = `r ftop_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
+Neither main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* `r if(ftop_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* `r if(ftop_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* `r if(ftop_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
 
@@
 
-Neither main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* = `r ptof_all_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* = `r ptof_all_aov$anova_table[2,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* = `r ptof_all_aov$anova_table[3,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
+Neither main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* `r if(ptof_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ptof_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* `r if(ptof_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* `r if(ptof_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
 
@@
 
-There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* = `r ioi_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (*Fig \ref{fig:ioi-2}*).
+There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* `r if(ioi_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (*Fig \ref{fig:ioi-2}*).
 
@@
 
-There was a significant main effect of Category (*F*(`r leg_ca_par_aov$anova_table[1,1]`, `r leg_ca_par_aov$anova_table[1,2]`) = `r leg_ca_par_aov$anova_table[1,4]`, *p* = `r leg_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', leg_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001 , baseline and dynamics-only: *p* = .008), suggesting that participants produced longer legato when they had an intention to teach (*Fig \ref{fig:kot-2}*, left).
+There was a significant main effect of Category (*F*(`r leg_ca_par_aov$anova_table[1,1]`, `r leg_ca_par_aov$anova_table[1,2]`) = `r leg_ca_par_aov$anova_table[1,4]`, *p* `r if(leg_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001 , baseline and dynamics-only: *p* .008), suggesting that participants produced longer legato when they had an intention to teach (*Fig \ref{fig:kot-2}*, left).
 
@@
 
-There was no significant main effect of Category (*F*(`r sta_ca_par_aov$anova_table[1,1]`, `r sta_ca_par_aov$anova_table[1,2]`) = `r sta_ca_par_aov$anova_table[1,4]`, *p* = `r sta_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', sta_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (*Fig \ref{fig:kot-2}*, right).
+There was no significant main effect of Category (*F*(`r sta_ca_par_aov$anova_table[1,1]`, `r sta_ca_par_aov$anova_table[1,2]`) = `r sta_ca_par_aov$anova_table[1,4]`, *p* `r if(sta_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (*Fig \ref{fig:kot-2}*, right).
 
@@
 
-There was no significant main effect of Category (*F*(`r for_ca_par_aov$anova_table[1,1]`, `r for_ca_par_aov$anova_table[1,2]`) = `r for_ca_par_aov$anova_table[1,4]`, *p* = `r for_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', for_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (*Fig \ref{fig:vel-2}*, left).
+There was no significant main effect of Category (*F*(`r for_ca_par_aov$anova_table[1,1]`, `r for_ca_par_aov$anova_table[1,2]`) = `r for_ca_par_aov$anova_table[1,4]`, *p* `r if(for_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (*Fig \ref{fig:vel-2}*, left).
 
@@
 
-There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* = `r pia_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05 , baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had an intention to teach (*Fig \ref{fig:vel-2}*, right).
+There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* `r if(pia_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', pia_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* .05 , baseline and articulation-only: *p* .03), suggesting that participants produced softer piano when they had an intention to teach (*Fig \ref{fig:vel-2}*, right).
 
@@
 
-There was no significant main effect of Category (*F*(`r ftop_ca_par_aov$anova_table[1,1]`, `r ftop_ca_par_aov$anova_table[1,2]`) = `r ftop_ca_par_aov$anova_table[1,4]`, *p* = `r ftop_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ftop_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
+There was no significant main effect of Category (*F*(`r ftop_ca_par_aov$anova_table[1,1]`, `r ftop_ca_par_aov$anova_table[1,2]`) = `r ftop_ca_par_aov$anova_table[1,4]`, *p* `r if(ftop_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
 
@@
 
-There was no significant main effect of Category (*F*(`r ptof_ca_par_aov$anova_table[1,1]`, `r ptof_ca_par_aov$anova_table[1,2]`) = `r ptof_ca_par_aov$anova_table[1,4]`, *p* = `r ptof_ca_par_aov$anova_table[1,6]`, $\eta_G^2$ = `r sprintf('%.3f', ptof_ca_par_aov$anova_table[1,5])`; Greenhouse–Geisser corrected).
+There was no significant main effect of Category (*F*(`r ptof_ca_par_aov$anova_table[1,1]`, `r ptof_ca_par_aov$anova_table[1,2]`) = `r ptof_ca_par_aov$anova_table[1,4]`, *p* `r if(ptof_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
 
@@
 
-```{r results-2, include = FALSE}
-source(here("analysis", "onlyfirst.R"), chdir = TRUE)
-```
+# Discussion
 
-# Discussion
+The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of students.
+We manipulated the implementation of articulation and dynamics in the piece to create four types of artificial recordings.
+These recordings were supposed to be the reflection of students' skill levels.
+In one quarter of the recordings, both articulation and dynamics were implemented
+[SUMMARY].
+We found that 
+Limitation
 

### 2022-10-31 paper/manuscript.tex
update results (analysis, floating points)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
-
-The authors made the following contributions.
-Atsuko Tominaga: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Günther Knoblich: Writing - Review \& Editing, Supervision; Natalie Sebanz: Writing - Review \& Editing, Supervision.
-
 Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria.
@@
 \abstract{%
-One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.
-
-Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.
-
-One sentence clearly stating the \textbf{general problem} being addressed by this particular study.
-
-One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).
-
-Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
-
-One or two sentences to put the results into a more \textbf{general context}.
-
-Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
+AAA
 }
@@
 The experiment was programmed in Max/MSP (8.1.11; \url{https://cycling74.com/products/max}) on a Mac Book Pro with Mac OS X Catalina 10.15.7.
-A weighted Yamaha MIDI digital piano was used to record participants' performance.
+A weighted Yamaha MIDI digital piano was used to record participants' performances.
 The pitch, onset and offset time of each note, and key velocity profiles were obtained from MIDI data using Max/MSP patchers.
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.11, \(\eta_G^2\) = 0.010) and Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.09, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
+Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) or Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
 Therefore, participants did not change the tempo depending on the types of the recordings (\emph{Fig \ref{fig:ioi-1}}).
@@
 
-There was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.03, \(\eta_G^2\) = 0.002).
+There was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
 However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002).
@@
 
-There was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.04, \(\eta_G^2\) = 0.009).
-However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.09, \(\eta_G^2\) = 0.001).
+There was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009).
+However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.093, \(\eta_G^2\) = 0.001).
 Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.21, \(\eta_G^2\) = 0.002) and Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
+Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) or Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
 Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (\emph{Fig \ref{fig:vel-1}}, left).
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.08, \(\eta_G^2\) = 0.012) and Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.34, \(\eta_G^2\) = 0.001).
+Neither main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) or Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.336, \(\eta_G^2\) = 0.001).
 
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.15, \(\eta_G^2\) = 0.005) and Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
+Neither main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.150, \(\eta_G^2\) = 0.005) or Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
 
@@
 
-There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse--Geisser corrected), suggesting that participants kept the tempo regardless of whether they had an intention to teach (\emph{Fig \ref{fig:ioi-2}}).
+There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (\emph{Fig \ref{fig:ioi-2}}).
 
@@
 
-There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} = 0.00, \(\eta_G^2\) = 0.060; Greenhouse--Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
+There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.060; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
 
@@
 
-There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse--Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, right).
+There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, right).
 
@@
 
-There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse--Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, left).
+There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, left).
 
@@
 
-There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.01, \(\eta_G^2\) = 0.072; Greenhouse--Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
+There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} .05 , baseline and articulation-only: \emph{p} .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
 
@@
 
-There was no significant main effect of Category (\emph{F}(1.53, 29.15) = 1.73, \emph{p} = 0.20, \(\eta_G^2\) = 0.014; Greenhouse--Geisser corrected).
+There was no significant main effect of Category (\emph{F}(1.53, 29.15) = 1.73, \emph{p} = 0.20, \(\eta_G^2\) = 0.014; Greenhouse-Geisser corrected).
 
@@
 
-There was no significant main effect of Category (\emph{F}(1.28, 24.39) = 2.27, \emph{p} = 0.14, \(\eta_G^2\) = 0.013; Greenhouse--Geisser corrected).
+There was no significant main effect of Category (\emph{F}(1.28, 24.39) = 2.27, \emph{p} = 0.14, \(\eta_G^2\) = 0.013; Greenhouse-Geisser corrected).
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comaprison with the baseline performance in terms of IOIs (ms).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comaprison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-2}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Comaprison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Comparison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-2}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Comaprison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Comparison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-2}
@@
 \section{Discussion}\label{discussion}}
+
+The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of students.
+We manipulated the implementation of articulation and dynamics in the piece to create four types of artificial recordings.
+These recordings were supposed to be the reflection of students' skill levels.
+In one quarter of the recordings, both articulation and dynamics were implemented
+{[}SUMMARY{]}.
+We found that
+Limitation
 

### 2022-10-31 analysis/baseline.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE------------------------------------------------------------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -12,7 +12,7 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE-------------------------------------------------------------------------------------
 data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
 data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
 data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
@@ -23,7 +23,7 @@ data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
 data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 
 
-## ----prep, include = FALSE----------------------------------------------------
+## ----prep, include = FALSE-------------------------------------------------------------------------------------
 # read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
 dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
@@ -70,7 +70,7 @@ dt_vel <- rbind(dt_vel, dt_bl_vel)
 dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
 
 
-## ----ioi----------------------------------------------------------------------
+## ----ioi-------------------------------------------------------------------------------------------------------
 # for each individual
 dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
 dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
@@ -82,17 +82,17 @@ setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
 ioi
 
 
-## ----ioi-plot, fig.width = 6--------------------------------------------------
+## ----ioi-plot, fig.width = 6-----------------------------------------------------------------------------------
[…]

### 2022-11-01 paper/manuscript.Rmd
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of students.
-We manipulated the implementation of articulation and dynamics in the piece to create four types of artificial recordings.
-These recordings were supposed to be the reflection of students' skill levels.
-In one quarter of the recordings, both articulation and dynamics were implemented
-[SUMMARY].
-We found that 
+We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
+
+We found that expert pianists did not modulate their tempo depending on the skill levels of students.
+Compared with the tempo of the baseline performance, they did not change the tempo either.
+These findings indicated that tempo is not employed specifically for teaching because tempo itself is relevant to neither articulation nor dynamics.
+In music performance, it is possible that changing tempo might give another interpretation of music (e.g., expressive timing).
+
+For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
+This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was not implemented in the recordings.
+However, we did not find the same results in terms of dynamics.
+Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings.
+These findings suggest that participants modulated the articulation-related aspect of performances only.
+
+There are two potential reasons that participants modulated articulation only for teaching purposes.
+One possibility would be that articulation was more important for the piece or the piece itself invites some implementations of articulation.
+In our previous experiment, we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).
+To make the implementation of articulation clear, we particularly selected detached performances for the recordings where articulation was missing.
+Another possibility is that dynamics implementation was not perceived by participants, therefore they did not modulate the dynamic-related aspect of performances for teaching.
+
+One of the limitations of the current study is that ---.
+
+Future research should investigate.
+
 Limitation

### 2022-11-01 paper/manuscript.tex
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of students.
-We manipulated the implementation of articulation and dynamics in the piece to create four types of artificial recordings.
-These recordings were supposed to be the reflection of students' skill levels.
-In one quarter of the recordings, both articulation and dynamics were implemented
-{[}SUMMARY{]}.
-We found that
+We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
+
+We found that expert pianists did not modulate their tempo depending on the skill levels of students.
+Compared with the tempo of the baseline performance, they did not change the tempo either.
+These findings indicated that tempo is not employed specifically for teaching because tempo itself is relevant to neither articulation nor dynamics.
+In music performance, it is possible that changing tempo might give another interpretation of music (e.g., expressive timing).
+
+For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
+This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was not implemented in the recordings.
+However, we did not find the same results in terms of dynamics.
+Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings.
+These findings suggest that participants modulated the articulation-related aspect of performances only.
+
+There are two potential reasons that participants modulated articulation only for teaching purposes.
+One possibility would be that articulation was more important for the piece or the piece itself invites some implementations of articulation.
+In our previous experiment, we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).
+To make the implementation of articulation clear, we particularly selected detached performances for the recordings where articulation was missing.
+Another possibility is that dynamics implementation was not perceived by participants, therefore they did not modulate the dynamic-related aspect of performances for teaching.
+
+One of the limitations of the current study is that ---.
+
+Future research should investigate.
+
 Limitation

### 2022-11-02 paper/manuscript.Rmd
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
+The recordings where the two techniques were implemented were supposed to replicate the performances of students with high skills.
+On the other hand, the recordings where none of the two techniques was implemented were supposed to replicate the performances of students with low skills.
+The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.
 
-We found that expert pianists did not modulate their tempo depending on the skill levels of students.
+We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
 Compared with the tempo of the baseline performance, they did not change the tempo either.
-These findings indicated that tempo is not employed specifically for teaching because tempo itself is relevant to neither articulation nor dynamics.
-In music performance, it is possible that changing tempo might give another interpretation of music (e.g., expressive timing).
+These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
+Especially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
 

### 2022-11-02 paper/manuscript.tex
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
+The recordings where the two techniques were implemented were supposed to replicate the performances of students with high skills.
+On the other hand, the recordings where none of the two techniques was implemented were supposed to replicate the performances of students with low skills.
+The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.
 
@@
 These findings indicated that tempo is not employed specifically for teaching because tempo itself is relevant to neither articulation nor dynamics.
-In music performance, it is possible that changing tempo might give another interpretation of music (e.g., expressive timing).
+Especially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
 

### 2022-11-02 paper/manuscript.Rmd
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 stim <- image_border(image_read(here("paper/image", "stim_m.png")), "#FFFFFF", "140x140")
+image_write(stim, path = here("paper/image", "Fig1.tiff"), format = "tiff")
 stim
+
 ```
@@
 ```{r ioi-bl-plot, include = FALSE}
-ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean, color = Category)) +
+ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
-  scale_color_aaas() +
+  geom_jitter(width = 0.15, alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 ```{r kot-bl-plot, include = FALSE}
-leg_ca_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category == "baseline" | Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean, color = Category)) +
+leg_ca_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category == "baseline" | Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
   labs(x = "Category", y = "KOT (ms)", title = "KOT - Legato") +
-  scale_color_aaas() +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
 
-sta_ca_plot <- ggplot(kot[Subcomponent1 == "Staccato" & Category == "baseline" | Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean, color = Category)) +
+sta_ca_plot <- ggplot(kot[Subcomponent1 == "Staccato" & Category == "baseline" | Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
   labs(x = "Category", y = "KOT (ms)", title = "KOT - Staccato") +
-  scale_color_aaas() +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 ```{r vel-bl-plot, include = FALSE}
-for_ca_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category == "baseline" | Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+for_ca_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category == "baseline" | Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
   labs(x = "Category", y = "KV (0-127)", title = "KV - Forte") +
-  scale_color_aaas() +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
 
-pia_ca_plot <- ggplot(vel[Subcomponent2 == "Piano" & Category == "baseline" | Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+pia_ca_plot <- ggplot(vel[Subcomponent2 == "Piano" & Category == "baseline" | Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
   labs(x = "Category", y = "KV (0-127)", title = "KV - Piano") +
-  scale_color_aaas() +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 ```{r vel-diff-bl-plot, include = FALSE}
-ftop_ca_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category == "baseline" | Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+ftop_ca_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category == "baseline" | Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
   labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - FtoP") +
-  scale_color_aaas() +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
 
-ptof_ca_plot <- ggplot(vel_diff[Subcomponent2 == "PtoF" & Category == "baseline" | Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean, color = Category)) +
+ptof_ca_plot <- ggplot(vel_diff[Subcomponent2 == "PtoF" & Category == "baseline" | Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
   geom_boxplot(outlier.shape = NA) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
   labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - PtoF") +
-  scale_color_aaas() +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
 We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
-Compared with the tempo of the baseline performance, they did not change the tempo either.
+Even compared with the tempo of the baseline performance, they did not change their tempo either.
 These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
@@
 For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
-This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was not implemented in the recordings.
-However, we did not find the same results in terms of dynamics.
+This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings.
+However, we did not find significant results in terms of dynamics.
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings.
-These findings suggest that participants modulated the articulation-related aspect of performances only.
+Participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
 
-There are two potential reasons that participants modulated articulation only for teaching purposes.
-One possibility would be that articulation was more important for the piece or the piece itself invites some implementations of articulation.
-In our previous experiment, we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).
+Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
+One possibility would be that articulation was more important for the piece we selected or the piece itself naturally invited some implementation of articulation.
+In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
+Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it.
+Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants.
+
+One of the limitation of the current study is that .
+One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it.
+After they had an idea of what the model performance (i.e., the goal of the performance) is, they have a chance to listen to the recordings produced by students with various skill levels as we did in the current experiment.
+
+we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).
 To make the implementation of articulation clear, we particularly selected detached performances for the recordings where articulation was missing.

### 2022-11-02 paper/manuscript.tex
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-We found that expert pianists did not modulate their tempo depending on the skill levels of students.
-Compared with the tempo of the baseline performance, they did not change the tempo either.
-These findings indicated that tempo is not employed specifically for teaching because tempo itself is relevant to neither articulation nor dynamics.
+We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
+Even compared with the tempo of the baseline performance, they did not change their tempo either.
+These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
 Especially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
@@
 For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
-This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was not implemented in the recordings.
-However, we did not find the same results in terms of dynamics.
+This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings.
+However, we did not find significant results in terms of dynamics.
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings.
-These findings suggest that participants modulated the articulation-related aspect of performances only.
+Participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
 
-There are two potential reasons that participants modulated articulation only for teaching purposes.
-One possibility would be that articulation was more important for the piece or the piece itself invites some implementations of articulation.
+Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
+One possibility would be that articulation was more important for the piece we selected or the piece itself naturally invited some implementation of articulation.
 In our previous experiment, we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).

### 2022-11-02 paper/manuscript.Rmd
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title             : "The title"
-shorttitle        : "Title"
+title             : "How do expert pianists adapt their didactic demonstration to novices’ skill levels?"
+shorttitle        : "How do expert pianists adapt their didactic demonstration to novices’ skill levels?"
 
@@
 abstract: |
-  AAA
+  Later
 
 keywords          : "keywords"
-wordcount         : "X"
 
-#bibliography      : ""
+bibliography      : "references.bib"
 
@@
 
-One of the limitation of the current study is that .
-One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it.
-After they had an idea of what the model performance (i.e., the goal of the performance) is, they have a chance to listen to the recordings produced by students with various skill levels as we did in the current experiment.
+One of the limitations of the current study is that participants did not have an idea about what the ideal performance was.
+As some participants reported that "(students had) very different interpretations of the melody" or "all kids played rhythmically correct" in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
+One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors in the recordings.
 
-we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).
-To make the implementation of articulation clear, we particularly selected detached performances for the recordings where articulation was missing.
-Another possibility is that dynamics implementation was not perceived by participants, therefore they did not modulate the dynamic-related aspect of performances for teaching.
+The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and performed the piece with the stable tempo.
+This was to make sure that we manipulated articulation and dynamics features of the recordings only.
+However, all the recordings might not have sounded like unskillful students were playing.
+In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.
 
-One of the limitations of the current study is that ---.
-
-Future research should investigate.
-
-Limitation
+Evidence from infant-directed action research has shown that caregivers are sensitive to the infants' skill levels and modified their demonstration according to infants' abilities to perform a motor task [@fukuyama_2015].
+In the current experiment, expert pianists needed to perform in the absence of actual students.
+Future research should investigate how expert pianists and students dynamically interact in a situation where they can communicate as a naturalistic lesson and examine how the performance of expert pianists and that of students are related.
 

### 2022-11-02 paper/manuscript.tex
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \ifLuaTeX
@@
 {}{\PatchFailed}
-\keywords{keywords\newline\indent Word count: X}
+\keywords{keywords}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \hypersetup{
-  pdftitle={The title},
+  pdftitle={How do expert pianists adapt their didactic demonstration to novices' skill levels?},
   pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
@@
 
-\title{The title}
+\title{How do expert pianists adapt their didactic demonstration to novices' skill levels?}
 \author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
@@
 
-\shorttitle{Title}
+\shorttitle{How do expert pianists adapt their didactic demonstration to novices' skill levels?}
 
@@
 \abstract{%
-AAA
+Later
 }
@@
 One possibility would be that articulation was more important for the piece we selected or the piece itself naturally invited some implementation of articulation.
-In our previous experiment, we found that even during measuring the baseline performance before the experiment, many participants tend to implement some articulation even though nothing was notated on the sheet music (see details: Supplementary Material).
-To make the implementation of articulation clear, we particularly selected detached performances for the recordings where articulation was missing.
-Another possibility is that dynamics implementation was not perceived by participants, therefore they did not modulate the dynamic-related aspect of performances for teaching.
+In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
+Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it.
+Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants.
 
-One of the limitations of the current study is that ---.
+One of the limitations of the current study is that participants did not have an idea about what the ideal performance was.
+As some participants reported that ``(students had) very different interpretations of the melody'' or ``all kids played rhythmically correct'' in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
+One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors in the recordings.
 
-Future research should investigate.
+The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and performed the piece with the stable tempo.
+This was to make sure that we manipulated articulation and dynamics features of the recordings only.
+However, all the recordings might not have sounded like unskillful students were playing.
+In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.
 
-Limitation
+Evidence from infant-directed action research has shown that caregivers are sensitive to the infants' skill levels and modified their demonstration according to infants' abilities to perform a motor task (Fukuyama et al., 2015).
+In the current experiment, expert pianists needed to perform in the absence of actual students.
+Future research should investigate how expert pianists and students dynamically interact in a situation where they can communicate as a naturalistic lesson and examine how the performance of expert pianists and that of students are related.
 
@@
 \hypertarget{acknowledgement}{%
-\section{Acknowledgement}\label{acknowledgement}}
+\section*{Acknowledgement}\label{acknowledgement}}
+\addcontentsline{toc}{section}{Acknowledgement}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
+Fukuyama, H., Qin, S., Kanakogi, Y., Nagai, Y., Asada, M., \& Myowa-Yamakoshi, M.
+(2015).
+Infant's action skill dynamically modulates parental action demonstration in the dyadic interaction.
+\emph{Developmental Science}, \emph{18}(6), 1006--1013.
+\url{https://doi.org/10.1111/desc.12270}
+
+\end{CSLReferences}
 

### 2022-11-07 paper/manuscript.Rmd
updated intro (tbc...)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 # Introduction
+
+Teaching is effective when teachers and learners intensively communicate.
+
+Exaggeration is used to communicate -. When experts have the intention to teach, they modulate their behaviour. Flexibly -.
+
+Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
+Fukuyama et al., (2014) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
+For example, the mothers exaggerated their movements when their infants failed to produce the expected actions whereas the mother did not produce such exaggeration when their infants did not have enough motor skills to produce the actions.
+The results indicated that adults monitored infants' actions and adapted their demonstration accordingly.
+
+In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance for teaching.
+
+Based on the literature on infant-directed action, teachers should be sensitive to the skill levels of students and such exaggeration should occur only when students fail to achieve a task.
+The current study investigated whether nad how expert pianists adapt their didactic performance depending on the skill levels of students.
+In order to manipulate
+We hypothesised that ---.
 

### 2022-11-07 paper/manuscript.tex
updated intro (tbc...)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \section{Introduction}\label{introduction}}
+
+Teaching is effective when teachers and learners intensively communicate.
+
+Exaggeration is used to communicate -. When experts have the intention to teach, they modulate their behaviour. Flexibly -.
+
+Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
+Fukuyama et al., (2014) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
+For example, the mothers exaggerated their movements when their infants failed to produce the expected actions whereas the mother did not produce such exaggeration when their infants did not have enough motor skills to produce the actions.
+The results indicated that adults monitored infants' actions and adapted their demonstration accordingly.
+
+In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance for teaching.
+
+Based on the literature on infant-directed action, teachers should be sensitive to the skill levels of students and such exaggeration should occur only when students fail to achieve a task.
+The current study investigated whether nad how expert pianists adapt their didactic performance depending on the skill levels of students.
+In order to manipulate
+We hypothesised that ---.
 

### 2022-11-08 paper/manuscript.Rmd
update intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Teaching is effective when teachers and learners intensively communicate.
+Teaching is effective when experts and novices dynamically interact with each other.
+In the literature of cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they acquired to the next generation [@mesoudi_2016a].
+However, information flows between experts and novices are not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive [@charbonneau_2022].
+Experts dynamically adapt their behaviour and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process [@okazaki_2019; @flynn_2022].
 
-Exaggeration is used to communicate -. When experts have the intention to teach, they modulate their behaviour. Flexibly -.
+Exaggeration of movement is often used to signal a communicative intent to others during ongoing interactions [@pezzulo_2019].
+When experts have the intention to teach, they modulate their behaviour.
+For instance, when care givers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002].
+Such exaggeration was observed not only in infant research but also when knowledgeable adults were teaching some knowledge and skills to adult learners [@mcellin_2017; @uther_2007].
 
 Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
-Fukuyama et al., (2014) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
+@fukuyama_2015 revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
 For example, the mothers exaggerated their movements when their infants failed to produce the expected actions whereas the mother did not produce such exaggeration when their infants did not have enough motor skills to produce the actions.

### 2022-11-08 paper/manuscript.tex
update intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Teaching is effective when teachers and learners intensively communicate.
+Teaching is effective when experts and novices dynamically interact with each other.
+In the literature of cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they acquired to the next generation (Mesoudi, 2016).
+However, information flows between experts and novices are not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022).
+Experts dynamically adapt their behaviour and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).
 
-Exaggeration is used to communicate -. When experts have the intention to teach, they modulate their behaviour. Flexibly -.
+Exaggeration of movement is often used to signal a communicative intent to others during ongoing interactions (Pezzulo et al., 2019).
+When experts have the intention to teach, they modulate their behaviour.
+For instance, when care givers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
+Such exaggeration was observed not only in infant research but also when knowledgeable adults were teaching some knowledge and skills to adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).
 
 Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
-Fukuyama et al., (2014) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
+Fukuyama et al.
+(2015) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
 For example, the mothers exaggerated their movements when their infants failed to produce the expected actions whereas the mother did not produce such exaggeration when their infants did not have enough motor skills to produce the actions.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-brand_2002}{}}%
+Brand, R.
+J., Baldwin, D.
+A., \& Ashburn, L.
+A.
+(2002).
+Evidence for {``motionese''}: Modifications in mothers' infant-directed action.
+\emph{Developmental Science}, \emph{5}(1), 72--83.
+\url{https://doi.org/10.1111/1467-7687.00211}
+
+\leavevmode\vadjust pre{\hypertarget{ref-charbonneau_2022}{}}%
+Charbonneau, M., \& Strachan, J.
+W.
+A.
+(2022).
+From {Copying} to {Coordination}: {An Alternative Framework} for {Understanding Cultural Learning Mechanisms}.
+\emph{Journal of Cognition and Culture}, \emph{22}(5), 451--466.
+\url{https://doi.org/10.1163/15685373-12340145}
+
+\leavevmode\vadjust pre{\hypertarget{ref-flynn_2022}{}}%
+Flynn, E.
+(2022).
+{``{Imitation Is} the {Sincerest Form} of''} \ldots{} {Cultural Evolution}, or {Is It}?
+\emph{Journal of Cognition and Culture}, \emph{22}(5), 436--450.
+\url{https://doi.org/10.1163/15685373-12340144}
+
 \leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mcellin_2017}{}}%
+McEllin, L., Knoblich, G., \& Sebanz, N.
+(2017).
+Distinct kinematic markers of demonstration and joint action coordination?
+{Evidence} from virtual xylophone playing.
+\emph{Journal of Experimental Psychology: Human Perception and Performance}, \emph{44}(6), 885.
+\url{https://doi.org/10.1037/xhp0000505}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mesoudi_2016a}{}}%
+Mesoudi, A. (2016). Cultural evolution: Integrating psychology, evolution and culture. \emph{Current Opinion in Psychology}, \emph{7}, 17--22. \url{https://doi.org/10.1016/j.copsyc.2015.07.001}
+
+\leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
+Okazaki, S., Muraoka, Y., \& Osu, R.
+(2019).
+Teacher-learner interaction quantifies scaffolding behaviour in imitation learning.
+\emph{Scientific Reports}, \emph{9}(1), 1--13.
+\url{https://doi.org/10.1038/s41598-019-44049-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-pezzulo_2019}{}}%
+Pezzulo, G., Donnarumma, F., Dindo, H., D'Ausilio, A., Konvalinka, I., \& Castelfranchi, C.
+(2019).
+The body talks: {Sensorimotor} communication and its brain and kinematic signatures.
+\emph{Physics of Life Reviews}, \emph{28}, 1--21.
+\url{https://doi.org/10.1016/j.plrev.2018.06.014}
+
+\leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
+Uther, M., Knoll, M.
+A., \& Burnham, D.
+(2007).
+Do you speak {E-NG-L-I-SH}?
+{A} comparison of foreigner- and infant-directed speech.
+\emph{Speech Communication}, \emph{49}(1), 2--7.
+\url{https://doi.org/10.1016/j.specom.2006.10.003}
+
 \end{CSLReferences}

### 2022-11-09 paper/manuscript.Rmd
update intro/discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2022-11-02 paper/manuscript.Rmd
@@
 ---
-title             : "How do expert pianists adapt their didactic demonstration to novices’ skill levels?"
-shorttitle        : "How do expert pianists adapt their didactic demonstration to novices’ skill levels?"
+title             : "How do expert pianists adapt their pedagogical demonstration to novices’ skill levels?"
+shorttitle        : "How do expert pianists adapt their pedagogical demonstration to novices’ skill levels?"
 
@@
 
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : no
@@
 # Introduction
+
+Teaching is effective when experts and novices dynamically interact with each other [@byrne_2011].
+In the literature on cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they have acquired to the next generation [@mesoudi_2016a].
+However, information flows between experts and novices are mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive [@charbonneau_2022].
+Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process [@okazaki_2019; @flynn_2022].
+
+Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019].
+When experts have an intention to teach, they modulate their behaviour.
+For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002].
+Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners [@mcellin_2017; @uther_2007].
+
+Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
+@fukuyama_2015 revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
+For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have enough motor skills to produce such actions.
+The results indicated that adults monitored infants' actions and abilities, and adapted their demonstration accordingly.
+
+In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
+For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had an intention to teach dynamics.
+Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance.
+If teaching is an adaptive and scaffolding process, it can be expected that experts should not use such exaggeration if novices became able to produce them.
+
+In the current study, we investigated whether and how expert pianists adapted their performance depending on the skill levels of novice learners.
+Instead of recruiting actual novice pianists, we created artificial recordings to manipulate the skill levels of novices to maximise the experimental control.
+The recordings were made based on the performance data from our previous experiment (Tominaga et al., under review).
+All the recordings had the same musical structure (i.e., the same piece) but had different implementations of articulation and dynamics.
+In the ideal performance, both articulation and dynamics should be implemented as notated on the sheet music (*Fig \ref{fig:stim}*).
+We made the recordings so that both techniques were implemented in the first quarter, only articulation was implemented in the second quarter, only dynamics was implemented in the third quarter and neither of them was implemented in the last quarter of the recordings.
+Therefore, in this experiment, the skill levels of novices corresponded to the implementation of articulation and dynamics.
+
+We hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
+More precisely, expert pianists would exaggerate articulation (i.e., producing longer legato and shorter staccato) when articulation was not implemented in the recordings whereas expert pianists would exaggerate dynamics (i.e., producing  louder forte and softer piano, and larger contrast between forte and piano) when dynamics was not implemented in the recordings.
+If both articulation and dynamics were missing, we predicted that expert pianists would exaggerate both aspects.
 
@@
 
-As stimuli, one piece of music, which was used in our previous experiment, was selected (for details, Tominaga et al., 2022).
+As stimuli, one piece of music, which was used in our previous experiment, was selected (Tominaga et al., under review).
 The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major.
@@
 
-We created artificial recordings of students to manipulate their skill levels.
-In this experiment, we made the recordings by changing two factors of articulation (present or absent) and dynamics (present or absent).
+We created artificial recordings of novices to manipulate their skill levels.
+In this experiment, we made the recordings by changing two factors Articulation (present or absent) and Dynamics (present or absent).
 Therefore, there were four types of recordings.
@@
 We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
-How we generated the stimuli is written in *Supplementary Material*.
+How we generated the recordings is written in *Supplementary Material*.
 
 ```{r stimli, out.width = "100%", fig.cap = "\\label{fig:stim}Sheet music.
-Articulation.
 For articulation notation, the curved line (slur) indicates legato and the dots indicate staccato.
 For dynamics notation, the symbol `f' denotes forte and the symbol `p' denotes piano.
-For data analysis, only the 8th notes were included.", echo = FALSE}
+For data analysis, only the 8th notes with expressive notations were included.", echo = FALSE}
 here::i_am("README.md")
@@
 
-First, we recorded participants' baseline performance by asking them to perform the piece without listening to any students' performance.
+First, we recorded participants' baseline performance by asking them to perform the piece without listening to any novices' performance.
 A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
@@
 Also, they were told to perform the piece expressively with their interpretation and do their best as a performer.
-This instruction was given to make sure that they paid attention to expressive aspects of the performance.
+This instruction was given to make sure that they paid attention to the expressive aspects of the performance.
 Each participant performed the piece twice.
@@
 Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to the student (i.e., recording).
-In total, there were 16 trials and participants played the piece for each recording only once.
+In total, there were 16 trials and participants played the piece for each student (i.e., recording) only once.
 The order of the recordings was randomised for each participant.
@@
 Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
-Key-overlap time (KOT) is the difference between the offset time of the current tone (i.e., key release time) and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
+Key-overlap time (KOT) is the difference between the offset time (i.e., key release time) of the current tone and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
 A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
@@
 Neither main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
-Therefore, participants did not change the tempo depending on the types of the recordings (*Fig \ref{fig:ioi-1}*).
+Therefore, participants did not change the tempo depending on the types of recordings (*Fig \ref{fig:ioi-1}*).
 
@@
 Neither main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
-Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (*Fig \ref{fig:vel-1}*, left).
+Participants did not change performances in terms of notated forte depending on the types of recordings (*Fig \ref{fig:vel-1}*, left).
 
@@
 
-These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of the recording (*Fig \ref{fig:vel-diff-1}*).
+These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of recordings (*Fig \ref{fig:vel-diff-1}*).
 
@@
 To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer.
-By comparing with the baseline performance, we can investigate how didactic intentions influence participants' performances.
+By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.
 
@@
 
-We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of students' skill levels.
+We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of novices' skill levels.
 We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
@@
 
-We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not on each recording.
+We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not in each recording.
 We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
@@
 There was a significant main effect of Category (*F*(`r leg_ca_par_aov$anova_table[1,1]`, `r leg_ca_par_aov$anova_table[1,2]`) = `r leg_ca_par_aov$anova_table[1,4]`, *p* `r if(leg_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001 , baseline and dynamics-only: *p* .008), suggesting that participants produced longer legato when they had an intention to teach (*Fig \ref{fig:kot-2}*, left).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001 , baseline and dynamics-only: *p* = .008), suggesting that participants produced longer legato when they had an intention to teach (*Fig \ref{fig:kot-2}*, left).
 
@@
 
-We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording.
+We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording.
 We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
@@
 There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* `r if(pia_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', pia_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* .05 , baseline and articulation-only: *p* .03), suggesting that participants produced softer piano when they had an intention to teach (*Fig \ref{fig:vel-2}*, right).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05 , baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had an intention to teach (*Fig \ref{fig:vel-2}*, right).
 
@@
 
+\clearpage
+
 # Discussion
 
-The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of students.
+The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of novices.
 We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
@@
 We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
-Even compared with the tempo of the baseline performance, they did not change their tempo either.
+Even compared with the tempo of the baseline performance, they did not change the tempo either.
 These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
-Especially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
+Although slower demonstration is generally considered to be used for teaching [@mcellin_2017; @saint-georges_2013; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
 
@@
 However, we did not find significant results in terms of dynamics.
-Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings.
-Participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
+Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
 
 Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
-One possibility would be that articulation was more important for the piece we selected or the piece itself naturally invited some implementation of articulation.
+One possibility would be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
 In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
 Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it.
-Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants.
+Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.
 
@@
 As some participants reported that "(students had) very different interpretations of the melody" or "all kids played rhythmically correct" in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
-One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors in the recordings.
+One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors by comparing each recording with the model performance.
 
-The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and performed the piece with the stable tempo.
+The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo.
 This was to make sure that we manipulated articulation and dynamics features of the recordings only.
-However, all the recordings might not have sounded like unskillful students were playing.
+However, all the recordings might not have sounded as if unskillful students were playing.
 In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.
 
-Evidence from infant-directed action research has shown that caregivers are sensitive to the infants' skill levels and modified their demonstration according to infants' abilities to perform a motor task [@fukuyama_2015].
-In the current experiment, expert pianists needed to perform in the absence of actual students.
-Future research should investigate how expert pianists and students dynamically interact in a situation where they can communicate as a naturalistic lesson and examine how the performance of expert pianists and that of students are related.
+Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place [@fukuyama_2015; @okazaki_2019].
+In the current experiment, expert pianists had only access to their imaginary students and needed to perform in the absence of actual students.
+Future research should investigate how expert pianists and novices dynamically interact in a situation where they can communicate in real-time and examine how the performance of expert pianists and that of novices are related to each other.
+
+\clearpage
 
@@
 
+\begingroup
+\setlength{\parindent}{-0in}
+\setlength{\leftskip}{0in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup
+
 # Acknowledgement
+This research was supported by the European Research Council under the European Union’s Seventh Framework Program (FP7/2007–2013)/ERC Grant agreement no.
+609819, SOMICS, and by ERC Grant agreement no.
+616072, JAXPERTISE.
+We thank Candasch Acar for his help with data collection.

### 2022-11-09 paper/manuscript.tex
update intro/discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \documentclass[
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 \keywords{keywords}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{csquotes}
@@
 \hypersetup{
-  pdftitle={How do expert pianists adapt their didactic demonstration to novices' skill levels?},
+  pdftitle={How do expert pianists adapt their pedagogical demonstration to novices' skill levels?},
   pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
@@
 
-\title{How do expert pianists adapt their didactic demonstration to novices' skill levels?}
+\title{How do expert pianists adapt their pedagogical demonstration to novices' skill levels?}
 \author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
@@
 
-\shorttitle{How do expert pianists adapt their didactic demonstration to novices' skill levels?}
+\shorttitle{How do expert pianists adapt their pedagogical demonstration to novices' skill levels?}
 
@@
 
-Teaching is effective when experts and novices dynamically interact with each other.
-In the literature of cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they acquired to the next generation (Mesoudi, 2016).
-However, information flows between experts and novices are not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022).
-Experts dynamically adapt their behaviour and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).
+Teaching is effective when experts and novices dynamically interact with each other (Byrne \& Rapaport, 2011).
+In the literature on cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they have acquired to the next generation (Mesoudi, 2016).
+However, information flows between experts and novices are mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022).
+Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).
 
-Exaggeration of movement is often used to signal a communicative intent to others during ongoing interactions (Pezzulo et al., 2019).
-When experts have the intention to teach, they modulate their behaviour.
-For instance, when care givers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
-Such exaggeration was observed not only in infant research but also when knowledgeable adults were teaching some knowledge and skills to adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).
+Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019).
+When experts have an intention to teach, they modulate their behaviour.
+For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
+Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).
 
@@
 (2015) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
-For example, the mothers exaggerated their movements when their infants failed to produce the expected actions whereas the mother did not produce such exaggeration when their infants did not have enough motor skills to produce the actions.
-The results indicated that adults monitored infants' actions and adapted their demonstration accordingly.
+For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have enough motor skills to produce such actions.
+The results indicated that adults monitored infants' actions and abilities, and adapted their demonstration accordingly.
 
-In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance for teaching.
+In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
+For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had an intention to teach dynamics.
+Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance.
+If teaching is an adaptive and scaffolding process, it can be expected that experts should not use such exaggeration if novices became able to produce them.
 
-Based on the literature on infant-directed action, teachers should be sensitive to the skill levels of students and such exaggeration should occur only when students fail to achieve a task.
-The current study investigated whether nad how expert pianists adapt their didactic performance depending on the skill levels of students.
-In order to manipulate
-We hypothesised that ---.
+In the current study, we investigated whether and how expert pianists adapted their performance depending on the skill levels of novice learners.
+Instead of recruiting actual novice pianists, we created artificial recordings to manipulate the skill levels of novices to maximise the experimental control.
+The recordings were made based on the performance data from our previous experiment (Tominaga et al., under review).
+All the recordings had the same musical structure (i.e., the same piece) but had different implementations of articulation and dynamics.
+In the ideal performance, both articulation and dynamics should be implemented as notated on the sheet music (\emph{Fig \ref{fig:stim}}).
+We made the recordings so that both techniques were implemented in the first quarter, only articulation was implemented in the second quarter, only dynamics was implemented in the third quarter and neither of them was implemented in the last quarter of the recordings.
+Therefore, in this experiment, the skill levels of novices corresponded to the implementation of articulation and dynamics.
+
+We hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
+More precisely, expert pianists would exaggerate articulation (i.e., producing longer legato and shorter staccato) when articulation was not implemented in the recordings whereas expert pianists would exaggerate dynamics (i.e., producing louder forte and softer piano, and larger contrast between forte and piano) when dynamics was not implemented in the recordings.
+If both articulation and dynamics were missing, we predicted that expert pianists would exaggerate both aspects.
 
@@
 We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
-How we generated the stimuli is written in \emph{Supplementary Material}.
+How we generated the recordings is written in \emph{Supplementary Material}.
 
@@
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimli-1} \caption{\label{fig:stim}Sheet music.
-Articulation.
 For articulation notation, the curved line (slur) indicates legato and the dots indicate staccato.
 For dynamics notation, the symbol `f' denotes forte and the symbol `p' denotes piano.
-For data analysis, only the 8th notes were included.}\label{fig:stimli}
+For data analysis, only the 8th notes with expressive notations were included.}\label{fig:stimli}
 \end{figure}
@@
 Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) or Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
-Therefore, participants did not change the tempo depending on the types of the recordings (\emph{Fig \ref{fig:ioi-1}}).
+Therefore, participants did not change the tempo depending on the types of recordings (\emph{Fig \ref{fig:ioi-1}}).
 
@@
 Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) or Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
-Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (\emph{Fig \ref{fig:vel-1}}, left).
+Participants did not change performances in terms of notated forte depending on the types of recordings (\emph{Fig \ref{fig:vel-1}}, left).
 
@@
 
-These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of the recording (\emph{Fig \ref{fig:vel-diff-1}}).
+These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of recordings (\emph{Fig \ref{fig:vel-diff-1}}).
 
@@
 There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.060; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
 
@@
 There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} .05 , baseline and articulation-only: \emph{p} .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
 
@@
 
+\clearpage
+
 \hypertarget{discussion}{%
@@
 
-The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of students.
+The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of novices.
 We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
@@
 We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
-Even compared with the tempo of the baseline performance, they did not change their tempo either.
+Even compared with the tempo of the baseline performance, they did not change the tempo either.
 These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
-Especially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
+Although slower demonstration is generally considered to be used for teaching (McEllin et al., 2017; Saint-Georges et al., 2013; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
 
@@
 As some participants reported that ``(students had) very different interpretations of the melody'' or ``all kids played rhythmically correct'' in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
-One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors in the recordings.
+One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors by comparing each recording with the model performance.
 
-The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and performed the piece with the stable tempo.
+The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo.
 This was to make sure that we manipulated articulation and dynamics features of the recordings only.
-However, all the recordings might not have sounded like unskillful students were playing.
+However, all the recordings might not have sounded as if unskillful students were playing.
 In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.
 
-Evidence from infant-directed action research has shown that caregivers are sensitive to the infants' skill levels and modified their demonstration according to infants' abilities to perform a motor task (Fukuyama et al., 2015).
-In the current experiment, expert pianists needed to perform in the absence of actual students.
-Future research should investigate how expert pianists and students dynamically interact in a situation where they can communicate as a naturalistic lesson and examine how the performance of expert pianists and that of students are related.
+Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place (Fukuyama et al., 2015; Okazaki et al., 2019).
+In the current experiment, expert pianists had only access to their imaginary students and needed to perform in the absence of actual students.
+Future research should investigate how expert pianists and novices dynamically interact in a situation where they can communicate in real-time and examine how the performance of expert pianists and that of novices are related to each other.
+
+\clearpage
 
@@
 
-\hypertarget{acknowledgement}{%
-\section*{Acknowledgement}\label{acknowledgement}}
-\addcontentsline{toc}{section}{Acknowledgement}
+\begingroup
+\setlength{\parindent}{-0in}
+\setlength{\leftskip}{0in}
 
@@
 \url{https://doi.org/10.1111/1467-7687.00211}
+
+\leavevmode\vadjust pre{\hypertarget{ref-byrne_2011}{}}%
+Byrne, R. W., \& Rapaport, L. G. (2011). What are we learning from teaching? \emph{Animal Behaviour}, \emph{82}(5), 1207--1211. \url{https://doi.org/10.1016/j.anbehav.2011.08.018}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-saint-georges_2013}{}}%
+Saint-Georges, C., Chetouani, M., Cassel, R., Apicella, F., Mahdhaoui, A., Muratori, F., \ldots{} Cohen, D.
+(2013).
+Motherese in {Interaction}: {At} the {Cross-Road} of {Emotion} and {Cognition}?
+({A Systematic Review}).
+\emph{PLOS ONE}, \emph{8}(10), e78103.
+\url{https://doi.org/10.1371/journal.pone.0078103}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schaik_2019}{}}%
+Schaik, J.
+E.
+van, Meyer, M., Ham, C.
+R.
+van, \& Hunnius, S.
+(2019).
+Motion tracking of parents' infant- versus adult-directed actions reveals general and action-specific modulations.
+\emph{Developmental Science}, \emph{0}(0), e12869.
+\url{https://doi.org/10.1111/desc.12869}
+
 \leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
@@
 
+\endgroup
+
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
+
+This research was supported by the European Research Council under the European Union's Seventh Framework Program (FP7/2007--2013)/ERC Grant agreement no.
+609819, SOMICS, and by ERC Grant agreement no.
+616072, JAXPERTISE.
+We thank Candasch Acar for his help with data collection.
+
 \end{document}

### 2022-11-29 paper/manuscript.Rmd
update methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The study (No.
-2020_05) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
+2020/05) was approved by CEU PU's Psychological Research Ethics Board (PREBO).
 
@@
 
-As stimuli, one piece of music, which was used in our previous experiment, was selected (Tominaga et al., under review).
-The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major.
-The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
+The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., under review).
+The piece was an excerpt from Clementi's Sonatina Op.36 (No.3) in C major.
+The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points (i.e., the number of notes, the number of intervals between notes) for each dependent variable.
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Two expressive notations (i.e., articulation and dynamics) were added as *Fig \ref{fig:stim}*.
+Two expressive notations (i.e., articulation and dynamics) were added, as shown in *Fig \ref{fig:stim}*.
 
-We created artificial recordings of novices to manipulate their skill levels.
-In this experiment, we made the recordings by changing two factors Articulation (present or absent) and Dynamics (present or absent).
-Therefore, there were four types of recordings.
-We aimed to create a quarter of the recordings where both articulation and dynamics were implemented.
-The second quarter included the recordings where only articulation was implemented whereas dynamics was missing.
-The third quarter included the recordings where only dynamics was implemented whereas articulation was missing.
-The last quarter included the recordings where neither articulation nor dynamics was implemented.
+We created artificial recordings of novices to manipulate the displayed proficiency with regard to the two relevant techniques.
+There were four types of recordings: 1) recordings where both articulation and dynamics were implemented, 2) recordings where only articulation was implemented whereas dynamics was missing, 3) recordings where only dynamics was implemented whereas articulation was missing, and 4) recordings where neither articulation nor dynamics was implemented.
 We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
-How we generated the recordings is written in *Supplementary Material*.
+How we generated the recordings is described in *Supplementary Material*.
 
@@
 
-Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and perform it without pitch errors while implementing notated expressions in the experiment.
+Prior to the experiment, participants were required to memorise the piece so that they could perform it with the notated expressions and without pitch errors in the experiment.
 
-First, we recorded participants' baseline performance by asking them to perform the piece without listening to any novices' performance.
+First, we recorded participants' baseline performance by asking them to perform the piece without having listened to any novices' performance.
 A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
 Sheet music (*Fig \ref{fig:stim}*) was displayed in front of the participants.
-Also, they were told to perform the piece expressively with their interpretation and do their best as a performer.
-This instruction was given to make sure that they paid attention to the expressive aspects of the performance.
+They were told to perform the piece expressively with their interpretation and to do their best as a performer.
+This instruction was given to make sure that they paid attention to the expressive aspects of the performance but did not have the intention to teach.
 Each participant performed the piece twice.
 
-After we recorded the  participants' baseline performance, participants were informed that they were going to listen to a number of recordings from 16 different students, who were learning musical expressive techniques.
-Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to the student (i.e., recording).
-In total, there were 16 trials and participants played the piece for each student (i.e., recording) only once.
+After we recorded the  participants' baseline performance, participants were told that they were going to listen to a number of recordings from 16 different students, who were learning musical expressive techniques.
+Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to that student.
+In total, there were 16 trials and participants played the piece for each student only once.
 The order of the recordings was randomised for each participant.
@@
 
-After participants completed the 16 trials, they were asked to perform the piece in the same situation as we recorded their baseline performance.
+After participants completed the 16 trials, they were asked to perform the piece with the same instruction as in their baseline performance.
 At the end of the experiment, participants filled in a questionnaire asking about their demographic information and experience in piano performance/teaching.
+All participants were fully debriefed at the end of the experiment and informed that the recordings had been artificially generated.
 
@@
 Key-overlap time (KOT) is the difference between the offset time (i.e., key release time) of the current tone and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
-A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
+A positive value indicates smooth legato style due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato style due to separation between the current and ensuing note.
 Tone intensity is assessed  by key velocity (KV) and measures the loudness of a musical note.
-A higher value indicates forte styles whereas a lower value indicates piano styles.
+A higher value indicates forte style whereas a lower value indicates piano style.
 The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
@@
 Pitch errors were identified by comparing the sequence of  musical notes produced by a participant with the sequence of musical notes according to the sheet music.
-Pitch errors included either, extra, missing or substituted tones and were manually removed by using the *editData* R package.
+Pitch errors included extra, missing or substituted tones and were manually removed by using the *editData* R package.
 For note onsets,  `r round(nrow(dt_error_onset)/(20*16)*100, 2)` % of the trials contained at least one pitch error (extra notes: `r round(nrow(error_extra_onset)/(20*16)*100, 2)` %, missing notes: `r round(nrow(error_missing_onset)/(20*16)*100, 2)` %, substituted notes: `r round(nrow(error_sub_onset)/(20*16)*100, 2)` %).
@@
 We considered these as errors and removed the erroneous notes even if the order of onsets was correct.
-As a result, less than 1 % of total responses were corrected.
+As a result, less than 1 % of total responses were removed.
 After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference, defined as values more than 3 standard deviations from the mean of each dependent variable.
@@
 We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.
-absent) and Dynamics (present or absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
+absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
 The *aov_ez* function in the *afex* R package was used for a repeated-measures ANOVA.
-For post-doc comparisons on the estimated marginal means, we used the  *emmeans* R package.
+For post-hoc comparisons on the estimated marginal means, we used the  *emmeans* R package.
 

### 2022-11-29 paper/manuscript.tex
update methods
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 All participants were recruited through an online participant platform (SONA system, \url{https://www.sona-systems.com}).
-The study (No.~2020\_05) was approved by the Psychological Research Ethics Board (PREBO) CEU PU in Austria.
+The study (No.~2020/05) was approved by CEU PU's Psychological Research Ethics Board (PREBO).
 
@@
 
-As stimuli, one piece of music, which was used in our previous experiment, was selected (Tominaga et al., under review).
-The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major.
-The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable.
+The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., under review).
+The piece was an excerpt from Clementi's Sonatina Op.36 (No.3) in C major.
+The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points (i.e., the number of notes, the number of intervals between notes) for each dependent variable.
 The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only.
-Two expressive notations (i.e., articulation and dynamics) were added as \emph{Fig \ref{fig:stim}}.
+Two expressive notations (i.e., articulation and dynamics) were added, as shown in \emph{Fig \ref{fig:stim}}.
 
-We created artificial recordings of novices to manipulate their skill levels.
-In this experiment, we made the recordings by changing two factors Articulation (present or absent) and Dynamics (present or absent).
-Therefore, there were four types of recordings.
-We aimed to create a quarter of the recordings where both articulation and dynamics were implemented.
-The second quarter included the recordings where only articulation was implemented whereas dynamics was missing.
-The third quarter included the recordings where only dynamics was implemented whereas articulation was missing.
-The last quarter included the recordings where neither articulation nor dynamics was implemented.
+We created artificial recordings of novices to manipulate the displayed proficiency with regard to the two relevant techniques.
+There were four types of recordings: 1) recordings where both articulation and dynamics were implemented, 2) recordings where only articulation was implemented whereas dynamics was missing, 3) recordings where only dynamics was implemented whereas articulation was missing, and 4) recordings where neither articulation nor dynamics was implemented.
 We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
-How we generated the recordings is written in \emph{Supplementary Material}.
+How we generated the recordings is described in \emph{Supplementary Material}.
 
@@
 
-Prior to the experiment, participants were required to memorise the piece so that they had enough time to practise and perform it without pitch errors while implementing notated expressions in the experiment.
+Prior to the experiment, participants were required to memorise the piece so that they could perform it with the notated expressions and without pitch errors in the experiment.
 
-First, we recorded participants' baseline performance by asking them to perform the piece without listening to any novices' performance.
+First, we recorded participants' baseline performance by asking them to perform the piece without having listened to any novices' performance.
 A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece.
 Sheet music (\emph{Fig \ref{fig:stim}}) was displayed in front of the participants.
-Also, they were told to perform the piece expressively with their interpretation and do their best as a performer.
-This instruction was given to make sure that they paid attention to the expressive aspects of the performance.
+They were told to perform the piece expressively with their interpretation and to do their best as a performer.
+This instruction was given to make sure that they paid attention to the expressive aspects of the performance but did not have the intention to teach.
 Each participant performed the piece twice.
 
-After we recorded the participants' baseline performance, participants were informed that they were going to listen to a number of recordings from 16 different students, who were learning musical expressive techniques.
-Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to the student (i.e., recording).
-In total, there were 16 trials and participants played the piece for each student (i.e., recording) only once.
+After we recorded the participants' baseline performance, participants were told that they were going to listen to a number of recordings from 16 different students, who were learning musical expressive techniques.
+Participants were required to listen to each student's recording first, and then to perform the same piece to teach musical expressive techniques to that student.
+In total, there were 16 trials and participants played the piece for each student only once.
 The order of the recordings was randomised for each participant.
@@
 Key-overlap time (KOT) is the difference between the offset time (i.e., key release time) of the current tone and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
-A positive value indicates smooth legato styles due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato styles due to separation between the current and ensuing note.
+A positive value indicates smooth legato style due to overlap between the current and ensuing tone whereas a negative value indicates sharp staccato style due to separation between the current and ensuing note.
 Tone intensity is assessed by key velocity (KV) and measures the loudness of a musical note.
-A higher value indicates forte styles whereas a lower value indicates piano styles.
+A higher value indicates forte style whereas a lower value indicates piano style.
 The value of KV in MIDI varies between 0 (minimum) and 127 (maximum).
@@
 Pitch errors were identified by comparing the sequence of musical notes produced by a participant with the sequence of musical notes according to the sheet music.
-Pitch errors included either, extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
+Pitch errors included extra, missing or substituted tones and were manually removed by using the \emph{editData} R package.
 For note onsets, 31.87 \% of the trials contained at least one pitch error (extra notes: 5.94 \%, missing notes: 24.38 \%, substituted notes: 1.56 \%).
@@
 We considered these as errors and removed the erroneous notes even if the order of onsets was correct.
-As a result, less than 1 \% of total responses were corrected.
+As a result, less than 1 \% of total responses were removed.
 After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference, defined as values more than 3 standard deviations from the mean of each dependent variable.
@@
 
-We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present or absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
+We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
 The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
-For post-doc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.
+For post-hoc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.
 

### 2022-11-29 paper/manuscript.Rmd
update results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 ---
-title             : "How do expert pianists adapt their pedagogical demonstration to novices’ skill levels?"
-shorttitle        : "How do expert pianists adapt their pedagogical demonstration to novices’ skill levels?"
+title             : "Do expert pianists adapt their pedagogical demonstrations to novices’ skills?"
+shorttitle        : "Do expert pianists adapt their pedagogical demonstrations to novices’ skills?"
 
@@
 
-Teaching is effective when experts and novices dynamically interact with each other [@byrne_2011].
-In the literature on cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they have acquired to the next generation [@mesoudi_2016a].
-However, information flows between experts and novices are mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive [@charbonneau_2022].
+Effective teaching involves experts and novices dynamically interacting with each other [@byrne_2011].
+Typically, information flows between experts and novices is mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive [@charbonneau_2022].
 Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process [@okazaki_2019; @flynn_2022].
 
-Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019].
-When experts have an intention to teach, they modulate their behaviour.
+Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts.
 For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002].
@@
 In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
-For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had an intention to teach dynamics.
+For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had the intention to teach dynamics.
 Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance.
@@
 
-We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.
+First, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.
 absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
@@
 For post-hoc comparisons on the estimated marginal means, we used the  *emmeans* R package.
+
+In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced for non-teaching purposes), we performed one-way repeated-measures ANOVAs.
+For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent (i.e., Legato, Staccato, Forte, Piano).
 
@@
 
-Neither main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
-Therefore, participants did not change the tempo depending on the types of recordings (*Fig \ref{fig:ioi-1}*).
+Neither the main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
+This indicates that participants did not change the tempo depending on the type of recordings (*Fig \ref{fig:ioi-1}*).
 
@@
   geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") +
+  geom_hline(yintercept = 300, linetype = "dashed") +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") + scale_y_continuous(breaks = seq(250, 300, 10), limits = c(250, 300)) +
   scale_color_aaas() +
@@
 
-There was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
-However, there was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* `r if(leg_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* `r if(leg_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
+As predicted, there was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
+There was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* `r if(leg_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* `r if(leg_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
 Participants produced longer legato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, left).
@@
 
-There was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* `r if(sta_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
-However, there was no significant main effect of Dynamics (*F*(`r sta_all_aov$anova_table[2,1]`, `r sta_all_aov$anova_table[2,2]`) = `r sta_all_aov$anova_table[2,4]`, *p* `r if(sta_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r sta_all_aov$anova_table[3,1]`, `r sta_all_aov$anova_table[3,2]`) = `r sta_all_aov$anova_table[3,4]`, *p* `r if(sta_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[3,5])`).
+In line with our prediction, there was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* `r if(sta_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
+There was no significant main effect of Dynamics (*F*(`r sta_all_aov$anova_table[2,1]`, `r sta_all_aov$anova_table[2,2]`) = `r sta_all_aov$anova_table[2,4]`, *p* `r if(sta_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r sta_all_aov$anova_table[3,1]`, `r sta_all_aov$anova_table[3,2]`) = `r sta_all_aov$anova_table[3,4]`, *p* `r if(sta_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[3,5])`).
 Participants produced shorter staccato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, right).
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KOT (ms)", title = "KOT - Legato") +
+  labs(y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-10, 120, 40)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KOT (ms)", title = "KOT - Staccato") +
+  labs(y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-310, -170, 40)) +
   scale_color_aaas() +
@@
 
-Neither main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
-Participants did not change performances in terms of notated forte depending on the types of recordings (*Fig \ref{fig:vel-1}*, left).
+Neither the main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
+Participants did not change their performances in terms of notated forte depending on the type of recordings (*Fig \ref{fig:vel-1}*, left).
 
@@
 There was a significant main effect of Articulation (*F*(`r pia_all_aov$anova_table[1,1]`, `r pia_all_aov$anova_table[1,2]`) = `r pia_all_aov$anova_table[1,4]`, *p* `r if(pia_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[1,5])`).
-However, there was no significant main effect of Dynamics (*F*(`r pia_all_aov$anova_table[2,1]`, `r pia_all_aov$anova_table[2,2]`) = `r pia_all_aov$anova_table[2,4]`, *p* `r if(pia_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[2,5])`) or interaction between Articulation and Dynamics (*F*(`r pia_all_aov$anova_table[3,1]`, `r pia_all_aov$anova_table[3,2]`) = `r pia_all_aov$anova_table[3,4]`, *p* `r if(pia_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[3,5])`).
+However, there was no significant main effect of Dynamics (*F*(`r pia_all_aov$anova_table[2,1]`, `r pia_all_aov$anova_table[2,2]`) = `r pia_all_aov$anova_table[2,4]`, *p* `r if(pia_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r pia_all_aov$anova_table[3,1]`, `r pia_all_aov$anova_table[3,2]`) = `r pia_all_aov$anova_table[3,4]`, *p* `r if(pia_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[3,5])`).
 Participants produced softer piano when listening to the recordings where articulation was implemented (*Fig \ref{fig:vel-1}*, right).
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV (0-127)", title = "KV - Forte") +
+  labs(y = "KV (0-127)", title = "KV - Forte") + scale_y_continuous(breaks = seq(60, 110, 10)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV (0-127)", title = "KV - Piano") +
+  labs(y = "KV (0-127)", title = "KV - Piano") + scale_y_continuous(breaks = seq(50, 70, 10)) +
   scale_color_aaas() +
@@
 
-Neither main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* `r if(ftop_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* `r if(ftop_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* `r if(ftop_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
+Neither the main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* `r if(ftop_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* `r if(ftop_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* `r if(ftop_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
 
@@
 
-Neither main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* `r if(ptof_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ptof_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) or Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* `r if(ptof_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* `r if(ptof_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
+Neither the main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* `r if(ptof_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ptof_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* `r if(ptof_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* `r if(ptof_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
 
-These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of recordings (*Fig \ref{fig:vel-diff-1}*).
+These results indicated that participants did not change their performances in terms of dynamics contrast depending on the type of recordings (*Fig \ref{fig:vel-diff-1}*).
 
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV Difference (-127-127)", title = "KV Difference - FtoP") +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - FtoP") + scale_y_continuous(breaks = seq(-35, 0, 10)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoP") +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoP") + scale_y_continuous(breaks = seq(-5, 45, 10)) +
   scale_color_aaas() +
@@
 
-```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}IOIs (ms). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
 img_ioi_1 <- image_scale(img_ioi_all_plot, "2200")
@@
 img_vel_diff_1
-```
-
-## Comparison with baseline performance
-In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced before listening to any of the recordings), we performed one-way repeated-measures ANOVAs.
-For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent.
-To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer.
-By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.
-
-## IOIs
-
-We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of novices' skill levels.
-We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
-
-There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* `r if(ioi_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (*Fig \ref{fig:ioi-2}*).
-
-```{r ioi-bl-plot, include = FALSE}
-ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
-  geom_jitter(width = 0.15, alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_ca_plot.png"), plot = ioi_ca_plot, dpi = 300, height = 5)
-img_ioi_ca_plot <- image_read(here("paper/image", "ioi_ca_plot.png"), density = 300)
-```
-
-## KOT
-
-We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not in each recording.
-We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
-
-### Legato
-
-There was a significant main effect of Category (*F*(`r leg_ca_par_aov$anova_table[1,1]`, `r leg_ca_par_aov$anova_table[1,2]`) = `r leg_ca_par_aov$anova_table[1,4]`, *p* `r if(leg_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001 , baseline and dynamics-only: *p* = .008), suggesting that participants produced longer legato when they had an intention to teach (*Fig \ref{fig:kot-2}*, left).
-
-### Staccato
-
-There was no significant main effect of Category (*F*(`r sta_ca_par_aov$anova_table[1,1]`, `r sta_ca_par_aov$anova_table[1,2]`) = `r sta_ca_par_aov$anova_table[1,4]`, *p* `r if(sta_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (*Fig \ref{fig:kot-2}*, right).
-
-```{r kot-bl-plot, include = FALSE}
-leg_ca_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category == "baseline" | Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA) +
-  geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KOT (ms)", title = "KOT - Legato") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-sta_ca_plot <- ggplot(kot[Subcomponent1 == "Staccato" & Category == "baseline" | Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA) +
-  geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KOT (ms)", title = "KOT - Staccato") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "leg_ca_plot.png"), plot = leg_ca_plot, dpi = 300, height = 5)
-img_leg_ca_plot <- image_read(here("paper/image", "leg_ca_plot.png"), density = 300)
-
-ggsave(here("paper/image", "sta_ca_plot.png"), plot = sta_ca_plot, dpi = 300, height = 5)
-img_sta_ca_plot <- image_read(here("paper/image", "sta_ca_plot.png"), density = 300)
-```
-
-## KV
-
-We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording.
-We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
-
-### Forte
-
-There was no significant main effect of Category (*F*(`r for_ca_par_aov$anova_table[1,1]`, `r for_ca_par_aov$anova_table[1,2]`) = `r for_ca_par_aov$anova_table[1,4]`, *p* `r if(for_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (*Fig \ref{fig:vel-2}*, left).
-
-### Piano
-
-There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* `r if(pia_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', pia_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05 , baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had an intention to teach (*Fig \ref{fig:vel-2}*, right).
-
-```{r vel-bl-plot, include = FALSE}
-for_ca_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category == "baseline" | Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA) +
-  geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV (0-127)", title = "KV - Forte") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-pia_ca_plot <- ggplot(vel[Subcomponent2 == "Piano" & Category == "baseline" | Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA) +
-  geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV (0-127)", title = "KV - Piano") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "for_ca_plot.png"), plot = for_ca_plot, dpi = 300, height = 5)
-img_for_ca_plot <- image_read(here("paper/image", "for_ca_plot.png"), density = 300)
-
-ggsave(here("paper/image", "pia_ca_plot.png"), plot = pia_ca_plot, dpi = 300, height = 5)
-img_pia_ca_plot <- image_read(here("paper/image", "pia_ca_plot.png"), density = 300)
-```
-
-## KV Difference
-
-We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording.
-We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
-
-### Forte to Piano
-
-There was no significant main effect of Category (*F*(`r ftop_ca_par_aov$anova_table[1,1]`, `r ftop_ca_par_aov$anova_table[1,2]`) = `r ftop_ca_par_aov$anova_table[1,4]`, *p* `r if(ftop_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-
-### Piano to Forte
-
-There was no significant main effect of Category (*F*(`r ptof_ca_par_aov$anova_table[1,1]`, `r ptof_ca_par_aov$anova_table[1,2]`) = `r ptof_ca_par_aov$anova_table[1,4]`, *p* `r if(ptof_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-
-These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had an intention to teach (*Fig \ref{fig:vel-diff-2}*).
-
-```{r vel-diff-bl-plot, include = FALSE}
-ftop_ca_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category == "baseline" | Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA) +
-  geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - FtoP") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-ptof_ca_plot <- ggplot(vel_diff[Subcomponent2 == "PtoF" & Category == "baseline" | Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA) +
-  geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - PtoF") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ftop_ca_plot.png"), plot = ftop_ca_plot, dpi = 300, height = 5)
-img_ftop_ca_plot <- image_read(here("paper/image", "ftop_ca_plot.png"), density = 300)
-
-ggsave(here("paper/image", "ptof_ca_plot.png"), plot = ptof_ca_plot, dpi = 300, height = 5)
-img_ptof_ca_plot <- image_read(here("paper/image", "ptof_ca_plot.png"), density = 300)
-```
-
-```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
-img_ioi_2 <- image_scale(img_ioi_ca_plot, "2200")
-image_write(img_ioi_2, path = here("paper/image", "Fig6.tiff"), format = "tiff")
-img_ioi_2
-```
-
-```{r plot-kot-2, out.width = "100%", fig.cap = "\\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
-# combine
-img_kot_2 <- image_append(c(img_leg_ca_plot, img_sta_ca_plot))
-img_kot_2 <- image_scale(img_kot_2, "2200")
-image_write(img_kot_2, path = here("paper/image", "Fig7.tiff"), format = "tiff")
-img_kot_2
-```
-
-```{r plot-vel-2, out.width = "100%", fig.cap = "\\label{fig:vel-2}Comparison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
-# combine
-img_vel_2 <- image_append(c(img_for_ca_plot, img_pia_ca_plot))
-img_vel_2 <- image_scale(img_vel_2, "2200")
-image_write(img_vel_2, path = here("paper/image", "Fig8.tiff"), format = "tiff")
-img_vel_2
-```
-
-```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Comparison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
-# combine
-img_vel_diff_2 <- image_append(c(img_ftop_ca_plot, img_ptof_ca_plot))
-img_vel_diff_2 <- image_scale(img_vel_diff_2, "2200")
-image_write(img_vel_diff_2, path = here("paper/image", "Fig9.tiff"), format = "tiff")
-img_vel_diff_2
 ```
@@
 We thank Candasch Acar for his help with data collection.
+
+\clearpage
+
+# Supplementary Material
+
+## 2. Comparison with baseline performance
+To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and to do their best as a performer.
+By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.
+
+## IOIs
+
+We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented.
+The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking.
+We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
+
+There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* `r if(ioi_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (*Fig \ref{fig:ioi-2}*).
+
+```{r ioi-bl-plot, include = FALSE}
+ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_hline(yintercept = 300, linetype = "dashed") + scale_y_continuous(breaks = seq(250, 300, 10), limits = c(250, 300)) +
+  geom_jitter(width = 0.15, alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_ca_plot.png"), plot = ioi_ca_plot, dpi = 300, height = 5)
+img_ioi_ca_plot <- image_read(here("paper/image", "ioi_ca_plot.png"), density = 300)
+```
+
+## KOT
+
+We compared the baseline performance with the performances in response to recordings where dynamics was implemented to examine whether participants performed differently depending on whether articulation was implemented or not in each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
+
+### Legato
+
+There was a significant main effect of Category (*F*(`r leg_ca_par_aov$anova_table[1,1]`, `r leg_ca_par_aov$anova_table[1,2]`) = `r leg_ca_par_aov$anova_table[1,4]`, *p* `r if(leg_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* < .001, baseline and dynamics-only: *p* = .008), suggesting that participants produced longer legato when they had the intention to teach (*Fig \ref{fig:kot-2}*, left).
+
+### Staccato
+
+There was no significant main effect of Category (*F*(`r sta_ca_par_aov$anova_table[1,1]`, `r sta_ca_par_aov$anova_table[1,2]`) = `r sta_ca_par_aov$anova_table[1,4]`, *p* `r if(sta_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had the intention to teach (*Fig \ref{fig:kot-2}*, right).
+
+```{r kot-bl-plot, include = FALSE}
+leg_ca_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category == "baseline" | Subcomponent1 == "Legato" & Category == "both" | Subcomponent1 == "Legato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
+  labs(x = "Category", y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-10, 120, 40)) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+sta_ca_plot <- ggplot(kot[Subcomponent1 == "Staccato" & Category == "baseline" | Subcomponent1 == "Staccato" & Category == "both" | Subcomponent1 == "Staccato" & Category == "dyn_only"], aes(x = factor(Category, level = c("baseline", "dyn_only", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
+  labs(x = "Category", y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-310, -170, 40)) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "leg_ca_plot.png"), plot = leg_ca_plot, dpi = 300, height = 5)
+img_leg_ca_plot <- image_read(here("paper/image", "leg_ca_plot.png"), density = 300)
+
+ggsave(here("paper/image", "sta_ca_plot.png"), plot = sta_ca_plot, dpi = 300, height = 5)
+img_sta_ca_plot <- image_read(here("paper/image", "sta_ca_plot.png"), density = 300)
+```
+
+## KV
+
+We compared the baseline performance with the performances in response to recordings where articulation was implemented to examine whether participants performed differently depending on whether dynamics was implemented or not in each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+### Forte
+
+There was no significant main effect of Category (*F*(`r for_ca_par_aov$anova_table[1,1]`, `r for_ca_par_aov$anova_table[1,2]`) = `r for_ca_par_aov$anova_table[1,4]`, *p* `r if(for_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had the intention to teach (*Fig \ref{fig:vel-2}*, left).
+
+### Piano
+
+There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* `r if(pia_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', pia_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05 , baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had the intention to teach (*Fig \ref{fig:vel-2}*, right).
+
+```{r vel-bl-plot, include = FALSE}
+for_ca_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category == "baseline" | Subcomponent2 == "Forte" & Category == "both" | Subcomponent2 == "Forte" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
+  labs(x = "Category", y = "KV (0-127)", title = "KV - Forte") + scale_y_continuous(breaks = seq(60, 110, 10)) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+pia_ca_plot <- ggplot(vel[Subcomponent2 == "Piano" & Category == "baseline" | Subcomponent2 == "Piano" & Category == "both" | Subcomponent2 == "Piano" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
+  labs(x = "Category", y = "KV (0-127)", title = "KV - Piano") + scale_y_continuous(breaks = seq(50, 70, 10)) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "for_ca_plot.png"), plot = for_ca_plot, dpi = 300, height = 5)
+img_for_ca_plot <- image_read(here("paper/image", "for_ca_plot.png"), density = 300)
+
+ggsave(here("paper/image", "pia_ca_plot.png"), plot = pia_ca_plot, dpi = 300, height = 5)
+img_pia_ca_plot <- image_read(here("paper/image", "pia_ca_plot.png"), density = 300)
+```
+
+## KV Difference
+
+We compared the baseline performance with the performances in response to recordings where articulation was implemented to examine whether participants performed differently depending on whether dynamics was implemented or not in each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+### Forte to Piano
+
+There was no significant main effect of Category (*F*(`r ftop_ca_par_aov$anova_table[1,1]`, `r ftop_ca_par_aov$anova_table[1,2]`) = `r ftop_ca_par_aov$anova_table[1,4]`, *p* `r if(ftop_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
+
+### Piano to Forte
+
+There was no significant main effect of Category (*F*(`r ptof_ca_par_aov$anova_table[1,1]`, `r ptof_ca_par_aov$anova_table[1,2]`) = `r ptof_ca_par_aov$anova_table[1,4]`, *p* `r if(ptof_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
+
+These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had the intention to teach (*Fig \ref{fig:vel-diff-2}*).
+
+```{r vel-diff-bl-plot, include = FALSE}
+ftop_ca_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category == "baseline" | Subcomponent2 == "FtoP" & Category == "both" | Subcomponent2 == "FtoP" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
+  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - FtoP") + scale_y_continuous(breaks = seq(-35, 0, 10)) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+ptof_ca_plot <- ggplot(vel_diff[Subcomponent2 == "PtoF" & Category == "baseline" | Subcomponent2 == "PtoF" & Category == "both" | Subcomponent2 == "PtoF" & Category == "art_only"], aes(x = factor(Category, level = c("baseline", "art_only", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(width = 0.15, alpha = 0.5) +
+  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - PtoF") + scale_y_continuous(breaks = seq(-5, 45, 10)) +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ftop_ca_plot.png"), plot = ftop_ca_plot, dpi = 300, height = 5)
+img_ftop_ca_plot <- image_read(here("paper/image", "ftop_ca_plot.png"), density = 300)
+
+ggsave(here("paper/image", "ptof_ca_plot.png"), plot = ptof_ca_plot, dpi = 300, height = 5)
+img_ptof_ca_plot <- image_read(here("paper/image", "ptof_ca_plot.png"), density = 300)
+```
+
+```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+img_ioi_2 <- image_scale(img_ioi_ca_plot, "2200")
+image_write(img_ioi_2, path = here("paper/image", "Fig6.tiff"), format = "tiff")
+img_ioi_2
+```
+
+```{r plot-kot-2, out.width = "100%", fig.cap = "\\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_kot_2 <- image_append(c(img_leg_ca_plot, img_sta_ca_plot))
+img_kot_2 <- image_scale(img_kot_2, "2200")
+image_write(img_kot_2, path = here("paper/image", "Fig7.tiff"), format = "tiff")
+img_kot_2
+```
+
+```{r plot-vel-2, out.width = "100%", fig.cap = "\\label{fig:vel-2}Comparison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_vel_2 <- image_append(c(img_for_ca_plot, img_pia_ca_plot))
+img_vel_2 <- image_scale(img_vel_2, "2200")
+image_write(img_vel_2, path = here("paper/image", "Fig8.tiff"), format = "tiff")
+img_vel_2
+```
+
+```{r plot-vel-diff-2, out.width = "100%", fig.cap = "\\label{fig:vel-diff-2}Comparison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+# combine
+img_vel_diff_2 <- image_append(c(img_ftop_ca_plot, img_ptof_ca_plot))
+img_vel_diff_2 <- image_scale(img_vel_diff_2, "2200")
+image_write(img_vel_diff_2, path = here("paper/image", "Fig9.tiff"), format = "tiff")
+img_vel_diff_2
+```

### 2022-11-29 paper/manuscript.tex
update results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 \hypersetup{
-  pdftitle={How do expert pianists adapt their pedagogical demonstration to novices' skill levels?},
+  pdftitle={Do expert pianists adapt their pedagogical demonstrations to novices' skills?},
   pdfauthor={Atsuko Tominaga1, Günther Knoblich1, \& Natalie Sebanz1},
@@
 
-\title{How do expert pianists adapt their pedagogical demonstration to novices' skill levels?}
+\title{Do expert pianists adapt their pedagogical demonstrations to novices' skills?}
 \author{Atsuko Tominaga\textsuperscript{1}, Günther Knoblich\textsuperscript{1}, \& Natalie Sebanz\textsuperscript{1}}
@@
 
-\shorttitle{How do expert pianists adapt their pedagogical demonstration to novices' skill levels?}
+\shorttitle{Do expert pianists adapt their pedagogical demonstrations to novices' skills?}
 
@@
 
-Teaching is effective when experts and novices dynamically interact with each other (Byrne \& Rapaport, 2011).
-In the literature on cultural evolution, it has been considered that humans have developed knowledge and skills by imitating knowledgeable others (experts) and transmitting what they have acquired to the next generation (Mesoudi, 2016).
-However, information flows between experts and novices are mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022).
+Effective teaching involves experts and novices dynamically interacting with each other (Byrne \& Rapaport, 2011).
+Typically, information flows between experts and novices is mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022).
 Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).
 
-Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019).
-When experts have an intention to teach, they modulate their behaviour.
+Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts.
 For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
@@
 
-We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
+First, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
 The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
 For post-hoc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.
+
+In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced for non-teaching purposes), we performed one-way repeated-measures ANOVAs.
+For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent (i.e., Legato, Staccato, Forte, Piano).
 
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) or Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
-Therefore, participants did not change the tempo depending on the types of recordings (\emph{Fig \ref{fig:ioi-1}}).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
+This indicates that participants did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).
 
@@
 
-There was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
-However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002).
+As predicted, there was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
+There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002).
 Participants produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).
@@
 
-There was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009).
-However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.093, \(\eta_G^2\) = 0.001).
+In line with our prediction, there was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009).
+There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.093, \(\eta_G^2\) = 0.001).
 Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) or Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
-Participants did not change performances in terms of notated forte depending on the types of recordings (\emph{Fig \ref{fig:vel-1}}, left).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) nor the main effect of Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
+Participants did not change their performances in terms of notated forte depending on the type of recordings (\emph{Fig \ref{fig:vel-1}}, left).
 
@@
 There was a significant main effect of Articulation (\emph{F}(1, 19) = 7.18, \emph{p} = 0.01, \(\eta_G^2\) = 0.060).
-However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000).
+However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000).
 Participants produced softer piano when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) or Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.336, \(\eta_G^2\) = 0.001).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) nor Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.336, \(\eta_G^2\) = 0.001).
 
@@
 
-Neither main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.150, \(\eta_G^2\) = 0.005) or Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.150, \(\eta_G^2\) = 0.005) nor Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
 
-These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of recordings (\emph{Fig \ref{fig:vel-diff-1}}).
+These results indicated that participants did not change their performances in terms of dynamics contrast depending on the type of recordings (\emph{Fig \ref{fig:vel-diff-1}}).
 
@@
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms).
+A dashed line represents the tempo given by a metronome.
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
@@
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-1}
-\end{figure}
-
-\hypertarget{comparison-with-baseline-performance}{%
-\subsection{Comparison with baseline performance}\label{comparison-with-baseline-performance}}
-
-In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced before listening to any of the recordings), we performed one-way repeated-measures ANOVAs.
-For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent.
-To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer.
-By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.
-
-\hypertarget{iois-1}{%
-\subsection{IOIs}\label{iois-1}}
-
-We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of novices' skill levels.
-We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
-
-There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had an intention to teach (\emph{Fig \ref{fig:ioi-2}}).
-
-\hypertarget{kot-1}{%
-\subsection{KOT}\label{kot-1}}
-
-We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not in each recording.
-We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
-
-\hypertarget{legato-1}{%
-\subsubsection{Legato}\label{legato-1}}
-
-There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.060; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
-
-\hypertarget{staccato-1}{%
-\subsubsection{Staccato}\label{staccato-1}}
-
-There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, right).
-
-\hypertarget{kv-1}{%
-\subsection{KV}\label{kv-1}}
-
-We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording.
-We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
-
-\hypertarget{forte-1}{%
-\subsubsection{Forte}\label{forte-1}}
-
-There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, left).
-
-\hypertarget{piano-1}{%
-\subsubsection{Piano}\label{piano-1}}
-
-There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
-
-\hypertarget{kv-difference-1}{%
-\subsection{KV Difference}\label{kv-difference-1}}
-
-We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not in each recording.
-We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
-
-\hypertarget{forte-to-piano-1}{%
-\subsubsection{Forte to Piano}\label{forte-to-piano-1}}
-
-There was no significant main effect of Category (\emph{F}(1.53, 29.15) = 1.73, \emph{p} = 0.20, \(\eta_G^2\) = 0.014; Greenhouse-Geisser corrected).
-
-\hypertarget{piano-to-forte-1}{%
-\subsubsection{Piano to Forte}\label{piano-to-forte-1}}
-
-There was no significant main effect of Category (\emph{F}(1.28, 24.39) = 2.27, \emph{p} = 0.14, \(\eta_G^2\) = 0.013; Greenhouse-Geisser corrected).
-
-These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-diff-2}}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Comparison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Comparison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-2}
 \end{figure}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mesoudi_2016a}{}}%
-Mesoudi, A. (2016). Cultural evolution: Integrating psychology, evolution and culture. \emph{Current Opinion in Psychology}, \emph{7}, 17--22. \url{https://doi.org/10.1016/j.copsyc.2015.07.001}
-
 \leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
@@
 
+\clearpage
+
+\hypertarget{supplementary-material}{%
+\section{Supplementary Material}\label{supplementary-material}}
+
+\hypertarget{comparison-with-baseline-performance}{%
+\subsection{2. Comparison with baseline performance}\label{comparison-with-baseline-performance}}
+
+To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and to do their best as a performer.
+By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.
+
+\hypertarget{iois-1}{%
+\subsection{IOIs}\label{iois-1}}
+
+We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented.
+The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking.
+We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
+
+There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (\emph{Fig \ref{fig:ioi-2}}).
+
+\hypertarget{kot-1}{%
+\subsection{KOT}\label{kot-1}}
+
+We compared the baseline performance with the performances in response to recordings where dynamics was implemented to examine whether participants performed differently depending on whether articulation was implemented or not in each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.
+
+\hypertarget{legato-1}{%
+\subsubsection{Legato}\label{legato-1}}
+
+There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.060; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001, baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had the intention to teach (\emph{Fig \ref{fig:kot-2}}, left).
+
+\hypertarget{staccato-1}{%
+\subsubsection{Staccato}\label{staccato-1}}
+
+There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse-Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had the intention to teach (\emph{Fig \ref{fig:kot-2}}, right).
+
+\hypertarget{kv-1}{%
+\subsection{KV}\label{kv-1}}
+
+We compared the baseline performance with the performances in response to recordings where articulation was implemented to examine whether participants performed differently depending on whether dynamics was implemented or not in each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+\hypertarget{forte-1}{%
+\subsubsection{Forte}\label{forte-1}}
+
+There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse-Geisser corrected), suggesting that participants did not play forte differently depending on whether they had the intention to teach (\emph{Fig \ref{fig:vel-2}}, left).
+
+\hypertarget{piano-1}{%
+\subsubsection{Piano}\label{piano-1}}
+
+There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had the intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
+
+\hypertarget{kv-difference-1}{%
+\subsection{KV Difference}\label{kv-difference-1}}
+
+We compared the baseline performance with the performances in response to recordings where articulation was implemented to examine whether participants performed differently depending on whether dynamics was implemented or not in each recording.
+We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.
+
+\hypertarget{forte-to-piano-1}{%
+\subsubsection{Forte to Piano}\label{forte-to-piano-1}}
+
+There was no significant main effect of Category (\emph{F}(1.53, 29.15) = 1.73, \emph{p} = 0.20, \(\eta_G^2\) = 0.014; Greenhouse-Geisser corrected).
+
+\hypertarget{piano-to-forte-1}{%
+\subsubsection{Piano to Forte}\label{piano-to-forte-1}}
+
+There was no significant main effect of Category (\emph{F}(1.28, 24.39) = 2.27, \emph{p} = 0.14, \(\eta_G^2\) = 0.013; Greenhouse-Geisser corrected).
+
+These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had the intention to teach (\emph{Fig \ref{fig:vel-diff-2}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Comparison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Comparison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right).
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-2}
+\end{figure}
+
 \end{document}

### 2022-12-01 paper/manuscript.Rmd
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   geom_hline(yintercept = 300, linetype = "dashed") +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") + scale_y_continuous(breaks = seq(250, 300, 10), limits = c(250, 300)) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") + scale_y_continuous(breaks = seq(240, 320, 20), limits = c(240, 320)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-10, 120, 40)) +
+  labs(y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-20, 140, 40), limits = c(-20, 140)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-310, -170, 40)) +
+  labs(y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-280, -120, 40), limits = c(-280, -120)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV (0-127)", title = "KV - Forte") + scale_y_continuous(breaks = seq(60, 110, 10)) +
+  labs(y = "KV (0-127)", title = "KV - Forte") + scale_y_continuous(breaks = seq(60, 100, 10), limits = c(60, 100)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV (0-127)", title = "KV - Piano") + scale_y_continuous(breaks = seq(50, 70, 10)) +
+  labs(y = "KV (0-127)", title = "KV - Piano") + scale_y_continuous(breaks = seq(40, 80, 10), limits = c(40, 80)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV Difference (-127-127)", title = "KV Difference - FtoP") + scale_y_continuous(breaks = seq(-35, 0, 10)) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - FtoP") + scale_y_continuous(breaks = seq(-40, 10, 10), limits = c(-40, 10)) +
   scale_color_aaas() +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoP") + scale_y_continuous(breaks = seq(-5, 45, 10)) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoP") + scale_y_continuous(breaks = seq(-10, 40, 10), limits = c(-10, 40)) +
   scale_color_aaas() +
@@
   geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
-  geom_hline(yintercept = 300, linetype = "dashed") + scale_y_continuous(breaks = seq(250, 300, 10), limits = c(250, 300)) +
+  geom_hline(yintercept = 300, linetype = "dashed") + scale_y_continuous(breaks = seq(200, 380, 20), limits = c(200, 380)) +
   geom_jitter(width = 0.15, alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
@@
   geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-10, 120, 40)) +
+  labs(x = "Category", y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-20, 140, 40), limits = c(-20, 140)) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-310, -170, 40)) +
+  labs(x = "Category", y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-320, -160, 40), limits = c(-320, -160)) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV (0-127)", title = "KV - Forte") + scale_y_continuous(breaks = seq(60, 110, 10)) +
+  labs(x = "Category", y = "KV (0-127)", title = "KV - Forte") + scale_y_continuous(breaks = seq(60, 100, 10), limits = c(60, 100)) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV (0-127)", title = "KV - Piano") + scale_y_continuous(breaks = seq(50, 70, 10)) +
+  labs(x = "Category", y = "KV (0-127)", title = "KV - Piano") + scale_y_continuous(breaks = seq(40, 80, 10), limits = c(40, 80)) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - FtoP") + scale_y_continuous(breaks = seq(-35, 0, 10)) +
+  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - FtoP") + scale_y_continuous(breaks = seq(-40, 10, 10), limits = c(-40, 10)) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
@@
   geom_jitter(width = 0.15, alpha = 0.5) +
-  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - PtoF") + scale_y_continuous(breaks = seq(-5, 45, 10)) +
+  labs(x = "Category", y = "KV Difference (-127-127)", title = "KV Difference - PtoF") + scale_y_continuous(breaks = seq(-10, 40, 10), limits = c(-10, 40)) +
   theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")

### 2022-12-01 paper/manuscript.Rmd
change chunk names
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r ioi-plot, include = FALSE}
+```{r ioi-plot-1, include = FALSE}
 ioi_all_plot <- ggplot(ioi[Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
@@
   geom_hline(yintercept = 300, linetype = "dashed") +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") + scale_y_continuous(breaks = seq(240, 320, 20), limits = c(240, 320)) +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") +
+  scale_y_continuous(breaks = seq(240, 320, 20), limits = c(240, 320)) +
   scale_color_aaas() +
@@
 
-```{r kot-plot, include = FALSE}
+```{r kot-plot-1, include = FALSE}
 leg_all_plot <- ggplot(kot[Subcomponent1 == "Legato" & Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
@@
 
-```{r vel-plot, include = FALSE}
+```{r vel-plot-1, include = FALSE}
 for_all_plot <- ggplot(vel[Subcomponent2 == "Forte" & Category != "baseline"], aes(x = Dynamics, y = Mean, color = Articulation)) +
@@
 
-```{r vel-diff-plot, include = FALSE}
+```{r vel-diff-plot-1, include = FALSE}
 ftop_all_plot <- ggplot(vel_diff[Subcomponent2 == "FtoP" & Category != "baseline"], aes(x = Dynamics, y = Mean, color = Articulation)) +

### 2022-12-01 analysis/baseline.Rmd
implement order to vel/vel-diff
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 vel <- vel_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
-setorder(vel, "Subcomponent2", "SubNr", "Dynamics", "Articulation", "Category")
+setorder(vel, "SubNr", "Subcomponent2", "SubNr", "Dynamics", "Articulation", "Category")
 vel
@@
 vel_diff <- vel_diff_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Category, Subcomponent2)]
-setorder(vel_diff, "SubNr", "Dynamics", "Articulation", "Category")
+setorder(vel_diff, "SubNr", "Subcomponent2", "Dynamics", "Articulation", "Category")
 vel_diff

### 2022-12-01 analysis/baseline.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE------------------------------------------------------------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -12,7 +12,7 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE-------------------------------------------------------------------------------------
+## ----file, include = FALSE----------------------------------------------------
 data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
 data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
 data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
@@ -23,7 +23,7 @@ data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
 data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 
 
-## ----prep, include = FALSE-------------------------------------------------------------------------------------
+## ----prep, include = FALSE----------------------------------------------------
 # read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
 dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
@@ -70,7 +70,7 @@ dt_vel <- rbind(dt_vel, dt_bl_vel)
 dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
 
 
-## ----ioi-------------------------------------------------------------------------------------------------------
+## ----ioi----------------------------------------------------------------------
 # for each individual
 dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
 dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
@@ -82,17 +82,17 @@ setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
 ioi
 
 
-## ----ioi-plot, fig.width = 6-----------------------------------------------------------------------------------
+## ----ioi-plot, fig.width = 6--------------------------------------------------
[…]

### 2022-12-01 paper/manuscript.Rmd
reflected some feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners [@mcellin_2017; @uther_2007].
+For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own [@mcellin_2017].
 
-Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
-@fukuyama_2015 revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
-For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have enough motor skills to produce such actions.
-The results indicated that adults monitored infants' actions and abilities, and adapted their demonstration accordingly.
+Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill levels of learners.
+@fukuyama_2015 found that mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation skills.
+For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
+To the best of our knowledge, it is unknown whether people teaching skills to adult learners also adjust their demonstrations as a function of the learners' demonstrated skills.
 
+To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
 In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
-For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had the intention to teach dynamics.
-Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance.
-If teaching is an adaptive and scaffolding process, it can be expected that experts should not use such exaggeration if novices became able to produce them.
+In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics.
+When they had the intention to demonstrate articulation, they exaggerated articulation by producing shorter staccato.
+In that study, teachers were instructed to demonstrate techniques to hypothetical learners and they did not hear novices' particular performances.
+This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way caregivers have been shown to adapt demonstrations of actions to their infants' skills.
+If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
 
-In the current study, we investigated whether and how expert pianists adapted their performance depending on the skill levels of novice learners.
-Instead of recruiting actual novice pianists, we created artificial recordings to manipulate the skill levels of novices to maximise the experimental control.
-The recordings were made based on the performance data from our previous experiment (Tominaga et al., under review).
-All the recordings had the same musical structure (i.e., the same piece) but had different implementations of articulation and dynamics.
-In the ideal performance, both articulation and dynamics should be implemented as notated on the sheet music (*Fig \ref{fig:stim}*).
-We made the recordings so that both techniques were implemented in the first quarter, only articulation was implemented in the second quarter, only dynamics was implemented in the third quarter and neither of them was implemented in the last quarter of the recordings.
-Therefore, in this experiment, the skill levels of novices corresponded to the implementation of articulation and dynamics.
+To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review).
+All the recordings were of the same same piece but the recordings differed in the implementation of articulation and dynamics.
+According to the notated expressions in the sheet music, proper performance of the piece required specific modulations with respect to both articulation and dynamics (*Fig \ref{fig:stim}*).
+We generated four types of recordings.
+In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation were implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
+This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.
 
@@
 
-The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of novices.
+The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novice whom they are trying to teach.
 We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
-The recordings where the two techniques were implemented were supposed to replicate the performances of students with high skills.
-On the other hand, the recordings where none of the two techniques was implemented were supposed to replicate the performances of students with low skills.
+The recordings where the two techniques were implemented were supposed to resemble the performances of students with good skills.
+The recordings where neither of the two techniques was implemented were supposed to resemble the performances of students with lower skills who had not yet learnt how to implement articulation and dynamics.
 The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.
 
-We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
-Even compared with the tempo of the baseline performance, they did not change the tempo either.
-These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
-Although slower demonstration is generally considered to be used for teaching [@mcellin_2017; @saint-georges_2013; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
-
-For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
+Expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
 This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings.
@@
 
+We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
+Also compared with the tempo of the baseline performance, they did not change the tempo in the teaching context.
+These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
+Although slower demonstration is generally considered to be used for teaching [@mcellin_2017; @saint-georges_2013; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
+
 Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
-One possibility would be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
-In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
-Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it.
+One possibility is that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
+In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
+Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to highlight it.
 Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.
 
-One of the limitations of the current study is that participants did not have an idea about what the ideal performance was.
+One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind.
 As some participants reported that "(students had) very different interpretations of the melody" or "all kids played rhythmically correct" in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
-One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors by comparing each recording with the model performance.
+One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants (expert pianists) to listen to it first so that they can detect errors by comparing each recording with the model performance.
 
 The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo.
-This was to make sure that we manipulated articulation and dynamics features of the recordings only.
-However, all the recordings might not have sounded as if unskillful students were playing.
-In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.
+This was to make sure that only articulation and dynamics features of varied.
+However, some of the recordings might not have sounded as if unskillful students were playing.
+In order to create more realistic novice performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask beginner pianists to play the piece and extract the temporal features of their performances.
 

### 2022-12-01 paper/manuscript.tex
reflected some feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).
+For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own (McEllin et al., 2017).
 
-Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners.
+Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill levels of learners.
 Fukuyama et al.
-(2015) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation.
-For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have enough motor skills to produce such actions.
-The results indicated that adults monitored infants' actions and abilities, and adapted their demonstration accordingly.
+(2015) found that mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation skills.
+For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
+To the best of our knowledge, it is unknown whether people teaching skills to adult learners also adjust their demonstrations as a function of the learners' demonstrated skills.
 
+To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
 In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
-For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had the intention to teach dynamics.
-Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance.
-If teaching is an adaptive and scaffolding process, it can be expected that experts should not use such exaggeration if novices became able to produce them.
+In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics.
+When they had the intention to demonstrate articulation, they exaggerated articulation by producing shorter staccato.
+In that study, teachers were instructed to demonstrate techniques to hypothetical learners and they did not hear novices' particular performances.
+This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way caregivers have been shown to adapt demonstrations of actions to their infants' skills.
+If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
 
-In the current study, we investigated whether and how expert pianists adapted their performance depending on the skill levels of novice learners.
-Instead of recruiting actual novice pianists, we created artificial recordings to manipulate the skill levels of novices to maximise the experimental control.
-The recordings were made based on the performance data from our previous experiment (Tominaga et al., under review).
-All the recordings had the same musical structure (i.e., the same piece) but had different implementations of articulation and dynamics.
-In the ideal performance, both articulation and dynamics should be implemented as notated on the sheet music (\emph{Fig \ref{fig:stim}}).
-We made the recordings so that both techniques were implemented in the first quarter, only articulation was implemented in the second quarter, only dynamics was implemented in the third quarter and neither of them was implemented in the last quarter of the recordings.
-Therefore, in this experiment, the skill levels of novices corresponded to the implementation of articulation and dynamics.
+To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review).
+All the recordings were of the same same piece but the recordings differed in the implementation of articulation and dynamics.
+According to the notated expressions in the sheet music, proper performance of the piece required specific modulations with respect to both articulation and dynamics (\emph{Fig \ref{fig:stim}}).
+We generated four types of recordings.
+In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation were implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
+This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.
 
@@
 
-The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of novices.
+The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novice whom they are trying to teach.
 We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
-The recordings where the two techniques were implemented were supposed to replicate the performances of students with high skills.
-On the other hand, the recordings where none of the two techniques was implemented were supposed to replicate the performances of students with low skills.
+The recordings where the two techniques were implemented were supposed to resemble the performances of students with good skills.
+The recordings where neither of the two techniques was implemented were supposed to resemble the performances of students with lower skills who had not yet learnt how to implement articulation and dynamics.
 The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.
 
-We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students.
-Even compared with the tempo of the baseline performance, they did not change the tempo either.
-These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics.
-Although slower demonstration is generally considered to be used for teaching (McEllin et al., 2017; Saint-Georges et al., 2013; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).
-
-For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
+Expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings.
 This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings.
@@
 
+We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
+Also compared with the tempo of the baseline performance, they did not change the tempo in the teaching context.
+These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
+Although slower demonstration is generally considered to be used for teaching (McEllin et al., 2017; Saint-Georges et al., 2013; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
+
 Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
-One possibility would be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
-In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
-Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it.
+One possibility is that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
+In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
+Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to highlight it.
 Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.
 
-One of the limitations of the current study is that participants did not have an idea about what the ideal performance was.
+One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind.
 As some participants reported that ``(students had) very different interpretations of the melody'' or ``all kids played rhythmically correct'' in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
-One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors by comparing each recording with the model performance.
+One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants (expert pianists) to listen to it first so that they can detect errors by comparing each recording with the model performance.
 
 The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo.
-This was to make sure that we manipulated articulation and dynamics features of the recordings only.
-However, all the recordings might not have sounded as if unskillful students were playing.
-In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.
+This was to make sure that only articulation and dynamics features of varied.
+However, some of the recordings might not have sounded as if unskillful students were playing.
+In order to create more realistic novice performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask beginner pianists to play the piece and extract the temporal features of their performances.
 

### 2022-12-02 paper/manuscript.Rmd
update intro/discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
     email         : "tominaga_atsuko@phd.ceu.edu"
-    # role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-    #   - "Conceptualization"
-    #   - "Writing - Original Draft Preparation"
-    #   - "Writing - Review & Editing"
+    role:         # Contributorship roles (e.g., CRediT, https://www.elsevier.com/authors/policies-and-guidelines/credit-author-statement)
+      - "Conceptualization"
+      - "Methodology"
+      - "Software"
+      - "Formal analysis"
+      - "Data curation"
+      - "Visualisation"
+      - "Writing - Original Draft Preparation"
+      - "Writing - Review & Editing"
   - name          : "Günther Knoblich"
     affiliation   : "1"
-    # role:
-    #   - "Writing - Review & Editing"
-    #   - "Supervision"
+    role:
+      - "Conceptualization"
+      - "Methodology"
+      - "Resources"
+      - "Writing - Review & Editing"
+      - "Supervision"
+      - "Funding acquisition"
   - name          : "Natalie Sebanz"
     affiliation   : "1"
-    # role:
-    #   - "Writing - Review & Editing"
-    #   - "Supervision"
+    role:
+      - "Conceptualization"
+      - "Methodology"
+      - "Resources"
+      - "Writing - Review & Editing"
+      - "Supervision"
+      - "Funding acquisition"
 
@@
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts.
-For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002].
+When caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002].
 Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners [@mcellin_2017; @uther_2007].
@@
 When they had the intention to demonstrate articulation, they exaggerated articulation by producing shorter staccato.
-In that study, teachers were instructed to demonstrate techniques to hypothetical learners and they did not hear novices' particular performances.
-This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way caregivers have been shown to adapt demonstrations of actions to their infants' skills.
+In that study, teachers were instructed to teach techniques to hypothetical learners and they did not hear novices' particular performances.
+This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
 If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
@@
 To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review).
-All the recordings were of the same same piece but the recordings differed in the implementation of articulation and dynamics.
-According to the notated expressions in the sheet music, proper performance of the piece required specific modulations with respect to both articulation and dynamics (*Fig \ref{fig:stim}*).
+All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
+According to the notated expressions in the sheet music, the proper performance of the piece required specific modulations with respect to both articulation and dynamics (*Fig \ref{fig:stim}*).
 We generated four types of recordings.
-In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation were implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
+In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation was implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
 This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.
@@
 
-The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novice whom they are trying to teach.
-We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
+The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novices whom they are trying to teach.
+We created artificial recordings to manipulate the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
 The recordings where the two techniques were implemented were supposed to resemble the performances of students with good skills.
@@
 
-The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo.
-This was to make sure that only articulation and dynamics features of varied.
+The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed with a stable tempo.
+This was to make sure that only articulation and dynamics features of the recordings varied.
 However, some of the recordings might not have sounded as if unskillful students were playing.

### 2022-12-02 paper/manuscript.tex
update intro/discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+The authors made the following contributions.
+Atsuko Tominaga: Conceptualization, Methodology, Software, Formal analysis, Data curation, Visualisation, Writing - Original Draft Preparation, Writing - Review \& Editing; Günther Knoblich: Conceptualization, Methodology, Resources, Writing - Review \& Editing, Supervision, Funding acquisition; Natalie Sebanz: Conceptualization, Methodology, Resources, Writing - Review \& Editing, Supervision, Funding acquisition.
+
 Correspondence concerning this article should be addressed to Atsuko Tominaga, Quellenstraße 51, 1100 Vienna, Austria.
@@
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts.
-For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
+When caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
 Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).
@@
 To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review).
-All the recordings were of the same same piece but the recordings differed in the implementation of articulation and dynamics.
-According to the notated expressions in the sheet music, proper performance of the piece required specific modulations with respect to both articulation and dynamics (\emph{Fig \ref{fig:stim}}).
+All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
+According to the notated expressions in the sheet music, the proper performance of the piece required specific modulations with respect to both articulation and dynamics (\emph{Fig \ref{fig:stim}}).
 We generated four types of recordings.
-In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation were implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
+In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation was implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
 This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.

### 2022-12-04 paper/manuscript.Rmd
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The recordings where neither of the two techniques was implemented were supposed to resemble the performances of students with lower skills who had not yet learnt how to implement articulation and dynamics.
-The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.
+The recordings where either of the two techniques was implemented but the other was missing were created to examine whether expert pianists could recognise a particular problem that students have (e.g., only articulation was missing) and adjust their performance accordingly (e.g., experts should modulate their performance relevant to articulation, but not dynamics).
 
@@
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
+This may imply that participants prioritised the teaching of articulation and did not modify two performance parameters at the same time.
+This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the action into sub-actions and adding complexity step by step, which enables infants to build up motor repertoires to achieve a goal [@zukow-goldring_2007].
 
-We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
-Also compared with the tempo of the baseline performance, they did not change the tempo in the teaching context.
-These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
-Although slower demonstration is generally considered to be used for teaching [@mcellin_2017; @saint-georges_2013; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
-
-Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
-One possibility is that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
+The reason why expert pianists seemed to prioritise the teaching of articulation might be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
 In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
@@
 Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.
+
+We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
+Also compared with the tempo of the baseline performance (see *Supplementary Material*), they did not change the tempo in the teaching context.
+These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
+Although slower demonstration is generally considered to be used for teaching [@brand_2002; @mcellin_2017; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
 

### 2022-12-04 paper/manuscript.tex
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
+This may imply that participants prioritised the teaching of articulation and did not modify two performance parameters at the same time.
+This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the action into sub-actions and adding complexity step by step, which enables infants to build up motor repertoires to achieve a goal (Zukow-Goldring \& Arbib, 2007).
 
-We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
-Also compared with the tempo of the baseline performance, they did not change the tempo in the teaching context.
-These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
-Although slower demonstration is generally considered to be used for teaching (McEllin et al., 2017; Saint-Georges et al., 2013; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
-
-Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation.
-One possibility is that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
+The reason why expert pianists seemed to prioritise the teaching of articulation might be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
 In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
@@
 Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.
+
+We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
+Also compared with the tempo of the baseline performance (see \emph{Supplementary Material}), they did not change the tempo in the teaching context.
+These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
+Although slower demonstration is generally considered to be used for teaching (Brand et al., 2002; McEllin et al., 2017; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-saint-georges_2013}{}}%
-Saint-Georges, C., Chetouani, M., Cassel, R., Apicella, F., Mahdhaoui, A., Muratori, F., \ldots{} Cohen, D.
-(2013).
-Motherese in {Interaction}: {At} the {Cross-Road} of {Emotion} and {Cognition}?
-({A Systematic Review}).
-\emph{PLOS ONE}, \emph{8}(10), e78103.
-\url{https://doi.org/10.1371/journal.pone.0078103}
-
 \leavevmode\vadjust pre{\hypertarget{ref-schaik_2019}{}}%
@@
 \url{https://doi.org/10.1016/j.specom.2006.10.003}
+
+\leavevmode\vadjust pre{\hypertarget{ref-zukow-goldring_2007}{}}%
+Zukow-Goldring, P., \& Arbib, M.
+A.
+(2007).
+Affordances, effectivities, and assisted imitation: {Caregivers} and the directing of attention.
+\emph{Neurocomputing}, \emph{70}(13), 2181--2193.
+\url{https://doi.org/10.1016/j.neucom.2006.02.029}
 

### 2022-12-04 paper/manuscript.Rmd
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 The recordings where neither of the two techniques was implemented were supposed to resemble the performances of students with lower skills who had not yet learnt how to implement articulation and dynamics.
-The recordings where either of the two techniques was implemented but the other was missing were created to examine whether expert pianists could recognise a particular problem that students have (e.g., only articulation was missing) and adjust their performance accordingly (e.g., experts should modulate their performance relevant to articulation, but not dynamics).
+The recordings where either of the two techniques was implemented but the other was missing were created to examine whether expert pianists could recognise a particular problem that students have (e.g., only articulation was missing) and adjust their performance accordingly (e.g., expert pianists should modulate their performance relevant to articulation, but not dynamics).
 
@@
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
-This may imply that participants prioritised the teaching of articulation and did not modify two performance parameters at the same time.
-This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the action into sub-actions and adding complexity step by step, which enables infants to build up motor repertoires to achieve a goal [@zukow-goldring_2007].
+This may imply that participants prioritised articulation over dynamics and did not modify two performance parameters at the same time.
+This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the sequential action into discrete sub-actions and adding complexity step by step [@zukow-goldring_2007].
+In our study, expert pianists might have focused on the teaching of articulation first.
 

### 2022-12-04 paper/manuscript.tex
update discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
-This may imply that participants prioritised the teaching of articulation and did not modify two performance parameters at the same time.
-This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the action into sub-actions and adding complexity step by step, which enables infants to build up motor repertoires to achieve a goal (Zukow-Goldring \& Arbib, 2007).
+This may imply that participants prioritised articulation over dynamics and did not modify two performance parameters at the same time.
+This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the sequential action into discrete sub-actions and adding complexity step by step (Zukow-Goldring \& Arbib, 2007).
+In our study, expert pianists might have focused on the teaching of articulation first.
 

### 2022-12-05 paper/manuscript.Rmd
change the order of dependent variables / update intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Effective teaching involves experts and novices dynamically interacting with each other [@byrne_2011].
-Typically, information flows between experts and novices is mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive [@charbonneau_2022].
-Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process [@okazaki_2019; @flynn_2022].
+During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities until novices can perform the actions independently without the help of experts [@mermelshtine_2017].
+Particularly, it is important for experts to find what novices have a problem from their failed attempt and to modify their behaviour [@ugur_2015; @zukow-goldring_2007].
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts.
-When caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002].
-Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners [@mcellin_2017; @uther_2007].
-For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own [@mcellin_2017].
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared with when they were demonstrating to adults [@brand_2002].
+Such exaggeration was observed not only towards infant learners but also when knowledgeable adults were interacting with adult learners [@mcellin_2017].
+For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own.
 
-Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill levels of learners.
-@fukuyama_2015 found that mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation skills.
+Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners [@fukuyama_2015] In this study, mothers were asked to demonstrate demonstrate a cup-nesting task to their infants.
+There were four cups with different colour and size.
+The mothers showed how to stack and fit the nesting cups so that infants could reproduce the actions by themselves.
+The findings showed that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
 For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
@@
 To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
-In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
+In our previous research (Tominaga et al., accepted), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
 In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics.
@@
 
-To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review).
+To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., accepted).
 All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
@@
 
-We hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
+Evidence from parental scaffolding research suggests that caregivers scaffold infants' learning when the infants failed to produce an expected action [@wood_1975; @zukow-goldring_2007].
+Therefore, we hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
 More precisely, expert pianists would exaggerate articulation (i.e., producing longer legato and shorter staccato) when articulation was not implemented in the recordings whereas expert pianists would exaggerate dynamics (i.e., producing  louder forte and softer piano, and larger contrast between forte and piano) when dynamics was not implemented in the recordings.
@@
 The dependent variables were computed from MIDI data for data analysis.
-Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
 Key-overlap time (KOT) is the difference between the offset time (i.e., key release time) of the current tone and the onset time of the ensuing tone and is a measure for the smoothness of musical sequences.
@@
 We particularly focused on specific points where each subcomponent changed from one to the other (i.e., forte to piano or piano to forte) to measure dynamics contrast between forte and piano.
+Interonset intervals (IOIs) are the intervals between onsets of adjacent notes and provide a measure of tempo.
 
@@
 As a result, less than 1 % of total responses were removed.
-After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference, defined as values more than 3 standard deviations from the mean of each dependent variable.
+After removing pitch errors, we removed outliers for KOT, KV, KV Difference and IOIs, defined as values more than 3 standard deviations from the mean of each dependent variable.
 For each dependent variable, this resulted in less than 2 % of overall responses being removed as outliers.
@@
 First, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.
-absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
+absent) and Dynamics (present vs absent) for each dependent variable (i.e.., KOT, KV, KV Difference, IOIs).
 The *aov_ez* function in the *afex* R package was used for a repeated-measures ANOVA.
@@
 source(here("analysis", "baseline.R"), chdir = TRUE)
-```
-
-## IOIs
-
-Neither the main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
-This indicates that participants did not change the tempo depending on the type of recordings (*Fig \ref{fig:ioi-1}*).
-
-```{r ioi-plot-1, include = FALSE}
-ioi_all_plot <- ggplot(ioi[Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
-  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
-  geom_hline(yintercept = 300, linetype = "dashed") +
-  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") +
-  scale_y_continuous(breaks = seq(240, 320, 20), limits = c(240, 320)) +
-  scale_color_aaas() +
-  theme_pubr(base_size = 20, base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_all_plot.png"), plot = ioi_all_plot, dpi = 300, height = 5)
-img_ioi_all_plot <- image_read(here("paper/image", "ioi_all_plot.png"), density = 300)
 ```
@@
 
+## IOIs
+
+Neither the main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
+This indicates that participants did not change the tempo depending on the type of recordings (*Fig \ref{fig:ioi-1}*).
+
+```{r ioi-plot-1, include = FALSE}
+ioi_all_plot <- ggplot(ioi[Category != "baseline"], aes(x = Articulation, y = Mean, color = Dynamics)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_hline(yintercept = 300, linetype = "dashed") +
+  geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) + labs(y = "IOI (ms)", title = "IOI") +
+  scale_y_continuous(breaks = seq(240, 320, 20), limits = c(240, 320)) +
+  scale_color_aaas() +
+  theme_pubr(base_size = 20, base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_all_plot.png"), plot = ioi_all_plot, dpi = 300, height = 5)
+img_ioi_all_plot <- image_read(here("paper/image", "ioi_all_plot.png"), density = 300)
+```
+
 ```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}IOIs (ms).
@@
 
-## IOIs
-
-We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented.
-The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking.
-We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
-
-There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* `r if(ioi_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (*Fig \ref{fig:ioi-2}*).
-
-```{r ioi-bl-plot, include = FALSE}
-ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean)) +
-  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
-  geom_hline(yintercept = 300, linetype = "dashed") + scale_y_continuous(breaks = seq(200, 380, 20), limits = c(200, 380)) +
-  geom_jitter(width = 0.15, alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
-  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
-
-# save plots
-ggsave(here("paper/image", "ioi_ca_plot.png"), plot = ioi_ca_plot, dpi = 300, height = 5)
-img_ioi_ca_plot <- image_read(here("paper/image", "ioi_ca_plot.png"), density = 300)
-```
-
 ## KOT
@@
 
+## IOIs
+
+We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented.
+The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking.
+We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
+
+There was no significant main effect of Category (*F*(`r ioi_ca_par_aov$anova_table[1,1]`, `r ioi_ca_par_aov$anova_table[1,2]`) = `r ioi_ca_par_aov$anova_table[1,4]`, *p* `r if(ioi_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (*Fig \ref{fig:ioi-2}*).
+
+```{r ioi-bl-plot, include = FALSE}
+ioi_ca_plot <- ggplot(ioi[Category == "none" | Category == "both" | Category == "baseline"], aes(x = factor(Category, level = c("baseline", "none", "both")), y = Mean)) +
+  geom_boxplot(outlier.shape = NA, position = position_dodge2(preserve = "single")) +
+  geom_hline(yintercept = 300, linetype = "dashed") + scale_y_continuous(breaks = seq(200, 380, 20), limits = c(200, 380)) +
+  geom_jitter(width = 0.15, alpha = 0.5) + labs(x = "Category", y = "IOI (ms)", title = "IOI") +
+  theme_pubr(base_size = 20, legend = "none", base_family = "Helvetica")
+
+# save plots
+ggsave(here("paper/image", "ioi_ca_plot.png"), plot = ioi_ca_plot, dpi = 300, height = 5)
+img_ioi_ca_plot <- image_read(here("paper/image", "ioi_ca_plot.png"), density = 300)
+```
 ```{r vel-diff-bl-plot, include = FALSE}

### 2022-12-05 paper/manuscript.tex
change the order of dependent variables / update intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Effective teaching involves experts and novices dynamically interacting with each other (Byrne \& Rapaport, 2011).
-Typically, information flows between experts and novices is mostly not unidirectional (i.e., from experts to novices) but rather bidirectional and interactive (Charbonneau \& Strachan, 2022).
-Experts adapt their behaviour to novices and scaffold the acquisition process of complex skills, and novices are actively participating in the learning process (Flynn, 2022; Okazaki, Muraoka, \& Osu, 2019).
+During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities until novices can perform the actions independently without the help of experts (Mermelshtine, 2017).
+Particularly, it is important for experts to find what novices have a problem from their failed attempt and to modify their behaviour (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts.
-When caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002).
-Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).
-For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own (McEllin et al., 2017).
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared with when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).
+Such exaggeration was observed not only towards infant learners but also when knowledgeable adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017).
+For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own.
 
-Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill levels of learners.
-Fukuyama et al.
-(2015) found that mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation skills.
+Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners (Fukuyama et al., 2015) In this study, mothers were asked to demonstrate demonstrate a cup-nesting task to their infants.
+There were four cups with different colour and size.
+The mothers showed how to stack and fit the nesting cups so that infants could reproduce the actions by themselves.
+The findings showed that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
 For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
@@
 
-We hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
+Evidence from parental scaffolding research suggests that caregivers scaffold infants' learning when the infants failed to produce an expected action (Wood \& Middleton, 1975; Zukow-Goldring \& Arbib, 2007).
+Therefore, we hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
 More precisely, expert pianists would exaggerate articulation (i.e., producing longer legato and shorter staccato) when articulation was not implemented in the recordings whereas expert pianists would exaggerate dynamics (i.e., producing louder forte and softer piano, and larger contrast between forte and piano) when dynamics was not implemented in the recordings.
@@
 As a result, less than 1 \% of total responses were removed.
-After removing pitch errors, we removed outliers for IOIs, KOT, KV and KV Difference, defined as values more than 3 standard deviations from the mean of each dependent variable.
+After removing pitch errors, we removed outliers for KOT, KV, KV Difference and IOIs, defined as values more than 3 standard deviations from the mean of each dependent variable.
 For each dependent variable, this resulted in less than 2 \% of overall responses being removed as outliers.
 
-First, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference).
+First, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present vs absent) for each dependent variable (i.e.., KOT, KV, KV Difference, IOIs).
 The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA.
@@
 All effects are reported as significant at \emph{p} \textless{} .05. For KOT, KV and KV Difference, we performed two-way ANOVAs separately for each subcomponent (i.e., legato, staccato, forte, piano).
-
-\hypertarget{iois}{%
-\subsection{IOIs}\label{iois}}
-
-Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
-This indicates that participants did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).
 
@@
 These results indicated that participants did not change their performances in terms of dynamics contrast depending on the type of recordings (\emph{Fig \ref{fig:vel-diff-1}}).
+
+\hypertarget{iois}{%
+\subsection{IOIs}\label{iois}}
+
+Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
+This indicates that participants did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).
 
@@
 
-Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place (Fukuyama et al., 2015; Okazaki et al., 2019).
+Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place (Fukuyama et al., 2015; Okazaki, Muraoka, \& Osu, 2019).
 In the current experiment, expert pianists had only access to their imaginary students and needed to perform in the absence of actual students.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-charbonneau_2022}{}}%
-Charbonneau, M., \& Strachan, J.
-W.
-A.
-(2022).
-From {Copying} to {Coordination}: {An Alternative Framework} for {Understanding Cultural Learning Mechanisms}.
-\emph{Journal of Cognition and Culture}, \emph{22}(5), 451--466.
-\url{https://doi.org/10.1163/15685373-12340145}
-
-\leavevmode\vadjust pre{\hypertarget{ref-flynn_2022}{}}%
-Flynn, E.
-(2022).
-{``{Imitation Is} the {Sincerest Form} of''} \ldots{} {Cultural Evolution}, or {Is It}?
-\emph{Journal of Cognition and Culture}, \emph{22}(5), 436--450.
-\url{https://doi.org/10.1163/15685373-12340144}
-
 \leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%
@@
 \url{https://doi.org/10.1037/xhp0000505}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mermelshtine_2017}{}}%
+Mermelshtine, R.
+(2017).
+Parent\textendash child learning interactions: {A} review of the literature on scaffolding.
+\emph{British Journal of Educational Psychology}, \emph{87}(2), 241--254.
+\url{https://doi.org/10.1111/bjep.12147}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-uther_2007}{}}%
-Uther, M., Knoll, M.
-A., \& Burnham, D.
-(2007).
-Do you speak {E-NG-L-I-SH}?
-{A} comparison of foreigner- and infant-directed speech.
-\emph{Speech Communication}, \emph{49}(1), 2--7.
-\url{https://doi.org/10.1016/j.specom.2006.10.003}
+\leavevmode\vadjust pre{\hypertarget{ref-ugur_2015}{}}%
+Ugur, E., Nagai, Y., Celikkanat, H., \& Oztop, E.
+(2015).
+Parental scaffolding as a bootstrapping mechanism for learning grasp affordances and imitation skills.
+\emph{Robotica}, \emph{33}(5), 1163--1180.
+\url{https://doi.org/10.1017/S0263574714002148}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wood_1975}{}}%
+Wood, D., \& Middleton, D. (1975). A {Study} of {Assisted Problem-Solving}. \emph{British Journal of Psychology}, \emph{66}(2), 181--191. \url{https://doi.org/10.1111/j.2044-8295.1975.tb01454.x}
 
@@
 By comparing with the baseline performance, we can investigate how pedagogical intentions influence participants' performances.
-
-\hypertarget{iois-1}{%
-\subsection{IOIs}\label{iois-1}}
-
-We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented.
-The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking.
-We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
-
-There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (\emph{Fig \ref{fig:ioi-2}}).
 
@@
 
+\hypertarget{iois-1}{%
+\subsection{IOIs}\label{iois-1}}
+
+We compared the baseline performance with the performances in response to recordings where either both articulation and dynamics were implemented or neither of them was implemented.
+The reason for choosing this comparison was that we wanted to test whether having the intention to teach would lead to slower performances, either regardless of the learner's skill or especially when skills appeared to be lacking.
+We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.
+
+There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse-Geisser corrected), suggesting that participants kept the same tempo regardless of whether they had the intention to teach (\emph{Fig \ref{fig:ioi-2}}).
+
 \begin{figure}

### 2022-12-05 paper/manuscript.Rmd
updated intro/discussion/supplementary material
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 abstract: |
-  Later
 
@@
 Effective teaching involves experts and novices dynamically interacting with each other [@byrne_2011].
-During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities until novices can perform the actions independently without the help of experts [@mermelshtine_2017].
-Particularly, it is important for experts to find what novices have a problem from their failed attempt and to modify their behaviour [@ugur_2015; @zukow-goldring_2007].
+First, experts demonstrate what to be acquired and then novices reproduce what the experts demonstrated.
+During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities [@mermelshtine_2017].
+This interactive learning process will be iterated until novices can perform independently without the help of experts.
+Experts scaffold novices' learning by finding the specific problem of each novice from their unsuccessful attempts and modifying their behaviour accordingly [@ugur_2015; @zukow-goldring_2007].
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts.
-When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared with when they were demonstrating to adults [@brand_2002].
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults [@brand_2002].
 Such exaggeration was observed not only towards infant learners but also when knowledgeable adults were interacting with adult learners [@mcellin_2017].
@@
 
-Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners [@fukuyama_2015] In this study, mothers were asked to demonstrate demonstrate a cup-nesting task to their infants.
-There were four cups with different colour and size.
+Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners [@fukuyama_2015].
+In this study, mothers were asked to demonstrate a cup-nesting task to their infants.
+There were four cups of different colours and sizes.
 The mothers showed how to stack and fit the nesting cups so that infants could reproduce the actions by themselves.
-The findings showed that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
+It was found that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
 For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
@@
 To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
-In our previous research (Tominaga et al., accepted), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
+In our previous research (Tominaga et al., in press), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
 In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics.
@@
 In that study, teachers were instructed to teach techniques to hypothetical learners and they did not hear novices' particular performances.
-This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
+This leaves open the question of whether pianists would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
 If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
 
-To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., accepted).
+To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
+The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
 All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
@@
 
-The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., under review).
+The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., in press).
 The piece was an excerpt from Clementi's Sonatina Op.36 (No.3) in C major.
@@
 We created artificial recordings of novices to manipulate the displayed proficiency with regard to the two relevant techniques.
-There were four types of recordings: 1) recordings where both articulation and dynamics were implemented, 2) recordings where only articulation was implemented whereas dynamics was missing, 3) recordings where only dynamics was implemented whereas articulation was missing, and 4) recordings where neither articulation nor dynamics was implemented.
+There were four types of recordings: 1) recordings where both articulation and dynamics were implemented (both), 2) recordings where only articulation was implemented whereas dynamics was missing (art only), 3) recordings where only dynamics was implemented whereas articulation was missing (dyn only), and 4) recordings where neither articulation nor dynamics was implemented (none).
 We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
@@
 stim
-
 ```
@@
 For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent (i.e., Legato, Staccato, Forte, Piano).
+The results are described in *Supplementary Material*.
 
@@
 
-```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}IOIs (ms).
-A dashed line represents the tempo given by a metronome.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
-img_ioi_1 <- image_scale(img_ioi_all_plot, "2200")
-image_write(img_ioi_1, path = here("paper/image", "Fig2.tiff"), format = "tiff")
-img_ioi_1
-```
-
 ```{r plot-kot-1, out.width = "100%", fig.cap = "\\label{fig:kot-1}KOT(ms) for each subcomponent; legato (left) and staccato (right).
@@
 
+```{r plot-ioi-1, out.width = "100%", fig.cap = "\\label{fig:ioi-1}IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+img_ioi_1 <- image_scale(img_ioi_all_plot, "2200")
+image_write(img_ioi_1, path = here("paper/image", "Fig2.tiff"), format = "tiff")
+img_ioi_1
+```
+
 \clearpage
@@
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
-This may imply that participants prioritised articulation over dynamics and did not modify two performance parameters at the same time.
+This may imply that participants prioritised the teaching of articulation over dynamics and did not modify two performance parameters at the same time.
 This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the sequential action into discrete sub-actions and adding complexity step by step [@zukow-goldring_2007].
-In our study, expert pianists might have focused on the teaching of articulation first.
 
-The reason why expert pianists seemed to prioritise the teaching of articulation might be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
-In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
+The reason why expert pianists seemed to prioritise the teaching of articulation might be that articulation was more important for the specific piece we selected or that the piece itself naturally invited some implementation of articulation.
+In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither articulation nor dynamics was notated on the sheet music.
 Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to highlight it.
@@
 
-We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
+We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill level of students.
 Also compared with the tempo of the baseline performance (see *Supplementary Material*), they did not change the tempo in the teaching context.
-These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
-Although slower demonstration is generally considered to be used for teaching [@brand_2002; @mcellin_2017; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
+These findings indicated that tempo was not employed specifically in teaching, likely because tempo itself was not relevant to either articulation or dynamics.
+Although slower demonstration is generally considered to be used for teaching [@brand_2002; @mcellin_2017; @schaik_2019], in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
 
@@
 As some participants reported that "(students had) very different interpretations of the melody" or "all kids played rhythmically correct" in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
-One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants (expert pianists) to listen to it first so that they can detect errors by comparing each recording with the model performance.
+One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants to listen to it first so that they can detect errors by comparing each recording with the model performance.
 
@@
 
-Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place [@fukuyama_2015; @okazaki_2019].
+Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place [@zukow-goldring_2007; @fukuyama_2015; @okazaki_2019].
 In the current experiment, expert pianists had only access to their imaginary students and needed to perform in the absence of actual students.
@@
 
+## 1. Artificial recordings
+
+Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al., in press).
+We used the data from Experiment 2.
+In Experiment 2, participants were asked to memorise one piece of music (*Fig 5* (A), Tominaga et al., in press) without any expressive notation before coming to the experiment.
+First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performance.
+After measuring the baseline performance, participants were asked to perform the piece with either articulation (*Fig 5* (B), Tominaga et al., in press) or dynamics (*Fig 5* (C), Tominaga et al., in press) in two conditions (teaching vs.
+performing).
+Here, we only used the data from the performing condition to make sure that the performances that we used did not have any particular modulations because of teaching intentions.
+Therefore, there were two types of performance data (articulation performance: performances played with notated articulation, dynamics performance: performances played with notated dynamics).
+We obtained 31 valid performances (i.g., performances without pitch errors) as baseline performances, 137 valid performances as articulation performances and 139 valid performances as dynamics performances.
+
+### 1. Procedure
+
+We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see *Apparatus and stimuli*) by using the performance data from Tominaga et al (in press).
+First, 16 instances for each type were generated.
+Second, for each type, we chose the four best instances, which exhibited the characteristic of each type well.
+
+### 2. Generation of 16 instances per each type
+
+In order to create recordings, we needed information about 1) the pitch order, 2) when each note was pressed (related to tempo), 3) when each note was released (related to articulation) and 4) how fast each note was hit (related to dynamics).
+Pitch information was already determined and the same for all structures because of the sheet music (*Fig \ref{fig:stim}*).
+
+First, we created 16 basic structures of recordings from the baseline performances.
+To determine the temporal feature of each structure (i.e., when each key was pressed), three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged.
+Due to the size of the dataset (i.e., 31 performances), some baseline performances were used twice.
+We made sure that the identical performance was not selected within each random selection (i.e., each selection consisted of three different performances).
+Thus, we created 16 basic structures for tempi.
+
+Next, we were going to follow the same procedure to determine the sound duration (i.e., when each key was released) of each note (related to articulation) and the key velocity profile (i.e., how fast each key was hit) of each note (related to dynamics).
+For the sound duration, however, we found that most participants implemented articulation (particularly legato) even in their baseline performances, where they should not implement any particular articulation or dynamics.
+Therefore, we decided to use the data from only two participants who did not show such tendencies.
+We obtained four performances from the two participants and used these four performances four times each without performing any averaging so as to have 16 instances in total.
+For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged.
+This is how we created 16 instances for three performance features (i.e., tempo, articulation and dynamics).
+We generated the 16 recordings where neither articulation nor dynamics was implemented (i.e., none) by combining the three performance features made from the baseline performances.
+
+In order to have other three types of recordings (i.e., both, art only, dyn only), we needed to create 16 instances for the sound duration when articulation was implemented and 16 instances for the key velocity profiles when dynamics was implemented.
+We followed the same procedure as we did to generate none recordings but we used articulation performances (not baseline performances) for the sound duration and dynamics performances (not baseline performances) for the key velocity profiles.
+To create art only recordings,
+
+### 3. Selection of four instances per type
+
+After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording and rate to what extent each expression was implemented between 0 (not at all) and 5 (fully).
+Based on these ratings, we selected four recordings for each type.
+
 ## 2. Comparison with baseline performance
+
 To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and to do their best as a performer.
@@
 There was a significant main effect of Category (*F*(`r pia_ca_par_aov$anova_table[1,1]`, `r pia_ca_par_aov$anova_table[1,2]`) = `r pia_ca_par_aov$anova_table[1,4]`, *p* `r if(pia_ca_par_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', pia_ca_par_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_ca_par_aov$anova_table[1,5])`; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05 , baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had the intention to teach (*Fig \ref{fig:vel-2}*, right).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: *p* = .05, baseline and articulation-only: *p* = .03), suggesting that participants produced softer piano when they had the intention to teach (*Fig \ref{fig:vel-2}*, right).
 
@@
 
-```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
-A dashed line represents the tempo given by a metronome.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
-img_ioi_2 <- image_scale(img_ioi_ca_plot, "2200")
-image_write(img_ioi_2, path = here("paper/image", "Fig6.tiff"), format = "tiff")
-img_ioi_2
-```
-
 ```{r plot-kot-2, out.width = "100%", fig.cap = "\\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
@@
 ```
+
+```{r plot-ioi-2, out.width = "100%", fig.cap = "\\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box."}
+img_ioi_2 <- image_scale(img_ioi_ca_plot, "2200")
+image_write(img_ioi_2, path = here("paper/image", "Fig6.tiff"), format = "tiff")
+img_ioi_2
+```

### 2022-12-05 paper/manuscript.tex
updated intro/discussion/supplementary material
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-\abstract{%
-Later
-}
-
 \begin{document}
@@
 Effective teaching involves experts and novices dynamically interacting with each other (Byrne \& Rapaport, 2011).
-During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities until novices can perform the actions independently without the help of experts (Mermelshtine, 2017).
-Particularly, it is important for experts to find what novices have a problem from their failed attempt and to modify their behaviour (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
+First, experts demonstrate what to be acquired and then novices reproduce what the experts demonstrated.
+During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities (Mermelshtine, 2017).
+This interactive learning process will be iterated until novices can perform independently without the help of experts.
+Experts scaffold novices' learning by finding the specific problem of each novice from their unsuccessful attempts and modify their behaviour accordingly (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts.
-When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared with when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).
 Such exaggeration was observed not only towards infant learners but also when knowledgeable adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017).
@@
 
-Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners (Fukuyama et al., 2015) In this study, mothers were asked to demonstrate demonstrate a cup-nesting task to their infants.
+Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners (Fukuyama et al., 2015) In this study, mothers were asked to demonstrate a cup-nesting task to their infants.
 There were four cups with different colour and size.
 The mothers showed how to stack and fit the nesting cups so that infants could reproduce the actions by themselves.
-The findings showed that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
+They found that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
 For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
@@
 
-To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., accepted).
+To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
 All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
@@
 For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent (i.e., Legato, Staccato, Forte, Piano).
+The results are described in \emph{Supplementary Material}.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms).
-A dashed line represents the tempo given by a metronome.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
-\end{figure}
-
-\begin{figure}
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-1-1} \caption{\label{fig:kot-1}KOT(ms) for each subcomponent; legato (left) and staccato (right).
@@
 Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-1}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
 \end{figure}
@@
 Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.
-This may imply that participants prioritised articulation over dynamics and did not modify two performance parameters at the same time.
+This may imply that participants prioritised the teaching of articulation over dynamics and did not modify two performance parameters at the same time.
 This is consistent with the findings in parental scaffolding where caregivers help infants acquire a complex action sequence by simplifying the sequential action into discrete sub-actions and adding complexity step by step (Zukow-Goldring \& Arbib, 2007).
-In our study, expert pianists might have focused on the teaching of articulation first.
 
@@
 As some participants reported that ``(students had) very different interpretations of the melody'' or ``all kids played rhythmically correct'' in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors.
-One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants (expert pianists) to listen to it first so that they can detect errors by comparing each recording with the model performance.
+One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants to listen to it first so that they can detect errors by comparing each recording with the model performance.
 
@@
 
-Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place (Fukuyama et al., 2015; Okazaki, Muraoka, \& Osu, 2019).
+Experiments which investigated dynamic interactions between experts and novices have been done with the physical presence of both an expert and a novice in the same place (Fukuyama et al., 2015; Okazaki, Muraoka, \& Osu, 2019; Zukow-Goldring \& Arbib, 2007).
 In the current experiment, expert pianists had only access to their imaginary students and needed to perform in the absence of actual students.
@@
 
+\hypertarget{artificial-recordings}{%
+\subsection{1. Artificial recordings}\label{artificial-recordings}}
+
+Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al.~, in press).
+We used the data from Experiment 2.
+In Experiment 2, participants were asked to memorise one piece of music (\emph{Fig 5} (A), Tominaga et al., in press) without any expressive notation before coming to the experiment.
+First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performance.
+After measuring the baseline performance, participants were asked to perform the piece with either articulation (\emph{Fig 5} (B), Tominaga et al., in press) or dynamics (\emph{Fig 5} (C), Tominaga et al., in press) in two conditions (teaching vs.~performing).
+Here, we only used the data from the performing condition to make sure that performances that we used did not have any particular modulations because of teaching intentions.
+Therefore, there were two types of performance data (articulation performance: performances played with notated articulation, dynamics performance: performances played with notated dynamics).
+We obtained 31 valid performances (i.g., performances without pitch errors) as baseline performances, 137 valid performances as articulation performances and 139 valid performances as dynamics performances.
+
+\hypertarget{procedure-1}{%
+\subsubsection{1. Procedure}\label{procedure-1}}
+
+We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga.
+First, 16 instances for each type were generated.
+Second, for each type, we chose the four best instances, which exhibited the characteristic of each type well.
+
+\hypertarget{generation-of-16-instances-per-each-type}{%
+\subsubsection{2. Generation of 16 instances per each type}\label{generation-of-16-instances-per-each-type}}
+
+In order to create recordings, we need information about 1) the pitch order, 2) when each note was pressed (related to tempo), 3) when each note was released (related to articulation) and 4) how fast each note was hit.
+Pitch information was already determined and the same for all structures because of the sheet music (\emph{Fig \ref{fig:stim}}).
+
+First, we created 16 basic structures of recordings from the baseline performances.
+To determine the temporal information of each structure (i.e., when each key was pressed), three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged.
+Due to the size of the dataset (i.e., 31 performances), some baseline performances were used twice.
+We made sure that the identical performance was not selected within each random selection (i.e., each selection consisted of three different performances).
+Thus, we created 16 basic structures for tempi.
+
+Next, we were going to follow the same procedure to determine the sound duration (i.e., when each key was released) of each note (related to articulation) and the key velocity profile (i.e., how fast each key was hit) of each note (related to dynamics).
+For the sound duration, however, we found that most participants implemented articulation (particularly legato) even in their baseline performances, where they should not implement any particular articulation or dynamics.
+Therefore, we decided to use the data from only two participants who did not show such tendencies.
+We obtained four performances from the two participants and used these four performances four times each without performing any averaging so as to have 16 instances in total.
+For the key velocity profiles, we followed the same procedure as determining tempi and three performances out of the 31 valid baseline performances were randomly selected and key velocity profiles of the three performances were averaged.
+This is how we created 16 instances for three performance features (i.e., tempo, articulation and dynamics).
+We generated the 16 recordings where neither articulation nor dynamics was implemented (i.e., none) by combining the three performance features.
+
+In order to have other three types of recordings (i.e., both, art only, dyn only), we needed to create 16 instances for the sound duration when articulation was implemented and 16 instances for the key velocity profiles when dynamics was implemented.
+We followed the same procedure as we did to generate none recordings but we used articulation performances (not baseline performances) for the sound duration and dynamics performances (not baseline performances) for the key velocity profiles.
+To create art only recordings,
+
+\hypertarget{selection-of-four-instances-per-type}{%
+\subsubsection{3. Selection of four instances per type}\label{selection-of-four-instances-per-type}}
+
+After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording and rate to what extent each expression was implemented between 0 (not at all) and 5 (fully).
+Based on these ratings, we selected four recordings for each type.
+
 \hypertarget{comparison-with-baseline-performance}{%
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
-A dashed line represents the tempo given by a metronome.
-Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
-\end{figure}
-
-\begin{figure}
 \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comparison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right).
@@
 \end{figure}
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comparison with the baseline performance in terms of IOIs (ms).
+A dashed line represents the tempo given by a metronome.
+Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
+\end{figure}
 

### 2023-01-15 paper/manuscript.Rmd
update intro/results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Effective teaching involves experts and novices dynamically interacting with each other [@byrne_2011].
-First, experts demonstrate what to be acquired and then novices reproduce what the experts demonstrated.
+First, experts demonstrate what novices are meant to acquired and then novices attempt to reproduce what the experts demonstrated.
 During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities [@mermelshtine_2017].
 This interactive learning process will be iterated until novices can perform independently without the help of experts.
-Experts scaffold novices' learning by finding the specific problem of each novice from their unsuccessful attempts and modifying their behaviour accordingly [@ugur_2015; @zukow-goldring_2007].
+Experts scaffold novices' learning by identifying novices' specific problems of each novice from their unsuccessful attempts and modifying their behaviour accordingly [@ugur_2015; @zukow-goldring_2007].
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts.
-When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults [@brand_2002].
-Such exaggeration was observed not only towards infant learners but also when knowledgeable adults were interacting with adult learners [@mcellin_2017].
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults [@brand_2002].Such pedagogical action modulations are sensitive to the skill level of learners.
+For instance, evidence from parental scaffolding research suggests that caregivers scaffold infants’ learning when the infants failed to produce an expected action [@wood_1975; @zukow-goldring_2007].
+@fukuyama_2015 performed a study, where mothers were asked to demonstrate a cup-nesting task to their infants.
+There were four cups of different colours and sizes.
+Mothers demonstrated how to stack and fit the nesting cups with the intention to teach their infants to reproduce the actions by themselves.
+It was found that the mothers dynamically modulated their demonstrations according to the infants' object manipulation skills.
+Specifically, mothers exaggerated their movements when their infants failed to produce expected actions but did not produce such exaggerations when their infants lacked the motor skills needed to produce the required actions.
+Movement exaggerations for the purpose of teaching are observed not only towards infant learners but also when knowledgeable adults are interacting with adult learners [@mcellin_2017].
 For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own.
+However, it has not been established whether and how experts teaching skills to adult learners adjust their demonstrations as a function of the learners’ demonstrated skills.
 
-Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners [@fukuyama_2015].
-In this study, mothers were asked to demonstrate a cup-nesting task to their infants.
-There were four cups of different colours and sizes.
-The mothers showed how to stack and fit the nesting cups so that infants could reproduce the actions by themselves.
-It was found that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
-For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
-To the best of our knowledge, it is unknown whether people teaching skills to adult learners also adjust their demonstrations as a function of the learners' demonstrated skills.
-
-To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
-In our previous research (Tominaga et al., in press), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
+To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
+In our previous research [@tominaga_2022], we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
 In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics.
@@
 
-To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
+To gain experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
 The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
@@
 
-Evidence from parental scaffolding research suggests that caregivers scaffold infants' learning when the infants failed to produce an expected action [@wood_1975; @zukow-goldring_2007].
-Therefore, we hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
+We hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
 More precisely, expert pianists would exaggerate articulation (i.e., producing longer legato and shorter staccato) when articulation was not implemented in the recordings whereas expert pianists would exaggerate dynamics (i.e., producing  louder forte and softer piano, and larger contrast between forte and piano) when dynamics was not implemented in the recordings.
@@
 
-As predicted, there was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
+As predicted, experts produced longer legato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, left).
+Accordingly, here was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
 There was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* `r if(leg_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* `r if(leg_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
-Participants produced longer legato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, left).
 
@@
 
-In line with our prediction, there was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* `r if(sta_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
+In line with our prediction, experts produced shorter staccato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, right).
+Accordingly, there was a significant main effect of Articulation (*F*(`r sta_all_aov$anova_table[1,1]`, `r sta_all_aov$anova_table[1,2]`) = `r sta_all_aov$anova_table[1,4]`, *p* `r if(sta_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[1,5])`).
 There was no significant main effect of Dynamics (*F*(`r sta_all_aov$anova_table[2,1]`, `r sta_all_aov$anova_table[2,2]`) = `r sta_all_aov$anova_table[2,4]`, *p* `r if(sta_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', sta_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r sta_all_aov$anova_table[3,1]`, `r sta_all_aov$anova_table[3,2]`) = `r sta_all_aov$anova_table[3,4]`, *p* `r if(sta_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', sta_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', sta_all_aov$anova_table[3,5])`).
-Participants produced shorter staccato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, right).
 
@@
 
+Unlike our prediction, experts did not change their performances in terms of notated forte depending on the type of recordings (*Fig \ref{fig:vel-1}*, left).
 Neither the main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
-Participants did not change their performances in terms of notated forte depending on the type of recordings (*Fig \ref{fig:vel-1}*, left).
 
@@
 
+Experts produced softer piano only when listening to the recordings where articulation was implemented (*Fig \ref{fig:vel-1}*, right).
 There was a significant main effect of Articulation (*F*(`r pia_all_aov$anova_table[1,1]`, `r pia_all_aov$anova_table[1,2]`) = `r pia_all_aov$anova_table[1,4]`, *p* `r if(pia_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[1,5])`).
 However, there was no significant main effect of Dynamics (*F*(`r pia_all_aov$anova_table[2,1]`, `r pia_all_aov$anova_table[2,2]`) = `r pia_all_aov$anova_table[2,4]`, *p* `r if(pia_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r pia_all_aov$anova_table[3,1]`, `r pia_all_aov$anova_table[3,2]`) = `r pia_all_aov$anova_table[3,4]`, *p* `r if(pia_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', pia_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', pia_all_aov$anova_table[3,5])`).
-Participants produced softer piano when listening to the recordings where articulation was implemented (*Fig \ref{fig:vel-1}*, right).
 
@@
 ## KV Difference
+
+Unlike our prediction, experts did not change their performances in terms of dynamics contrast depending on the type of recordings (*Fig \ref{fig:vel-diff-1}*).
+
 ### Forte to Piano
@@
 Neither the main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* `r if(ptof_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ptof_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* `r if(ptof_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* `r if(ptof_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
-
-These results indicated that participants did not change their performances in terms of dynamics contrast depending on the type of recordings (*Fig \ref{fig:vel-diff-1}*).
 
@@
 
+Unlike our prediction, experts did not change the tempo depending on the type of recordings (*Fig \ref{fig:ioi-1}*).
 Neither the main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
-This indicates that participants did not change the tempo depending on the type of recordings (*Fig \ref{fig:ioi-1}*).
 

### 2023-01-15 paper/manuscript.tex
update intro/results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Effective teaching involves experts and novices dynamically interacting with each other (Byrne \& Rapaport, 2011).
-First, experts demonstrate what to be acquired and then novices reproduce what the experts demonstrated.
+First, experts demonstrate what novices are meant to acquired and then novices attempt to reproduce what the experts demonstrated.
 During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities (Mermelshtine, 2017).
 This interactive learning process will be iterated until novices can perform independently without the help of experts.
-Experts scaffold novices' learning by finding the specific problem of each novice from their unsuccessful attempts and modify their behaviour accordingly (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
+Experts scaffold novices' learning by identifying novices' specific problems of each novice from their unsuccessful attempts and modifying their behaviour accordingly (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts.
-When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).
-Such exaggeration was observed not only towards infant learners but also when knowledgeable adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017).
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).Such pedagogical action modulations are sensitive to the skill level of learners.
+For instance, evidence from parental scaffolding research suggests that caregivers scaffold infants' learning when the infants failed to produce an expected action (Wood \& Middleton, 1975; Zukow-Goldring \& Arbib, 2007).
+Fukuyama et al.
+(2015) performed a study, where mothers were asked to demonstrate a cup-nesting task to their infants.
+There were four cups of different colours and sizes.
+Mothers demonstrated how to stack and fit the nesting cups with the intention to teach their infants to reproduce the actions by themselves.
+It was found that the mothers dynamically modulated their demonstrations according to the infants' object manipulation skills.
+Specifically, mothers exaggerated their movements when their infants failed to produce expected actions but did not produce such exaggerations when their infants lacked the motor skills needed to produce the required actions.
+Movement exaggerations for the purpose of teaching are observed not only towards infant learners but also when knowledgeable adults are interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017).
 For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own.
+However, it has not been established whether and how experts teaching skills to adult learners adjust their demonstrations as a function of the learners' demonstrated skills.
 
-Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill level of learners (Fukuyama et al., 2015) In this study, mothers were asked to demonstrate a cup-nesting task to their infants.
-There were four cups with different colour and size.
-The mothers showed how to stack and fit the nesting cups so that infants could reproduce the actions by themselves.
-They found that the mothers dynamically modulated their demonstration according to the infants' object manipulation skills.
-For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions.
-To the best of our knowledge, it is unknown whether people teaching skills to adult learners also adjust their demonstrations as a function of the learners' demonstrated skills.
-
-To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
-In our previous research (Tominaga et al., in press), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
+To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
+In our previous research (Tominaga, Knoblich, \& Sebanz, 2022), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
 In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics.
@@
 In that study, teachers were instructed to teach techniques to hypothetical learners and they did not hear novices' particular performances.
-This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
+This leaves open the question of whether pianists would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
 If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
 
-To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
+To gain experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
+The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
 All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
@@
 
-Evidence from parental scaffolding research suggests that caregivers scaffold infants' learning when the infants failed to produce an expected action (Wood \& Middleton, 1975; Zukow-Goldring \& Arbib, 2007).
-Therefore, we hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
+We hypothesised that expert pianists would exaggerate their performance only when specific techniques (i.e., either articulation or dynamics, or both) were missing in the recordings.
 More precisely, expert pianists would exaggerate articulation (i.e., producing longer legato and shorter staccato) when articulation was not implemented in the recordings whereas expert pianists would exaggerate dynamics (i.e., producing louder forte and softer piano, and larger contrast between forte and piano) when dynamics was not implemented in the recordings.
@@
 
-As predicted, there was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
+As predicted, experts produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).
+Accordingly, here was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
 There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002).
-Participants produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).
 
@@
 
-In line with our prediction, there was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009).
+In line with our prediction, experts produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).
+Accordingly, there was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.040, \(\eta_G^2\) = 0.009).
 There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.093, \(\eta_G^2\) = 0.001).
-Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).
 
@@
 
+Unlike our prediction, experts did not change their performances in terms of notated forte depending on the type of recordings (\emph{Fig \ref{fig:vel-1}}, left).
 Neither the main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) nor the main effect of Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
-Participants did not change their performances in terms of notated forte depending on the type of recordings (\emph{Fig \ref{fig:vel-1}}, left).
 
@@
 
+Experts produced softer piano only when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).
 There was a significant main effect of Articulation (\emph{F}(1, 19) = 7.18, \emph{p} = 0.01, \(\eta_G^2\) = 0.060).
 However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000).
-Participants produced softer piano when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).
 
@@
 \subsection{KV Difference}\label{kv-difference}}
+
+Unlike our prediction, experts did not change their performances in terms of dynamics contrast depending on the type of recordings (\emph{Fig \ref{fig:vel-diff-1}}).
 
@@
 
-These results indicated that participants did not change their performances in terms of dynamics contrast depending on the type of recordings (\emph{Fig \ref{fig:vel-diff-1}}).
-
 \hypertarget{iois}{%
@@
 
+Unlike our prediction, experts did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).
 Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
-This indicates that participants did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).
 
@@
 
-The reason why expert pianists seemed to prioritise the teaching of articulation might be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation.
-In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music.
+The reason why expert pianists seemed to prioritise the teaching of articulation might be that articulation was more important for the specific piece we selected or that the piece itself naturally invited some implementation of articulation.
+In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither articulation nor dynamics was notated on the sheet music.
 Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to highlight it.
@@
 
-We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students.
+We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill level of students.
 Also compared with the tempo of the baseline performance (see \emph{Supplementary Material}), they did not change the tempo in the teaching context.
-These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics.
-Although slower demonstration is generally considered to be used for teaching (Brand et al., 2002; McEllin et al., 2017; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
+These findings indicated that tempo was not employed specifically in teaching, likely because tempo itself was not relevant to either articulation or dynamics.
+Although slower demonstration is generally considered to be used for teaching (Brand et al., 2002; McEllin et al., 2017; Schaik, Meyer, Ham, \& Hunnius, 2019), in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-tominaga_2022}{}}%
+Tominaga, A., Knoblich, G., \& Sebanz, N.
+(2022).
+Expert pianists make specific exaggerations for teaching.
+\emph{Scientific Reports}, \emph{12}(1), 21296.
+\url{https://doi.org/10.1038/s41598-022-25711-3}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ugur_2015}{}}%
@@
 
-Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al.~, in press).
+Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al., in press).
 We used the data from Experiment 2.
@@
 After measuring the baseline performance, participants were asked to perform the piece with either articulation (\emph{Fig 5} (B), Tominaga et al., in press) or dynamics (\emph{Fig 5} (C), Tominaga et al., in press) in two conditions (teaching vs.~performing).
-Here, we only used the data from the performing condition to make sure that performances that we used did not have any particular modulations because of teaching intentions.
+Here, we only used the data from the performing condition to make sure that the performances that we used did not have any particular modulations because of teaching intentions.
 Therefore, there were two types of performance data (articulation performance: performances played with notated articulation, dynamics performance: performances played with notated dynamics).
@@
 
-We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga.
+We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga et al (in press).
 First, 16 instances for each type were generated.
@@
 
-In order to create recordings, we need information about 1) the pitch order, 2) when each note was pressed (related to tempo), 3) when each note was released (related to articulation) and 4) how fast each note was hit.
+In order to create recordings, we needed information about 1) the pitch order, 2) when each note was pressed (related to tempo), 3) when each note was released (related to articulation) and 4) how fast each note was hit (related to dynamics).
 Pitch information was already determined and the same for all structures because of the sheet music (\emph{Fig \ref{fig:stim}}).
@@
 First, we created 16 basic structures of recordings from the baseline performances.
-To determine the temporal information of each structure (i.e., when each key was pressed), three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged.
+To determine the temporal feature of each structure (i.e., when each key was pressed), three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged.
 Due to the size of the dataset (i.e., 31 performances), some baseline performances were used twice.
@@
 We obtained four performances from the two participants and used these four performances four times each without performing any averaging so as to have 16 instances in total.
-For the key velocity profiles, we followed the same procedure as determining tempi and three performances out of the 31 valid baseline performances were randomly selected and key velocity profiles of the three performances were averaged.
+For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged.
 This is how we created 16 instances for three performance features (i.e., tempo, articulation and dynamics).
-We generated the 16 recordings where neither articulation nor dynamics was implemented (i.e., none) by combining the three performance features.
+We generated the 16 recordings where neither articulation nor dynamics was implemented (i.e., none) by combining the three performance features made from the baseline performances.
 
@@
 There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.009, \(\eta_G^2\) = 0.072; Greenhouse-Geisser corrected).
-Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had the intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
+Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05, baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had the intention to teach (\emph{Fig \ref{fig:vel-2}}, right).
 
@@
 \end{figure}
+
 \begin{figure}

### 2023-01-16 paper/manuscript.Rmd
update figures / reflected feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 if (!require("ggsci")) {install.packages("ggsci"): require("ggsci")}
+if (!require("ggsignif")) {install.packages("ggsignif"): require("ggsignif")}
 ```
@@
 Effective teaching involves experts and novices dynamically interacting with each other [@byrne_2011].
-First, experts demonstrate what novices are meant to acquired and then novices attempt to reproduce what the experts demonstrated.
+First, experts demonstrate what novices are meant to acquire and then novices attempt to reproduce what the experts demonstrated.
 During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities [@mermelshtine_2017].
 This interactive learning process will be iterated until novices can perform independently without the help of experts.
-Experts scaffold novices' learning by identifying novices' specific problems of each novice from their unsuccessful attempts and modifying their behaviour accordingly [@ugur_2015; @zukow-goldring_2007].
+Experts scaffold novices' learning by identifying novices' specific problems from their unsuccessful attempts and modifying their behaviour accordingly [@ugur_2015; @zukow-goldring_2007].
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts.
-When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults [@brand_2002].Such pedagogical action modulations are sensitive to the skill level of learners.
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and to produce wide and expansive movements compared to when they were demonstrating to adults [@brand_2002].
+Such pedagogical action modulations are sensitive to the skill level of learners.
 For instance, evidence from parental scaffolding research suggests that caregivers scaffold infants’ learning when the infants failed to produce an expected action [@wood_1975; @zukow-goldring_2007].
@@
 Mothers demonstrated how to stack and fit the nesting cups with the intention to teach their infants to reproduce the actions by themselves.
-It was found that the mothers dynamically modulated their demonstrations according to the infants' object manipulation skills.
-Specifically, mothers exaggerated their movements when their infants failed to produce expected actions but did not produce such exaggerations when their infants lacked the motor skills needed to produce the required actions.
+It was found that mothers dynamically modulated their demonstrations according to the infants' object manipulation skills.
+Specifically, mothers exaggerated their movements when their infants failed to produce expected actions but did not produce such exaggerations when their infants lacked the motor skills to produce the required actions.
 Movement exaggerations for the purpose of teaching are observed not only towards infant learners but also when knowledgeable adults are interacting with adult learners [@mcellin_2017].
@@
 
-To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
+To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on learners' skill level.
 In our previous research [@tominaga_2022], we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
@@
 
-To gain experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
-The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
+To obtain experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
+The recordings were made based on the performance data from our previous study [@tominaga_2022].
 All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
@@
 
-The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., in press).
+The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study [@tominaga_2022].
 The piece was an excerpt from Clementi's Sonatina Op.36 (No.3) in C major.
@@
 As predicted, experts produced longer legato when listening to the recordings where articulation was not implemented (*Fig \ref{fig:kot-1}*, left).
-Accordingly, here was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
+Accordingly, there was a significant main effect of Articulation (*F*(`r leg_all_aov$anova_table[1,1]`, `r leg_all_aov$anova_table[1,2]`) = `r leg_all_aov$anova_table[1,4]`, *p* `r if(leg_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', leg_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[1,5])`).
 There was no significant main effect of Dynamics (*F*(`r leg_all_aov$anova_table[2,1]`, `r leg_all_aov$anova_table[2,2]`) = `r leg_all_aov$anova_table[2,4]`, *p* `r if(leg_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[2,5])`) and no significant interaction between Articulation and Dynamics (*F*(`r leg_all_aov$anova_table[3,1]`, `r leg_all_aov$anova_table[3,2]`) = `r leg_all_aov$anova_table[3,4]`, *p* `r if(leg_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', leg_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', leg_all_aov$anova_table[3,5])`).
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_signif(
+    comparisons = list(c("Present", "Absent")),
+    y_position = 120, tip_length = 0, textsize = 10, color = "black",
+    annotation = c("*"),
+  ) +
   labs(y = "KOT (ms)", title = "KOT - Legato") + scale_y_continuous(breaks = seq(-20, 140, 40), limits = c(-20, 140)) +
@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
+  geom_signif(
+    comparisons = list(c("Present", "Absent")),
+    y_position = -140, tip_length = 0, textsize = 10, color = "black",
+    annotation = c("*"),
+  ) +
   labs(y = "KOT (ms)", title = "KOT - Staccato") + scale_y_continuous(breaks = seq(-280, -120, 40), limits = c(-280, -120)) +
@@
 Unlike our prediction, experts did not change their performances in terms of notated forte depending on the type of recordings (*Fig \ref{fig:vel-1}*, left).
-Neither the main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
+Neither the main effect of Articulation (*F*(`r for_all_aov$anova_table[1,1]`, `r for_all_aov$anova_table[1,2]`) = `r for_all_aov$anova_table[1,4]`, *p* `r if(for_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.2f', for_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r for_all_aov$anova_table[2,1]`, `r for_all_aov$anova_table[2,2]`) = `r for_all_aov$anova_table[2,4]`, *p* `r if(for_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r for_all_aov$anova_table[3,1]`, `r for_all_aov$anova_table[3,2]`) = `r for_all_aov$anova_table[3,4]`, *p* `r if(for_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', for_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', for_all_aov$anova_table[3,5])`).
 
@@
 
-The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novices whom they are trying to teach.
+The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novices whom they are intending to teach.
 We created artificial recordings to manipulate the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music.
@@
 
-Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al., in press).
+Artificial recordings were generated from performance data that we obtained in our previous study [@tominaga_2022].
 We used the data from Experiment 2.
-In Experiment 2, participants were asked to memorise one piece of music (*Fig 5* (A), Tominaga et al., in press) without any expressive notation before coming to the experiment.
+In Experiment 2, participants were asked to memorise one piece of music (*Fig 5* (A), @tominaga_2022) without any expressive notation before coming to the experiment.
 First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performance.
-After measuring the baseline performance, participants were asked to perform the piece with either articulation (*Fig 5* (B), Tominaga et al., in press) or dynamics (*Fig 5* (C), Tominaga et al., in press) in two conditions (teaching vs.
+After measuring the baseline performance, participants were asked to perform the piece with either articulation (*Fig 5* (B), @tominaga_2022) or dynamics (*Fig 5* (C), @tominaga_2022) in two conditions (teaching vs.
 performing).
@@
 
-We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see *Apparatus and stimuli*) by using the performance data from Tominaga et al (in press).
+We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see *Apparatus and stimuli*) by using the performance data from @tominaga_2022.
 First, 16 instances for each type were generated.

### 2023-01-16 paper/manuscript.tex
update figures / reflected feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 Effective teaching involves experts and novices dynamically interacting with each other (Byrne \& Rapaport, 2011).
-First, experts demonstrate what novices are meant to acquired and then novices attempt to reproduce what the experts demonstrated.
+First, experts demonstrate what novices are meant to acquire and then novices attempt to reproduce what the experts demonstrated.
 During a skill acquisition process, experts regularly monitor novices' actions and intervene by adapting their behaviour to novices' observed abilities (Mermelshtine, 2017).
 This interactive learning process will be iterated until novices can perform independently without the help of experts.
-Experts scaffold novices' learning by identifying novices' specific problems of each novice from their unsuccessful attempts and modifying their behaviour accordingly (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
+Experts scaffold novices' learning by identifying novices' specific problems from their unsuccessful attempts and modifying their behaviour accordingly (Ugur, Nagai, Celikkanat, \& Oztop, 2015; Zukow-Goldring \& Arbib, 2007).
 
 Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts.
-When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and produce wide and expansive movements compared to when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).Such pedagogical action modulations are sensitive to the skill level of learners.
+When mothers were teaching how to use a novel toy to their infants, they tended to demonstrate slowly and to produce wide and expansive movements compared to when they were demonstrating to adults (Brand, Baldwin, \& Ashburn, 2002).
+Such pedagogical action modulations are sensitive to the skill level of learners.
 For instance, evidence from parental scaffolding research suggests that caregivers scaffold infants' learning when the infants failed to produce an expected action (Wood \& Middleton, 1975; Zukow-Goldring \& Arbib, 2007).
@@
 Mothers demonstrated how to stack and fit the nesting cups with the intention to teach their infants to reproduce the actions by themselves.
-It was found that the mothers dynamically modulated their demonstrations according to the infants' object manipulation skills.
-Specifically, mothers exaggerated their movements when their infants failed to produce expected actions but did not produce such exaggerations when their infants lacked the motor skills needed to produce the required actions.
+It was found that mothers dynamically modulated their demonstrations according to the infants' object manipulation skills.
+Specifically, mothers exaggerated their movements when their infants failed to produce expected actions but did not produce such exaggerations when their infants lacked the motor skills to produce the required actions.
 Movement exaggerations for the purpose of teaching are observed not only towards infant learners but also when knowledgeable adults are interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017).
@@
 
-To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners.
+To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on learners' skill level.
 In our previous research (Tominaga, Knoblich, \& Sebanz, 2022), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
@@
 
-To gain experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
-The recordings were made based on the performance data from our previous study (Tominaga et al., in press).
+To obtain experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists.
+The recordings were made based on the performance data from our previous study (Tominaga et al., 2022).
 All the recordings were of the same piece but the recordings differed in the implementation of articulation and dynamics.
@@
 
-The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., in press).
+The recorded performances serving as stimuli were all of one piece of music that had already been used in our previous study (Tominaga et al., 2022).
 The piece was an excerpt from Clementi's Sonatina Op.36 (No.3) in C major.
@@
 As predicted, experts produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).
-Accordingly, here was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
+Accordingly, there was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.029, \(\eta_G^2\) = 0.002).
 There was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) and no significant interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002).
@@
 Unlike our prediction, experts did not change their performances in terms of notated forte depending on the type of recordings (\emph{Fig \ref{fig:vel-1}}, left).
-Neither the main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.213, \(\eta_G^2\) = 0.00) nor the main effect of Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.21, \(\eta_G^2\) = 0.00) nor the main effect of Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000).
 
@@
 
-Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al., in press).
+Artificial recordings were generated from performance data that we obtained in our previous study (Tominaga et al., 2022).
 We used the data from Experiment 2.
-In Experiment 2, participants were asked to memorise one piece of music (\emph{Fig 5} (A), Tominaga et al., in press) without any expressive notation before coming to the experiment.
+In Experiment 2, participants were asked to memorise one piece of music (\emph{Fig 5} (A), Tominaga et al.
+(2022)) without any expressive notation before coming to the experiment.
 First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performance.
-After measuring the baseline performance, participants were asked to perform the piece with either articulation (\emph{Fig 5} (B), Tominaga et al., in press) or dynamics (\emph{Fig 5} (C), Tominaga et al., in press) in two conditions (teaching vs.~performing).
+After measuring the baseline performance, participants were asked to perform the piece with either articulation (\emph{Fig 5} (B), Tominaga et al.
+(2022)) or dynamics (\emph{Fig 5} (C), Tominaga et al.
+(2022)) in two conditions (teaching vs.~performing).
 Here, we only used the data from the performing condition to make sure that the performances that we used did not have any particular modulations because of teaching intentions.
@@
 
-We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga et al (in press).
+We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga et al.
+(2022).
 First, 16 instances for each type were generated.

### 2023-01-16 paper/manuscript.Rmd
reflected comments on intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on learners' skill level.
+To address this question, we investigated whether and how experts can exaggerate specific parameters of the action to be acquired depending on learners' skill level.
+We chose to study expressive music performance, in which experts need to modulate multiple performance parameters such as articulation and dynamics to implement expression [@cancino-chacon_2018].
+Musicians can create their unique expression and interpretation of music even if they play exactly the same piece because when and how they change multiple performance parameters are different.
+By studying expressive performance, we can investigate how precisely experts can highlight specific parameters in a multidimentional space of expression (e.g., tempo, articulation dynamics) depending on the goal of teaching (e.g., to teach articulation or dynamics) even when they are performing the same piece.
+
 In our previous research [@tominaga_2022], we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
@@
 When they had the intention to demonstrate articulation, they exaggerated articulation by producing shorter staccato.
-In that study, teachers were instructed to teach techniques to hypothetical learners and they did not hear novices' particular performances.
-This leaves open the question of whether pianists would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
-If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
+In that study, experts were instructed to teach techniques to hypothetical learners and they did not hear novices' particular performances.
+This leaves open the question of whether experts would make specific adjustments depending on learners' skills in the way mothers have been shown to adapt demonstrations of actions to their infants' skills.
+If experts modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.
+For example, if learners did not implement notated articulation during their performance, experts should exaggerate the performance relevant to articulation.
+However, if learners implemented notated articulation correctly, experts should not exaggerate the performance relevant to articulation.
 

### 2023-01-16 paper/manuscript.tex
reflected comments on intro
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-To address this question, we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on learners' skill level.
+To address this question, we investigated whether and how experts can exaggerate specific parameters of the action to be acquired depending on learners' skill level.
+We chose to study expressive music performance, in which experts need to modulate multiple performance parameters such as articulation and dynamics to implement expression (Cancino-Chacón, Grachten, Goebl, \& Widmer, 2018).
+Musicians can create their unique expression and interpretation of music even if they play exactly the same piece because when and how they change multiple performance parameters are different.
+By studying expressive performance, we can investigate how precisely experts can highlight specific parameters in a multidimentional space of expression (e.g., tempo, articulation dynamics) depending on the goal of teaching (e.g., to teach articulation or dynamics) even when they are performing the same piece.
+
 In our previous research (Tominaga, Knoblich, \& Sebanz, 2022), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cancino-chacon_2018}{}}%
+Cancino-Chacón, C.
+E., Grachten, M., Goebl, W., \& Widmer, G.
+(2018).
+Computational {Models} of {Expressive Music Performance}: {A Comprehensive} and {Critical Review}.
+\emph{Frontiers in Digital Humanities}, \emph{5}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-fukuyama_2015}{}}%

### 2023-01-17 paper/manuscript.Rmd
updated discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 We generated four types of recordings.
-In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation was implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
+In one quarter of the recordings, both techniques were implemented; in one quarter, neither articulation nor dynamics was implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented.
 This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.
@@
 
+There are possibilities that it may be difficult in general for experts to flexibly adjust to different action parameters while teaching or they might have chosen to focus on the most salient parameters or the ones that they find most relevant.
+For example, when teaching dance, it is considered that good teachers need to identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation [@mainwaring_2010].
+For example, when teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha [@lobel_2021].
+Karl Leimer, a German music teacher and pianist, wrote in his booked that when doing finger exercises, students should perform with legato at first [@gieseking_1972].
+Legato creates round and mellow tones, which is one of the most essential elements of piano touches in performance.
+In both dance and music examples, the ultimate goal of skill acquisition is that students are able to use multiple techniques flexibly while performing.
+However, it seems that students need to build complex skills based on basic ones during the learning process.
+Therefore, identifying and focusing on the most relevant problem for each student may be more effective to facilitate learning.
+
 One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind.

### 2023-01-17 paper/manuscript.tex
updated discussion
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
+There are possibilities that it may be difficult in general for experts to flexibly adjust to different action parameters while teaching or they might have chosen to focus on the most salient parameters or the ones that they find most relevant.
+For example, when teaching dance, it is considered that good teachers need to identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation (Mainwaring \& Krasnow, 2010).
+For example, when teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha (Lobel, 2021).
+Karl Leimer, a German music teacher and pianist, wrote in his booked that when doing finger exercises, students should perform with legato at first (Gieseking \& Leimer, 1972).
+Legato creates round and mellow tones, which is one of the most essential elements of piano touches in performance.
+In both dance and music examples, the ultimate goal of skill acquisition is that students are able to use multiple techniques flexibly while performing.
+However, it seems that students need to build complex skills based on basic ones during the learning process.
+Therefore, identifying and focusing on the most relevant problem for each student may be more effective to facilitate learning.
+
 One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind.
@@
 \url{https://doi.org/10.1111/desc.12270}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gieseking_1972}{}}%
+Gieseking, W., \& Leimer, K. (1972). \emph{Piano {Technique}}. {Dover Publications}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lobel_2021}{}}%
+Lobel, E.
+E.
+(2021).
+A {Model} for {Integrating Social} and {Creative Dance Forms} in a {Dance Education Course}.
+\emph{Journal of Dance Education}, \emph{21}(1), 44--47.
+\url{https://doi.org/10.1080/15290824.2020.1732984}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mainwaring_2010}{}}%
+Mainwaring, L.
+M., \& Krasnow, D.
+H.
+(2010).
+Teaching the {Dance Class}: {Strategies} to {Enhance Skill Acquisition}, {Mastery} and {Positive Self-Image}.
+\emph{Journal of Dance Education}, \emph{10}(1), 14--21.
+\url{https://doi.org/10.1080/15290824.2010.10387153}
 

### 2023-01-17 paper/manuscript.Rmd
add some figures for supplementary materials
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2023-01-16 paper/manuscript.Rmd
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_docx
 ---
@@
 
-To address this question, we investigated whether and how experts can exaggerate specific parameters of the action to be acquired depending on learners' skill level.
+To address this question, we investigated whether and how experts can exaggerate specific parameters of the action to be acquired depending on learners' skill levels.
 We chose to study expressive music performance, in which experts need to modulate multiple performance parameters such as articulation and dynamics to implement expression [@cancino-chacon_2018].
 Musicians can create their unique expression and interpretation of music even if they play exactly the same piece because when and how they change multiple performance parameters are different.
-By studying expressive performance, we can investigate how precisely experts can highlight specific parameters in a multidimentional space of expression (e.g., tempo, articulation dynamics) depending on the goal of teaching (e.g., to teach articulation or dynamics) even when they are performing the same piece.
+By studying expressive performance, we can investigate how precisely experts can highlight specific parameters in a multidimensional space of expression (e.g., tempo, articulation dynamics) depending on the goal of teaching (e.g., to teach articulation or dynamics) even when they are performing the same piece.
 
@@
 
-The experiment was programmed in Max/MSP (8.1.11; https://cycling74.com/products/max) on a Mac Book Pro with Mac OS X Catalina 10.15.7.
+The experiment was programmed in Max/MSP (8.1.11; https://cycling74.com/products/max) on a MacBook Pro with Mac OS X Catalina 10.15.7.
 A weighted Yamaha MIDI digital piano was used to record participants' performances.
@@
 Two expressive notations (i.e., articulation and dynamics) were added, as shown in *Fig \ref{fig:stim}*.
+The fingering was assigned but participants were told that they did not have to follow the fingering indicated if they were not comfortable with it.
 
 We created artificial recordings of novices to manipulate the displayed proficiency with regard to the two relevant techniques.
-There were four types of recordings: 1) recordings where both articulation and dynamics were implemented (both), 2) recordings where only articulation was implemented whereas dynamics was missing (art only), 3) recordings where only dynamics was implemented whereas articulation was missing (dyn only), and 4) recordings where neither articulation nor dynamics was implemented (none).
-We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
+There were four types of recordings: 1) recordings where both articulation and dynamics were implemented (both), 2) recordings where only articulation was implemented whereas dynamics was missing (art-only), 3) recordings where only dynamics was implemented whereas articulation was missing (dyn-only), and 4) recordings where neither articulation nor dynamics was implemented (none).
+We generated four instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
 How we generated the recordings is described in *Supplementary Material*.
@@
 
-After participants completed the 16 trials, they were asked to perform the piece with the same instruction as in their baseline performance.
+After participants completed the 16 trials, they were asked to perform the piece with the same instruction as in their baseline performances.
 At the end of the experiment, participants filled in a questionnaire asking about their demographic information and experience in piano performance/teaching.
@@
 
+There are possibilities that it may be difficult in general for experts to flexibly adjust to different action parameters while teaching or they might have chosen to focus on the most salient parameters or the ones that they find most relevant.
+For example, when teaching dance, it is considered that good teachers identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation [@mainwaring_2010].
+When teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha [@lobel_2021].
+Karl Leimer, a German music teacher and pianist, wrote in his booked that students should perform with legato at first in finger exercises such as the playing of scales [@gieseking_1972].
+In both dance and music examples, the ultimate goal of skill acquisition is that students are able to use multiple techniques flexibly while performing.
+However, it seems that students need to build complex skills progressively based on the basics during the learning process.
+Therefore, identifying and focusing on the most relevant problem for each student may be more effective to facilitate skill acquisition.
+
 One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind.
@@
 We used the data from Experiment 2.
-In Experiment 2, participants were asked to memorise one piece of music (*Fig 5* (A), @tominaga_2022) without any expressive notation before coming to the experiment.
-First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performance.
+In Experiment 2, participants were asked to memorise one piece of music (*Fig 5* (A), @tominaga_2022) without any expressive notation before coming to the lab.
+First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performances.
 After measuring the baseline performance, participants were asked to perform the piece with either articulation (*Fig 5* (B), @tominaga_2022) or dynamics (*Fig 5* (C), @tominaga_2022) in two conditions (teaching vs.
@@
 
-We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see *Apparatus and stimuli*) by using the performance data from @tominaga_2022.
+We aimed to generate four types of recordings (i.e., both, art-only, dyn-only, none) and each type had four instances (see *Apparatus and stimuli*) by using the performance data from @tominaga_2022.
 First, 16 instances for each type were generated.
@@
 
-In order to create recordings, we needed information about 1) the pitch order, 2) when each note was pressed (related to tempo), 3) when each note was released (related to articulation) and 4) how fast each note was hit (related to dynamics).
-Pitch information was already determined and the same for all structures because of the sheet music (*Fig \ref{fig:stim}*).
+In order to create recordings, we needed information about 1) the pitch order (the order of each key press), 2) the onset time of each key press (related to tempo), 3) the offset time of each key press (related to articulation) and 4) the key velocity of each key press (related to dynamics).
+Pitch information was already determined and the same for all instances based on sheet music (*Fig \ref{fig:stim}*).
 
-First, we created 16 basic structures of recordings from the baseline performances.
-To determine the temporal feature of each structure (i.e., when each key was pressed), three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged.
-Due to the size of the dataset (i.e., 31 performances), some baseline performances were used twice.
-We made sure that the identical performance was not selected within each random selection (i.e., each selection consisted of three different performances).
-Thus, we created 16 basic structures for tempi.
+First, we created 16 temporal structures of recordings from the baseline performances.
+Three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged for each interval.
+Due to the dataset size (i.e., 31 performances), some baseline performances were used twice.
+We made sure that the identical performance was not selected within each random selection of three performances.
 
-Next, we were going to follow the same procedure to determine the sound duration (i.e., when each key was released) of each note (related to articulation) and the key velocity profile (i.e., how fast each key was hit) of each note (related to dynamics).
+Next, we were going to follow the same procedure to determine the sound duration of each note (related to articulation) and the key velocity profile of each note (related to dynamics).
 For the sound duration, however, we found that most participants implemented articulation (particularly legato) even in their baseline performances, where they should not implement any particular articulation or dynamics.
 Therefore, we decided to use the data from only two participants who did not show such tendencies.
-We obtained four performances from the two participants and used these four performances four times each without performing any averaging so as to have 16 instances in total.
-For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged.
+We obtained four performances from the two participants and used these four times each without performing any averaging so as to have 16 instances in total.
+For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged for each key.
 This is how we created 16 instances for three performance features (i.e., tempo, articulation and dynamics).
@@
 
-In order to have other three types of recordings (i.e., both, art only, dyn only), we needed to create 16 instances for the sound duration when articulation was implemented and 16 instances for the key velocity profiles when dynamics was implemented.
-We followed the same procedure as we did to generate none recordings but we used articulation performances (not baseline performances) for the sound duration and dynamics performances (not baseline performances) for the key velocity profiles.
-To create art only recordings,
+In order to have other three types of recordings (i.e., both, art-only, dyn-only), we needed to create 16 instances for the sound duration when articulation was implemented and 16 instances for the key velocity profiles when dynamics was implemented.
+We followed the same procedure of averaging as we did to generate none recordings but we used articulation performances to determine the sound duration and dynamics performances to determine the key velocity profiles.
+To create art-only recordings, we replaced the articulation performance feature of none recordings with the one generated from articulation performances.
+Similarly for dyn-only recordings, we replaced the dynamics performance feature of none recordings with the one generated from dynamics performances.
+Lastly, to create both recordings, we replaced the articulation and dynamics performance features of none recordings with the ones generated from articulation and dynamics performances (*Fig \ref{fig:supplementary}*).
+Thus, we generated 16 recordings for each type.
 
@@
 
-After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording and rate to what extent each expression was implemented between 0 (not at all) and 5 (fully).
-Based on these ratings, we selected four recordings for each type.
+After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording while sheet music (*Fig \ref{fig:stim}*) was displayed and rate to what extent each notated expression (i.e., articulation, dynamics) was implemented between 0 (not at all) and 5 (fully).
+Based on these ratings, we selected four recordings, which represented each type of recordings the most.
+In order to fix the temporal feature of performance for four selected instances per type, we randomly selected 4 tempo performance features from 16 instances and then replaced the tempo performance features of each recording type with the ones we selected.
+
+### 4. Tweaking selected recordings
+
+For dyn-only and none recordings, the recordings sounded too much staccato.
+We calculated the difference between the mean of each recording for dyn-only and none recordings and the grand mean of art-only and both recordings in terms of KOT and add that difference to all KOT values for dyn_only and low, so that the KOT values should shift closer to the grand mean of art_only and high stimuli.
+
+```{r supplementary, out.width = "100%", fig.cap = "\\label{fig:supplementary}Schematic representation of stimuli generation processes", echo = FALSE}
+here::i_am("README.md")
+stim <- image_border(image_read(here("paper/image", "supplementary.png")), "#FFFFFF", "140x140")
+image_write(stim, path = here("paper/image", "Fig10.tiff"), format = "tiff")
+stim
+```
+
+```{r analysis-supplementary, include = FALSE}
+here::i_am("README.md")
+source(here("material/expstimuli/audio/tweaked", "analysis.R"), chdir = TRUE)
+
+ioi_profiles <- gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
+
+kot_profiles <- gghistogram(dt_kot[SubArt != "NA"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
+
+vel_profiles <- gghistogram(dt_vel[SubArt != "NA"], x = "Velocity", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
+
+vel_diff_profiles <- gghistogram(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "Diff", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
+
+# save plots
+ggsave(here("paper/image", "s_ioi_profiles.png"), plot = ioi_profiles, dpi = 300, width = 7, height = 5)
+img_s_ioi_profiles <- image_read(here("paper/image", "s_ioi_profiles.png"), density = 300)
+
+ggsave(here("paper/image", "s_kot_profiles.png"), plot = kot_profiles, dpi = 300, width = 7, height = 5)
+img_s_kot_profiles <- image_read(here("paper/image", "s_kot_profiles.png"), density = 300)
+
+ggsave(here("paper/image", "s_vel_profiles.png"), plot = vel_profiles, dpi = 300, width = 7, height = 5)
+img_s_vel_profiles <- image_read(here("paper/image", "s_vel_profiles.png"), density = 300)
+
+ggsave(here("paper/image", "s_vel_diff_profiles.png"), plot = vel_diff_profiles, dpi = 300, width = 7, height = 5)
+img_s_vel_diff_profiles <- image_read(here("paper/image", "s_vel_diff_profiles.png"), density = 300)
+```
+
+```{r plot-s-ioi, out.width = "100%", fig.cap = "\\label{fig:s-ioi}A"}
+img_s_ioi_profiles
+```
+
+```{r plot-s-kot, out.width = "100%", fig.cap = "\\label{fig:s-kot}A"}
+img_s_kot_profiles
+```
+
+```{r plot-s-vel, out.width = "100%", fig.cap = "\\label{fig:s-vel}A"}
+img_s_vel_profiles
+```
+
+```{r plot-s-vel-diff, out.width = "100%", fig.cap = "\\label{fig:s-vel-diff}A"}
+img_s_vel_diff_profiles
+```
 

### 2023-01-17 paper/manuscript.tex
add some figures for supplementary materials
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 There are possibilities that it may be difficult in general for experts to flexibly adjust to different action parameters while teaching or they might have chosen to focus on the most salient parameters or the ones that they find most relevant.
-For example, when teaching dance, it is considered that good teachers need to identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation (Mainwaring \& Krasnow, 2010).
-For example, when teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha (Lobel, 2021).
-Karl Leimer, a German music teacher and pianist, wrote in his booked that when doing finger exercises, students should perform with legato at first (Gieseking \& Leimer, 1972).
-Legato creates round and mellow tones, which is one of the most essential elements of piano touches in performance.
+For example, when teaching dance, it is considered that good teachers identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation (Mainwaring \& Krasnow, 2010).
+When teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha (Lobel, 2021).
+Karl Leimer, a German music teacher and pianist, wrote in his booked that students should perform with legato at first in finger exercises such as the playing of scales (Gieseking \& Leimer, 1972).
 In both dance and music examples, the ultimate goal of skill acquisition is that students are able to use multiple techniques flexibly while performing.
-However, it seems that students need to build complex skills based on basic ones during the learning process.
-Therefore, identifying and focusing on the most relevant problem for each student may be more effective to facilitate learning.
+However, it seems that students need to build complex skills progressively based on the basics during the learning process.
+Therefore, identifying and focusing on the most relevant problem for each student may be more effective to facilitate skill acquisition.
 

### 2023-01-17 material/expstimuli/audio/tweaked/analysis.R
change labelling for manuscript
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,18 +1,13 @@
----
-title: "Music Adaptation: Analysis for selected stimuli"
-output: html_notebook
----
-
-```{r packages, include = FALSE}
+## ----packages, include = FALSE------------------------------
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 # midi
 if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
 # plot
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
-```
 
-```{r midi, include = FALSE}
+
+## ----midi, include = FALSE----------------------------------
 # create a list of data file names
 lf <- list.files("../selected/mid/", pattern = "mid")
 
@@ -57,10 +52,9 @@ ls_piano_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(5
 change_1 <- c(5, 17, 47)
 # define Component Change (StoL, PtoF)
 change_2 <- c(8, 20, 39)
-```
 
-# IOI
-```{r ioi, include = FALSE}
+
+## ----ioi, include = FALSE-----------------------------------
 data_onset <- data[event == "Note On"]
 data_offset <- data[event == "Note Off"]
 
@@ -104,12 +98,9 @@ for (number in change_2){
  dt_ioi[Interval == number]$SubArt <- "StoL"
  dt_ioi[Interval == number]$SubDyn <- "PtoF"
 }
-```
 
-- There are 4 base tempi.
 
-## Histogram
-```{r ioi-distribution, echo = FALSE}
+## ----ioi-distribution, echo = FALSE-------------------------
 # change labelling
 dt_ioi[MidFile == "high_1"]$MidFile <- "both_1"
 dt_ioi[MidFile == "high_2"]$MidFile <- "both_2"
@@ -121,24 +112,16 @@ dt_ioi[MidFile == "low_3"]$MidFile <- "none_3"
 dt_ioi[MidFile == "low_4"]$MidFile <- "none_4"
 
 gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
-```
 
-## Mean
-- The red line indicates the ideal tempo (100bpm - IOI: 300 ms)
 
-```{r ioi-mean, echo = FALSE}
[…]

### 2023-01-17 material/expstimuli/audio/tweaked/analysis.Rmd
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

--- 2021-02-15 material/expstimuli/tweaked/analysis.Rmd
@@
 ```{r ioi-distribution, echo = FALSE}
+# change labelling
+dt_ioi[MidFile == "high_1"]$MidFile <- "both_1"
+dt_ioi[MidFile == "high_2"]$MidFile <- "both_2"
+dt_ioi[MidFile == "high_3"]$MidFile <- "both_3"
+dt_ioi[MidFile == "high_4"]$MidFile <- "both_4"
+dt_ioi[MidFile == "low_1"]$MidFile <- "none_1"
+dt_ioi[MidFile == "low_2"]$MidFile <- "none_2"
+dt_ioi[MidFile == "low_3"]$MidFile <- "none_3"
+dt_ioi[MidFile == "low_4"]$MidFile <- "none_4"
+
 gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
@@
 # KOT
-
 ```{r kot, include = FALSE}
@@
 ## Histogram
+
+- For art_only and both stimuli, there should be a bimodal distribution to represent the implementation of legato and staccato.
+
+- For dyn_only and none stimuli, there should be a normal distribution.
+
 ```{r kot-distribution, echo = FALSE}
-gghistogram(dt_kot[SubArt != "NA"], x = "KOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
+# change labelling
+dt_kot[MidFile == "high_1"]$MidFile <- "both_1"
+dt_kot[MidFile == "high_2"]$MidFile <- "both_2"
+dt_kot[MidFile == "high_3"]$MidFile <- "both_3"
+dt_kot[MidFile == "high_4"]$MidFile <- "both_4"
+dt_kot[MidFile == "low_1"]$MidFile <- "none_1"
+dt_kot[MidFile == "low_2"]$MidFile <- "none_2"
+dt_kot[MidFile == "low_3"]$MidFile <- "none_3"
+dt_kot[MidFile == "low_4"]$MidFile <- "none_4"
+
+gghistogram(dt_kot[SubArt != "NA" & SubArt != "LtoS" & SubArt != "StoL"], x = "KOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
 ```
@@
 ## Tweak
+- For dyn_only and none stimuli, it seems the KOT values are too close to the staccato range.
+We calculated the difference between the mean of each stimulus and the grand mean of art_only and both stimuli and add that difference to all KOT values for dyn_only and none stimuli, so that the KOT values should shift closer to the grand mean of art_only and both stimuli.
+
+## Histogram (Fixed KOT)
 ```{r kot-tweak, echo = FALSE}
-# grandMean of art_only & high stimuli
-kot_grandMean <- mean(dt_kot[grepl("art_only", MidFile) | grepl("high", MidFile) & SubArt != "NA"]$KOT)
+# grandMean of art_only & both stimuli
+kot_grandMean <- mean(dt_kot[grepl("art_only", MidFile) | grepl("both", MidFile) & SubArt != "NA"]$KOT)
 
 # calculate the diff between each KOT mean and grandMean
-tweak_data <- dt_kot[grepl("dyn_only", MidFile) | grepl("low", MidFile) & SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]
-tweak_data$Diff <- abs(tweak_data$Mean-kot_grandMean)
+tweak_data <- dt_kot[grepl("dyn_only", MidFile) | grepl("none", MidFile) & SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]
+tweak_data$Diff <- round(abs(tweak_data$Mean-kot_grandMean), 0)
 tweak_data
@@
 
-## Histogram (Fixed KOT)
 ```{r kot-distribution-fixed, echo = FALSE}
-gghistogram(dt_kot[SubArt != "NA"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
+gghistogram(dt_kot[SubArt != "NA" & SubArt != "LtoS" & SubArt != "StoL"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
 ```
@@
 # KV
-
 ```{r vel, include = FALSE}
@@
 ## Histogram
+
+- For dyn_only and both stimuli, there should be a bimodal distribution to represent the implementation of forte and piano.
+
+- For art_only and none stimuli, there should be a normal distribution.
+
 ```{r vel-distribution, echo = FALSE}
+# change labelling
+dt_vel[MidFile == "high_1"]$MidFile <- "both_1"
+dt_vel[MidFile == "high_2"]$MidFile <- "both_2"
+dt_vel[MidFile == "high_3"]$MidFile <- "both_3"
+dt_vel[MidFile == "high_4"]$MidFile <- "both_4"
+dt_vel[MidFile == "low_1"]$MidFile <- "none_1"
+dt_vel[MidFile == "low_2"]$MidFile <- "none_2"
+dt_vel[MidFile == "low_3"]$MidFile <- "none_3"
+dt_vel[MidFile == "low_4"]$MidFile <- "none_4"
+
 gghistogram(dt_vel[SubArt != "NA"], x = "Velocity", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
@@
 ## Histogram
+
+- For dyn_only and both stimuli, there should be a bimodal distribution to represent the transition from forte to piano or from piano to forte.
+
+- For art_only and none stimuli, there should be a normal distribution.
+
 ```{r vel-diff-distribution, echo = FALSE}
+# change labelling
+dt_vel_diff[MidFile == "high_1"]$MidFile <- "both_1"
+dt_vel_diff[MidFile == "high_2"]$MidFile <- "both_2"
+dt_vel_diff[MidFile == "high_3"]$MidFile <- "both_3"
+dt_vel_diff[MidFile == "high_4"]$MidFile <- "both_4"
+dt_vel_diff[MidFile == "low_1"]$MidFile <- "none_1"
+dt_vel_diff[MidFile == "low_2"]$MidFile <- "none_2"
+dt_vel_diff[MidFile == "low_3"]$MidFile <- "none_3"
+dt_vel_diff[MidFile == "low_4"]$MidFile <- "none_4"
+
 gghistogram(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "Diff", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
@@
 ```
+
+```{r export, include = FALSE}
+knitr::purl("analysis.Rmd")
+```
+
+<!-- # Export stimuli with fixed KOT as mid files -->
+<!-- ```{r export, include = FALSE} -->
+<!-- # remove unnecessary columns -->
+<!-- data_onset$IOI <- NULL -->
+<!-- data_onset$KOT <- NULL -->
+<!-- data_onset$Diff <- NULL -->
+<!-- # fix time for onset and offset data -->
+<!-- data_onset$timeFixed <- data_onset$time -->
+<!-- data_offset$timeFixed <- data_offset$time -->
+<!-- for (midfile in unique(tweak_data$MidFile)){ -->
+<!--    # offset[i] = onset[i+1]+KOT[i] -->
+<!--    for (i in 1:nrow(data_onset[MidFile == midfile])-1){ -->
+<!--       data_offset[MidFile == midfile][i]$timeFixed <- data_onset[MidFile == midfile][i+1]$time+dt_kot[MidFile == midfile][i]$FixedKOT -->
+<!--    } -->
+<!-- } -->
+
+<!-- # export each instance as txt file -->
+<!-- for (stim in unique(data_onset$MidFile)){ -->
+<!--    onset <- data_onset[MidFile == stim] -->
+<!--    offset <- data_offset[MidFile == stim] -->
+<!--    instance <- rbind(onset, offset) -->
+<!--    # Note_On/Off >> 1 or 0 -->
+<!--    instance$KeyOnOff <- 0 -->
+<!--    instance[event == "Note On"]$KeyOnOff <- 1 -->
+<!--    # sort by timeFixed -->
+<!--    instance <- instance[order(timeFixed)] -->
+
+<!--    filename = paste(stim, ".txt", sep = "") -->
+<!--    fwrite(instance, filename) -->
+<!-- } -->
+<!-- ``` -->
+
+<!-- ```{bash, echo = TRUE} -->
+<!-- # python -->
+<!-- source /Users/atsukotominaga/Pyenv/bin/activate -->
+<!-- python mid_export.py -->
+<!-- ``` -->

### 2023-01-17 paper/manuscript.Rmd
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 img_kot_1 <- image_scale(img_kot_1, "2200")
-image_write(img_kot_1, path = here("paper/image", "Fig3.tiff"), format = "tiff")
+image_write(img_kot_1, path = here("paper/image", "Fig2.tiff"), format = "tiff")
 img_kot_1
@@
 img_vel_1 <- image_scale(img_vel_1, "2200")
-image_write(img_vel_1, path = here("paper/image", "Fig4.tiff"), format = "tiff")
+image_write(img_vel_1, path = here("paper/image", "Fig3.tiff"), format = "tiff")
 img_vel_1
@@
 img_vel_diff_1 <- image_scale(img_vel_diff_1, "2200")
-image_write(img_vel_diff_1, path = here("paper/image", "Fig5.tiff"), format = "tiff")
+image_write(img_vel_diff_1, path = here("paper/image", "Fig4.tiff"), format = "tiff")
 img_vel_diff_1
@@
 img_ioi_1 <- image_scale(img_ioi_all_plot, "2200")
-image_write(img_ioi_1, path = here("paper/image", "Fig2.tiff"), format = "tiff")
+image_write(img_ioi_1, path = here("paper/image", "Fig5.tiff"), format = "tiff")
 img_ioi_1
@@
 
-```{r supplementary, out.width = "100%", fig.cap = "\\label{fig:supplementary}Schematic representation of stimuli generation processes", echo = FALSE}
+```{r supplementary, out.width = "100%", fig.cap = "\\label{fig:supplementary}Schematic representation of the stimuli generation process.", echo = FALSE}
 here::i_am("README.md")
-stim <- image_border(image_read(here("paper/image", "supplementary.png")), "#FFFFFF", "140x140")
-image_write(stim, path = here("paper/image", "Fig10.tiff"), format = "tiff")
-stim
+supplementary <- image_border(image_read(here("paper/image", "supplementary.png")), "#FFFFFF", "140x140")
+image_write(supplementary, path = here("paper/image", "Fig6.tiff"), format = "tiff")
+supplementary
 ```
@@
 
-kot_profiles <- gghistogram(dt_kot[SubArt != "NA"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
+kot_profiles <- gghistogram(dt_kot[SubArt != "NA" & SubArt != "LtoS" & SubArt != "StoL"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
 
@@
 
-```{r plot-s-ioi, out.width = "100%", fig.cap = "\\label{fig:s-ioi}A"}
+```{r plot-s-ioi, out.width = "100%", fig.cap = "\\label{fig:s-ioi}Histogram of interonset intervals (IOIs) for each recording.
+Since we fixed the tempo performance feature across 4 instances, IOI profiles were identical for each type."}
+image_write(img_s_ioi_profiles, path = here("paper/image", "Fig7.tiff"), format = "tiff")
 img_s_ioi_profiles
@@
 
-```{r plot-s-kot, out.width = "100%", fig.cap = "\\label{fig:s-kot}A"}
+```{r plot-s-kot, out.width = "100%", fig.cap = "\\label{fig:s-kot}Histogram of Key-overlap time (KOT) for each recording.
+For art-only and both recordings, there is a bimodal distribution to represent the implementation of legato and staccato.
+For dyn-only and none recordings, there is no bimodal distribution (no difference between legato and staccato parts of performance).
+Red bars indicate KOT values where legato is notated.
+Blue bars indicate KOT values where staccato is notated."}
+image_write(img_s_kot_profiles, path = here("paper/image", "Fig8.tiff"), format = "tiff")
 img_s_kot_profiles
@@
 
-```{r plot-s-vel, out.width = "100%", fig.cap = "\\label{fig:s-vel}A"}
+```{r plot-s-vel, out.width = "100%", fig.cap = "\\label{fig:s-vel}Histogram of key velocity (KV) for each recording.
+For dyn-only and both recordings, there is a bimodal distribution to represent the implementation of forte and piano.
+For art-only and none recordings, there is no bimodal distribution (no difference between forte and piano parts of performance).
+Red bars indicate KV values where forte is notated.
+Blue bars indicate KV values where piano is notated."}
+image_write(img_s_vel_profiles, path = here("paper/image", "Fig9.tiff"), format = "tiff")
 img_s_vel_profiles
@@
 
-```{r plot-s-vel-diff, out.width = "100%", fig.cap = "\\label{fig:s-vel-diff}A"}
+```{r plot-s-vel-diff, out.width = "100%", fig.cap = "\\label{fig:s-vel-diff}Histogram of KV difference for each recording.
+For dyn-only and both recordings, there is a bimodal distribution to represent the transition from forte to piano or from piano to forte.
+For art-only and none recordings, there is no bimodal distribution (no contrast difference in transition parts between forte and piano).
+Red bars indicate KV difference of the transition from forte to piano.
+Blue bars indicate KV difference of the transition from piano to forte."}
+image_write(img_s_vel_diff_profiles, path = here("paper/image", "Fig10.tiff"), format = "tiff")
 img_s_vel_diff_profiles
@@
 img_kot_2 <- image_scale(img_kot_2, "2200")
-image_write(img_kot_2, path = here("paper/image", "Fig7.tiff"), format = "tiff")
+image_write(img_kot_2, path = here("paper/image", "Fig11.tiff"), format = "tiff")
 img_kot_2
@@
 img_vel_2 <- image_scale(img_vel_2, "2200")
-image_write(img_vel_2, path = here("paper/image", "Fig8.tiff"), format = "tiff")
+image_write(img_vel_2, path = here("paper/image", "Fig12.tiff"), format = "tiff")
 img_vel_2
@@
 img_vel_diff_2 <- image_scale(img_vel_diff_2, "2200")
-image_write(img_vel_diff_2, path = here("paper/image", "Fig9.tiff"), format = "tiff")
+image_write(img_vel_diff_2, path = here("paper/image", "Fig13.tiff"), format = "tiff")
 img_vel_diff_2
@@
 img_ioi_2 <- image_scale(img_ioi_ca_plot, "2200")
-image_write(img_ioi_2, path = here("paper/image", "Fig6.tiff"), format = "tiff")
+image_write(img_ioi_2, path = here("paper/image", "Fig14.tiff"), format = "tiff")
 img_ioi_2

### 2023-01-17 material/expstimuli/audio/tweaked/analysis.R

@@ -1,4 +1,4 @@
-## ----packages, include = FALSE------------------------------
+## ----packages, include = FALSE------------------------------------------------
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 # midi
@@ -7,7 +7,7 @@ if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----midi, include = FALSE----------------------------------
+## ----midi, include = FALSE----------------------------------------------------
 # create a list of data file names
 lf <- list.files("../selected/mid/", pattern = "mid")
 
@@ -54,7 +54,7 @@ change_1 <- c(5, 17, 47)
 change_2 <- c(8, 20, 39)
 
 
-## ----ioi, include = FALSE-----------------------------------
+## ----ioi, include = FALSE-----------------------------------------------------
 data_onset <- data[event == "Note On"]
 data_offset <- data[event == "Note Off"]
 
@@ -100,7 +100,7 @@ for (number in change_2){
 }
 
 
-## ----ioi-distribution, echo = FALSE-------------------------
+## ----ioi-distribution, echo = FALSE-------------------------------------------
 # change labelling
 dt_ioi[MidFile == "high_1"]$MidFile <- "both_1"
 dt_ioi[MidFile == "high_2"]$MidFile <- "both_2"
@@ -114,14 +114,14 @@ dt_ioi[MidFile == "low_4"]$MidFile <- "none_4"
 gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
 
 
-## ----ioi-mean, echo = FALSE---------------------------------
+## ----ioi-mean, echo = FALSE---------------------------------------------------
 # mean
 dt_ioi[SubArt != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), "MidFile"]
 
 ggboxplot(dt_ioi[SubArt != "NA"], x = "MidFileShort", y = "IOI", add = "jitter", width = 0.8) + geom_hline(yintercept = 300, linetype = "dashed", color = "red")
 
 
[…]

### 2023-01-17 paper/manuscript.Rmd
second draft / remove fig6 for now
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 classoption       : "man"
-output            : papaja::apa6_docx
+output            : papaja::apa6_pdf
 ---
@@
 For example, when teaching dance, it is considered that good teachers identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation [@mainwaring_2010].
-When teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha [@lobel_2021].
-Karl Leimer, a German music teacher and pianist, wrote in his booked that students should perform with legato at first in finger exercises such as the playing of scales [@gieseking_1972].
+When teaching social dance, a teacher started with a basic step as the hustle and then moved on more complex steps progressively like the waltz and the cha cha [@lobel_2021].
+Karl Leimer, a German music teacher and pianist, wrote in his book that students should perform with legato at first in finger exercises such as the playing of scales before learning other techniques [@gieseking_1972].
 In both dance and music examples, the ultimate goal of skill acquisition is that students are able to use multiple techniques flexibly while performing.
@@
 Due to the dataset size (i.e., 31 performances), some baseline performances were used twice.
-We made sure that the identical performance was not selected within each random selection of three performances.
+We made sure that the identical performance was not selected within each random selection of three performances (*Fig \ref{fig:supplementary}* I).
 
@@
 We obtained four performances from the two participants and used these four times each without performing any averaging so as to have 16 instances in total.
-For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged for each key.
+For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged for each key (*Fig \ref{fig:supplementary}* I).
 This is how we created 16 instances for three performance features (i.e., tempo, articulation and dynamics).
@@
 Similarly for dyn-only recordings, we replaced the dynamics performance feature of none recordings with the one generated from dynamics performances.
-Lastly, to create both recordings, we replaced the articulation and dynamics performance features of none recordings with the ones generated from articulation and dynamics performances (*Fig \ref{fig:supplementary}*).
+Lastly, to create both recordings, we replaced the articulation and dynamics performance features of none recordings with the ones generated from articulation and dynamics performances (*Fig \ref{fig:supplementary}* II).
 Thus, we generated 16 recordings for each type.
@@
 After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording while sheet music (*Fig \ref{fig:stim}*) was displayed and rate to what extent each notated expression (i.e., articulation, dynamics) was implemented between 0 (not at all) and 5 (fully).
-Based on these ratings, we selected four recordings, which represented each type of recordings the most.
-In order to fix the temporal feature of performance for four selected instances per type, we randomly selected 4 tempo performance features from 16 instances and then replaced the tempo performance features of each recording type with the ones we selected.
+Based on these ratings, we selected four recordings, which represented each type of recordings the most (*Fig \ref{fig:supplementary}* III).
+In order to fix the temporal feature of performance for four selected instances per type, we randomly selected four tempo performance features from 16 instances and then replaced the tempo performance features of each recording type with the ones we selected (*Fig \ref{fig:s-ioi}*).
 
@@
 For dyn-only and none recordings, the recordings sounded too much staccato.
-We calculated the difference between the mean of each recording for dyn-only and none recordings and the grand mean of art-only and both recordings in terms of KOT and add that difference to all KOT values for dyn_only and low, so that the KOT values should shift closer to the grand mean of art_only and high stimuli.
+We calculated the difference between the mean of each recording for dyn-only and none recordings and the grand mean of art-only and both recordings in terms of KOT and add that difference to all KOT values for dyn_only and none recordings, so that the KOT values should shift closer to the grand mean of art_only and both recordings (*Fig \ref{fig:s-kot}*).
 
@@
 here::i_am("README.md")
-supplementary <- image_border(image_read(here("paper/image", "supplementary.png")), "#FFFFFF", "140x140")
+supplementary <- image_border(image_read(here("paper/image", "supplementary.jpg")), "#FFFFFF", "140x140")
 image_write(supplementary, path = here("paper/image", "Fig6.tiff"), format = "tiff")
@@
 
-```{r plot-s-kot, out.width = "100%", fig.cap = "\\label{fig:s-kot}Histogram of Key-overlap time (KOT) for each recording.
+```{r plot-s-kot, out.width = "100%", fig.cap = "\\label{fig:s-kot}Histogram of key-overlap time (KOT) for each recording.
 For art-only and both recordings, there is a bimodal distribution to represent the implementation of legato and staccato.

### 2023-01-17 paper/manuscript.tex
second draft / remove fig6 for now
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-To address this question, we investigated whether and how experts can exaggerate specific parameters of the action to be acquired depending on learners' skill level.
+To address this question, we investigated whether and how experts can exaggerate specific parameters of the action to be acquired depending on learners' skill levels.
 We chose to study expressive music performance, in which experts need to modulate multiple performance parameters such as articulation and dynamics to implement expression (Cancino-Chacón, Grachten, Goebl, \& Widmer, 2018).
 Musicians can create their unique expression and interpretation of music even if they play exactly the same piece because when and how they change multiple performance parameters are different.
-By studying expressive performance, we can investigate how precisely experts can highlight specific parameters in a multidimentional space of expression (e.g., tempo, articulation dynamics) depending on the goal of teaching (e.g., to teach articulation or dynamics) even when they are performing the same piece.
+By studying expressive performance, we can investigate how precisely experts can highlight specific parameters in a multidimensional space of expression (e.g., tempo, articulation dynamics) depending on the goal of teaching (e.g., to teach articulation or dynamics) even when they are performing the same piece.
 
@@
 
-The experiment was programmed in Max/MSP (8.1.11; \url{https://cycling74.com/products/max}) on a Mac Book Pro with Mac OS X Catalina 10.15.7.
+The experiment was programmed in Max/MSP (8.1.11; \url{https://cycling74.com/products/max}) on a MacBook Pro with Mac OS X Catalina 10.15.7.
 A weighted Yamaha MIDI digital piano was used to record participants' performances.
@@
 Two expressive notations (i.e., articulation and dynamics) were added, as shown in \emph{Fig \ref{fig:stim}}.
+The fingering was assigned but participants were told that they did not have to follow the fingering indicated if they were not comfortable with it.
 
 We created artificial recordings of novices to manipulate the displayed proficiency with regard to the two relevant techniques.
-There were four types of recordings: 1) recordings where both articulation and dynamics were implemented (both), 2) recordings where only articulation was implemented whereas dynamics was missing (art only), 3) recordings where only dynamics was implemented whereas articulation was missing (dyn only), and 4) recordings where neither articulation nor dynamics was implemented (none).
-We generated 4 instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
+There were four types of recordings: 1) recordings where both articulation and dynamics were implemented (both), 2) recordings where only articulation was implemented whereas dynamics was missing (art-only), 3) recordings where only dynamics was implemented whereas articulation was missing (dyn-only), and 4) recordings where neither articulation nor dynamics was implemented (none).
+We generated four instances for each type, therefore there were 16 stimuli (i.e., recordings) in total.
 How we generated the recordings is described in \emph{Supplementary Material}.
@@
 For example, when teaching dance, it is considered that good teachers identify and build the basic knowledge and skills first and then add complexity progressively from the basic foundation (Mainwaring \& Krasnow, 2010).
-When teaching social dance, a teacher started with a basic step as the hustle and then moved onto more complex steps progressively like the waltz and the cha cha (Lobel, 2021).
-Karl Leimer, a German music teacher and pianist, wrote in his booked that students should perform with legato at first in finger exercises such as the playing of scales (Gieseking \& Leimer, 1972).
+When teaching social dance, a teacher started with a basic step as the hustle and then moved on more complex steps progressively like the waltz and the cha cha (Lobel, 2021).
+Karl Leimer, a German music teacher and pianist, wrote in his book that students should perform with legato at first in finger exercises such as the playing of scales before learning other techniques (Gieseking \& Leimer, 1972).
 In both dance and music examples, the ultimate goal of skill acquisition is that students are able to use multiple techniques flexibly while performing.
@@
 In Experiment 2, participants were asked to memorise one piece of music (\emph{Fig 5} (A), Tominaga et al.
-(2022)) without any expressive notation before coming to the experiment.
-First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performance.
+(2022)) without any expressive notation before coming to the lab.
+First, we asked participants to perform the memorised piece before starting the experiment to measure their baseline performances.
 After measuring the baseline performance, participants were asked to perform the piece with either articulation (\emph{Fig 5} (B), Tominaga et al.
@@
 
-We aimed to generate four types of recordings and each type (i.e., both, art only, dyn only, none) had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga et al.
+We aimed to generate four types of recordings (i.e., both, art-only, dyn-only, none) and each type had four instances (see \emph{Apparatus and stimuli}) by using the performance data from Tominaga et al.
 (2022).
@@
 
-In order to create recordings, we needed information about 1) the pitch order, 2) when each note was pressed (related to tempo), 3) when each note was released (related to articulation) and 4) how fast each note was hit (related to dynamics).
-Pitch information was already determined and the same for all structures because of the sheet music (\emph{Fig \ref{fig:stim}}).
+In order to create recordings, we needed information about 1) the pitch order (the order of each key press), 2) the onset time of each key press (related to tempo), 3) the offset time of each key press (related to articulation) and 4) the key velocity of each key press (related to dynamics).
+Pitch information was already determined and the same for all instances based on sheet music (\emph{Fig \ref{fig:stim}}).
 
-First, we created 16 basic structures of recordings from the baseline performances.
-To determine the temporal feature of each structure (i.e., when each key was pressed), three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged.
-Due to the size of the dataset (i.e., 31 performances), some baseline performances were used twice.
-We made sure that the identical performance was not selected within each random selection (i.e., each selection consisted of three different performances).
-Thus, we created 16 basic structures for tempi.
+First, we created 16 temporal structures of recordings from the baseline performances.
+Three performances out of the 31 valid baseline performances were randomly selected and interonset intervals (IOIs) of the three performances were averaged for each interval.
+Due to the dataset size (i.e., 31 performances), some baseline performances were used twice.
+We made sure that the identical performance was not selected within each random selection of three performances (\emph{Fig \ref{fig:supplementary}} I).
 
-Next, we were going to follow the same procedure to determine the sound duration (i.e., when each key was released) of each note (related to articulation) and the key velocity profile (i.e., how fast each key was hit) of each note (related to dynamics).
+Next, we were going to follow the same procedure to determine the sound duration of each note (related to articulation) and the key velocity profile of each note (related to dynamics).
 For the sound duration, however, we found that most participants implemented articulation (particularly legato) even in their baseline performances, where they should not implement any particular articulation or dynamics.
 Therefore, we decided to use the data from only two participants who did not show such tendencies.
-We obtained four performances from the two participants and used these four performances four times each without performing any averaging so as to have 16 instances in total.
-For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged.
+We obtained four performances from the two participants and used these four times each without performing any averaging so as to have 16 instances in total.
+For the key velocity profiles, we followed the same procedure as we determined tempi and three performances out of the 31 valid baseline performances were randomly selected and the key velocity profiles of the three performances were averaged for each key (\emph{Fig \ref{fig:supplementary}} I).
 This is how we created 16 instances for three performance features (i.e., tempo, articulation and dynamics).
@@
 
-In order to have other three types of recordings (i.e., both, art only, dyn only), we needed to create 16 instances for the sound duration when articulation was implemented and 16 instances for the key velocity profiles when dynamics was implemented.
-We followed the same procedure as we did to generate none recordings but we used articulation performances (not baseline performances) for the sound duration and dynamics performances (not baseline performances) for the key velocity profiles.
-To create art only recordings,
+In order to have other three types of recordings (i.e., both, art-only, dyn-only), we needed to create 16 instances for the sound duration when articulation was implemented and 16 instances for the key velocity profiles when dynamics was implemented.
+We followed the same procedure of averaging as we did to generate none recordings but we used articulation performances to determine the sound duration and dynamics performances to determine the key velocity profiles.
+To create art-only recordings, we replaced the articulation performance feature of none recordings with the one generated from articulation performances.
+Similarly for dyn-only recordings, we replaced the dynamics performance feature of none recordings with the one generated from dynamics performances.
+Lastly, to create both recordings, we replaced the articulation and dynamics performance features of none recordings with the ones generated from articulation and dynamics performances (\emph{Fig \ref{fig:supplementary}} II).
+Thus, we generated 16 recordings for each type.
 
@@
 
-After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording and rate to what extent each expression was implemented between 0 (not at all) and 5 (fully).
-Based on these ratings, we selected four recordings for each type.
+After we generated 16 recordings for each type, we asked 3 musician raters including the first author to listen to each generated recording while sheet music (\emph{Fig \ref{fig:stim}}) was displayed and rate to what extent each notated expression (i.e., articulation, dynamics) was implemented between 0 (not at all) and 5 (fully).
+Based on these ratings, we selected four recordings, which represented each type of recordings the most (\emph{Fig \ref{fig:supplementary}} III).
+In order to fix the temporal feature of performance for four selected instances per type, we randomly selected four tempo performance features from 16 instances and then replaced the tempo performance features of each recording type with the ones we selected (\emph{Fig \ref{fig:s-ioi}}).
+
+\hypertarget{tweaking-selected-recordings}{%
+\subsubsection{4. Tweaking selected recordings}\label{tweaking-selected-recordings}}
+
+For dyn-only and none recordings, the recordings sounded too much staccato.
+We calculated the difference between the mean of each recording for dyn-only and none recordings and the grand mean of art-only and both recordings in terms of KOT and add that difference to all KOT values for dyn\_only and none recordings, so that the KOT values should shift closer to the grand mean of art\_only and both recordings (\emph{Fig \ref{fig:s-kot}}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/supplementary-1} \caption{\label{fig:supplementary}Schematic representation of the stimuli generation process.}\label{fig:supplementary}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-s-ioi-1} \caption{\label{fig:s-ioi}Histogram of interonset intervals (IOIs) for each recording.
+Since we fixed the tempo performance feature across 4 instances, IOI profiles were identical for each type.}\label{fig:plot-s-ioi}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-s-kot-1} \caption{\label{fig:s-kot}Histogram of key-overlap time (KOT) for each recording.
+For art-only and both recordings, there is a bimodal distribution to represent the implementation of legato and staccato.
+For dyn-only and none recordings, there is no bimodal distribution (no difference between legato and staccato parts of performance).
+Red bars indicate KOT values where legato is notated.
+Blue bars indicate KOT values where staccato is notated.}\label{fig:plot-s-kot}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-s-vel-1} \caption{\label{fig:s-vel}Histogram of key velocity (KV) for each recording.
+For dyn-only and both recordings, there is a bimodal distribution to represent the implementation of forte and piano.
+For art-only and none recordings, there is no bimodal distribution (no difference between forte and piano parts of performance).
+Red bars indicate KV values where forte is notated.
+Blue bars indicate KV values where piano is notated.}\label{fig:plot-s-vel}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-s-vel-diff-1} \caption{\label{fig:s-vel-diff}Histogram of KV difference for each recording.
+For dyn-only and both recordings, there is a bimodal distribution to represent the transition from forte to piano or from piano to forte.
+For art-only and none recordings, there is no bimodal distribution (no contrast difference in transition parts between forte and piano).
+Red bars indicate KV difference of the transition from forte to piano.
+Blue bars indicate KV difference of the transition from piano to forte.}\label{fig:plot-s-vel-diff}
+\end{figure}
 

### 2023-01-31 paper/manuscript.Rmd
update figure / fix floating pointst
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Neither the main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* `r if(ftop_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* `r if(ftop_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* `r if(ftop_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
+Neither the main effect of Articulation (*F*(`r ftop_all_aov$anova_table[1,1]`, `r ftop_all_aov$anova_table[1,2]`) = `r ftop_all_aov$anova_table[1,4]`, *p* `r if(ftop_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ftop_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ftop_all_aov$anova_table[2,1]`, `r ftop_all_aov$anova_table[2,2]`) = `r ftop_all_aov$anova_table[2,4]`, *p* `r if(ftop_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ftop_all_aov$anova_table[3,1]`, `r ftop_all_aov$anova_table[3,2]`) = `r ftop_all_aov$anova_table[3,4]`, *p* `r if(ftop_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ftop_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ftop_all_aov$anova_table[3,5])`).
 
@@
 
-Neither the main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* `r if(ptof_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ptof_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* `r if(ptof_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* `r if(ptof_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
+Neither the main effect of Articulation (*F*(`r ptof_all_aov$anova_table[1,1]`, `r ptof_all_aov$anova_table[1,2]`) = `r ptof_all_aov$anova_table[1,4]`, *p* `r if(ptof_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[1,5])`) nor Dynamics (*F*(`r ptof_all_aov$anova_table[2,1]`, `r ptof_all_aov$anova_table[2,2]`) = `r ptof_all_aov$anova_table[2,4]`, *p* `r if(ptof_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ptof_all_aov$anova_table[3,1]`, `r ptof_all_aov$anova_table[3,2]`) = `r ptof_all_aov$anova_table[3,4]`, *p* `r if(ptof_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ptof_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ptof_all_aov$anova_table[3,5])`).
 
@@
 Unlike our prediction, experts did not change the tempo depending on the type of recordings (*Fig \ref{fig:ioi-1}*).
-Neither the main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
+Neither the main effect of Articulation (*F*(`r ioi_all_aov$anova_table[1,1]`, `r ioi_all_aov$anova_table[1,2]`) = `r ioi_all_aov$anova_table[1,4]`, *p* `r if(ioi_all_aov$anova_table[1,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[1,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[1,5])`) nor the main effect of Dynamics (*F*(`r ioi_all_aov$anova_table[2,1]`, `r ioi_all_aov$anova_table[2,2]`) = `r ioi_all_aov$anova_table[2,4]`, *p* `r if(ioi_all_aov$anova_table[2,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.3f', ioi_all_aov$anova_table[2,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[2,5])`) nor the interaction between Articulation and Dynamics was significant (*F*(`r ioi_all_aov$anova_table[3,1]`, `r ioi_all_aov$anova_table[3,2]`) = `r ioi_all_aov$anova_table[3,4]`, *p* `r if(ioi_all_aov$anova_table[3,6] < 0.001){"< 0.001"} else {paste("= ", sprintf('%.2f', ioi_all_aov$anova_table[3,6]))}`, $\eta_G^2$ = `r sprintf('%.3f', ioi_all_aov$anova_table[3,5])`).
 

### 2023-01-31 paper/manuscript.tex
update figure / fix floating pointst
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-Neither the main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) nor Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.336, \(\eta_G^2\) = 0.001).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.081, \(\eta_G^2\) = 0.012) nor Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.34, \(\eta_G^2\) = 0.001).
 
@@
 
-Neither the main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.150, \(\eta_G^2\) = 0.005) nor Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.15, \(\eta_G^2\) = 0.005) nor Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).
 
@@
 Unlike our prediction, experts did not change the tempo depending on the type of recordings (\emph{Fig \ref{fig:ioi-1}}).
-Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.111, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
+Neither the main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.11, \(\eta_G^2\) = 0.010) nor the main effect of Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.086, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002).
 

### 2023-03-21 analysis/demographics/questionnaire.R
remove unrelated comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@ -1,14 +1,14 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------
 filename_q = "./questionnaire.csv"
 
 
-## ----extract, echo = FALSE----------------------------------------------------
+## ----extract, echo = FALSE----------
 # read csv files
 dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
 dt <- dt[-c(1,2),] # exclude test/pilot data
@@ -34,10 +34,10 @@ dt_included <- dt
 print(dt_included)
 
 
-## ----2, echo = FALSE----------------------------------------------------------
+## ----2, echo = FALSE----------------
 data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
 
 
-## ----export, include = FALSE--------------------------------------------------
+## ----export, include = FALSE--------
 knitr::purl("questionnaire.Rmd")
 

### 2023-04-29 analysis/baseline.Rmd
made some plots for my defense
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
 
-```{r ioi-plot, fig.width = 6}
+```{r ioi-plot, fig.width = 10}
 ggboxplot(ioi_trial, "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
@@
 
-```{r seq-ioi, fig.width = 7, fig.height = 2}
+```{r seq-ioi, fig.width = 9, fig.height = 4}
 ioi_seq <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Category, Interval)]
@@
 
-```{r kot-plot,  echo = FALSE, fig.width = 6}
+```{r kot-plot,  echo = FALSE, fig.width = 10}
 ggboxplot(kot_trial[Subcomponent1 == "Legato"], "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
@@
 
+<!-- #### Bar graph for main effect (for presentation purposes) -->
+
+```{r kot-bar-plot, include = FALSE}
+ggbarplot(kot[Subcomponent1 == "Legato" & Articulation != "Baseline"], x = "Articulation", y = "Mean", add = "mean_se", ylab = "KOT (ms)", ylim = c(40, 60), size = 2) + theme(text = element_text(size = 40))
+
+ggbarplot(kot[Subcomponent1 == "Staccato" & Articulation != "Baseline"], x = "Articulation", y = "Mean", add = "mean_se", ylab = "KOT (ms)", ylim = c(-220, -200), size = 2) + theme(text = element_text(size = 40))
+```
+
 ## 3. Sequence plots
 
-```{r seq-kot, fig.width = 7, fig.height = 2}
+```{r seq-kot, fig.width = 9, fig.height = 4}
 kot_seq <- dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Category, Interval)]
@@
 
-```{r vel-plot, echo = FALSE, fig.width = 6}
+```{r vel-plot, echo = FALSE, fig.width = 10}
 ggboxplot(vel_trial[Subcomponent2 == "Forte"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
@@
 
+<!-- #### Bar graph for main effect (for presentation purposes) -->
+
+```{r vel-bar-plot, include = FALSE}
+ggbarplot(vel[Subcomponent2 == "Forte" & Dynamics != "Baseline"], x = "Dynamics", y = "Mean", add = "mean_se", ylab = "KV (0-127)", ylim = c(75, 85), size = 2) + theme(text = element_text(size = 40)) + scale_y_continuous(breaks=seq(75, 85, 5))
+
+ggbarplot(vel[Subcomponent2 == "Piano" & Dynamics != "Baseline"], x = "Dynamics", y = "Mean", add = "mean_se", ylab = "KV (0-127)", ylim = c(55, 65), size = 2) + theme(text = element_text(size = 40)) + scale_y_continuous(breaks=seq(55, 65, 5))
+
+ggbarplot(vel[Subcomponent2 == "Piano" & Articulation != "Baseline"], x = "Articulation", y = "Mean", add = "mean_se", ylab = "KV (0-127)", ylim = c(55, 65), size = 2) + theme(text = element_text(size = 40)) + scale_y_continuous(breaks=seq(55, 65, 5))
+```
+
 ## 3. Sequence plots
 
-```{r seq-vel, fig.width = 7, fig.height = 2}
+```{r seq-vel, fig.width = 9, fig.height = 4}
 vel_seq <- dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Category, RowNr)]
@@
 
-```{r vel-diff-plot, echo = FALSE, fig.width = 6}
+```{r vel-diff-plot, echo = FALSE, fig.width = 10}
 ggboxplot(vel_diff_trial[Subcomponent2 == "FtoP"], "Dynamics", "Mean", color = "Articulation", add = "jitter", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (FtoP)")
@@
 
-```{r seq-vel-diff, fig.width = 7, fig.height = 2}
+```{r seq-vel-diff, fig.width = 9, fig.height = 4}
 vel_diff_seq <- dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Category, Interval)]

### 2023-04-29 analysis/baseline.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------------------------
+## ----setup, include = FALSE---------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -12,7 +12,7 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE----------------------------------------------------
+## ----file, include = FALSE----------------------------------
 data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
 data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
 data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
@@ -23,7 +23,7 @@ data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
 data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 
 
-## ----prep, include = FALSE----------------------------------------------------
+## ----prep, include = FALSE----------------------------------
 # read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
 dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
@@ -70,7 +70,7 @@ dt_vel <- rbind(dt_vel, dt_bl_vel)
 dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
 
 
-## ----ioi----------------------------------------------------------------------
+## ----ioi----------------------------------------------------
 # for each individual
 dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
 dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
@@ -82,17 +82,17 @@ setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
 ioi
 
 
-## ----ioi-plot, fig.width = 6--------------------------------------------------
+## ----ioi-plot, fig.width = 8--------------------------------
 ggboxplot(ioi_trial, "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
 
 
[…]

### 2023-07-06 paper/manuscript.Rmd
correct title mistake for figure
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

@@
   geom_point(position = position_jitterdodge(jitter.width = 0.25), alpha = 0.5) +
-  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoP") + scale_y_continuous(breaks = seq(-10, 40, 10), limits = c(-10, 40)) +
+  labs(y = "KV Difference (-127-127)", title = "KV Difference - PtoF") + scale_y_continuous(breaks = seq(-10, 40, 10), limits = c(-10, 40)) +
   scale_color_aaas() +

### 2023-07-06 analysis/baseline.R

@@ -1,4 +1,4 @@
-## ----setup, include = FALSE---------------------------------
+## ----setup, include = FALSE---------------------------------------------------
 # set chunk option
 knitr::opts_chunk$set(echo = FALSE)
 
@@ -12,7 +12,7 @@ if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
 if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
 
 
-## ----file, include = FALSE----------------------------------
+## ----file, include = FALSE----------------------------------------------------
 data_ioi = "preprocessor/trimmed/dt_ioi_trimmed.txt"
 data_kot = "preprocessor/trimmed/dt_kot_trimmed.txt"
 data_vel = "preprocessor/trimmed/dt_vel_trimmed.txt"
@@ -23,7 +23,7 @@ data_bl_vel = "preprocessor/predata/trimmed/dt_vel_trimmed.txt"
 data_bl_vel_diff = "preprocessor/predata/trimmed/dt_vel_diff_trimmed.txt"
 
 
-## ----prep, include = FALSE----------------------------------
+## ----prep, include = FALSE----------------------------------------------------
 # read data
 dt_ioi <- fread(data_ioi, header = T, sep = ",", dec = ".")
 dt_kot <- fread(data_kot, header = T, sep = ",", dec = ".")
@@ -70,7 +70,7 @@ dt_vel <- rbind(dt_vel, dt_bl_vel)
 dt_vel_diff <- rbind(dt_vel_diff, dt_bl_vel_diff)
 
 
-## ----ioi----------------------------------------------------
+## ----ioi----------------------------------------------------------------------
 # for each individual
 dt_ioi$Articulation <- factor(dt_ioi$Articulation, c("Present", "Absent", "Baseline"))
 dt_ioi$Dynamics <- factor(dt_ioi$Dynamics, c("Present", "Absent", "Baseline"))
@@ -82,17 +82,17 @@ setorder(ioi, "SubNr", "Articulation", "Dynamics", "Category")
 ioi
 
 
-## ----ioi-plot, fig.width = 8--------------------------------
+## ----ioi-plot, fig.width = 10-------------------------------------------------
 ggboxplot(ioi_trial, "Articulation", "Mean", color = "Dynamics", add = "jitter", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
 
 
-## ----ioi-all------------------------------------------------
[…]

### 2023-07-06 analysis/demographics/questionnaire.R

@@ -1,14 +1,14 @@
-## ----setup, include = FALSE---------
+## ----setup, include = FALSE---------------------------------------------------
 # packages
 # data manipulation
 if (!require("data.table")) {install.packages("data.table"); require("data.table")}
 
 
-## ----file, include = FALSE----------
+## ----file, include = FALSE----------------------------------------------------
 filename_q = "./questionnaire.csv"
 
 
-## ----extract, echo = FALSE----------
+## ----extract, echo = FALSE----------------------------------------------------
 # read csv files
 dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
 dt <- dt[-c(1,2),] # exclude test/pilot data
@@ -34,10 +34,10 @@ dt_included <- dt
 print(dt_included)
 
 
-## ----2, echo = FALSE----------------
+## ----2, echo = FALSE----------------------------------------------------------
 data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
 
 
-## ----export, include = FALSE--------
+## ----export, include = FALSE--------------------------------------------------
 knitr::purl("questionnaire.Rmd")
 
